### 2021-08-07 dft/notes/notes.tex
auto push

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prb,aps,11pt,superscriptaddress,floatfix]{revtex4-2} 
\usepackage{graphicx}
\usepackage{color} 
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Modelling crystals from first principles}

\author{Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

\begin{abstract}
This is a \textbf{work-in-progess} document intended as an introduction to density functional theory (DFT) for my research group. The idea is to make this an approachable document designed to help my experimentalist-peers become more familiar with the methods and lingo used in computational condensed matter physics. The \emph{flavor} of this project will be lattice dynamics, which is what we study. For my own benefit, I also plan to include the technical details of building a full-potential linearized augemented-projector-wave code (FLAPW or just LAPW) similar to the \textsc{elk} FORTAN code. I am working on that in parallel to writing this document, so this text will be closely linked with my code.
\end{abstract}

\maketitle

%\listoffigures
%\listoftables
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

This document is a work-in-progess project intended to introduce my research group to density functional theory (DFT). The idea is to make this an approachable document designed to help my experimentalist-peers become more familiar with the methods and lingo used in modern computational physics. The \emph{flavor} of this project will be lattice dynamics, which is what we study in my research group. To that end, I plan to cover both finite-difference and density functional perturbation theory (DFPT) methods. For my own benefit, I also plan to include the technical details of building a full-potential linearized augemented-projector-wave code (FLAPW or just LAPW) similar to the \textsc{elk} FORTAN code. I am working on that in parallel to writing this document, so this text will be closely linked with my code.

My end goal is to have a code that can find a ground-state potential that is then frozen and new model-Hamiltonian style potentials can be added; i.e. instead of fitting a tight-binding model to the DFT bands, simply start from a rigid potential and add new terms. This might work as a simple modification to a non-self-consistent (NSCF) calculation or might be more involved. I don't know yet. Furthermore, it may not work at all and I may write a code that is designed to fit tight-binding models. It may make more sense to fit the potential rather than the bands anyway ... I don't know this yet either. To that end, my code will probably depend heavily on reading the potential and wave functions from other DFT packages (e.g. \textsc{elk}, \textsc{abinit}, \textsc{vasp}, ...) since these are much faster and more feature rich than my code ever will be. I am mainly writing the DFT part to get a better grasp of the methods and implementation because I hope to work on one of the packages one day!

A broad outline (\emph{to be updated as I go!}) is as follows (note, this is NOT a table of contents but merely what I plan to put in this document):
\begin{enumerate}
  \item \textbf{What is density-functional-theory}: The Hohenberg-Kohn theorems. The Hohenberg-Kohn theorems prove that the ground-state density of a many-body system can be mapped to a non-interacting problem with the many-body effects included in a mean-field potential (that is an explicit functional of the ground state density). A second theorem also proves that the ground-state density that minimizes the total energy is the true ground-state density and is the only density that does it. \emph{Refs: "Modern Condensed Matter Physics" by Girvin and Yang, "Computational Physics" by Jos Thijssen, "Electronic Structure" by Richard Martin. Probably also the Hohenberg and Kohn papers.}
  \item \textbf{DFT in practice}: the Kohn-Sham Hamiltonian. The Kohn-Sham (KS) Hamiltonian, also called the KS equations is a set of independent-particle Schr\"odinger like equations that can derived using the Hohenberg-Kohn theorems. These equations provide the starting point for building a DFT code. Typically a variational basis set is assumed and the KS equations are solved, varying the expansion coefficients until the energy is minimized. \emph{"Computational Physics" by Jos Thijssen, Paolo Gianozzi's lecture notes. Probably need to look at Goldstein and Afken-Weber-Harris to get the Largrange multiplier part right. I think "Modern Consdensed Matter Physics" has an appendix on functional derivatives.}

  \item \textbf{DFT for atoms}: a fundamental task in all (well, most) DFT codes is solving the radial Schr\"odinger's equation for a spherically symmetric potential. This is the case for atoms, etc. I think this is really only necesssary in muffin-tin type codes (e.g. LAPW, PAW, and USPP methods) and not in pseudopotential (PP) codes... but I am not sure. I will revisit this claim when I get to it. I also should introduce pseudopotentials and other methods here. \emph{Refs: "Computational Physics" by Jos Thijssen, "Electronic Structure" by Richard Martin.}
  \item \textbf{A brief review of of electronic structure theory}: DFT methods for crystals, like any electronic structure method for periodic external potentials, make extensive use of Bloch's theorem. Use of Bloch's theorem results in it being necessary only to solve the KS equations in the first Brillouin zone of reciprocal space (as opposed to real space in e.g. quantum chemical or molecular methods). Since crystals are assumed to extend over \emph{all} of space, it intractable to solve the equations in real space. Instead, the whole problem is formulated in reciprocal space and the first Brillouin zone is sampled on a fine grid of $\pmb{k}$ points. The total energy $E$ is calculated by integrating the Brillouin zone over all occupied KS eigenvalues $\epsilon_{n}(\pmb{k})$ \emph{Refs: "Computational Physics" by Jos Thijssen, Vanderbilt's "Berry Phase in Electronic Structure" chapter on Electronic Structure Theory, David Tong's lecture notes..}
  \item \textbf{DFT for crystals}: the KS equations in a crystal potential. The natural basis set for a reciprocal space code is plane waves (PW). However, due to the requirement of orthogonality of the KS Bloch states $\psi_n(\pmb{k})$, an impractically high number of PWs is required to achieve reasonable precision (remember, the basis set is variational!). Instead, two methods are usually used: 
    \begin{enumerate}
      \item In the core-region close to the atomic nuclei, local orbitals are used instead of PWs (which are still used in between nuclei), and the PWs are matched to the local orbitals are the core-region boundaries. Thes is the idea behind the APW and LAPW methods as well the old-school, less sophisticated OPW and muffin-tin methods. 
      \item Instead of modifying the wave-functions, we can modify the \emph{potential}. This is the idea behind the pseudopotential (PP) method. In the PP method, the potential due to the density of the core states, as seen by the valence states, is transformed away so that, outside of some specified cutoff region, the valence wave functions are still the same as they would be if the potential were not modified. In the core-region, the potential is such that the wave function is as smooth as possible, reducing the number of PWs required to get good convergence. 
    \end{enumerate}
    It is also possible to combine both, as in the PAW or mixed-basis pseudopotential methods (cite Rolf Heid's work). The wave functions are transformed \emph{and} the potential is modified so that as few plane waves as possible are required. Generally, the number of PWs required by the methods applying local orbitals are much lower. This is favorable as the KS equation is still expressed in terms of the PWs: the matrix elements are just calculated using the transformation to local orbitals. With fewer basis functions, diagonalizing the Hamiltonian (an operation that scales $\mathcal{O}(N^3)$ is much easier. However, in the case of all electron (AE) methods like LAPW, there are many states present to treat so the computational effort is still large. However, AE methods are the most accurate methods available for crystals. On the otherhand, frozen-core methods like PAW benefit from both fewer basis functions and fewer states in the secular problem. In contrast, PP methods require more basis functions but the codes are much easier to write resulting in more properties being available from PP methods (though PAW and LAPW codes calculating the same properties usually come around eventually). Additionally, pure PWs are a good starting basis for more advanced many-body perturbation methods so PP methods are sometimes used as starting points for those codes. \emph{Refs: "Computational Physics" by Jos Thijssen, "Electronic Structure" by Martin, "Plane Waves and Pseudopotentials" by Singh and Nordstrom, Paolo Gianozzi's lecture notes.}
  \item \textbf{Review of lattice dynamical calculations}: Here, I will review how the lattice dynamical equations are derived and solved. Ultimately, the goal of a DFT code applied to the lattice dynamical problem is to calculate the inter-atomic force constants (IFCs) as done in the finite-difference method or to calculate the dynamical matrics (which are the space Fourier transform of the IFCs) as done in the DFPT method. \emph{Refs: "Lattice Dynamics" by Dove and "Solid State Physics" by Ashcroft and Mermin. Girvin and Yang do a good job too.}
  \item \textbf{Finite-diffence}: It is a looong-way away to write this stuff... lets finish the DFT and lattice dynamical calculation parts first. 
  \item \textbf{DFPT}: It is a looong-way away to write this stuff... lets finish the DFT and lattice dynamical calculation parts first. \emph{Refs: "Phonons and related crystal properties from density-functional perturbation theory" by Baroni and Paul Tulip's dissertation "Dielectric and Lattice Dynamical Properties of Molecular Crystals via Density Functional PerturbationTheory: Implementation within a First Principles Code" (this thesis does a good job reviewing Gonze's papers deriving the DFPT equations and discusses the implementation of DFPT in the PWPP code \textsc{CASTEP}. Also Rolf Heid's notes in Pavarini's Correlated-Electons compilations.)}
  \item \textbf{Electron-phonon coupling}: Second quantization of phonons etc. It is a looong-way away to write this stuff... lets finish the DFT and lattice dynamical calculation parts first. \emph{Refs: "Electron-phonon interactions from first principles" by Feliciano Giustino. Also Rolf Heid's notes in Pavarini's Correlated-Electons compilations.}
\end{enumerate}

That is enough rambling for now. Let's get to work!

\section{scattering from atoms}

We will need to know how to solve the Schr\"odinger's equation for arbitrary (but well behaved!) spherically symmetrical potenials. The goal later on will be to apply this to solving the potential in \emph{core} regions near nuclei in crystals. However, we aren't there yet and numerically solving the radial Schr\"odinger's equation is a good warm-up to computational physics. So in this section, we will look at a problem we don't care that much about but that gives us a nice example of the connection between computational physics and experiment.

We will study the elastic scattering of (spinless) quantum particles from spherically symmetric potential. We will, in a sense, assume that we already know the scattering potential $V(\pmb{r}) \equiv V(r)$ so that the problem is reduced simply to solving the radial Schr\"odinger's equation:
\begin{equation}
  H\psi(\pmb{r})= \left[ \frac{\hbar^2}{2m} \nabla^2 + V(r)\right]\psi(\pmb{r}) = E \psi(\pmb{r})
\end{equation}

This is useful in numerous circumstances: e.g. to solve the KS equations for core-states in a DFT code, for building pseudopotentials, or for studying a real scattering potential in an experiment. In the later case, the resulting eigenvectors $\psi$ and eigenvalues $E$ are used to model measured spectra and the parameters used to to define the potential $V(r)$, are varied to fit the experiment. We will model the scattering of hydrogen off krypton for an example.

As we will see, solving the spherically symmetric Schr\"odinger's equation ultimately results in numerically solving a 1-dimensional Schr\"odinger's equation. But before we dive into the numerical solution, lets review the analytical solution for simple cases.

In spherical coordinates, the Schr\"odinger's reads:
\begin{equation}
  -\frac{\hbar^2}{2m} \left[  \frac{1}{r^2} \frac{\partial}{\partial r}  \left( r^2 \frac{\partial \psi }{\partial r} \right) 
 + \frac{1}{r^2 \sin\theta} \frac{ \partial}{ \partial \theta} \left( \sin \theta \frac{ \partial \psi }{ \partial \theta} \right) 
  +  \frac{1}{r^2 \sin^2\theta} \left( \frac{\partial^2 \psi}{\partial \phi^2} \right) \right]+V(r)\psi=E\psi
  \label{eq:sph_se}
\end{equation}

We can seperate the solution into a radial part and a spherical part: $\psi(\pmb{r})=R(r)Y(\theta,\phi)$. Now we insert this ansatz into eq. \ref{eq:sph_se} and seperate the radial and angular parts to get:
\begin{equation}
  \Big\{ \frac{1}{R} \frac{d}{dr} \left( r^2 \frac{dR}{dr} \right) - \frac{2mr^2}{\hbar^2} \left[V(r)-E \right] \Big\} + \frac{1}{Y} \Big\{ \frac{1}{\sin \theta} \frac{\partial}{\partial \theta} \left( \sin \theta \frac{\partial Y}{\partial \theta} \right) + \frac{1}{\sin^2 \theta}\frac{\partial^2 Y}{\partial \phi^2} \Big\}=0
\end{equation}
Since each part is independent of the other, for the right hand side to be true, they must both be equal to $\pm$ the same constant. Anticipating the solution, lets call the constant $l(l+1)$. Then we need to solve the two inhomogeneous equations seperately.

The radial equation is
\begin{equation}
  \frac{1}{R} \frac{d}{dr} \left( r^2 \frac{dR}{dr} \right) - \frac{2mr^2}{\hbar^2} \left[V(r)-E \right] = l(l+1)
  \label{eq:radial}
\end{equation}
and the angular equation is
\begin{equation}
  \frac{1}{Y} \Big\{ \frac{1}{\sin \theta} \frac{\partial}{\partial \theta} \left( \sin \theta \frac{\partial Y}{\partial \theta} \right) + \frac{1}{\sin^2 \theta}\frac{\partial^2 Y}{\partial \phi^2} \Big\}=-l(l+1)
  \label{eq:angular}
\end{equation}

\subsection{The Angular Equation}

The solution to the angular equation follows the standard prescription in all kinds of books and, since the external potential $V(r)$ doesn't appear in the angular part, isn't particularly useful to repeat here. Instead, we just quoute the answer: $Y(\theta,\phi)\equiv Y_l^m(\theta,\phi)\equiv|l,m\rangle$ are the well-known spherical harmonics (see Sakurai's book). Note that they depend on $l$ which shows up in the seperation constant $l(l+1)$ that we chose before; we chose this since it is the eigenvalue of $\pmb{L}^2|l,m\rangle = l(l+1)\hbar^2|l,m\rangle$ which we would have found solving the angular equation anyway.

There is more than one way to define spherical harmonics, but the convention we will use is that the spherical harmonics are normalized and satisfy $\langle l',m'|l,m\rangle=\delta_{l,l'}\delta_{m,m'}$. (See e.g. \emph{Classical Electrodynamics} by Jackson for lots of details). We will quote other properties of the spherical harmonics as we go. \textcolor{red}{I might come back here and update this section if I have to.}

\subsection{The Radial Equation}

The radial equation is the interesting bit since, as long as the potential is spherically symmetrical, the Schr\"odinger's equation is seperable and the angular part is always spherical harmonics. The potential $V(r)$ affects only radial part of the wave function. Let us change variables (again...) and let $u(r)=rR(r)$. Then the radial equation becomes
\begin{equation}
  -\frac{\hbar^2}{2m}\frac{d^2u}{dr^2}+\left[ V+\frac{\hbar^2}{2m}\frac{l(l+1)}{r^2}\right] u=Eu
  \label{eq:radial_u}
\end{equation}

As promised, this equation is equivalent to a 1-dimensional Schr\"odinger's equation with wave function $u$ and effective potential $V^l_{eff}(r)=\left[V(r)+\frac{\hbar^2}{2m}\frac{l(l+1)}{r^2}\right]$. The supercript $l$, which I will drop from now on unless I need it again for some reason, is to denote the dependence on angular momentum. The $\propto l(l+1)/r^2$ term is called the \emph{centrigual term} since it tends to repel particles outwards. Note that since the spherical harmonics are (ortho)-noramlized, $u(r)$ satisfies $\int_0^{\infty} |u(r)|^2 dr=1$.

Without specifying properties of the potential, this is as far as we can go. However, if we make the assumption $r^2V(r)\rightarrow 0$ as $r\rightarrow 0$, (i.e. the potential diverges more slowly than the centrigual term), then for small $r$, $\frac{d^2u}{dr^2} \sim \frac{l(l+1)}{r^2}$. The solution is $ u \sim A r^{l+1}+Br^{-l}$. To avoid diverging wave functions (and probabilities!), we have to discard the $\sim r^{-l}$ solution so that $u \propto r^{l+1}$. Recalling that $R=u/r$, we see that $R\propto r^l$ which means that aside from $s$ angular momentum states, larger angular momentum tends to repel the wave function from the origin (or core, scattering center, etc.). On the otherhand, if we look far from the origin (assuming the potential is localized), we find $\frac{d^2u}{dr^2} \sim \kappa^2 u$ where $\kappa^2=2mE/\hbar^2$. To conserve probability, the solution is $u\propto \exp(-\kappa r)$.

\subsection{Example of Spherical Potentials}

Now let us look at some examples of spherical potentials. We will study no-potential ($V(r)=0$, i.e. the free particle) and the infinite spherical well ($V(r)=\infty$ for $r \geq a$). Let us look at the free particle case first. Inspired by the free particle in cartesian coordinates, we can define $E=\hbar^2 k^2/2m$ and change variables so that $\rho=kr$. The radial equation for $u$ becomes 
\begin{equation}
  \frac{d^2R}{d\rho^2} + \frac{2}{\rho}\frac{dR}{d\rho}+\left[1-\frac{l(l+1)}{\rho^2}\right] R=0
  \label{eq:free_particle_spherical}
\end{equation}

This is the \emph{Bessel equation} and the solutions are available willy-nilly in books or you can solve it yourself like I did in the math-methods class (see e.g. \emph{Advanced mathematical methods for scientists and engineers} by Bender and Orszag for how!). The solutions in this context are called the \emph{spherical Bessel functions} and are 
\begin{equation}
  j_l(\rho)=(-\rho)^l\left[\frac{1}{\rho} \frac{d}{d\rho} \right]^l \left(\frac{\sin \rho}{\rho} \right)
\end{equation}
and 
\begin{equation}
  n_l(\rho)=-(-\rho)^l\left[\frac{1}{\rho} \frac{d}{d\rho} \right]^l \left(\frac{\cos \rho}{\rho} \right)
\end{equation}

As $\rho \rightarrow 0$, $j_l(\rho)\rightarrow \rho^l$ and $n_l(\rho)\rightarrow \rho^{-l-1}$ so that only the $j_l$ solution is kept if the origin is included in the problem. For the infinite spherical well, we require that $j_l(ka)=0$, which gives $E_{l=0}= \frac{\hbar^2}{2ma^2}\left[\pi^2,(2\pi)^2,...\right]$. For higher $l$, a computer can be used to invert the Bessel functions to fing e.g. $E_{l=1}= \frac{\hbar^2}{2ma^2}\left[(4.49)^2,(7.73)^2,...\right]$.

\subsection{Numerical Solution to the Radial Schr\"odinger's Equation}

We could go on to solve the hydrogen atom and other wacky potentials by hand like is done in quantum textbooks, or we could move on to something that isn't boring and solve the radial equation on a computer. Let's do that!

\subsection{Further reading}

\begin{itemize}
  \item \emph{Introduction to Quantum Mechanics} by David Griffiths
  \item \emph{Modern Quantum Mechanics} by Jun Sakurai
  \item \emph{Computational Physics} by Jos Thijssen
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ref}

\end{document}

### 2021-08-09 dft/programs/numerov_1d.py
asdasd

@@ -1,0 +1,73 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# --------------------------------------------------------------------------------------------------------
+
+class radial_equation:
+
+    """
+    set up radial equation and integrate numerical to get the solution.
+    """
+
+    # ----------------------------------------------------------------------------------------------------
+
+    def __init__(self,ang_mom_l,input_energy,radial_grid_step,radial_grid_max,radial_grid_min=0):
+
+        """
+        get and set the input options
+        """
+        
+        # set the variables
+        self.ang_mom_l = ang_mom_l
+        self.input_energy = input_energy
+        self.radial_grid_step = radial_grid_step
+        self.radial_grid_max = radial_grid_max
+        self.radial_grid_min = radial_grid_min # defaults to the origin, r=0.
+
+        # set up the radial grid
+        self._set_radial_grid()
+
+    # ----------------------------------------------------------------------------------------------------
+
+    def _set_radial_grid(self):
+
+        """
+        set up the grid
+        """
+
+        # the radial grid
+        self.radial_grid = np.arange(self.radial_grid_min,self.radial_grid_max+self.radial_grid_step,
+                              self.radial_grid_step)
+        self.num_grid_points = self.radial_grid.shape[0]
+
+        print(self.num_grid_points)
+
+    # ----------------------------------------------------------------------------------------------------
+
+    def numerov(self):
+
+        """
+        integrate the radial equation using numerovs method
+        """
+
+        pass
+
+    # ----------------------------------------------------------------------------------------------------
+
+
+# ---------------------------------------------------------------------------------------------------------
+
+if __name__ == "__main__":
+
+    ang_mom_l = 0
+    input_energy = 1
+    radial_grid_step = 0.01
+    radial_grid_max = 10
+
+    rad_eq = radial_equation(ang_mom_l=ang_mom_l,input_energy=input_energy,radial_grid_step=radial_grid_step,
+                            radial_grid_max=radial_grid_max)
+
+
+
+
+

### 2021-08-20 model_builder/notes/notes.tex
asdasd

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prb,aps,11pt,superscriptaddress,floatfix]{revtex4-2} 
\usepackage{graphicx}
\usepackage{color} 
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Tight Binding Notes}

\author{Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

\begin{abstract}
This is a \textbf{work-in-progess} document intended as a clean way to keep notes while writing my tight binding (TB) code. I am writing a \textsc{python} software package that will be used build tight binding models by fitting to density functional theory (DFT) outputs. It can then be used to solve for the valence band energies at arbitrary points in the Brillouin zone and (eventually) to build electron-phonon Hamiltonians using first-principles electron-phonon coupling parameters. But first, I need to figure out how to write the TB code; that's what these notes are for!
\end{abstract}

\maketitle

%\listoffigures
%\listoftables
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

These notes are intended to track my progress and record the relevant theory whie building a tight binding (TB) code. I am writing a \textsc{python} software package that will be used build tight binding models by fitting to density functional theory (DFT) outputs. It can then be used to solve for the valence band energies at arbitrary points in the Brillouin zone and (eventually) to build electron-phonon Hamiltonians using first-principles electron-phonon coupling parameters. The code will be \emph{modular} in the sense that a script will return the Hamiltonian that can be solved at arbitrary \textbf{k}-points. The Hamiltonian can be put into other scripts that tune the parameters to fit density functional theory (DFT) outputs, add more model terms (e.g. Hubbard U or spin-orbit coupling), calculate properties from the energies, etc. It should be more-or-less plug and play.

First-things-first, I need to actually figure out how to build the TB part of the code. I need matrix elements between orbitals in a unitcell. I can then transform the Hamiltonian to \textbf{k}-space, diagonalize to get the energies and eigenvalues, and profit. I want it to work with very little user input: mainly just give positions of atoms, what orbitals are valence on each, and what other atoms they couple to. The code should then fit the energies to DFT outputs, write the parameters, then return a model that can be used willy-nilly. The hard part as of now is calculating the matrix elements for arbitrary angular momentum. I need to figure out how to represent the matrix elements between arbitrary orbitals. It seems like the most reasonable way to implement in a code is to use rotation matrices and rotate orbitals to be quantized along a common z-axis; i.e. the vector that seperates them. Then their matrix elements are simpler and have a common form that can be tabulated for all $Y^\dagger_{l',m'}Y_{l,m}$ pairs. I also need the radial functions, but these will probably be Gaussians and thier integrals are easy.

I plan to fit the hopping parameters to DFT calculations: either to band-structures or, somehow, to the potential. Fitting to the potential seems optimal since we dont have to care about entangled bands or band crossings, etc. We just set up and diagonalize the Hamiltonian and get the bands and calculate stuff from them. However, the potential from a DFT code usually includes the core and exchange-correlation (XC) potential which are (generally) not present in TB. Since we are only including the valence bands, we would have to subtract the core and XC from the total DFT potential OR maybe include the core potential in the TB calculation. Alternatively, we just fit the bands which is the usual technique. This is the simplest way: we just define the orbtials we care about, solve the model to get the valence bands, and fit these to the DFT bands. Either way, this code needs to set up the model, return it, and different modules can be used to solve and fit it to different DFT calculations.

A broad outline (\emph{to be updated as I go!}) is as follows (note, this is NOT a table of contents but merely what I plan to put in this document):

\begin{enumerate}
  \item \textbf{Introduction to tight binding}: Introduce the tight-binding method, derive the equations, etc.
  \item \textbf{tight binding matrix elements}: Derive the tight binding matrix elements for arbitrary angular momentum. We will use the general rotation matrix formalism valid for any angular momentum.
  \item \textbf{Example?}: I dunno
\end{enumerate}

That is enough rambling for now. Let's get to work!

\section{Introduction to tight binding}

The idea of tight binding (TB) is to have a minimal model that is easy enough to solve and that we can still gain physical intuition from. Originally, Slater and Kosters' idea was that a model should easily solvable by hand. Now that we have modern computers, we really just want a model that can very quickly solved on a consumer PC. Regarding physical intuition, we don't really care that much about the rigour of our model so much as that the terms in the Hamiltonian are physically transparent (e.g. Hubbard U) and that they can be rapidly added to a code and, tuned, and many calculations performed. This is, in principle, possible with DFT. For example, Hubbard-U like corrections to DFT are already availble. However, the terms are \emph{not} transparent but are rather black-box like. Similarly implementation is \emph{not} straightforward and running many calculations is \emph{not} cheap.

To keep TB as transparent and cheap as possible, we have to accept it for what it is: a cheap and simple approximation scheme (Slater and Koster called it an 'interpolation' technique, but I don't really care about interpolating band structure ). To that end, we ignore all the terms that add complexity (e.g. 3-center integrals) and simply treat the remaining terms as fitting parameters that are tuned to best reproduce some physical properties: e.g. DFT band structures, experimental optical gaps, etc.

We also work in real space (atleast, we parameterize our Hamiltonian in real space) so that we can use localized, atomic-like orbitals that we are comfortable with. We usually assume some radial function times a spherical harmonic. This is transparent in the sense that we can easily investage angular momentum related properties like spin-orbit coupling. More over, most correlated calculations focus on highly localized states like $d$ electrons.

Our starting point is atomic-like orbtials $\phi_{\{i,l,m\}}(\pmb{r}-(\pmb{R}_p+\pmb{\tau}_i))$. $\phi$ is an atomic-like orbital which we will take to be the product of a localized radial function $\chi_l$ and a spherical harmonic $Y_{l,m}$. The indices $l$ and $m$ denote the orbital angular momentum quantum numbers ($l(l+1)\hbar^2$ is the eigenvalue of $\pmb{L}^2$ and $m\hbar$ is the eigenvalue of $L_z$). $i$ labels which atom in the unitcell we are looking at. The curly brace $\{i,l,m\}$ is to indicate that a particular $l,m$ orbtital is associated with atom $i$. $\pmb{R}_p$ is a lattice vector with $p$ indicating the unitcell and $\pmb{\tau}_i$ is the basis vector of atom $i$ relative to the unitcell origin.

The Hamiltonian can written:
\begin{equation}
  \hat{H}=  \frac{\hbar^2}{2m} \nabla^2 + V(\pmb{r})
\end{equation}
In a crystal, the potential satisfies $V(\pmb{r})=V(\pmb{r}+\pmb{R}_p)$ and we know that the solutions of this Hamiltonian are Bloch functions.

\section{Further reading}

\begin{itemize}
  \item \emph{Modern Quantum Mechanics} by Jun Sakurai
  \item \emph{Computational Physics} by Jos Thijssen
  \item \emph{Berry Phases in Electronic Structure Theory} by David Vanderbilt
  \item \emph{Electronic Structure: Basic Theory and Practical Methods} by Richard Martin
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ref}

\end{document}

### 2021-08-20 model_builder/mod_hamiltonian.py

@@ -1,0 +1,51 @@
+# system modules
+import numpy as np
+
+# custom modules
+from mod_utils import raise_error, print_message
+
+# --------------------------------------------------------------------------------------------------
+# (electronic) hamiltonian class
+
+class hamiltonian:
+
+    """
+    hamiltonian takes lattice, atoms, bonds, etc. as arguments and sets up hamiltonian matrix at 
+    whatever requested k-points. returns hamiltonian or diagonalizes it and returns eigenvals, 
+    eigenvectors etc.
+    """
+
+    def __init__(self,lattice):
+
+        """
+        save the classes needed to build the hamiltonian
+        """
+        
+        self.lattice = lattice
+
+    # ----------------------------------------------------------------------------------------------
+
+    def get_hamiltonian(self,k_point=[0,0,0]):
+
+        """
+        return the hamiltonian matrix at a given k-point
+        """
+
+        pass
+
+    # ----------------------------------------------------------------------------------------------
+
+    def get_eigenvals(self,k_point=[0,0,0]):
+
+        """
+        return the eigenvalues and eigenvectors at a given k-point
+        """
+
+        ham_k = self.get_hamitonian(k_point)
+        eigvals, eigvecs = np.linalg.eigh(ham_k)
+        
+    # ----------------------------------------------------------------------------------------------
+
+
+
+

### 2021-08-20 model_builder/mod_lattice.py

@@ -1,0 +1,150 @@
+# system modules
+import numpy as np
+
+# custom modules
+from mod_utils import raise_error
+
+# --------------------------------------------------------------------------------------------------
+# the lattice class
+
+class lattice:
+    
+    """
+    lattice and supercell vectors, reciprocal lattice, basis atom positions, etc.
+    """
+
+    # ----------------------------------------------------------------------------------------------
+
+    def __init__(self):
+
+        """
+        nothing yet
+        """
+
+        pass
+
+    # ----------------------------------------------------------------------------------------------
+
+    def set_prim_vecs(self,prim_vecs,prim_scale=1):
+        
+        """
+        set up primitive vectors and reciprocal lattice
+        """
+
+        try:
+            self.prim_vecs = np.array(prim_vecs)*prim_scale
+        except:
+            message = 'prim_vecs should be a 3x3 list of floats, prim_scale should be a float'
+            raise_error(message)
+        if len(self.prim_vecs.shape) != 2:
+            message = 'prim_vecs should be a 3x3 list of floats'
+            raise_error(message)
+        if self.prim_vecs.shape[0] != 3 and self.prim_vecs.shape[1] != 3:
+            message = 'prim_vecs should be a 3x3 list of floats'
+            raise_error(message)
+        
+        self.r_lat_vecs, self.prim_cell_vol = self._compute_reciprocal_lattice(self.prim_vecs)
+
+    # ----------------------------------------------------------------------------------------------
+
+    def set_basis_positions(self,basis_pos,basis_labels):
+
+        """
+        set atom positions, convert to cartesian coords, and find unique basis labels.
+        """
+
+        try:
+            self.basis_pos = np.array(basis_pos)
+        except:
+            message = 'basis_pos should be a n_atom x 3 list of floats'
+            raise_error(message)
+        if len(self.basis_pos.shape) != 2:
+            message = 'basis_pos should be a n_atom x 3 list of floats'
+            raise_error(message)
+        if self.basis_pos.shape[1] != 3:
+            message = 'basis_pos should be a n_atom x 3 list of floats'
+            raise_error(message)
+        self.num_atoms = self.basis_pos.shape[0]
+
+        if len(basis_labels) != self.num_atoms:
[…]

### 2021-08-20 model_builder/tb_builder.py

@@ -1,0 +1,64 @@
+# system modules
+
+# custom modules 
+import mod_lattice
+import mod_hamiltonian
+
+# --------------------------------------------------------------------------------------------------
+# build a model hamiltonian
+
+def model_init():
+
+    """
+    set up and return a model to do stuff with
+    """
+    
+    # set up the crystal lattice and define atom positions 
+    lattice = mod_lattice.lattice()
+
+    # set primitive vectors
+    prim_scale = 3.929
+    prim_vecs = [[1.000, 0.000, 0.000], # row vectors in angstroms
+                 [0.000, 1.000, 0.000],
+                 [0.000, 0.000, 2.496]]
+    lattice.set_prim_vecs(prim_vecs,prim_scale)
+
+    # set basis atoms 
+    basis_pos = [[0.0, 0.0, 0.0], # row vectors in reduced (crystal) coordinates
+                 [0.5, 0.0, 0.5],
+                 [0.0, 0.5, 0.0]]
+    basis_labels = ['Cu','O','O'] # can be anything but will be used to identify inequivalent sites
+    lattice.set_basis_positions(basis_pos,basis_labels)
+
+    # setup supercell
+    sc_scale = 3.929
+    sc_vecs = [[2.000, 0.000, 0.000], # row vectors in angstroms
+               [0.000, 2.000, 0.000],
+               [0.000, 0.000, 2.496]]
+    lattice.set_supercell(sc_vecs,sc_scale)
+
+
+    # not as sure what to do here. need to set 'orbitals' on each site and define the coupling
+    # between them. dunno how i want to do it.  
+
+
+    # setup hamiltonian
+    hamiltonian = mod_hamiltonian.hamiltonian(lattice)
+
+    return hamiltonian
+
+# --------------------------------------------------------------------------------------------------
+# if not imported as a module, do the stuff above
+
+if __name__ == '__main__':
+
+    model = model_init()
+
+
+
+
+
+
+
+
+

### 2021-08-20 tb_modelling/mod_model.py

@@ -1,0 +1,147 @@
+# system modules
+import numpy as np 
+
+# custom modules
+from mod_utils import print_message, raise_error
+
+
+# -------------------------------------------------------------------------------------------------
+# lattice class
+
+class lattice:
+
+    """
+    holds lattice vectors and methods to do lattice vectory stuff
+    """
+
+    #----------------------------------------------------------------------------------------------
+
+    def __init__(self):
+
+        """
+        nothing yet
+        """
+
+        pass
+
+    # ---------------------------------------------------------------------------------------------
+
+    def set_lat_vectors(self,lat_vecs,lat_scale):
+
+        """
+        set the real space lattice vectors and call method to calculate reciprocal lattice vectors
+        """
+
+        try:
+            self.lat_vecs = np.array(lat_vecs)*lat_scale
+        except:
+            message = 'lattice vectors should be a 3x3 list of floats'
+            raise_error(message)
+        if self.lat_vecs.shape[0] != 3 or self.lat_vecs.shape[1] != 3:
+            message = 'lattice vectors should be a 3x3 list of floats'
+            raise_error(message)
+
+        self.r_lat_vecs, self.lat_cell_vol = self._compute_reciprocal_lattice(self.lat_vecs)
+
+    # --------------------------------------------------------------------------------------------- 
+    # private methods
+
+    def _compute_reciprocal_lattice(self,lat_vecs):
+
+        """
+        convert real space to reciprocal lattice vectors
+        """
+
+        r_lat_vecs = np.zeros((3,3))
+        cell_vol = lat_vecs[0,:].dot(np.cross(lat_vecs[1,:],lat_vecs[2,:]))
+        r_lat_vecs[0,:] = 2*np.pi*np.cross(lat_vecs[1,:],lat_vecs[2,:])/cell_vol
+        r_lat_vecs[1,:] = 2*np.pi*np.cross(lat_vecs[2,:],lat_vecs[0,:])/cell_vol
+        r_lat_vecs[2,:] = 2*np.pi*np.cross(lat_vecs[0,:],lat_vecs[1,:])/cell_vol
+
+        return r_lat_vecs, cell_vol
+
+    # --------------------------------------------------------------------------------------------- 
+
+    def _convert_to_cartesian(self,reduced_position):
+
+        """
+        convert from rduced (crystal) to cartesian coordinates
+        """
+
+        return np.matmul(self.lat_vecs.T,reduced_pos)
+
+    # ---------------------------------------------------------------------------------------------
+
+
[…]

### 2021-08-20 tb_modelling/mod_orbitals.py

@@ -1,42 +1,0 @@
-import numpy as np
-from mod_utils import print_message, raise_error
-
-class orbitals:
-
-    """
-    assign orbtials to each atom in the primitive and supercells and set up hopping elements from 
-    atoms in primitive cell to neighbors in supercell
-    """
-
-    # -----------------------------------------------------------------------------------------------
-
-    def __init__(self):
-
-        """
-        nothing for now
-        """
-
-        pass
-
-    # -----------------------------------------------------------------------------------------------
-
-    # -----------------------------------------------------------------------------------------------
-    # public methods go below here
-    # -----------------------------------------------------------------------------------------------
-
-    def set_orbital_on_atom(self):
-
-        """
-        TO DO
-        """
-
-        pass
-
-    # -----------------------------------------------------------------------------------------------
-
-    # -----------------------------------------------------------------------------------------------
-    # private methods go below here
-    # -----------------------------------------------------------------------------------------------
-
-
-

### 2021-08-20 tb_modelling/model_building.py

@@ -1,0 +1,41 @@
+# system modules
+
+# custom modules
+from mod_model import model
+
+
+# -------------------------------------------------------------------------------------------------
+# return a model hamiltonian
+
+def model_init():
+    
+    """
+    sets up and returns a 'model' hamiltonian
+    """
+
+    tb_model = model()
+
+    # set lattice vectors
+    prim_scale = 3.929
+    prim_vecs = [[1.000, 0.000, 0.000], # row vectors in angstroms
+                 [0.000, 1.000, 0.000],
+                 [0.000, 0.000, 2.496]]
+    tb_model.set_lattice(prim_vecs,prim_scale)
+
+    return tb_model
+
+
+# -------------------------------------------------------------------------------------------------
+# if not being called from another file, do the thing
+
+if __name__ == '__main__':
+
+    cuprate = model_init()
+
+
+
+
+
+
+
+

### 2021-08-20 tb_modelling/tb_modelling.py

@@ -1,70 +1,0 @@
-# system modules
-import numpy as np
-
-# custom modules
-import mod_lattice 
-import mod_orbitals
-
-
-# --------------------------------------------------------------------------------------------------
-# set up the primitive cell of the crystal
-# --------------------------------------------------------------------------------------------------
-
-lattice = mod_lattice.lattice()
-
-prim_scale = 3.929 # scaling factor for primitive vectors 
-prim_vecs = [[1.000,0.000,0.000], # row vectors in angstroms
-             [0.000,1.000,0.000],
-             [0.000,0.000,2.496]]
-lattice.set_primitive_vectors(prim_vecs,prim_scale) # set the primitive lattice vectors
-
-# basis positions in crystal coords
-basis_pos = [[0.000,0.000,0.000], # Cu
-             [0.500,0.000,0.000], # O
-             [0.000,0.500,0.000]] # O
-basis_labels = ['Cu','O','O'] # optional metadata about the atoms
-lattice.set_basis_positions(basis_pos,basis_labels)
-
-
-# --------------------------------------------------------------------------------------------------
-# set up the supercell that includes nearest neighbors
-# --------------------------------------------------------------------------------------------------
-
-sc_scale = 3.929 # scaling factor for supercell
-sc_vecs = [[2.000,0.000,0.000], # row vectors angstroms
-           [0.000,2.000,0.000],
-           [0.000,0.000,2.496]]
-lattice.set_supercell_vectors(sc_vecs,sc_scale) # set the supercell lattice vectors
-
-# supercell positions in crystal coords
-sc_pos = [[0.000,0.000,0.000], # Cu  
-          [0.250,0.000,0.000], # O
-          [0.000,0.250,0.000], # O
-          [0.500,0.000,0.000], # Cu
-          [0.750,0.000,0.000], # O
-          [0.500,0.250,0.000], # O
-          [0.500,0.500,0.000], # Cu
-          [0.750,0.000,0.000], # O
-          [0.500,0.750,0.000], # O
-          [0.000,0.500,0.000], # Cu
-          [0.250,0.500,0.000], # O
-          [0.000,0.750,0.000]] # O
-sc_labels = ['Cu','O','O','Cu','O','O','Cu','O','O','Cu','O','O'] # optional metadata about the atoms
-lattice.set_supercell_positions(sc_pos,sc_labels)
-
-
-# --------------------------------------------------------------------------------------------------
-# find the neighbor lists in the supercell
[…]

### 2021-08-20 tb_modelling/v2/mod_input.py

@@ -1,148 +1,0 @@
-import os
-from mod_utils import raise_error
-
-class invars:
-
-    # ------------------------------------------------------------------------------------------------
-
-    """
-    input variables for the calculation are stored here. defaults are set, input file is parsed and 
-    some checking on varibles is done. 
-    """
-
-    def __init__(self):
-
-        """
-        set defaults and store some other stuff. initialize all variables here
-        """
-        
-        # list of variables with defaults or that arent required
-        self.optional_variables = ['task'] # default
-
-        # list of mandatory variables
-        self.mandatory_variables = ['prim_scale',
-                                    'prim_vectors',
-                                    'num_basis_atoms',
-                                    'basis_atom_positions']
-
-        # allowed variables is union of optional and mandatory variables
-        self.allowed_variables = self.optional_variables+self.mandatory_variables
-
-        # default variables go below here
-        self.task = 'solve_k_grid' # what the code is gonna do
-
-    # ----------------------------------------------------------------------------------------------
-
-    def parse_input_file(self,input_file='input_params'):
-
-        """
-        read the input file 
-        """
-
-        # check if input file exists, if not exit
-        if not os.path.exists(input_file):
-            message = f'input file \'{input_file}\' not found'
-            raise_error(message)
-
-        # get the text from the input file
-        with open(input_file,'r') as f_input:
-            self.input_text = f_input.readlines() # read all the lines
-
-        # now format the lines and check that all mandatory variables exist
-        self._preprocess_input()
-
-        # -----------------------------------------------------------------------------------------
-        # now get the variables. if new vars are added that depend on previous variables, be careful
-        # to add them in the logical order
-        # -----------------------------------------------------------------------------------------
-
-    # -----------------------------------------------------------------------------------------------
-
-    # -----------------------------------------------------------------------------------------------
-    # private methods go below here 
-    # -----------------------------------------------------------------------------------------------
-
-    def _preprocess_input(self):
-
-        """
-        strip all the blank lines, comments, and check that all mandatroy variables are defined. 
-        """
-
-        duplicates = [] # store duplicate variables
-        trimmed_text = [] # stripped of comments and blank lines
-        for line in self.input_text:
-
-            # if comment line or blank, skip. 
-            if len(line.split()) == 0 or line.strip().startswith('#'):
-                continue
-
-            # otherwise, process the line further
-            else: 
-     
-                # strip comment off end of line
-                tmp_line = line.split('#')[0].strip()
-
[…]

### 2021-08-20 tb_modelling/v2/mod_utils.py

@@ -1,25 +1,0 @@
-"""
-various utility stuff like wrappers to print to screen, raise error, etc
-"""
-
-# -----------------------------------------------------------------------------------------------------
-
-def print_stdout(message,msg_type='NOTE'):
-
-    """
-    print to the screen using common formatting
-    """
-    
-    print(f'\n ** {msg_type} **')
-    print(f' {message}\n',flush=True)
-
-# -----------------------------------------------------------------------------------------------------
-
-def raise_error(message):
-
-    print_stdout(message,msg_type='ERROR')
-    exit()
-
-# -----------------------------------------------------------------------------------------------------
-
-

### 2021-08-20 tb_modelling/v2/tbm.py

@@ -1,18 +1,0 @@
-# import system modules
-import numpy as np
-import sys
-
-# import custom modules
-import mod_input
-
-# get input file from command line or use default
-if len(sys.argv) == 1:
-    input_file = 'input_params'
-else:
-    input_file = sys.argv[1]
-
-# read the input file
-invars = mod_input.invars()
-invars.parse_input_file(input_file)
-
-

### 2021-08-20 tb_modelling/v5/mod_atoms.py

@@ -1,0 +1,90 @@
+# system modules
+import numpy as np
+
+# custom modules
+from mod_utils import raise_error
+
+# -------------------------------------------------------------------------------------------------
+
+class _atom:
+    
+    """
+    atom object. it holds each atom's positions, orbitals, etc
+    """
+
+    def __init__(self):
+
+        pass
+
+# -------------------------------------------------------------------------------------------------
+# -------------------------------------------------------------------------------------------------
+
+class atoms:
+
+    """
+    set atom positions, orbitals, etc
+    """
+
+    # ---------------------------------------------------------------------------------------------
+
+    def __init__(self,lattice):
+
+        """
+        hold list of 'atoms' and number of atoms
+        """
+
+        self.lattice = lattice 
+
+        self.basis_atoms = []
+        self.num_basis_atoms = 0
+        self.basis_types = []
+        self.num_basis_types = 0
+
+        self.sc_atoms = []
+        self.num_sc_atoms = 0
+        self.sc_types = []
+        self.num_sc_types = 0
+
+        self.num_atoms = 0
+
+    # ---------------------------------------------------------------------------------------------
+
+    def set_atom(self,reduced_pos,atom_type='none'):
+
+        """
+        creates an instance of '_atom' class and sets some attributes
+        """
+
+        try:
+            self.reduced_pos = np.array(reduced_pos).flatten()
+        except:
+            message = 'reduced position should be a 3 element list of floats'
+            raise_error(message)
+        if self.reduced_pos.size != 3:
+            message = 'reduced position should be a 3 element list of floats'
+            raise_error(message)
+
+        atom = _atom()
+        atom.reduced_pos = reduced_pos
+        atom.atom_type = atom_type
+        
+        if self.reduced_pos[0] >= 1 or self.reduced_pos[1] >= 1 or self.reduced_pos[2] >= 1:
+            atom.is_basis = True
+            atom.atom_index = self.num_atoms
+            self.basis_atoms.append(atom)
+            self.num_atoms = self.num_atoms+1
+            self.num_basis_atoms = self.num_basis_atoms+1
+        else:
+            atom.is_basis = False
+            atom.atom_index = self.num_atoms
+            self.sc_atoms.append(atom)
+            self.num_atoms = self.num_atoms+1
+            self.num_sc_atoms = self.num_sc_atoms+1
+
+    # ---------------------------------------------------------------------------------------------
+
+
+    # ---------------------------------------------------------------------------------------------
+
+
+    # ---------------------------------------------------------------------------------------------

### 2021-08-20 tb_modelling/v5/mod_lattice.py

@@ -1,0 +1,73 @@
+# system modules
+import numpy as np
+
+# custom modules
+from mod_utils import raise_error
+
+# -------------------------------------------------------------------------------------------------
+
+class lattice:
+
+    """
+    stores primitive and supercell lattice vectors, converts atom positions to cartesian coords, 
+    etc. also computes neighbor vectors using minimum image convention
+    """
+
+    # ---------------------------------------------------------------------------------------------
+
+    def __init__(self):
+
+        """
+        nothing yet
+        """
+
+        pass
+
+    # ---------------------------------------------------------------------------------------------
+
+    def set_prim_cell(self,prim_vecs,prim_scale):
+
+        """
+        set the primitive cell lattice vectors and get the reciprocal lattice
+        """
+
+        try:
+            self.prim_vecs = np.array(prim_vecs)*prim_scale
+        except:
+            message = 'primitive vectors should be 3x3 list of floats'
+            raise_error(message)
+        if self.prim_vecs.shape[0] != 3 and self.prim_vecs.shape[1] != 3:
+            message = 'primitive vectors should be 3x3 list of floats'
+            raise_error(message)
+
+        self.r_prim_vecs, self.prim_cell_vol = self._compute_reciprocal_lattice(self.prim_vecs)
+
+    # ---------------------------------------------------------------------------------------------
+
+    def convert_to_cart(self,reduced_position):
+
+        return np.matmul()
+
+    # ---------------------------------------------------------------------------------------------
+    # private methods
+
+    def _compute_reciprocal_lattice(self,lat_vecs):
+
+        """
+        convert real space to reciprocal lattice vectors
+        """
+
+        r_lat_vecs = np.zeros((3,3))
+        cell_vol = lat_vecs[0,:].dot(np.cross(lat_vecs[1,:],lat_vecs[2,:]))
+        r_lat_vecs[0,:] = 2*np.pi*np.cross(lat_vecs[1,:],lat_vecs[2,:])/cell_vol
+        r_lat_vecs[1,:] = 2*np.pi*np.cross(lat_vecs[2,:],lat_vecs[0,:])/cell_vol
+        r_lat_vecs[2,:] = 2*np.pi*np.cross(lat_vecs[0,:],lat_vecs[1,:])/cell_vol
+
+        return r_lat_vecs, cell_vol
+
+    # ---------------------------------------------------------------------------------------------
+
+
+
+
+

### 2021-08-20 tb_modelling/v5/tb_modelling.py

@@ -1,0 +1,54 @@
+# system modules
+import sys
+
+# custom modules
+import mod_lattice
+import mod_atoms
+from mod_utils import raise_error
+
+# -------------------------------------------------------------------------------------------------
+# get which task to do
+
+if len(sys.argv) != 1:
+    try:
+        task = int(sys.argv[1])
+    except:
+        message = f'task \'{task}\' is not one of the allowed options'
+        raise_error(task)
+else:
+    task = 1 # do band structure calculation
+
+# -------------------------------------------------------------------------------------------------
+# set up primitive cell
+
+lattice = mod_lattice.lattice()
+
+prim_scale = 3.929 # scaling factor for primitive vectors
+prim_vecs = [[1.000, 0.000, 0.000], # row vectors in angstroms
+             [0.000, 1.000, 0.000],
+             [0.000, 0.000, 2.496]]
+lattice.set_prim_cell(prim_vecs,prim_scale)
+
+basis_atoms = []
+basis_inds = []
+basis_type = []
+
+sc_atoms = []
+sc_inds = []
+sc_types = []
+
+# -------------------------------------------------------------------------------------------------
+# set atoms in primitive cell
+
+atoms = mod_atoms.atoms(lattice)
+
+reduced_pos = [0.0, 0.0, 0.0]
+atom_type = 'Cu'
+
+atoms.set_atom(reduced_pos,atom_type)
+
+
+
+
+
+

### 2021-08-20 model_builder/modules/m_electron_hamiltonian.py
asdsaasdf

@@ -2,7 +2,7 @@
 import numpy as np
 
 # custom modules
-from mod_utils import raise_error, print_message
+from m_utils import raise_error, print_message
 
 # --------------------------------------------------------------------------------------------------
 # (electronic) hamiltonian class
@@ -15,35 +15,44 @@ class hamiltonian:
     eigenvectors etc.
     """
 
-    def __init__(self,lattice):
+    def __init__(self,lattice,orbitals):
 
         """
         save the classes needed to build the hamiltonian
         """
         
         self.lattice = lattice
+        self.orbitals = orbitals
 
     # ----------------------------------------------------------------------------------------------
 
-    def get_hamiltonian(self,k_point=[0,0,0]):
+    def get_hamiltonian_k(self,k_point=[0,0,0]):
 
         """
         return the hamiltonian matrix at a given k-point
         """
 
+        # this should do the space FT and return the hamiltonian at a given k-point
+        # do something like this
+        #ham_k = self._do_space_ft(k_point)
+        #return ham_k
+
         pass
 
     # ----------------------------------------------------------------------------------------------
 
-    def get_eigenvals(self,k_point=[0,0,0]):
+    def get_eigenvals_k(self,k_point=[0,0,0]):
 
         """
         return the eigenvalues and eigenvectors at a given k-point
         """
 
-        ham_k = self.get_hamitonian(k_point)
-        eigvals, eigvecs = np.linalg.eigh(ham_k)
+        # do something like this
+        #ham_k = self.get_hamitonian_k(k_point)
+        #eigvals, eigvecs = np.linalg.eigh(ham_k)
         
+        pass
+
     # ----------------------------------------------------------------------------------------------
 
 

### 2021-08-20 model_builder/modules/m_lattice.py

@@ -2,7 +2,7 @@
 import numpy as np
 
 # custom modules
-from mod_utils import raise_error
+from m_utils import raise_error
 
 # --------------------------------------------------------------------------------------------------
 # the lattice class

### 2021-08-20 model_builder/modules/m_orbitals.py

@@ -1,0 +1,64 @@
+# system modules
+
+# custom modules
+from m_utils import raise_error
+
+# --------------------------------------------------------------------------------------------------
+# orbitals class. should hold all the orbitals and hopping/onsite params
+
+class orbitals:
+
+    """
+    hold the orbital definitions and hopping/onsite params
+    """
+
+    # ----------------------------------------------------------------------------------------------
+
+    def __init__(self,lattice):
+
+        """
+        copy classes and define set of allowed orbitals
+        """
+
+        # copy the class(es) over
+        self.lattice = lattice
+
+        # a list of the allowed orbital types
+        self.allowed_orbitals = ['s',                                   # the s orbital
+                                 'p_x','p_y','p_z',                     # the p orbitals
+                                 'd_z2','d_xz','d_yz','d_xy','d_x2-y2'] # the d orbitals
+
+        # a dictionary that holds which orbital is on which site
+        self.orbs = {} 
+
+    # ----------------------------------------------------------------------------------------------
+
+    def set_orbital(self,site,n,orb):
+
+        # check that the requested orbital is allowed
+        if orb not in allowed_orbitals:
+            message = f'the orbital type {orb} is not allowed'
+            raise_error(message)
+
+        # check the the requested orbital doesnt already exist
+        handle = f'site_{site}'
+        if handle in self.orbs.keys():
+            print(handle)
+        else:
+            self.orbs[handle]['site'] = site
+            self.orbs[handle]['n'] = n
+            self.orbs[handle]['orb'] = orb
+
+    # ----------------------------------------------------------------------------------------------
+
+    def set_hopping(self):
+
+        # should set the hopping param between the 2 orbitals on 2 sites. they can be the same 
+        # site as long as they are different {l,m}, e.g. like molecular hopping. should only allow
+        # hopping between sites and orbitals that exist and should not allow hopping between the 
+        # same pairs
+
+        pass
+
+    # ----------------------------------------------------------------------------------------------
+

### 2021-08-20 model_builder/tb_builder.py

@@ -1,64 +1,0 @@
-# system modules
-
-# custom modules 
-import mod_lattice
-import mod_hamiltonian
-
-# --------------------------------------------------------------------------------------------------
-# build a model hamiltonian
-
-def model_init():
-
-    """
-    set up and return a model to do stuff with
-    """
-    
-    # set up the crystal lattice and define atom positions 
-    lattice = mod_lattice.lattice()
-
-    # set primitive vectors
-    prim_scale = 3.929
-    prim_vecs = [[1.000, 0.000, 0.000], # row vectors in angstroms
-                 [0.000, 1.000, 0.000],
-                 [0.000, 0.000, 2.496]]
-    lattice.set_prim_vecs(prim_vecs,prim_scale)
-
-    # set basis atoms 
-    basis_pos = [[0.0, 0.0, 0.0], # row vectors in reduced (crystal) coordinates
-                 [0.5, 0.0, 0.5],
-                 [0.0, 0.5, 0.0]]
-    basis_labels = ['Cu','O','O'] # can be anything but will be used to identify inequivalent sites
-    lattice.set_basis_positions(basis_pos,basis_labels)
-
-    # setup supercell
-    sc_scale = 3.929
-    sc_vecs = [[2.000, 0.000, 0.000], # row vectors in angstroms
-               [0.000, 2.000, 0.000],
-               [0.000, 0.000, 2.496]]
-    lattice.set_supercell(sc_vecs,sc_scale)
-
-
-    # not as sure what to do here. need to set 'orbitals' on each site and define the coupling
-    # between them. dunno how i want to do it.  
-
-
-    # setup hamiltonian
-    hamiltonian = mod_hamiltonian.hamiltonian(lattice)
-
-    return hamiltonian
-
-# --------------------------------------------------------------------------------------------------
-# if not imported as a module, do the stuff above
-
-if __name__ == '__main__':
-
-    model = model_init()
-
-
-
-
-
-
-
-
-

### 2021-08-20 model_builder/tb_fit.py

@@ -1,0 +1,99 @@
+# system modules
+import sys
+
+# set path to custom modules
+module_path = './modules'
+sys.path.append(module_path)
+
+# custom modules 
+import m_lattice
+import m_orbitals
+import m_electron_hamiltonian
+
+# --------------------------------------------------------------------------------------------------
+# build a model hamiltonian
+
+def model_init():
+
+    """
+    set up and return a model to do stuff with
+    """
+
+    # ----------------------------------------------------------------------------------------------
+    # set up the crystal lattice and define atom positions 
+
+    lattice = m_lattice.lattice()
+
+    # set primitive vectors
+    prim_scale = 3.929
+    prim_vecs = [[1.000, 0.000, 0.000], # row vectors in angstroms
+                 [0.000, 1.000, 0.000],
+                 [0.000, 0.000, 2.496]]
+    lattice.set_prim_vecs(prim_vecs,prim_scale)
+
+    # set basis atoms 
+    basis_pos = [[0.0, 0.0, 0.0], # row vectors in reduced (crystal) coordinates
+                 [0.5, 0.0, 0.5],
+                 [0.0, 0.5, 0.0]]
+    basis_labels = ['Cu','O','O'] # can be anything but will be used to identify inequivalent sites
+    lattice.set_basis_positions(basis_pos,basis_labels)
+
+    # ----------------------------------------------------------------------------------------------
+    # setup supercell
+
+    sc_scale = 3.929
+    sc_vecs = [[2.000, 0.000, 0.000], # row vectors in angstroms
+               [0.000, 2.000, 0.000],
+               [0.000, 0.000, 2.496]]
+    lattice.set_supercell(sc_vecs,sc_scale)
+
+    # ----------------------------------------------------------------------------------------------
+    # set up class to hold all the orbitals etc.
+
+    orbitals = m_orbitals.orbitals(lattice)
+
+    # now set the orbtials
+    orbitals.set_orbital(site=0,n=3,orb='d_x2-y2')
+    orbitals.set_orbital(site=0,n=3,orb='d_x2-y2')
+    orbitals.set_orbital(site=1,n=2,orb='px')
+    orbitals.set_orbital(site=1,n=2,orb='py')
+    orbitals.set_orbital(site=2,n=2,orb='px')
+    orbitals.set_orbital(site=2,n=2,orb='py')
+
+    # define irreducible set of hoppings between orbitals and onsite energies 
[…]

### 2021-08-25 dft/mixed_basis/notes/notes.tex
asdsaasdf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prb,aps,11pt,superscriptaddress,floatfix]{revtex4-2} 
\usepackage{graphicx}
\usepackage{color} 
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Mixed-basis (semi-empirical) pseudopotential method}

\author{Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}
\affiliation{Center for Experiments on Quantum Materials, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\author{Dmitry Reznik}
\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}
\affiliation{Center for Experiments on Quantum Materials, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

\begin{abstract}
	This is a \textbf{work-in-progess} document where I am keeping notes for a project I am working on. I am writing a density functional theory (DFT) package that will use a combination of atom-centered Gaussian functions and plane-waves (PW) as the basis-set. The atom-cores will be replaced with a \emph{semi-empirical} pseudopotential. Semi-empirical, in this context, means that I will treat the pseudopotentials as fitting parameters. I plan to fit bands calculated from my code to all-electron (or other high quality) DFT calculations. This is really no more 'empirical' than the typical ab-initio pseudopotential method which involves fitting the pseudopotential for the isolated atom to all-electron calculations. The ab-initio pseudopotential is more ab-initio in the sense that it does not depend on the atom's environment and can be 'plugged-in' to any crystal. (Well, not really: there has been considerable effort in making them as \emph{transferrable} as possible). In my case, the \emph{semi-empirical} pseudopotentials will be fit so that the smallest basis set possible can be used in a particular configuration and still reproduce the ab-intio band structure. Think of it like a tight-binding method, but instead of throwing away all the bonus multi-center integrals, I keep everything and work in reciprocal space. The are 2 primary reasons behind this project: (1) I want to learn more about writing a DFT code and there is no better way than writing a DFT code. (2) If this works well and is efficient, the pseudopotentials and even the SCF potential can be plugged into to larger supercell models that include electron-phonon coupling explicity. I hope this works ... 
\end{abstract}

\maketitle

%\listoffigures
%\listoftables
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

This is a \textbf{work-in-progess} document where I am keeping notes for a project I am working on. I am writing a density functional theory (DFT) package that will use a combination of atom-centered Gaussian functions and plane-waves (PW) as the basis-set. This is the so called \emph{mixed-basis pseudopotential method} \cite{louie1979self}.

The atom-cores will be replaced with a \emph{semi-empirical} pseudopotential. Semi-empirical, in this context, means that I will treat the pseudopotentials as fitting parameters. I plan to fit bands calculated from my code to all-electron (or other high quality) DFT calculations. This is really no more 'empirical' than the typical ab-initio pseudopotential method which involves fitting the pseudopotential for the isolated atom to all-electron calculations. The ab-initio pseudopotential is more ab-initio in the sense that it does not depend on the atom's environment and can be 'plugged-in' to any crystal. (Well, not really: there has been considerable effort in making them as \emph{transferrable} as possible). In my case, the \emph{semi-empirical} pseudopotentials will be fit so that the smallest basis set possible can be used in a particular configuration and still reproduce the ab-intio band structure. Think of it like a tight-binding method, but instead of throwing away all the bonus multi-center integrals, I keep everything and work in reciprocal space.

The pseudpotentials will be non-local (i.e. angular momentum $l \neq 0$). Non-local pseudopotentials usually look like:
\begin{equation}
	\hat{V}^{pp}_{\tau} = V_{\tau,0}(|\pmb{r}|)+\sum_{l=1}^{l_c} V_{\tau,l}(|\pmb{r}|) \hat{P}_l
\end{equation}

Here, $l$ denotes the orbital angular momentum quantum number, $\tau$ denotes the atom in the unitcell, and $\hat{P}_l$ is a projector onto angular momentum character $l$. The radial functions $V_{\tau,l}(|\pmb{r}|)$ are spherically symetric. I will probably pick them to be Gaussians too since the matrix elements will be simple. Maybe this will be a bad idea, I don't know yet.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ref}

\end{document}

### 2021-08-25 dft/mixed_basis/m_lattice.py

@@ -1,0 +1,43 @@
+# system modules
+import numpy as np
+
+# custom modules
+from m_utils import raise_error
+
+class lattice:
+
+    def __init__(self):
+
+        pass
+
+    def set_lattice_vectors(self,lat_vecs,lat_scale=1):
+
+        # check that the lattice vectors make sense
+        err_msg = 'lat_vecs should be a 3x3 list of floats'
+        try:
+            self.lat_vecs = np.array(lat_vecs,dtype=float)
+        except:
+            raise_error(err_msg)
+        if len(self.lat_vecs.shape) != 2:
+            raise_error(err_msg)
+        if self.lat_vecs.shape[0] != 3 or self.lat_vecs.shape[1] != 3:
+            raise_error(err_msg)
+
+        # check that the lat_scale makes sense
+        err_msg = 'lat_scale should be a positive, non-zero float'
+        try:
+            lat_scale = float(lat_scale)
+        except:
+            raise_error(err_msg)
+        if lat_scale <= 0: 
+            raise_error(err_msg)
+
+        # now apply the scale
+        self.lat_vecs = self.lat_vecs*lat_scale
+
+        
+
+
+
+
+

### 2021-08-25 dft/mixed_basis/main.py

@@ -1,0 +1,34 @@
+# system modules
+
+# custom modules
+import m_lattice
+import m_atoms
+
+def hamiltonian_init():
+
+    # ----------------------------------------------------------------------------------------------
+    # set up the lattice
+
+    lattice = m_lattice.lattice()
+    
+    lat_vecs = [[0.0, 1.0, 1.0], # row vectors in angstrom
+                [1.0, 0.0, 1.0],
+                [1.0, 1.0, 0.0]]
+    lat_scale = 1.98918 # lattice vector scaling factor
+    lattice.set_lattice_vectors(lat_vecs,lat_scale)
+
+    # ----------------------------------------------------------------------------------------------
+    # set atom positions
+
+    atoms = m_atoms.atoms()
+
+    atom_pos = [[0.0, 0.0, 0.0]]
+
+
+if __name__ == '__main__':
+
+    hamlt = hamiltonian_init()
+
+
+
+

### 2021-08-25 dft/programs/numerov.py

@@ -1,98 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-# ---------------------------------------------------------------------------------------------------------
-
-class numerov_integrator:
-
-    # -----------------------------------------------------------------------------------------------------
-
-    def __init__(self):
-        
-        self.const = 1 # 2m/hbar**2
-
-    # -----------------------------------------------------------------------------------------------------
-
-    def init_grid(self,x_max,x_min=0,dx=0.01):
-
-        self.x_min = x_min
-        self.x_max = x_max
-        self.dx = dx
-        self.x_grid = np.arange(x_min,x_max+dx,dx)
-        self.num_grid = self.x_grid.size
-
-    # -----------------------------------------------------------------------------------------------------
-
-    def numerov(self,pot,x0=0,y0=2,y1=1,xf=8,e_val=1):
-
-        """
-        integrates to the right
-        """
-
-        f = 1+self.const*(e_val-pot)*self.dx**2/12
-
-        self.sol = np.zeros(self.num_grid)
-
-        n = np.argwhere(self.x_grid >= x0).flatten().min()
-        nf = np.argwhere(self.x_grid <= xf).flatten().max()
-
-        n_1 = n+1
-        n_2 = n_1+1
-
-        self.start = n 
-        self.end = nf
-
-        self.sol[n] = y0
-        self.sol[n_1] = y1
-
-        num_to_do = nf-n-2
-        print(num_to_do)
-
-        for ii in range(num_to_do):
-
-            self.sol[n_2] = ((12-10*f[n_1])*self.sol[n_1]-f[n]*self.sol[n])/f[n_2]
-            n = n_1
-            n_1 = n_2
-            n_2 = n_2+1
-
-
-    # -----------------------------------------------------------------------------------------------------
-
-
-# ---------------------------------------------------------------------------------------------------------
-# run the thing
-
-if __name__ == '__main__':
-
-    numerov = numerov_integrator()
-    numerov.init_grid(x_max=10,x_min=-10,dx=0.001)
-
-    k = 1
-    pot = k/2*numerov.x_grid**2
-
-    numerov.numerov(pot,x0=0,y0=20,y1=15,e_val=10,xf=8)
-
-    plt.plot(numerov.x_grid,pot,color='b')
-    plt.plot(numerov.x_grid,numerov.sol,color='r')
-    plt.show()
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2021-08-25 dft/pw_pp/mod_basis.py

@@ -1,79 +1,0 @@
-import numpy as np
-from mod_utils import print_stdout, raise_error
-from mod_constants import hbar_sq_over_2m
-
-
-# ====================================================================================================
-# ----------------------------------------------------------------------------------------------------
-# ====================================================================================================
-
-class pw:
-
-    # ------------------------------------------------------------------------------------------------
-
-    def __init__(self):
-        pass
-
-    # ------------------------------------------------------------------------------------------------
-
-    def set_plane_waves(self,crystal,pw_cut_off):
-
-        # cut off should be a number
-        try:
-            pw_cut_off = float(pw_cut_off)
-        except:
-            message = 'pw_cutoff should a be a floating point number that is larger than 0'
-            raise_error(message)
-
-        # cut off should be larger than 0
-        if pw_cut_off <= 0:
-            message = 'pw_cutoff should a be a floating point number that is larger than 0'
-            raise_error(message)
-
-        self.pw_cut_off = pw_cut_off
-       
-        # find shortest recip. lat vec
-        b0 = self._norm(crystal.r_lat_vecs[0,:])
-        for ii in range(1,3):
-            bi = self._norm(crystal.r_lat_vecs[ii,:])
-            if bi <= b0:
-                b0 = bi
-
-        # find largest n to check
-        n_max = int(1.5*np.sqrt(pw_cut_off/hbar_sq_over_2m)/b0)
-
-        # create list of G to check
-        num_test = (2*n_max+1)**3 # i.e. [-n_max,n_max]*[-n_max,n_max]*[-n_max,n_max]
-        test_G = np.zeros((num_test,3))
-        test_E = np.zeros(num_test)
-    
-        # loop over G vectors and find which ones are below cutoff
-        ind = 0
-        for nx in range(-n_max,n_max+1): # [-n_max,n_max]
-            for ny in range(-n_max,n_max+1): # [-n_max,n_max]
-                for nz in range(-n_max,n_max+1): # [-n_max,n_max]
-                    test_G[ind,:] = [nx,ny,nz]
-                    test_E[ind] = self._norm(nx*crystal.r_lat_vecs[0,:]+ny*crystal.r_lat_vecs[1,:]+
-                                             nz*crystal.r_lat_vecs[2,:])**2 # G**2
-                    ind = ind+1
-        
-        # find which are below cutoff
-        test_E = test_E*hbar_sq_over_2m # KE = hbar**2/2/m * G**2
[…]

### 2021-08-25 dft/pw_pp/mod_crystal.py

@@ -1,105 +1,0 @@
-import numpy as np
-from mod_utils import raise_error, print_stdout
-
-
-# ====================================================================================================
-# ----------------------------------------------------------------------------------------------------
-# ====================================================================================================
-
-class crystal:
-
-    # -------------------------------------------------------------------------------------------------
-
-    def __init__(self):
-        pass
-
-    # -------------------------------------------------------------------------------------------------
-
-    def set_lattice(self,lat_vecs,lat_scale):
-
-        # check that lattice vector matrix is correct dims
-        self.lat_vecs = np.array(lat_vecs)
-        if self.lat_vecs.shape[0] !=3 or self.lat_vecs.shape[1] !=3:
-            message = 'lattice vectors should be a 3x3 matrix'
-            raise_error(message)
-        self.lat_vecs = self.lat_vecs*lat_scale
-
-        # compute reciprocal lattice
-        self._compute_reciprocal_lattice()
-
-        # print lattice info
-        message = 'real space lattice (Angstrom):\n'
-        for ii in range(3):
-            message = message+f'  {self.lat_vecs[ii,0]: 2.4f} {self.lat_vecs[ii,1]: 2.4f}' \
-                    f' {self.lat_vecs[ii,2]: 2.4f}\n'
-        message = message+' reciprocal lattice (1/Angstrom):\n'
-        for ii in range(3):
-            message = message+f'  {self.r_lat_vecs[ii,0]: 2.4f} {self.r_lat_vecs[ii,1]: 2.4f}' \
-                    f' {self.r_lat_vecs[ii,2]: 2.4f}\n'
-        print_stdout(message,msg_type='LATTICE')
-
-    # -------------------------------------------------------------------------------------------------
-
-    def _compute_reciprocal_lattice(self):
-
-        self.r_lat_vecs = np.zeros((3,3))
-        self.cell_vol = self.lat_vecs[0,:].dot(np.cross(self.lat_vecs[1,:],self.lat_vecs[2,:]))
-        self.r_lat_vecs[0,:] = 2*np.pi*np.cross(self.lat_vecs[1,:],self.lat_vecs[2,:])/self.cell_vol
-        self.r_lat_vecs[1,:] = 2*np.pi*np.cross(self.lat_vecs[2,:],self.lat_vecs[0,:])/self.cell_vol
-        self.r_lat_vecs[2,:] = 2*np.pi*np.cross(self.lat_vecs[0,:],self.lat_vecs[1,:])/self.cell_vol
-
[…]

### 2021-08-25 dft/pw_pp/mod_kpoints.py

@@ -1,74 +1,0 @@
-import numpy as np
-from mod_utils import print_stdout, raise_error
-from mod_constants import hbar_sq_over_2m
-
-
-# ====================================================================================================
-# ----------------------------------------------------------------------------------------------------
-# ====================================================================================================
-
-class kpoints:
-
-    # ------------------------------------------------------------------------------------------------
-
-    def __init__(self):
-        
-        self.allowed_reductions = ['full']
-
-    # ------------------------------------------------------------------------------------------------
-
-    def set_grid(self,kpt_grid,grid_reduction='full'):
-
-        # check that kpoint grid makes sense
-        try:
-            self.kpt_grid = np.array(kpt_grid,dtype=int)
-        except:
-            message = 'the kpoint grid should be a list of 3 integers'
-            raise_error(message)
-        if self.kpt_grid.shape[0] != 3:
-            message = 'the kpoint grid should be a list of 3 integers'
-            raise_error(message)
-
-        # check that requested method makes sense
-        if grid_reduction not in self.allowed_reductions:
-            message = f'kpoint grid reduction method \'{grid_reduction}\' not one of the' \
-                    ' allowed methods.'
-            raise_error(message)
-
-        # grid reduction method
-        self.grid_reduction = grid_reduction
-
-        # set up grid based on requested method
-        if self.grid_reduction == 'full':
-            self._set_full_grid()
-
-        # print the kpt grid
-        message = f'there are {self.num_k} kpoints: (only printing the 1st 50)\n'
-        message = message+f' index, weight, kx ky kz (crystal coords)\n'
-        n_print = min(self.num_k,51)
-        for kk in range(n_print):
-            message = message+f'   {kk} {self.weights[kk]:2.3f}   {self.kpts[kk,0]: 2.4f}' \
-                    f' {self.kpts[kk,1]: 2.4f} {self.kpts[kk,2]: 2.4f}\n'
-        print_stdout(message,msg_type='KPOINTS')
-
-    # ------------------------------------------------------------------------------------------------
-
-    def _set_full_grid(self):
-
-        # full grid from -1/2 to 1/2
-        self.num_k = self.kpt_grid[0]*self.kpt_grid[1]*self.kpt_grid[2]
-        self.kpts = np.zeros((self.num_k,3))
-        self.weights = np.ones(self.num_k)
[…]

### 2021-08-25 dft/pw_pp/pw_pp.py

@@ -1,49 +1,0 @@
-import mod_crystal
-import mod_basis 
-import mod_kpoints
-#import mod_pp
-#import mod_hamiltonian
-#import mod_scf
-
-# -----------------------------------------------------------------------------------------------------
-
-crystal = mod_crystal.crystal()
-
-# set lattice vectors and compute reciprocal lattice
-lat_vecs = [[0.0,1.0,1.0],      # in angstrom
-            [1.0,0.0,1.0],
-            [1.0,1.0,0.0]]
-lat_scale = 2.01                # scale for lattice vectors
-crystal.set_lattice(lat_vecs,lat_scale)
-
-# set atom positions and convert to cartesian coords
-atom_pos = [[0.00,0.00,0.00],   # in crystal coords
-            [0.25,0.25,0.25]]
-atom_types = [1,1]              # 1 per atom
-atomic_numbers = [14]           # 1 per type
-crystal.set_atoms(atom_pos,atom_types,atomic_numbers)
-
-# -----------------------------------------------------------------------------------------------------
-
-pw = mod_basis.pw()
-
-# set up plane waves, max G vectors, etc
-pw_cut_off = 250 # eV
-pw.set_plane_waves(crystal,pw_cut_off)
-
-# -----------------------------------------------------------------------------------------------------
-
-kpts = mod_kpoints.kpoints()
-
-# set kpoint grid 
-kpt_grid = [7,7,7]
-grid_reduction = 'full' 
-kpts.set_grid(kpt_grid,grid_reduction)
- 
-# -----------------------------------------------------------------------------------------------------
-
-
-
-
-
-

### 2021-08-25 model_builder/bak/tb_modelling/v5/mod_utils.py

@@ -1,0 +1,26 @@
+
+# -------------------------------------------------------------------------------------------------
+
+def print_message(message,message_type='MESSAGE'):
+
+    """
+    print stuff to the screen using a standard formatting
+    """
+
+    print(f'\n ** {message_type} **\n')
+    print(f' {message}\n',flush=True)
+
+# -------------------------------------------------------------------------------------------------
+
+def raise_error(error_message):
+
+    """
+    print error message and exit
+    """
+
+    print_message(error_message,message_type='ERROR')
+    exit()
+
+# -------------------------------------------------------------------------------------------------
+
+

### 2021-08-25 phonon_model/mod_dynmat.py

@@ -1,55 +1,0 @@
-import numpy as np
-from mod_utils import raise_error
-
-class dynamical_matrices:
-
-    # --------------------------------------------------------------------------------------------------
-
-    def __init__(self,lattice,qpoints):
-
-        # dynamical matrices
-        self.dyn_mats = np.zeros((qpoints.num_q,lattice.num_basis*3,lattice.num_basis*3),dtype=complex)
-
-        # eigenvals
-        self.freq = np.zeros((qpoints.num_q,lattice.num_basis*3))
-        self.eigenvecs = np.zeros((qpoints.num_q,lattice.num_basis*3,lattice.num_basis*3),dtype=complex)
-
-    # --------------------------------------------------------------------------------------------------
-
-    def calculate_phonons(self,lattice,fc_mats,qpoints):
-
-        for qq in range(qpoints.num_q):
-
-            q_vec = qpoints.q_cart[qq,:]
-            self.dyn_mats[qq,:,:] = self._compute_dynamical_matrix(q_vec,fc_mats,lattice)
-
-            self.freq[qq,:] = np.sqrt(np.abs(np.linalg.eigvalsh(self.dyn_mats[qq,:,:])))
-
-        np.savetxt('freqs',self.freq,fmt='%2.4f')
-
-    # --------------------------------------------------------------------------------------------------
-
-    def _compute_dynamical_matrix(self,q_vec,fc_mats,lattice):
-
-        dyn_mat = np.zeros((lattice.num_basis*3,lattice.num_basis*3),dtype=complex)
-
-        for pair in range(lattice.num_sc*lattice.num_basis):
-
-            ii = fc_mats.inds[pair,0]
-            jj = fc_mats.inds[pair,1]
-            type_i = lattice.basis_types[ii]
-            type_j = lattice.basis_types[jj]
-            sc_ii = fc_mats.inds[pair,2]
-            sc_jj = fc_mats.inds[pair,3]
-
-            fc = fc_mats.fc_matrices[pair,:,:]
-            q_dot_R = lattice.sc_rel_cart[sc_ii,sc_jj]*q_vec
-            mass_fac = 1/np.sqrt(lattice.atom_masses[type_i]*lattice.atom_masses[type_j])
-            fc = fc*np.exp(1j*q_dot_R.sum())/mass_fac
-            dyn_mat[ii*3:(ii+1)*3,jj*3:(jj+1)*3] = dyn_mat[ii*3:(ii+1)*3,jj*3:(jj+1)*3]+fc
-
[…]

### 2021-08-25 phonon_model/mod_fc.py

@@ -1,55 +1,0 @@
-import numpy as np
-from mod_utils import raise_error
-
-class force_constant_matrices:
-    
-    # -------------------------------------------------------------------------------------------------
-
-    def __init__(self,lattice):
-
-        # fc matrices between each atom and the atoms in the SC
-        self.fc_matrices = np.zeros((lattice.num_basis*lattice.num_sc,3,3))
-        self.inds = np.zeros((lattice.num_basis*lattice.num_sc,4),dtype=int)
-
-        # copy some stuff
-        self.num_basis = lattice.num_basis
-        self.num_sc = lattice.num_sc
-
-    # -------------------------------------------------------------------------------------------------
-
-    def parse_force_constants_file(self,fc_file='force_constants'):
-
-        with open(fc_file,'r') as f_fc:
-            
-            for pair in range(self.num_sc*self.num_basis):
-                    
-                # get the inds from the file
-                try:
-                    inds = f_fc.readline().strip().split() 
-                    inds = [int(x) for x in inds]
-                except:
-                    message = f'indices in file \'{fc_file}\ seem wrong'
-                    raise_error(message)
-                basis_ind_i = inds[0]
-                basis_ind_j = inds[1]
-                sc_ind_i = inds[2]
-                sc_ind_j = inds[3]
-
-                # get the fc from the file
-                fc = np.zeros((3,3))
-                try:
-                    for kk in range(3):
-                        line = f_fc.readline().strip().split()
-                        fc[kk,:] = [float(x) for x in line]
-                except:
-                    message = 'force constants in file \'{fc_file}\ seem wrong'
-                    raise_error(message)
-
-                self.fc_matrices[pair,:,:] = fc
-                self.inds[pair,:] = [basis_ind_i,basis_ind_j,sc_ind_i,sc_ind_j]
-
-    # -------------------------------------------------------------------------------------------------
-
-
-
-

### 2021-08-25 phonon_model/mod_lattice.py

@@ -1,332 +1,0 @@
-import numpy as np
-from mod_utils import print_stdout, raise_error
-
-
-class lattice:
-
-    # ------------------------------------------------------------------------------------------------
-        
-    def __init__(self):
-
-        pass
-        
-    # ------------------------------------------------------------------------------------------------
-
-    def set_atom_data(self,atom_masses):
-
-        """
-        set data for atoms. 
-        """
-        self.atom_masses = np.array(atom_masses) # atomic mass in amu
-
-    # ------------------------------------------------------------------------------------------------
-
-    def set_primitive_cell(self,prim_cell_file='prim_cell'):
-
-        """
-        parse file with atom data in it and set up primitive cell
-        """
-
-        # parse the file
-        lat, nat, pos, types = self._parse_cell_file(prim_cell_file)
-
-        # set the data
-        self.prim_vecs = lat
-        self.num_basis = nat
-        self.basis_pos = pos
-        self.basis_types = types
-
-        # get the reciprocal lattice vectors
-        self.r_prim_vecs, self.prim_vol = self._compute_reciprocal_lattice(self.prim_vecs)
-
-        # get basis pos in cartesian coords
-        self.basis_pos_cart = self._crystal_to_cart(self.prim_vecs,self.basis_pos)
-
-        # check number of types mathces number of masses and number of atomic numbers
-        self.num_types = np.unique(self.basis_types).shape[0]
-        if self.num_types != self.atom_masses.shape[0]:
-            message = 'number of basis types doesnt match number of masses'
-            raise_error(message)
-
-    # ------------------------------------------------------------------------------------------------
-
-    def set_super_cell(self,super_cell_file='super_cell'):
-
-        """
-        parse file with atom data in it and set up super cell
-        """
-
-        # parse the file
-        lat, nat, pos, types = self._parse_cell_file(super_cell_file)
-
-        # set the data
-        self.sc_vecs = lat
-        self.num_sc = nat
-        self.sc_pos = pos
-        self.sc_types = types
-
-        # get the reciprocal lattice vectors
[…]

### 2021-08-25 phonon_model/mod_qpoints.py

@@ -1,65 +1,0 @@
-import numpy as np
-from mod_utils import raise_error
-
-
-class qpoints:
-
-    # ------------------------------------------------------------------------------------------------
-
-    def __init__(self):
-    
-        pass
-
-    # ------------------------------------------------------------------------------------------------
-
-    def q_path(self,lattice,q_vert=[[0,0,0],[0.5,0,0]],num_q_sm=101):
-
-        # vertices of the q path
-        q_vert = np.array(q_vert)
-        if q_vert.shape[1] != 3:
-            message = 'q vertices should be 3D'
-            raise_error(message)
-        num_vert = q_vert.shape[0]
-        num_path = num_vert-1
-
-        # vertices of the q path in cartesian coords
-        q_vert_cart = self._crystal_to_cart(lattice.r_prim_vecs,q_vert)
-
-        self.q_dist = np.zeros(num_path)
-        for qq in range(num_path):
-            q_dist = q_vert_cart[qq+1,:]-q_vert_cart[qq,:]
-            q_dist = np.sqrt((q_dist**2).sum())
-            self.q_dist[qq] = q_dist
-
-        num_per = self.q_dist/self.q_dist.min()*num_q_sm
-        num_per = np.round(num_per).astype(int)
-        
-        self.num_q = num_per.sum()
-        self.q_cart = np.zeros((self.num_q,3))
-
-        shift = 0
-        for qq in range(num_path):
-            for ii in range(3):
-                self.q_cart[shift:shift+num_per[qq],ii] = np.linspace(q_vert_cart[qq,ii],
-                        q_vert_cart[qq+1,ii],num_per[qq])
-            shift = shift+num_per[qq]
-
-    # -----------------------------------------------------------------------------------------------
-
-    def _crystal_to_cart(self,lat_vecs,crystal_coords):
-
-        """
-        wrapper to just matrix multiply each atoms vector by the fiven transformation matrix.
-        """
-
-        nat = crystal_coords.shape[0]
-        cart_coords = np.zeros((nat,3))
-        for atom in range(nat):
-            cart_coords[atom,:] = np.matmul(lat_vecs,crystal_coords[atom,:])
-
-        return cart_coords
-
-    # ------------------------------------------------------------------------------------------------
-
-
-

### 2021-08-25 phonon_model/phonon_model.py

@@ -1,61 +1,0 @@
-import mod_lattice
-import mod_fc
-import mod_qpoints
-import mod_dynmat 
-
-# ---------------------------------------------------------------------------------------------------
-
-lattice = mod_lattice.lattice()
-
-### set atom data
-atom_masses = [51.9961] 
-lattice.set_atom_data(atom_masses)
-
-### read primtive cell vectors and positions
-prim_cell_file = 'prim_cell'
-lattice.set_primitive_cell(prim_cell_file)
-
-### read super cell vectors and positions
-super_cell_file = 'super_cell'
-lattice.set_super_cell(super_cell_file)
-
-### find nearest neighbor vectors 
-lattice.find_neighbors()
-
-### write lattice info to files
-#lattice.write_lattice()
-#lattice.write_neighbors()
-
-
-# -------------------------------------------------------------------------------------------------
-
-fc_mats = mod_fc.force_constant_matrices(lattice)
-
-### read the force constant file 
-fc_mats.parse_force_constants_file('force_constants')
-
-
-# ---------------------------------------------------------------------------------------------------
-
-qpoints = mod_qpoints.qpoints()
-
-### set up q points path
-q_vert = [[0.000,0.000,0.000],
-          [0.500,0.000,0.000],
-          [0.500,0.500,0.000],
-          [0.000,0.000,0.000],
-          [0.500,0.500,0.500]]
-qpoints.q_path(lattice,q_vert,num_q_sm=151)
-
-# ---------------------------------------------------------------------------------------------------
-
-dyn_mats = mod_dynmat.dynamical_matrices(lattice,qpoints)
-
-### diagonalize dynamical mats
-dyn_mats.calculate_phonons(lattice,fc_mats,qpoints)
-
-# ---------------------------------------------------------------------------------------------------
-
-
-
-

### 2021-08-25 phonon_model/tmp.py

@@ -1,27 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-
-freq = np.loadtxt('freqs')
-
-nb = freq.shape[1]
-nq = freq.shape[0]
-
-for bb in range(nb):
-
-    plt.plot(freq[:,bb],marker='o',ms=2,mfc='w',mec='k',mew=0.5,ls='')
-
-plt.show()
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2021-11-09 comps_2/outline.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prb,aps,11pt,superscriptaddress,floatfix]{revtex4-2} 
\usepackage{graphicx}
\usepackage{color} 
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{The Y-files}

\author{Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

\begin{abstract}
This is my proposal for the project/paper portion of the \emph{Comps 2} (C2) exam in the physics graduate program at the University of Colorado Boulder (CU). I breifly summarize my history in the graduate program in order to inform the reader of my previous and current research activites; then I introduce my proposed topic. I provide historical context and a breif outline of the structure of my paper. The planned contents can be loosely group into the following 5 sections: (i) introduction and historical context, (ii) important theoretical details, (iii) experimental sucesses and failures, (iv) current state of the art, and (v) summary.  
\end{abstract}

\maketitle

%\listoffigures
%\listoftables
%\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Previous Research History}
I have always been interested in doo-dads, gadgets, and gizmos. I grew up on a poor-ish farm in rural Texas. We couldn't afford to hire mechanics, plumbers, or electricitians, so we nearly everything ourselves. I was a curious enough kid, so I learned how to do these things too.

My interest in mechanics, machninery, and DIY lead me to Texas State University to study Manufacturing. Skip forward a few years. I worked in a few labs as a undergraduate assistant: in my first gig, I programmed industrial robots to TIG weld test-specimens, ran the procedure, then measured the weld strengths. My lab had written some machine-learing type algorithms to automate the various parameters relevant to the weld strength (e.g. electrode travel speed, electrode distance from sample, current) and I was generating the training data (i.e. got two IEEE conference papers out of this work!).

My next job was to manufacture some prototypes of heat sinks for Hyundai headlights. The eventual mass-produced things would be 3D printed, but we were going to machine them from fancy high-thermal conductivity alloys. I raised the question that the transport properties of these materials would be vastly different due to the different microstructures. I thought we should account for that and was given the go-ahead to start working on that problem. Unfortunately I graduated shorty after and left Texas State... but I didn't left my question behind.

Skip forward another few years. I joined the materials science PhD program here at CU to study heat transport at the nano-scale. I wanted to understand how different crystalline structures in a material affect the flow of heat. I joined a project working on thermoelectric superlattices and worked with molecular dyanmics simulations to model and quantify the flow of heat in these structures. I eventually left this lab too (with one co-author publication in PRB), but not before I learned that the question I brought with me was probably not going to be easy to answer. Furthermore, I had learned that the lattice also exchanges heat with the electrons through electron-phonon coupling: I ended up with another question!

Finally, I landed in Dmitry Reznik's lab. My study in heat transport in crystals led me to focus a lot of my attention on phonons and I had become interetsted in electron-phonon coupling by now too. Dmitry and his lab are experts in this stuff and I was lucky enough to be invited to work with them. I wanted (and badly needed) to learn more advanced physics than my engineering background had provided, so Dmitry encouraged me to transfer into the physics PhD program. I was lucky enough to be let in... and here I am. In my time there, I have used density functional theory, molecular dynamics, and inelastic neutron scattering to study even wackier materials than I had ever guessed had existed. My personal and academic interests have evolved and by now I am heavily invested in applying (continously more advanced) computational methods to interesting materials science problems. I have calculated neutron scattering spectra from density functional theory to study electron-phonon effects in cuprates (two co-authro papers in PRB, first author paper in progress). In another project, I used time-of-flight inelastic neutron scattering data and data science-esque techniques and software developed by my lab to determine the phonon dispersions a spin-spiral ordered antiferromagnet (one co-authro Phys. Rev. Research paper). My current projects are involved in using molecular dynamics simulations to understand the lattice dynamics and electron-phonon effects in dynamically disordered CH$_3$NH$_3$PbI$_3$, which has a nearly-freely rotating molecules trapped in its lattice. We are also looking at the effects of static disorder on the lattice dynamics in a thermoelectric clathrate and the results are exciting. On the back-burner are experimental investigations (data already taken!) on nickelates and manganates. I have some ideas to use even more advanced theoretical/computation methods to include the non-trivial, dynamical magnetic ordering that couples with the lattice in these materials.

Now lets move on to my topic. Obviously, my interests and experiences lie in condensed matter so nearly anything in condensed matter is exlcuded as a topic choice. There are many fascinating and modern things here at CU that would probably be natural choices: e.g. quantum information, AMO physics, high-energy computational methods, etc. However, I suspect that a vast and many students jump onto these hot-topics and I have never been one to get-in-line. T that end, I have made it my goal to pick something that it is both fun and unusual. I have no personal interests in the \emph{paranormal}

\section{Paper Topic}

I am still tossing around 2 ideas: (i) physics of the paranormal and (ii) harvesting the zero-point energy.

\subsection{The Y-Files (or the physics of the paranormal)}
My proposed idea is to discuss the actual physics of the paranormal. Specifically, I would focus on only a few phenomena that really merit investigation. Things that I consider \emph{beyond realistic} (e.g. ghosts, monsters, and a whole slew of other things crammed into modern 'scary' movies) will not be considered. On the other hand, things that could convceivably fall under the realms of quantum mechanics, statistical mechanics, electrodynamics, thermodynamics, etc. (e.g. telekenesis, telepathy, etc.) will be looked at in some detail. The Y in the title is just the next letter in the alphabet after X: \emph{The X-files} was a TV show that followed an FBI Agent tasked with investigating reports of paramormal phenomena. Unlike the sometimes delusional, conspirary obsessed agent leading the show, we will not entertain explanations based on consipracies, aliens, or other unprovable or unverifiable non-sense.

Instead, we will undertake rigourous investigations grounded in physics. We will clearly define the phenomena under study. We will look at detailed experiments with controlled conditions as well as serious theories proposing plausibale explanations. A recurring theme will be to assess the thermodynamics of a phenomena: claims of occurences that violate our beloved laws of thermodynamics are probably false. On the otherhand, I wouldn't be so sure about phenomena like telekinesis and telepathy: we regularly experience and observe forces applied by 'invisible hands' in the form of electromagnetic interactions. Electric and magnetic fields can move massive amounts of energy and exert incredible forces. Yet, in many cases, we cannot detect the source of these phenomena unless we employ specialized instruments. Similarly, we send and recieve 'telepathic' messages in the form texts and emails on a, lets face it, nearly constant basis.

To an ignorant observer these phenomena could seem unnatural, like science fiction. \emph{give a parable about our ancestors observing electricity, fire, etc.}.

\section{layout}
Start with a discussion of how mythology and claims of the paranormal emerged as an explanation for things like electromagnetism, fire, etc. Maybe start with a neat parable. Move onto to highlight how rigourous physical emerged in the course of serious scientific investigation. Give examples of similar things in modern life and discuss the current state of the art theories about these things and point out the experimental situation and indicate the things that have been proved and disproven for-sure. Actually, maybe don't stick to 'paranormal' stuff but look at anything that is for certain real but unexplained. E.g. flashing/glowing lights in swamps? We will take a Sherlock Holme's in the one where goes to hunt the wolf in the bog approach. Maybe even like Fox Moulder.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ref}

\end{document}

### 2021-11-09 sonoluminescence/paper/notes.tex
asdsad

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prb,aps,12pt,superscriptaddress,floatfix]{revtex4-2} 
\usepackage{graphicx}
\usepackage{color} 
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{\huge Shrimpoluminescence}

\author{\large Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
%\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

%\begin{abstract}
%    abstract
%\end{abstract}

\maketitle

%\listoffigures
%\listoftables
%\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
\includegraphics[width=1\linewidth]{figs/shrimpy.pdf}
%    \caption{}
\label{fig:shrimpy}
\end{figure}

\newpage

\section{Details of the project}
Pistol shrimp, like the red-banded pistol shrimp shown in Fig. \ref{fig:shrimpy}, are capabale of generating a jet of water with high enough velocity that a \emph{cavitation} bubble forms behind it. When the bubble collpases under the pressure of the sea, enough energy is released to produce a shock-wave that can kill the shrimp's prey \cite{}. If the shrimp's prey had very sensitive eyes (and also weren't dead) they might also see a flash of light produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp, but more generally known as \emph{sonoluminescence} \cite{}.

\section{Paper Topic}

Some items to consider in the introduction of the actual paper. Caviation bubbles capabale of sonoluminesce occur in numerous places in nature:
\begin{itemize}
    \item the snapping and mantis shrimps us caviation bubbles to create shock waves that kill prey
    \item cavitation bubbles occur in pumps and after propellers and the shock damages the machinery
    \item ...
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ref}

\end{document}

### 2021-11-09 sonoluminescence/proposal/notes.tex
asdsad

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prb,aps,12pt,superscriptaddress,floatfix]{revtex4-2} 
\usepackage{graphicx}
\usepackage{color} 
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{\huge On the Origin of Shrimpoluminescence }

\author{\large Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
%\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

%\begin{abstract}
%    abstract
%\end{abstract}

\maketitle

%\listoffigures
%\listoftables
%\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
\includegraphics[width=1\linewidth]{figs/shrimpy2.pdf}
%    \caption{}
\label{fig:shrimpy}
\end{figure}

\newpage

\section{Details...}
Pistol shrimp, like the red-banded pistol shrimp shown on the cover, are capabale of generating a jet of water with high enough velocity that a cavitation bubble forms behind it. When the bubble collpases under the pressure of the sea, enough energy is released to produce a shock-wave that can kill the shrimp's prey \cite{}. If the shrimp's prey had very sensitive eyes (and also weren't dead) they might notice a flash of light produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{}, but more generally known as \emph{sonoluminescence} \cite{}.

Sonolumiscence is more precisely defined as ... \cite{} and occurs in a number of scenarios where cavitation occurs ... \cite{}. Yet the physical origin of light production in sonolumiscence is not understood. To-date, there are two \emph{prevailing} theories: (i) it is a thermally induced phenomena \cite{} and (ii) it is an electrical effect \cite{}. I will review both of these ideas, describing both the theory and experiments. The thermal school of thought has focused a lot of effort on measuring the temperature of the bubble when light is emmited and estimates are in the \emph{thousands} of Kelvin (the surface of the Sun is ... K! \cite{}) so some time will be spent discussing this.

On the other hand, I can't ignore an ``alternative" theory which is actually what led me to this topic: a phenomenological theory based on quantum electrodynamics. This subject will get us out of the main scope of the paper (and into an area that I probably can't answer questions about) but I plan to spend a small amount of time on this too because, well... it sounds interesting!

\section{About me}

My undergraduate degree is in manufacturing where I worked with industrial robots on manufacturing process optimization. I also completed a masters degree in materials science at CU before transferring into the physics PhD program. In the materials science program, I worked in the aerospace department using molecular dynamics to study (lattice) thermal conductivity in thermoelectrics. I work for Dmitry Reznik so my research is still focused on lattice-dynamics, but my methods are based on computational electronic-structure theory(i.e DFT). We usually look at fancy materials like cuprates or magnets, but we also study energy materials like thermoelectrics and hybrid perovskites. The point is, none of this is even closely related to shrimpoluminscence. I plan to have fun writing this paper and I hope you will have fun reading it!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ref}

\end{document}

### 2021-11-10 sonoluminescence/paper/paper.tex
asdsad

--- 2021-11-09 sonoluminescence/proposal/notes.tex
@@
 
-\title{\huge On the Origin of Shrimpoluminescence }
+\title{\huge On the Origin of Shrimpoluminescence}
 
@@
 %\begin{abstract}
-%    abstract
 %\end{abstract}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{figs/shrimpy2.pdf}
+\includegraphics[width=1\linewidth]{../figs/shrimpy2.pdf}
 %    \caption{}
@@
 
-\section{Details...}
-Pistol shrimp, like the red-banded pistol shrimp shown on the cover, are capabale of generating a jet of water with high enough velocity that a cavitation bubble forms behind it.
-When the bubble collpases under the pressure of the sea, enough energy is released to produce a shock-wave that can kill the shrimp's prey \cite{}.
-If the shrimp's prey had very sensitive eyes (and also weren't dead) they might notice a flash of light produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{}, but more generally known as \emph{sonoluminescence} \cite{}.
+\section{QED}
+the ref \cite{liberati2000sonoluminescence} has lots of good refs.
+they have follow up papers extending Schwingers papers (which are refd therein) and referencing a bunch of other relevant literature.
+I should look for refs citing Schwingers paper to try to figure out the modern state of the art of this thin.
 
-Sonolumiscence is more precisely defined as ...
-\cite{} and occurs in a number of scenarios where cavitation occurs ...
-\cite{}.
-Yet the physical origin of light production in sonolumiscence is not understood.
-To-date, there are two \emph{prevailing} theories: (i) it is a thermally induced phenomena \cite{} and (ii) it is an electrical effect \cite{}.
-I will review both of these ideas, describing both the theory and experiments.
-The thermal school of thought has focused a lot of effort on measuring the temperature of the bubble when light is emmited and estimates are in the \emph{thousands} of Kelvin (the surface of the Sun is ...
-K!
-\cite{}) so some time will be spent discussing this.
+\section{Modelling}
+in the ref \cite{schanz2012molecular}, they use MD to modelling the equation of state for a collapsing bubble. they calculate the temperature and study the conditions for emitting light.
 
-On the other hand, I can't ignore an ``alternative" theory which is actually what led me to this topic: a phenomenological theory based on quantum electrodynamics.
-This subject will get us out of the main scope of the paper (and into an area that I probably can't answer questions about) but I plan to spend a small amount of time on this too because, well...
-it sounds interesting!
-
-\section{About me}
-
-My undergraduate degree is in manufacturing where I worked with industrial robots on manufacturing process optimization.
-I also completed a masters degree in materials science at CU before transferring into the physics PhD program.
-In the materials science program, I worked in the aerospace department using molecular dynamics to study (lattice) thermal conductivity in thermoelectrics.
-I work for Dmitry Reznik so my research is still focused on lattice-dynamics, but my methods are based on computational electronic-structure theory(i.e DFT).
-We usually look at fancy materials like cuprates or magnets, but we also study energy materials like thermoelectrics and hybrid perovskites.
-The point is, none of this is even closely related to shrimpoluminscence.
-I plan to have fun writing this paper and I hope you will have fun reading it!
+\section{Refs. to different arguments}
+ref \cite{didenko2000molecular} gives examples of other origin theories (molecular excited states, black-body radiation, bremsstrahlung, ion-electron recombination, confined electrons)
 
@@
 
-\bibliography{ref}
+\bibliography{../ref}
 

### 2021-11-10 sonoluminescence/proposal/prop.tex
asdsad

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[prl,aps,11pt,superscriptaddress,floatfix]{revtex4-2} 
\usepackage{graphicx}
\usepackage{color} 
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{bm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{\huge On the Origin of Shrimpoluminescence }

\author{\large Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
%\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

%\begin{abstract}
%    abstract
%\end{abstract}

\maketitle

%\listoffigures
%\listoftables
%\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
\includegraphics[width=1\linewidth]{../figs/shrimpy2.pdf}
%    \caption{}
\label{fig:shrimpy}
\end{figure}

\newpage

\section{Details...}
Pistol shrimp, like the red-banded pistol shrimp shown on the cover, are capabale of generating a jet of water with high enough velocity that a cavitation bubble forms behind it. When the bubble collpases, a powerful shock-wave is produced that can kill the shrimp's prey \cite{versluis2000snapping}. If the shrimp's prey had very sensitive eyes (and also weren't dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.

Sonoluminescence can be more precisely defined as the effect in which light is produced when an ultrasonically driven bubble collapses \cite{borisenok2020mechanisms}. Unlike the shrimp's method, laboratory techniques can produce stable, single bubbles that can be driven and collapsed with high precision. As such, quite a bit of effort has been spent studying \emph{single-bubble sonoluminescence} (SBSL) \cite{lohse2018bubble,brenner2002single}. Yet the physical origin of the light production has eluded understanding. There are two prevailing theories: (i) the effect is of a thermal origin and (ii) it is an electrical effect \cite{borisenok2020mechanisms}. The thermal argument is based on the observation that the temperature of the gas trapped in the bubble when light is emmited is $\sim$10000 K! \cite{flannigan2005plasma,lohse2018bubble} (for reference, the surface of the Sun is about 6000 K). It is beleived that thermal radiation (among other mechanisms) could be the source of the light. The electrical arguments are instead based on the belief that the collapsing bubble does not remain spherical: the asymmetrical shape of the bubble leads to the formations of \emph{jets} that cause non-Newtonian fracturing in the liquid \cite{prosperetti1997new}. There is subsequently a dielectric break-down at the fracture and the electrical spark is the source of the light. Along similar lines, it has been argued that the rapidly deforming bubble polarizes the water at the bubble's surface, leading to an (electric) polarization field. Assuming the thermal effects are also true, it was argued that free electrons from the ionized gas in the bubble enter the polarized water layer and form a plasma that emmits the light.

On the other hand, I can't ignore an ``alternative" theory which is actually what led me to this topic: a phenomenological theory based on quantum electrodynamics. This subject will get us pretty far afield but I will try to spend some time on this too because, well, it sounds interesting! In short, Schwinger \cite{schwinger1993casimir} and subsequently others \cite{liberati2000sonoluminescence,eberlein1996sonoluminescence} have argued that changes in the \emph{Casimir energy} due to changing the volume of a dielectric sphere can lead to production of photons.

\section{About me}

My undergraduate degree is in manufacturing; I worked with industrial robots on manufacturing process optimization. I also completed a masters degree in materials science at CU before transferring into the physics PhD program. In the materials science program, I worked in the aerospace department using molecular dynamics to study (lattice) thermal conductivity in thermoelectrics. Now I work for Dmitry Reznik so my research is still focused on lattice-dynamics but my methods are based on computational electronic-structure theory (i.e DFT). We usually look at fancy materials like cuprates or magnets, but we also study energy materials like thermoelectrics and hybrid perovskites. The point is, none of this is even closely related to shrimpoluminscence.

I plan to have fun writing this paper and I hope you will have fun reading it!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{../ref}

\end{document}

### 2021-11-10 sonoluminescence/proposal/prop.tex
asdsad

@@
 \section{Details...}
-Pistol shrimp, like the red-banded pistol shrimp shown on the cover, are capabale of generating a jet of water with high enough velocity that a cavitation bubble forms behind it.
+Pistol shrimp, like the red-banded pistol shrimp shown on the cover, are capabale of generating a jet of water with high enough velocity that a cavitation bubble forms in its wake.
 When the bubble collpases, a powerful shock-wave is produced that can kill the shrimp's prey \cite{versluis2000snapping}.

### 2021-11-29 sonoluminescence/paper/paper.tex
assda

@@
 
-\documentclass[prb,aps,12pt,superscriptaddress,floatfix]{revtex4-2} 
+\documentclass[prb,aps,11pt,superscriptaddress,floatfix]{revtex4-2} 
 \usepackage{graphicx}
-\usepackage{color} 
+\usepackage{xcolor} 
 \usepackage{amsmath}
@@
 
+\emph{some abbreviations}
+\begin{itemize}
+    \item SL: sonoluminescence
+    \item SBSL: single-bubble sonoluminesence
+    \item MBSL: multi-bubble sonoluminescence
+    \item EOM: equation(s) of motion
+\end{itemize}
+
+\section{rough-outline}
+\begin{itemize}
+    \item start the introduction with captivating example, i.e.
+shrimpoluminescence.
+describe what it is, how it was found, then transition into sonolouminescence.
+move on to explain how its not understood, various theories try.
+highlight these.
+then move on to discuss the experimental situation etc.
+    \item Section: experiment. basically layout what actually is known (by observation) to happen during sonoluminescence. facts only, not trying to explain.
+    \item explanations: then move on to theories that try to explain it: thermal, electrical. cover each of these, highlighting thier sucesses and failures.
+    \item QED theory: breifly cover this one
+\end{itemize}
+
+\section{random notes}
+apparently a shitty movie called 'Chain Reaction' with Keanu Reeves was centered around the premise of MBSL as a means of clean energy production.
+it was based on the speculation that temperatures inside the bubbles ciould reach up to millions of Kelvin and lead to fusion.
+this was debunked by Prosperetti (see the pop sci article \cite{chainreaction}).
+He points out that the temp is speculative and is based on assumptions of a perfect sphere collapsing.
+he notes that its probably not spherical when light is produced and even points out that groups have tried to produce fusion power by intentionally collapsing gas filled micro baloons with lasers...
+and failed.
+Tatartcheceko also shows that this is horsehit \cite{tatartchenko2017sonoluminescence}.
+in this article, Prosperetti also explains how a 'jet' of water whooshing across the bubble could be going faster that 4000 mph, fast enough to cause non-netwownian fracturing of the water leading to light producuction through (apparently) the same mechanism of light emession when ice or wint-o-green mints are snapped in half...
+huh.
+this is good introduction fodder.
+need to find the primary lit.
+
+\emph{heurestic explanation of bubble trapping given by Brenner in \cite{chainreaction}.}: the bonds between molecules in a fluid are very strong, so instead of 'tearing', the fluid stretches.
+however, if there is a bubble present, the local dialation of the fluid can change the bubbles diameter by $\sim$ 1000 times to avoid breaking bonds at the interface between bubble and fluid.
+But when the compression waves comes by, the 'gas' in the bubble, which has boiled due to low pressure, offers very little resistance to compression.
+the wave compresses the bubble rapidly, which spikes the temp in the gas to incandescent temperatueres.
+this heuristic explanation is on-track with Crum \cite{crum1994sonoluminescence} below.
+
+\section{1st discovery}
+MBSL was discovered first.
+its (obviously) hard to create and control single bubbles so this wasnt feasible until the 1990s, attributed to Gaitan (see below).
+allegedly, it was Marinesco et al \cite{marinesco1933actions} that discovered MBSL in 1933 (according to \cite{gaitan1990experimental} below and to \cite{tatartchenko2017sonoluminescence}) but I cant find this paper (or understand it since its in french).
+Other sources (e.g..
+wikipedia, \cite{tatartchenko2017sonoluminescence,crum1994sonoluminescence}) credit Frenzel et al \cite{frenzel1934luminescenz} as the discoverers.
+I cant German or find this paper either, but in anycase, Tatartchenko summarized thier discovery: Freznel and coworkers were trying to speed up photo development by 'insonating' i.e.
+subjected developer-fluid to sound.
+they found that when they immersed photo-plates or whatever in the fluid, it became 'foggy' and overexposed, as if subjected to diffuse light.
+A bunch of other papers followed up on this.
+get the deets from \cite{gaitan1990experimental} below.
+
+Update: Brenner \cite{brenner2002single} consisely summarizes the history: Marinesco notes the photographic plates fogs in an ultrasonic bath \cite{marinesco1933actions} and Frenzel et al \cite{frenzel1934luminescenz} deliberately studied the light emission.
+According to brenner, no one was surprised by this since it was already known that cavitating bubbles fuck shit up.
+
+SBSL was discovered by Gaitan \cite{gaitan1990experimental,gaitan1992sonoluminescence} etc.
+who trapped a single air bubble in a water filled flask using piezoacoutics stuck to the outside.
+(note: the trapping force is called \emph{Bjerknes force}, refs in \cite{lohse2018bubble}).
+
+\section{notes from \cite{crum1994sonoluminescence}}
+pop sci type of magazine article.
+Gives heurestic expanation: When an intense sound field is produced within a liquid, microscopic cavities of gas or vapor can be generated when the liquid fails under tensile stress.
+The subsequent acoustic compression cycle forces these cavities to collapse violently, which results in a remarkable concentration of mechanical energy, estimated to be as high as 12orders of magnitude (refs are given too).
+Also notes that MBSL is very hard to study since the bubbles occur randomly and transiently.
+Describes SBSL as the \emph{hydrogen atom of sonoluminescence}.also discusses that attempts to measure the (time) interval over which light is produced have found it be less than $\sim$50 ps which is 1/1000 times smaller than theory predicts.
+They also found the light to be more 'stable' than the acoustic generator, i.e.
+more regular with more controlled intervals.
+an old but probably still relevant result.
+mentions a neat claim that the conditions of the gas at the center of the bubble are probably more like a metal than a gas.
+\textcolor{red}{hell, this article basically summarizes all of the anomalies present in 1994.
+a very good resource.}
+
+\section{notes from ref \cite{lohse2018bubble}} 
+SBSL was discovered by Giatan et.
+al.
+in 1990s at U.
+Mississippi \cite{gaitan1990experimental,crum1994sonoluminescence,gaitan1992sonoluminescence} (note, \cite{ggaitan1992sonoluminescence} has $\sim$1000 citations..).
+The SBSL experiments were done by sticking peizoacoustic transducers to the outside of a water filled glass.
+the 'Bjerknes' force trapps the bubble and the acoustic wave drives its expansion/collapse.
+What Gaitan saw was that even the single bubble emits light.
+Wow!
+This was baffling since 'acoustic energy' \textcolor{red}{(i dont understand this)} is on the order of $\sim10^{-12}$ eV/molecule while visible light is $\sim1$ eV.
+NVM: I do get this.
+Typical acoustic waves, e.g.
+in air, only displace molecules on the order of nanometers, which aquires an elastic energy gaine between molecules of $\sim10^{-12}$ eV/molecule: certainly not enough to excite an electronic transition.
+
+This paper refs the review by Propsperetti: \cite{plesset1977bubble}.
+This paper gives EOM for the bubbles radius, the so called \emph{Plesset-Rayleight} equation which I have seen everywhere related to bubbles.
+Talks about solving it, some other crap related to bubble dynamics.
+this requires a detailed re-read.
+
+Later in the paper, also revisists snapping shrimp.
+
+\section{notes from ref \cite{prosperetti2004bubbles}}
+has shit tons of relevant math. take a look. a pretty paper overall
+
+\section{notes from \cite{gaitan1990experimental}}
+disseration by Giatan who is credited with discovering SBSL.
+says apparently SL was discovered in 1933.
+need to go read this crap.
+Apparently this was by Marinesco et.
+al.
+in 1933.
+Zimakov tried to explain it in 1934 and concluded that the light was due to electric discharge between vapor cavities and the glass container.
+Frenzel and Schultes in 1935 instead claimed it was from friction between cavitating bubbles and the water.
+Chamber in 1936 studied SL in other nonwater liquids.
+Chamber also postulated the light was from 'tribolumiscence' which occurs when the crystalline structure of the liquid-bubble lattice collapses?
+Apparently somthing similar occurs in real crystals.
+...
+shit, a bunch more models are given.
+give this dis.
+a lot of attention.
+
+also says caviation was predicted by 'Euler' in 1754: \emph{if the velocity of a fluid is high enouhg, negative pressures can form and the liquid can 'break'.} the 'breaking' was called cavitation in 895 by R.E.
+Froude, whoever the fuck this is, when studying bubbles near propellers.
+Rayleigh, unsurpisingly, was the first to give a rigorous theoretical treatment of cavitation in 1917.
+
+\section{notes from \cite{brenner2002single}}
+defines SBSL as ``Single-bubble sonoluminescence occurs when an acoustically trapped and periodically driven gas
+bubble collapses so strongly that the energy focusing at collapse leads to light emission".
+
+\newpage
+
 \section{QED}

### 2021-12-07 sonoluminescence/paper/paper.tex
asdsad

@@
 
-\documentclass[prb,aps,11pt,superscriptaddress,floatfix]{revtex4-2} 
+\documentclass[rmp,aps,superscriptaddress,floatfix]{revtex4-2} 
 \usepackage{graphicx}
@@
 
-\title{\huge On the Origin of Shrimpoluminescence}
+\title{On the Origin of Shrimpoluminescence}
 
-\author{\large Tyler C. Sterling}
+\author{Tyler C. Sterling}
 \email{ty.sterling@colorado.edu}
@@
 
-%\begin{abstract}
-%\end{abstract}
+\begin{abstract}
+shrimpity shrimp shrimp shrimp.
+abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract
+\end{abstract}
 
@@
 
+%\emph{Frequently Used Abbreviations}
+%\begin{itemize}
+%    \item SL: sonoluminescence
+%    \item SBSL: single-bubble sonoluminesence
+%    \item MBSL: multi-bubble sonoluminescence
+%\end{itemize}
+
+\section{Introduction}
+Snapping shrimp, like our cute friend in Fig.
+\ref{fig:shrimp_claw} (left), have long been known to produce cavitating bubbles by \emph{snapping} their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
+They have a strong appendage (the \emph{dactyl}) in thier claw [Fig.
+\ref{fig:shrimp_claw} (center)] that is used to create a very high-velocity jet of water.
+The low pressure region in the jet's wake forms a bubble [Fig.
+\ref{fig:shrimp_claw} (t=0 ms)] that, among other things, produces an exceptionally loud noise when it collapses.
+The sound (i.e.
+the pressure wave) produced by even a \emph{single} shrimp's snap is detectable over a mile away \cite{everest1948acoustical}.
+The noise produced by groups of shrimp is so intense that the U.S.
+Navy used them as ``sonar-camoflauge" in the Pacific ocean during World War II \cite{versluis2000snapping}.
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figs/shrimpy2.pdf}
-%    \caption{}
-\label{fig:shrimpy}
+\includegraphics[width=0.9\linewidth]{figs/shrimp_claw.pdf}
+    \caption{(left) Snapping shrimp (\emph{Alpheus heterochaelis}).
+(center) Blown-up view of the shrimp's claw.
+The \emph{plunger} (pl) on the \emph{dactyl} (d) rapidly enters the \emph{socket} (s), ejecting a high-velocity jet of water.
+The water ejection and subsequent bubble formation and, finally, bubble collape (at t=0) are shown on the right.
+The units are ms.
+Adapted from ref.
+\cite{versluis2000snapping}}
+\label{fig:shrimp_claw}
 \end{figure}
+
+The shrimp were not patriots helping the war-effort, however; they snapped for food.
+The shock-wave produced by the cavitating bubble is used to stun and even kill prey \cite{versluis2000snapping}.
+If the shrimp's prey had very sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.
+
+Sonoluminescence (SL) is more precisely defined as the process by which a ``driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission" \cite{brenner2002single}.
+Sonoluminescence comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonolumiscence.
+The disticntion is self-explanatory: multi-bubble sonolumiscence (MBSL) consists of  ``the simultaneous creation and destruction of many seperate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
+The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light very high precisions \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
+We will discuss MBSL breifly later, but for now it suffices to say that nearly all theoretical and experimental progress has been made using SBSL, with some authors even calling it ``the hydrogen atom of sonoluminescence" \cite{lohse2018bubble,crum1994sonoluminescence}.
+
+The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \cite{} to more exotic explanations based of quantum field theory \cite{schwinger1993casimir,eberlein1996sonoluminescence,liberati2000sonoluminescence}.
+
+But besides shrimpoluminescence, why is SL interesting?
+Well, at a glance, it is not obvious why SL occurs.
+The acoustic wave in the liquid displaces molecules on the order of nanometers, costing an elastic energy of  $\sim1\times 10^{-12}$ eV/molecule while emitting visible light (through e.g.
+electronic transitions) costs an energy $\sim1$ eV/molecule \cite{lohse2018bubble,crum1994sonoluminescence}...
+there is a 12-orders-of-magnitude concentration of energy.
+That is huge!
+After a moments thought, however, the massive energy focusing is not that surprising.
+For example, the energy in the shock wave produced by a cavitating bubble, regardless of the light production, has been known to damage ship propellers since atleast 1917 when Lord Rayleigh's gave the first rigorous treatment of caviating bubbles \cite{gaitan1990experimental}.
+A heurestic argument goes like this: \cite{plesset1949dynamics}.
+
+So again, why is the light emmission interesting? In short, after 3 decades of studying SBSL, the mechanism of light production is still not understood.
+
+The excitment wasn't restricted to academics however, with SBSL catching Hollywood's attention leading to an unpopular movie called \emph{Chain Reaction} where t
+
+So again, why is the light emmission interesting?
+After 3 decades of studying SBSL, the physical mechanism of light production in SL is still not understood.
+This can be seen by glancing at the literature and noticing that there are many recent papers with different arguments for \emph{where} the light comes from \cite{borisenok2020mechanisms,flannigan2013non,flannigan2012temperature,tatartchenko2017sonoluminescence}.
+We emphasive ``where" since it's not even certain if the light is \emph{surface} or \emph{volumetric} emission, i.e.
+it is literally not known \emph{where} the light comes from \cite{}.
+
+...
+and others stating that ``SBSL has become a rather sophisticated testing ground for the ability of mathematical models and numerical simulations to explain detailed experimental data from a complicated physical process" \cite{brenner2002single}.
+
+The goal of this paper is to catch the reader up on recent efforts to explain SBSL.
+While the author of this paper is particularly excited about shrimpoluminescence, it is important to stress that this work will focus on SBSL in general.
+It is apparently simpler to create and characterize SBSL in the laboratory without involving shrimp (for example, using shrimp involves convinving them to snap by tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}), so practically all work to study SL has not involved shrimp ( ...
+with notable exceptions \cite{tang2019bioinspired}).
+
+\section{Historical Overview}
+
+\section{What do we know?}
+\subsection{The Bubble Wall}
+\subsection{Experimental Results}
+what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}):
+\begin{itemize}
+    \item Ar content
+    \item driving frequency
+    \item P$_{max}$
+    \item ...
+\end{itemize}
+
+\section{What don't we know}
+\subsection{The Bubble Wall (Again)}
+\subsection{Experimental Anomalies}
+
+\section{Let there be light!}
+\subsection{Thermal Arguments}
+\subsection{Electric Arguments}
+
+\subsection{The Casimir Effect}
+\subsection{The Casimir Argument}
+
+\section{Summary}
+
+\section{Acknowledgments}
 
@@
 
-\emph{some abbreviations}
-\begin{itemize}
-    \item SL: sonoluminescence
-    \item SBSL: single-bubble sonoluminesence
-    \item MBSL: multi-bubble sonoluminescence
-    \item EOM: equation(s) of motion
-\end{itemize}
+\section{random notes}
 
-\section{rough-outline}
-\begin{itemize}
-    \item start the introduction with captivating example, i.e.
-shrimpoluminescence.
-describe what it is, how it was found, then transition into sonolouminescence.
-move on to explain how its not understood, various theories try.
-highlight these.
-then move on to discuss the experimental situation etc.
-    \item Section: experiment. basically layout what actually is known (by observation) to happen during sonoluminescence. facts only, not trying to explain.
-    \item explanations: then move on to theories that try to explain it: thermal, electrical. cover each of these, highlighting thier sucesses and failures.
-    \item QED theory: breifly cover this one
-\end{itemize}
+argon flash: a method of 'imaging explosions' similar to why argon is more useful for SBSL.
+argon has no internal DOF whereas air (i.e.
+N2, O2, water vapor) does and energy from explosions is absorbed dissociating bonds.
+argon, instead, becomes 'hotter' and emits more light, making photographing explosions easier.
+see wikipedia article "argon flash"
 
-\section{random notes}
+some dickheads 3d printed a shrimp claw and it also shrimpoluminesces \cite{tang2019bioinspired}
+
 apparently a shitty movie called 'Chain Reaction' with Keanu Reeves was centered around the premise of MBSL as a means of clean energy production.
@@
 need to find the primary lit.
+
+for fusions, also see wikipedia article on 'bubble fusion' and 'Rusi P Taleyarkhan'.
+Ruse was some asshole prof at purdue that claimed fusion occured inside cavitating bubbles, but it was later found out that the faked data and interfered with peer reviews leading to him being fucking fired.
 
@@
 in 1990s at U.
-Mississippi \cite{gaitan1990experimental,crum1994sonoluminescence,gaitan1992sonoluminescence} (note, \cite{ggaitan1992sonoluminescence} has $\sim$1000 citations..).
+Mississippi \cite{gaitan1990experimental,crum1994sonoluminescence,gaitan1992sonoluminescence} (note, \cite{gaitan1992sonoluminescence} has $\sim$1000 citations..).
 The SBSL experiments were done by sticking peizoacoustic transducers to the outside of a water filled glass.
@@
 \section{notes from \cite{brenner2002single}}
-defines SBSL as ``Single-bubble sonoluminescence occurs when an acoustically trapped and periodically driven gas
-bubble collapses so strongly that the energy focusing at collapse leads to light emission".
-
-\newpage
+defines SBSL as ``Single-bubble sonoluminescence occurs when an acoustically trapped and periodically driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission".
 

### 2021-12-08 sonoluminescence/paper/paper.tex
asdsad

@@
 
-\documentclass[rmp,aps,superscriptaddress,floatfix]{revtex4-2} 
+\documentclass[rmp,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
 \usepackage{graphicx}
@@
 \usepackage{bm}
+\bibliographystyle{apsrev4-2}
+\setcitestyle{numbers,square}
 
@@
 
-The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \cite{} to more exotic explanations based of quantum field theory \cite{schwinger1993casimir,eberlein1996sonoluminescence,liberati2000sonoluminescence}.
+The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are far too many to try list and we will not be interested in the vast majority of early models which are now irrelevant.
+References to these can be found in Gaitan \cite{gaitan1990experimental} and Brenner et al.
+\cite{brenner2002single}.} to more exotic explanations based of quantum field theory \cite{schwinger1993casimir,eberlein1996sonoluminescence,liberati2000sonoluminescence}.
 
-But besides shrimpoluminescence, why is SL interesting?
+But besides the obvious case of shrimpoluminescence, why is SL interesting?
 Well, at a glance, it is not obvious why SL occurs.
@@
 That is huge!
-After a moments thought, however, the massive energy focusing is not that surprising.
+Estimates of the temperature at the center bubble are ...
+
+%After a moments thought, however, the massive energy focusing is not that surprising.
 For example, the energy in the shock wave produced by a cavitating bubble, regardless of the light production, has been known to damage ship propellers since atleast 1917 when Lord Rayleigh's gave the first rigorous treatment of caviating bubbles \cite{gaitan1990experimental}.
-A heurestic argument goes like this: \cite{plesset1949dynamics}.
 
-So again, why is the light emmission interesting? In short, after 3 decades of studying SBSL, the mechanism of light production is still not understood.
+%So again, why is the light emmission interesting? In short, after 3 decades of studying SBSL, the mechanism of light production is still not understood.
 
-The excitment wasn't restricted to academics however, with SBSL catching Hollywood's attention leading to an unpopular movie called \emph{Chain Reaction} where t
+%The excitment wasn't restricted to academics however, with SBSL catching Hollywood's attention leading to an unpopular movie called \emph{Chain Reaction} where t
 
-So again, why is the light emmission interesting?
 After 3 decades of studying SBSL, the physical mechanism of light production in SL is still not understood.
-This can be seen by glancing at the literature and noticing that there are many recent papers with different arguments for \emph{where} the light comes from \cite{borisenok2020mechanisms,flannigan2013non,flannigan2012temperature,tatartchenko2017sonoluminescence}.
-We emphasive ``where" since it's not even certain if the light is \emph{surface} or \emph{volumetric} emission, i.e.
-it is literally not known \emph{where} the light comes from \cite{}.
-
-...
-and others stating that ``SBSL has become a rather sophisticated testing ground for the ability of mathematical models and numerical simulations to explain detailed experimental data from a complicated physical process" \cite{brenner2002single}.
+This can be seen by glancing at the literature and noticing that there are many recent papers with different arguments for \emph{where} \footnote{We emphasive ``where" since it's not even certain if the light is \emph{surface} or \emph{volumetric} emission, i.e.
+it is literally not known \emph{where} the light comes from \cite{}.} the light comes from \cite{borisenok2020mechanisms,flannigan2013non,flannigan2012temperature,tatartchenko2017sonoluminescence}.
+Numerous aspects of the process are conveniently accessible to the experimentalist, while at the same time the theory of the bubble's interior is quite mature.
+Many theories require experimental inputs as parameters, while in other cases experimental results are indirectly \emph{inferred} by fitting to a theory.
+Put simply, the history of studying SBSL is a rather beautiful example of the scientific process of explaining nature.
+Accorind to Brenner ``SBSL has become a rather sophisticated testing ground for the ability of mathematical models and numerical simulations to explain detailed experimental data from a complicated physical process" \cite{brenner2002single}.
 
@@
 While the author of this paper is particularly excited about shrimpoluminescence, it is important to stress that this work will focus on SBSL in general.
-It is apparently simpler to create and characterize SBSL in the laboratory without involving shrimp (for example, using shrimp involves convinving them to snap by tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}), so practically all work to study SL has not involved shrimp ( ...
-with notable exceptions \cite{tang2019bioinspired}).
+It is apparently simpler to create and characterize SBSL in the laboratory without involving shrimp (for example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}), so practically all work to study SL has not involved shrimp \footnote{ With notable exceptions \cite{tang2019bioinspired}}.
+Also notably absent will be any detailed discussion of MBSL since this will take us too far afield.
+Instead, we will summarize the history of SBSL research up to now starting at the discovery of MBSL leading to the subsequent discovery of SBSL and a rush of effort to explain it.
+We will explore both the experimental and theoretical discoveries along the way.
+Finally, we will review the current state of agreement between what is known experimentally and theoretically about SBSL.
 
@@
 
+SL was discovered by accident in 1933 (in the form of MBSL) by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz} \footnote{Neither the paper due to Marinesco et al.
+nor the one due to Frenzel et al.
+can be found by the author; in any case he can't read French or German so having the papers wouldn't be much use.
+As such, the story of how the effect was discovered is taken from more recent sources whose authors hopefully could read French and German \cite{brenner2002single,gaitan1990experimental,crum1994sonoluminescence}} Marinesco et al.
+were trying to accelerate photo development by \emph{insonating} developing fluid.
+They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy" which they attributed to exposure to light.
+Shortly after, Frenzel et al.
+repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
+
+This result was not particularly surprising to the community since it had been known for a while that cavitating bubbles could do tremendous damage to e.g.
+ships propellers.
+The discovery's impact is consicesly summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
+In fact, cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
+The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures coud become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence} and was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
+Shortly after, Lord Rayliegh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigoruous theoretical treatment of cavitation \cite{rayleigh1917pressure}.
+Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
+Still, very little information was accessible about the light emission until the 1990's when stable, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
+With this discovery, serious interest took root.
+
+Details of a typical SBSL experiment will be given later, but for now it suffices to note that SBSL, unlike previous studies on MBSL, allowed very precise control and measurment of the SL process.
+With unprecidented experimental control, many discrepencies in previous assumptions about SL were discovered.
+Experiments measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
+This discovery implied that SL was nearly decoupled from the bubble's dynamics, contradicting models based on Rayleigh's equation.
+New models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubbles $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
+At the same time, experimentalists fit the bubble's light emission spectra as a \emph{black body emitter} and concluded that the temperature in the bubble was atleast 25,000 K \cite{hiller1992spectrum}.
+
+These very high estimates for the temperature in the bubbles lead to a spark of interest more broadly: if the temerature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
+\textcolor{red}{ talk about the stupid ass movie and that asshole that faked science data.}
+Now talk about sonochemistry etc.
+
 \section{What do we know?}
 \subsection{The Bubble Wall}
+As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics.
+This isn't particularly surprising since the light emission and bubble collapse occur at different timescales \cite{}.
+The starting point for any theory of the bubble dynamics is Rayleigh's original expounded upon by many others \cite{}: the so called ``the Rayleigh-Plesset" (RP) equation \footnote{The RP equation is derived from the Navier-Stokes equation in the appendix}:
+\begin{equation}
+B=D lololol
+\label{eq:RP}
+\end{equation}
+
 \subsection{Experimental Results}
@@
 \section{Acknowledgments}
+
+\appendix
+
+\section{The Rayleigh-Plesset Equation}
+Derive the RP equation here
 

### 2021-12-13 klotski/klotski.py
assda

@@ -1,0 +1,272 @@
+"""
+note: still need to add logic for it to check if the current phase space coordinate is the solution 
+or not. actually, need to check if any of the moves for current coord are. if so, move there.
+"""
+
+import numpy as np
+
+# --------------------------------------------------------------------------------------------------
+
+class config:
+
+    def __init__(self,coords,ensemble):
+
+        """
+        hold block coords for this particular phase space coordinate. find neibhoring 
+        configurations and check if this config is connected to the solution
+        """
+
+        # this particular configs phase space coordinates
+        self.coords = coords
+        self.num_blocks = ensemble.num_blocks
+        self.blocks = ensemble.blocks
+        self.shapes = ensemble.shapes
+        self.puzzle_shape = ensemble.puzzle_shape
+        self.solution_block = ensemble.solution_block
+
+        # the one to check against the solution
+        self.masked = self.coords*(self.coords == self.solution_block).astype(int)- \
+                            (self.coords != self.solution_block).astype(int)
+
+        # check if we can move, check if is the solution 
+        self._check_moves(ensemble.solution,ensemble.history)
+
+    # ---------------------------------------------------------------------------------------------
+
+    def _check_moves(self,solution,history):
+        
+        """
+        find all neighboring configurations. not that since the coord we are moving from is 
+        in the history, we wont go back to it
+        """
+
+        self.moves = []
+        self.can_move = False
+        self.solved = False
+        self.solution = None
+
+        # loop over the blocks and see if they can be moved
+        for block in self.blocks:
+
+            # spaces occupied by this block
+            inds = np.argwhere(self.coords == block)
+            rows = inds[:,0].reshape(self.shapes[block])
+            cols = inds[:,1].reshape(self.shapes[block])
+
+            # check moves for this block
+            self._check_move('down',inds,rows,cols,block,history,solution)
+            self._check_move('up',inds,rows,cols,block,history,solution)
+            self._check_move('right',inds,rows,cols,block,history,solution)
+            self._check_move('left',inds,rows,cols,block,history,solution)
+
+    # ------------------------------------------------------------------------------------------
+
+    def _check_move(self,direction,inds,rows,cols,block,history,solution):
+
+        """
+        check if the any moves are possible
+        """
+
+        if direction == 'down':
+            shift = [1,0]
+            bound = rows.max()
+            shape = self.puzzle_shape[0]-1
+        elif direction == 'up':
+            shift = [-1,0]
[…]

### 2021-12-13 sonoluminescence/paper/paper.tex
asdsad

@@
 \section{The Rayleigh-Plesset Equation}
-Derive the RP equation here
+The Rayleigh-Plesset (RP) equation can be deduced from the Navier-Stokes equations \cite{}. To that end, we first remind the reader of the Navier-Stokes equations.
+
+The mass in an arbitrary element of a fluid (whose coordinates and boundary may depend on time) is given by  $\frac{d}{dt} \int_{\Omega_t} \rho(\bm{x},t) d\bm{x}$ where $\Omega_t$ is the domain of the fluid element under consdiration and $\rho(\bm{x},t)$ is the local, time-dependent density of the fluid.
+Using the transport theorem of fluid dynamics \cite{mcdonough2009lectures}, Gauss's theorem, and (with only some loss of generality) taking $\Omega_t$ to be a \emph{fluid parcel} (so that the boundary of the fluid parcel, $\partial \Omega_t$, moves with the fluid's velocity, $\bm{u}(\bm{x},t)$), this becomes
+\begin{equation}
+    \frac{d}{dt} \int_{\Omega_t} \rho dV = \int_{\Omega_t} \frac{\partial \rho}{\partial t} dV+\int_{\partial \Omega_t} \rho \bm{u} \cdot \bm{n} dA = \int_{\Omega_t} \frac{\partial \rho}{\partial t}+\nabla\cdot(\rho \bm{u})dV.
+    \label{eq:mass_continuity}
+\end{equation}
+Now to have \emph{mass-continuity}, which we will require on physical grounds, we set eq.
+\ref{eq:mass_continuity} to 0.
+Basically, we are saying no additional material is being added or removed from the fluid.
+Since the fluid parcel's boundary $\Omega_t$ is arbitrary, we find that $\frac{\partial \rho}{\partial t}+\nabla\cdot(\rho \bm{u})\equiv 0$ which is true for every point in the fluid volume.
+
+Newton's second law for fluids (\emph{momentum-continuity}) is defined similarly.
+We write the momentum-density as $\rho \bm{u}$.
+Following the same steps as above, we find for e.g.
+the $i^{th}$ component of momentum
+\begin{equation}
+    \frac{d}{dt} \int_{\Omega_t} \rho u_x dV = \int_{\Omega_t} \frac{\partial (\rho u_x)}{\partial t} dV+\int_{\partial \Omega_t} \rho \bm{u} \cdot \bm{n} dA = \int_{\Omega_t} \frac{\partial \rho}{\partial t}+\nabla\cdot(\rho \bm{u})dV.
+    \label{eq:mom_continuity}
+\end{equation}
+\textcolor{red}{note, in Schoeffel's notes, he writes down an equation involving the curl.
+he also points out that, for irrotational flow, we can write v as the gradient of a scalar function and derive a simple eq.
+this is what is done in Brenner's review.}
 

### 2021-12-15 sonoluminescence/paper/paper.tex
asdsad

--- 2021-12-08 sonoluminescence/paper/paper.tex
@@
 \section{The Rayleigh-Plesset Equation}
-Derive the RP equation here
+The Rayleigh-Plesset (RP) equation can be deduced from the compressible Navier-Stokes equations \cite{}. To that end, we first remind the reader of the Navier-Stokes equations.
+
+The \emph{total} change (in time) of mass in an arbitrary element of a fluid (whose coordinates and boundary may depend on time) is given by  $\frac{d}{dt} \int_{\Omega_t} \rho dV$ where $\Omega_t$ is the domain of the fluid element under consdiration and $\rho(\bm{x},t)$ is the local, time-dependent density of the fluid.
+Using the transport theorem of fluid dynamics \cite{mcdonough2009lectures}, Gauss's theorem, and (with only some loss of generality) taking $\Omega_t$ to be a \emph{fluid parcel} (so that the boundary of the fluid parcel, $\partial \Omega_t$, moves with the fluid's velocity, $\bm{u}(\bm{x},t)$), this becomes
+\begin{equation}
+    \frac{d}{dt} \int_{\Omega_t} \rho dV = \int_{\Omega_t} \frac{\partial \rho}{\partial t} dV+\int_{\partial \Omega_t} \rho \bm{u} \cdot \bm{n} dA = \int_{\Omega_t} \frac{\partial \rho}{\partial t}+\nabla \cdot(\rho \bm{u})dV.
+    \label{eq:mass_continuity}
+\end{equation}
+Now to have \emph{mass-continuity}, which we will require on physical grounds, we set eq.
+\ref{eq:mass_continuity} to 0.
+Basically, we are saying no additional material is being added or removed from the fluid.
+Since the fluid parcel's boundary $\Omega_t$ is arbitrary, we find that $\frac{\partial \rho}{\partial t}+\nabla\cdot(\rho \bm{u})\equiv 0$ which is true for every point in the fluid volume.
+
+Newton's second law for fluids (\emph{momentum-continuity}) is defined similarly.
+We write the momentum-density as $\rho \bm{u}$ with momentum given by $\int_{\Omega_t} \rho \bm{u} dV$.
+To simiplify what follows, we focus on only the $i^{th}$ component of velocity at a time.
+Then Newton's second law can be written as
+\begin{equation}
+    \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \frac{\partial (\rho u_i)}{\partial t} dV+\int_{\partial \Omega_t} \rho u_i ( \bm{u} \cdot \bm{n}) dA = F_i.
+\end{equation}
+Now using Gauss's theorem, the product rule, and mass-continuity we find $ \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \rho \frac{d u_i}{dt} dV = F_i $. In vector notation this reads
+\begin{equation}
+    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt} = \bm{F}
+    \label{eq:mom_continuity}
+\end{equation}
+
+Now let us seperate the forces into \emph{body-} and \emph{surface-forces}: $\bm{F}=\int_{\Omega_t} \bm{f}_B dV+ \int_{\partial \Omega_t} \bm{f}_S dA $.
+If we only consider an ideal fluid, the surface part can be written as $\int_{\partial \Omega_t} \bm{f}_S dA =-\int_{\partial \Omega_t} p \bm{n} dA = -\int_{\Omega_t}\nabla p dV$ with $p$ the pressure  and the minus sign being due to the sign convention: \emph{out} of the surface $\partial \Omega_t$ is positive.
+If instead the fluid is real, we can't neglect viscous forces.
+We suppose that viscous surface term can be represented by a tensor, $\bm{\tau}$, such that $\bm{f}_S = -p\bm{n}+\bm{\tau} \cdot \bm{n}$ with diagonal elements of $\bm{\tau}$ the \emph{normal} viscous stresses and the other components being \emph{shear} stresses.
+Note here that $\bm{\tau} \cdot \bm{n}$ denotes matrix multipication.
+It is conventional to combine the pressure and viscous stresses into a single entity called the \emph{stress-tensor}.
+We denote it by $\bm{\sigma}=-p\bm{I}+\bm{\tau}$.
+
+The components of the viscous stress tensor $\bm{\tau}$ can be deduced from Newton's law of viscosity and combined with symmetry of the tensor itself \cite{mcdonough2009lectures}.
+We quote the results here \footnote{We are assuming an \emph{isotropic} fluid so that the the viscosities $\eta$, $\mu$ are the same in all directions.
+This is true for water.}:
+\begin{align}
+    \tau_{ii} & = 2 \eta \frac{\partial u_i}{\partial x_i} \\
+    \tau_{ij} & =  \mu \left( \frac{\partial u_i}{\partial x_j}+\frac{\partial u_j}{\partial x_i} \right).
+\end{align}
+
+The body-force density, $\bm{f}_B$, is due to forces acting in the bulk of the liquid e.g.
+gravity, electrostatic forces, etc.
+and is typically from sources external to the liquid.
+Then the total force on a fluid element is $\bm{F}=\int_{\Omega_t} \bm{f}_B+\nabla \cdot \bm{\sigma} dV$.
+We can succinctly summarize the momentum-continuity equation as
+\begin{equation}
+    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt}-\bm{f}_B-\nabla \cdot \bm{\sigma} dV =0
+\end{equation}
+which is equivalent to $\frac{d \bm{u}}{dt}=\frac{1}{\rho}\left[ \bm{f}_B+\nabla \cdot \bm{\sigma} \right]$.
+
+We would rather have an equation explicitly involving the velocity field $\bm{u}$. To that end, let us work out the stress term.
 

### 2021-12-15 sonoluminescence/paper/paper.tex
asdsad

--- 2021-12-08 sonoluminescence/paper/paper.tex
@@
 
-The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are far too many to try list and we will not be interested in the vast majority of early models which are now irrelevant.
+{\color{red}According to Suslick et al \cite{suslick2008inside}, SBSL makes it difficult to directly measure temperature and pressure in the bubble due to its tiny volume.
+On the otherhand, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volume emission from the bubble cloud, etc.
+Similiarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls.}
+
+The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are far too many to try list here and we will not be interested in the vast majority of early models which are now irrelevant.
 References to these can be found in Gaitan \cite{gaitan1990experimental} and Brenner et al.
@@
 The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures coud become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence} and was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
-Shortly after, Lord Rayliegh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigoruous theoretical treatment of cavitation \cite{rayleigh1917pressure}.
+Shortly after, Lord Rayliegh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigoruous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
 Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
@@
 The starting point for any theory of the bubble dynamics is Rayleigh's original expounded upon by many others \cite{}: the so called ``the Rayleigh-Plesset" (RP) equation \footnote{The RP equation is derived from the Navier-Stokes equation in the appendix}:
-\begin{equation}
-B=D lololol
-\label{eq:RP}
-\end{equation}
 
@@
 \section{The Rayleigh-Plesset Equation}
-Derive the RP equation here
+The Rayleigh-Plesset (RP) equation can be deduced from the compressible Navier-Stokes equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
+To that end, we first remind the reader of the Navier-Stokes equations.
+
+The \emph{total} change (in time) of mass in an arbitrary element of a fluid (whose coordinates and boundary may depend on time) is given by  $\frac{d}{dt} \int_{\Omega_t} \rho dV$ where $\Omega_t$ is the domain of the fluid element under consdiration and $\rho(\bm{x},t)$ is the local, time-dependent density of the fluid.
+Using the transport theorem of fluid dynamics \cite{mcdonough2009lectures}, Gauss's theorem, and (with only some loss of generality) taking $\Omega_t$ to be a \emph{fluid parcel} (so that the boundary of the fluid parcel, $\partial \Omega_t$, moves with the fluid's velocity, $\bm{u}(\bm{x},t)$), this becomes
+\begin{equation}
+    \frac{d}{dt} \int_{\Omega_t} \rho dV = \int_{\Omega_t} \partial_t \rho dV+\int_{\partial \Omega_t} \rho \bm{u} \cdot \bm{n} dA = \int_{\Omega_t} \partial_t \rho+\nabla \cdot(\rho \bm{u})dV.
+    \label{eq:mass_continuity}
+\end{equation}
+Now to have \emph{mass-continuity}, which we will require on physical grounds, we set eq.
+\ref{eq:mass_continuity} to 0.
+Basically, we are saying no additional material is being added or removed from the fluid.
+Since the fluid parcel's boundary $\Omega_t$ is arbitrary, we find that $\partial_t \rho+\nabla\cdot(\rho \bm{u})\equiv 0$ which is true for every point in the fluid volume.
+
+Newton's second law for fluids (\emph{momentum-continuity}) is defined similarly \cite{schoeffel2014lecture,mcdonough2009lectures}.
+We write the momentum-density as $\rho \bm{u}$ with momentum given by $\int_{\Omega_t} \rho \bm{u} dV$.
+To simiplify what follows, we focus on only the $i^{th}$ component of velocity at a time.
+Then Newton's second law can be written as
+\begin{equation}
+    \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \partial_t (\rho u_i) dV+\int_{\partial \Omega_t} \rho u_i ( \bm{u} \cdot \bm{n}) dA = F_i.
+\end{equation}
+Now using Gauss's theorem, the product rule, and mass-continuity we find $ \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \rho \frac{d u_i}{dt} dV = F_i $. In vector notation this reads
+\begin{equation}
+    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt} = \bm{F}
+    \label{eq:mom_continuity}
+\end{equation}
+
+We now turn to the forces.
+Let us seperate the forces into \emph{body-} and \emph{surface-forces}: $\bm{F}=\int_{\Omega_t} \bm{f}_B dV+ \int_{\partial \Omega_t} \bm{f}_S dA $.
+The body-force density, $\bm{f}_B$, is due to forces acting in the bulk of the liquid e.g.
+gravity, electrostatic forces, etc.
+and is typically from sources external to the liquid.
+We won't be concerned with these forces here.
+Instead, let us concentrate on the suface part.
+If we only consider an ideal fluid, the surface part can be written as $\int_{\partial \Omega_t} \bm{f}_S dA =-\int_{\partial \Omega_t} p \bm{n} dA = -\int_{\Omega_t}\nabla p dV$ with $p$ the pressure  and the minus sign being due to the sign convention: \emph{out} of the surface $\partial \Omega_t$ is positive.
+
+If instead the fluid is real, we can't neglect viscous forces (i.e.
+dissipation).
+We suppose that the viscous surface term can be represented by a tensor, $\hat{\bm{\tau}}$ such that $\bm{F}_v=\int_{\partial \Omega_t} \hat{\bm{\tau}}\cdot \bm{n} dA = \int_{\Omega_t} \nabla \cdot \hat{\bm{\tau}} dV$.
+(Note, here the dot product is to be understood as matrix multiplication).
+To leading order in gradients of the velocity, (and under the physical constraints that the tensor be symmetric), the viscous-stress tensor is $\hat{\bm{\tau}}=\hat{\bm{\mu}}\cdot \hat{\bm{\epsilon}}$ where $\hat{\bm{\mu}}$ is the viscosity-tensor and $\hat{\bm{\epsilon}}$ is the strain-rate tensor: $\epsilon_{ij}=\frac{1}{2}\left[ \partial_i u_j+ \partial_j u_i \right]$.
+
+In general, $\hat{\bm{\mu}}$ is a symmetric rank-4 tensor.
+However, if we assume an isotropic fluid \footnote{Definitely true for water which is usually the fluid used in sonoluminescence experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
+\begin{equation}
+\begin{split}
+    \hat{\bm{\tau}} & = \xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)} \\
+    \hat{\bm{\epsilon}}^{(v)} & = \delta_{ij}\epsilon_{kk} \\
+    \hat{\bm{\epsilon}}^{(s)} & = \hat{\bm{\epsilon}}-\frac{1}{3}\delta_{ij}\epsilon_{kk}.
+    \label{eq:viscous_stress_tensor}
+\end{split}
+\end{equation}
+Thus, we see that the viscosity-tensor only has two free components: $\xi$, the normal viscosity and $\mu$ the shear viscosity.
+It is usual to combine the pressure (i.e.
+elastic stresses) and the viscous-stresses into the stress-tensor $\hat{\bm{\sigma}}=-p\hat{\bm{I}}+\xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)}$.
+Combined with this, Newton's second law becomes
+\begin{equation}
+    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt}-\bm{F} = \int_{\Omega_t} \rho \frac{d \bm{u}}{dt} - \nabla \cdot \hat{\bm{\sigma}}-\bm{f}_B dV = 0
+\end{equation}
+Again, since we allow the volume of integration to be arbitrary, the integrand must vanish everywhere: $\rho \frac{d \bm{u}}{dt} - \nabla \cdot \hat{\bm{\sigma}}-\bm{f}_B=0$.
+This is the momentum-continuity equation for a fluid.
+Now we would rather have this explicitly in terms of the velocity $\bm{u}$.
+To that end, we may subsitute eq.
+\ref{eq:viscous_stress_tensor} into the momentum-continuity equation.
+For e.g.
+the $i^{th}$ component we may expand out the divergence part as
+\begin{equation}
+\begin{split}
+    \left(\nabla \cdot \hat{\bm{\sigma}}\right)_i = -\partial_i p + \left[\xi +\frac{2}{3}\mu \right] \partial_i(\nabla \cdot \bm{u})+\mu \nabla^2 u_i \equiv -\partial_i p + \eta \partial_i(\nabla \cdot \bm{u})+\mu \nabla^2 u_i
+\end{split}
+\end{equation}
+where we have introduced the \emph{bulk-viscosity} $\eta=\xi+\frac{2}{3}\mu$.
+Collecting all of the components into vector form and, combined with the mass-continuity equation \ref{eq:mass_continuity}, we arrive at the \emph{Navier-Stokes} (NS) equations for a compressible fluid:
+\begin{equation}
+\begin{split}
+    \rho \frac{d \bm{u}}{dt} = \rho \left[ \partial_t \bm{u}+\left(\bm{u}\cdot \nabla \right)\bm{u} \right] & =-\nabla p + \eta \nabla \left(\nabla \cdot \bm{u} \right)+\mu \nabla^2 \bm{u} + \bm{f}_B \\
+     \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
+     \label{eq:ns_equations}
+\end{split}
+\end{equation}
+
+Some comments are in order.
+The total-derivative of a moving fluid $\frac{d}{dt}\equiv D_t$ is sometimes called the \emph{material derivative}.
+In the case of velocity $D_t \bm{u}$ is the \emph{acceleration} of the fluid.
+It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies \emph{in space} as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
+The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
+On the other hand, the terms $\sim \bm{u}$ are viscous (damping) terms that tend to make the velocity field spatially uniform: the vanish when the (spatial) derivatives of the velocity vanish.
+The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
+dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
+The body term $\bm{f}_B$ is from external sources: from here on, we will assume it is zero.
+
+Even with these simplifying assumptions, solving eqs.
+\ref{eq:ns_equations} is a formidable task \footnote{In fact, it's not even clear that the NS equations \emph{can} be solved even for incompressible fluids.
+Proving that smooth, sensible solutions exist earns \$1,000,000 \cite{fefferman2006existence}.
+Better hurry though...
+inflation was 6.8\% in Nov.
+2021} To that end, several more drastic approximations have to be made that allow us to make progess.
+\begin{enumerate}
+    \item We will assume irrotational flow (i.e.
+$\nabla\times\bm{u}\equiv 0$) and only radial motion in the liquid (i.e.
+$\bm{u}=u\bm{r}$).
+Note that, since the flow is assumed irrotation, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$.
+This amounts to assuming that the bubble is always spherical.
+This seems like a rather drastic approximation but has been validated experimentally in many cases \cite{}.
+This can be understood by accounting for surface-tension at the liquid-bubble interface \cite{}.
+    \item Next, we assume that the viscous terms are negligible in the bulk dyanimcs of the liquid.
+We will acount, to some extent, for viscosity later when looking at the bubble-liquid interface.
+Moreover, we are assuming that the flow is \emph{isentropic}, i.e.
+that it is reversible (no damping) and that no heat is exchanged between fluid parcels.
+We will further assume that the liquid is isothermal, i.e.
+its temperature is constant (in space).
+Then pressure $p$ is determined from an instantaneous equation-of-state $p=p(\rho,T)$.
+In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and, as we will later see, heat-transport across the liquid-bubble interfacing is usually neglected anyway (i.e.
+the bubble is compressed adiabatically in the thermodynamic sense \footnote{According to Wikipedia, \emph{adiabatic} means \emph{fast} in thermodynamics lingo.
+This is relevant to us since we are claiming the bubble wall moves so quickly that it is compressed to its maximum pressure before any heat can flow out.
+On the otherhand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
+This is, e.g.
+the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its groundstate at all times.}
+    \item Related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
+Of course the bubble-liquid interface enters both systems as a boundary condition.
+\end{enumerate}
 
@@
 see wikipedia article "argon flash"
+
+Brennen defines caviation as (bubble) nucleation when the pressure drops below the vapor pressure...
+as compared to boiling in which the temperature goes above the saturated liquid/vapor temp.
+\cite{brennen2014cavitation}.
+
+Yasui defines \emph{acoustic cavitation} as the formation and collapse of a bubble driven by a powerful ultrasonic wave \cite{yasui2018acoustic}.
+He defines bubble dynamics as bubble pulsation under the driving sound.
+\textcolor{red}{read this book}
 

### 2021-12-15 sonoluminescence/paper/paper.tex
asdsad

@@
 These very high estimates for the temperature in the bubbles lead to a spark of interest more broadly: if the temerature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
-\textcolor{red}{ talk about the stupid ass movie and that asshole that faked science data.}
-Now talk about sonochemistry etc.
+\textcolor{red}{ talk about the stupid ass movie and that asshole that faked science data.
+and then about sonochemistry etc}
 
@@
 \subsection{The Bubble Wall}
-As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics.
-This isn't particularly surprising since the light emission and bubble collapse occur at different timescales \cite{}.
-The starting point for any theory of the bubble dynamics is Rayleigh's original expounded upon by many others \cite{}: the so called ``the Rayleigh-Plesset" (RP) equation \footnote{The RP equation is derived from the Navier-Stokes equation in the appendix}:
+As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
+This isn't particularly surprising since the light emission and bubble collapse occur at different timescales \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
+The starting point for any theory of the bubble dynamics is Rayleigh's original expounded upon by many others \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}: the so called ``the Rayleigh-Plesset" (RP) equation \footnote{The RP equation is derived from the Navier-Stokes equation in the appendix}:
 
 \subsection{Experimental Results}
-what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}):
-\begin{itemize}
-    \item Ar content
-    \item driving frequency
-    \item P$_{max}$
-    \item ...
-\end{itemize}
+what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}) (Ar content, driving frequency, P$_{max}$)
 
@@
 inflation was 6.8\% in Nov.
-2021} To that end, several more drastic approximations have to be made that allow us to make progess.
+2021} To that end, several more-drastic approximations have to be made that allow us to make progess.
 \begin{enumerate}
@@
 $\bm{u}=u\bm{r}$).
-Note that, since the flow is assumed irrotation, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$.
+Note that, since the flow is assumed irrotation, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
 This amounts to assuming that the bubble is always spherical.
@@
     \item Next, we assume that the viscous terms are negligible in the bulk dyanimcs of the liquid.
+For (relatively) low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation and is widely used \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single}.
 We will acount, to some extent, for viscosity later when looking at the bubble-liquid interface.
@@
 \end{enumerate}
+
+With these simplifying assumptions (and a vector identity \footnote{$\left( \nabla \cdot \bm{u} \right) \bm{u} = \frac{1}{2} \nabla (u^2) - \bm{u}\times \left( \nabla \times \bm{u} \right)$}) we rewrite eqs.
+\ref{eq:ns_equations} as a system of $1+1$ dimensional partial differential equations:
+\begin{equation}
+\begin{split}
+     \partial_r \left[ \partial_t \phi +\frac{1}{2} \left( \partial_r \phi \right)^2 \right] & = - \frac{1}{\rho} \partial_r p  \\ 
+     \partial_t \rho+ \partial_r \phi \partial_r \rho + \rho \partial^2_r \phi & = 0
+%     \label{eq:ns_equations}
+\end{split}
+\end{equation}
 

### 2022-01-04 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 
-\documentclass[rmp,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
+\documentclass[rmp,aps,nofootinbib,superscriptaddress,floatfix,10pt]{revtex4-2} 
 \usepackage{graphicx}
@@
 The water ejection and subsequent bubble formation and, finally, bubble collape (at t=0) are shown on the right.
-The units are ms.
+The time-offset are shown in the panels in ms.
 Adapted from ref.
@@
 Sonoluminescence comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonolumiscence.
-The disticntion is self-explanatory: multi-bubble sonolumiscence (MBSL) consists of  ``the simultaneous creation and destruction of many seperate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
-The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light very high precisions \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
-We will discuss MBSL breifly later, but for now it suffices to say that nearly all theoretical and experimental progress has been made using SBSL, with some authors even calling it ``the hydrogen atom of sonoluminescence" \cite{lohse2018bubble,crum1994sonoluminescence}.
+The distinction is self-explanatory: multi-bubble sonolumiscence (MBSL) consists of  ``the simultaneous creation and destruction of many seperate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
+The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
+We will discuss MBSL breifly throughout this paper, but for now it suffices to say that nearly all theoretical and experimental progress to explain sonoluminescence has been made using SBSL, with some authors even calling it ``the hydrogen atom of sonoluminescence" \cite{lohse2018bubble,crum1994sonoluminescence}.
 
-{\color{red}According to Suslick et al \cite{suslick2008inside}, SBSL makes it difficult to directly measure temperature and pressure in the bubble due to its tiny volume.
-On the otherhand, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volume emission from the bubble cloud, etc.
-Similiarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls.}
-
-The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are far too many to try list here and we will not be interested in the vast majority of early models which are now irrelevant.
+In SBSL, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volumetric-light emission from the bubble cloud, etc.
+Similiarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls and the theory is greatly simplified compared to MBSL.
+SBSL, while offering advantages, also comes with complications.
+Due to the tiny volume of the bubble, directly measuring the temperature or pressure \emph{inside} the bubble is practically impossible \cite{suslick2008inside}.
+Still, the discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are far too many to try list here and we will not be interested in the vast majority of early models which are now irrelevant.
 References to these can be found in Gaitan \cite{gaitan1990experimental} and Brenner et al.
@@
 This result was not particularly surprising to the community since it had been known for a while that cavitating bubbles could do tremendous damage to e.g.
-ships propellers.
+ship's propellers \footnote{A more modern example could be my van's oil pump, which is in need of replacing.}.
 The discovery's impact is consicesly summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
 In fact, cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
-The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures coud become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence} and was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
+The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures could become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence}.
+Cavitation was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
 Shortly after, Lord Rayliegh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigoruous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
 Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
-Still, very little information was accessible about the light emission until the 1990's when stable, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
+Still, very little information was accessible about the light emission until the 1990's when stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing caviation.
+A stable bubble persists through multiple cycles of cavitation, oscillating nonlinearly around an equilibrium size.
+Transient bubbles appear and then collapse to dissapear}, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
 With this discovery, serious interest took root.
@@
 With unprecidented experimental control, many discrepencies in previous assumptions about SL were discovered.
-Experiments measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
+Experimental Results measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
 This discovery implied that SL was nearly decoupled from the bubble's dynamics, contradicting models based on Rayleigh's equation.
@@
 These very high estimates for the temperature in the bubbles lead to a spark of interest more broadly: if the temerature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
-\textcolor{red}{ talk about the stupid ass movie and that asshole that faked science data.
-and then about sonochemistry etc}
+An action movie featuring Keanu Reeves was even made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
+In the early 2000's, several papers were even published claiming that stable nuclear fusion was possible in the lab using a method based on SBSL \cite{}.
+Unfortunately these results turned out to be fraudulent \cite{}.
+Shortly after this fiasco, evidenced emerged that there are shape instabilities in the bubbles dynamics indicating that the converging shock-wave hypothesis is probably not right.
+This severely reduced the estimates for the temperature to below $10,000$ K, ruling out the possibility of fusion occuring at all \cite{}.
+Moreover, it was shown that the light is most likely emitted from the \emph{bulk} of the bubble as opposed to its surface, making fits to a black body specturm (i.e.
+surface emission) incorrect \cite{}.
+Current estimates for the temperature at the center of a sonoluminescing bubble are $...$ \cite{}
 
-\section{What do we know?}
+Still, the very high pressures and temperatures in the bubble during sonoluminescence see technological applications in chemistry and biosciences. ...
+
+History has shown that the most important part of understanding SBSL is understanding the dyanmics of a single acoustically driven bubble.
+We will see that the dynamics of the gas \emph{inside} the bubble is nearly decoupled from the dyanmics of the \emph{fluid} containing the bubble so that we may look at these two problems seperately.
+Firstly, we look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
+Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
+After that, we will combine these two and discuss solutions of the full problem.
+The ultimate result of this will turn out to be an estimate of the \emph{temperature} in the bubble.
+We will then look at theories presenting the origin of light from the very hot contents of the bubble.
+Finally, we will wrap this whole thing up with a summary of unsolved problems and an outlook to the future.
+
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figs/flask.pdf}
+    \caption{(a) Schematic of a spherical acoustic levitation cell.
+The piezoelectirc transducers that drive the bubble trapped in the flask are labeled in the diagram.
+(b) Photograph of the same.
+Here, the fluid is H$_2$SO$_4$ instead of water.
+A sonoluminescing bubble is clearly visible in the center of the flask.
+The photo was taken in a fully lit room with exposure time 2 s.
+(c) Schematic of a Mie scattering experimental setup.
+A similar acoustic levitation cell as (a) is shown with an added laser source and photomultiplier tubes (PMT) to analyze the scattered and emitted light.
+(a), (b), and (c) are from refs.
+\cite{brenner2002single}, \cite{suslick2008inside}, and \cite{gompf2000mie} respectively.}
+\label{fig:flask}
+\end{figure}
+
+\section{Bubble Dyanmics}
+The conditions in the bubble's \emph{interior} are still not entirely understood and will be the subject of an entire subsection of this paper.
+Before we get to this, we will at an older problem: the dynamics of a periodically driven bubble in a liquid.
+An important aspect of the bubble's dyanmics is that they are experimentally accessible.
+Moreover, as we shall see, a (relatively) simple theory describles the dynamics of the bubble quite well.
+On the other hand, measuring what is going on in the bubble's interior is not at all simple though much progress has been made theoretically.
+
+%Before we look at what we don't know, it is useful to summarize what we \emph{do} know about SBSL.
+This has two parts: on the one hand, the dynamics of the bubble are experimentally accessible and have been well studied by \emph{Mie scattering} experiments.
+As we shall see, a (relatively) simple theory describles the dynamics of the bubble quite well.
+On the other hand, a great deal of effort has been invested in measuring what is going on in the bubble's interior.
+We will look at what we have been able to figure out.
+
 \subsection{The Bubble Wall}
 As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
-This isn't particularly surprising since the light emission and bubble collapse occur at different timescales \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
-The starting point for any theory of the bubble dynamics is Rayleigh's original expounded upon by many others \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}: the so called ``the Rayleigh-Plesset" (RP) equation \footnote{The RP equation is derived from the Navier-Stokes equation in the appendix}:
+This isn't particularly surprising since the light emission and bubble collapse occur at different timescales (the duration of maximum compressions is $\sim 0.003\%$ of the bubble's period) \cite{suslick2008inside,prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
+The starting point for any theory of the bubble dynamics is Rayleigh's original work \cite{rayleigh1917pressure}, now expounded upon by many others.
+The most widely known modern work is that of Plesset \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}, resulting in the so called ``the Rayleigh-Plesset" (RP) equation.
+The derivation assumes the host liquid is incompressible, ultimately resulting in us neglecting the effects of sound radiated by the bubble.
+Corrections exist that improve upon the RP equation in various ways and we will discuss these below \cite{}.
+However, for brevity's sake, we will only derive the RP equation here and merely quote the others as they are very similar in form to the RP equation.
+The field of \emph{bubble dynamics} is quite mature and attempting to review it here would be out of scope of this paper; instead, the reader is referred to numerous books and reviews \cite{}.
+
+The RP equation can be deduced from the \emph{Navier-Stokes} (NS) equations \footnote{The NS equations are derived in the appendix} \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
+The NS equations for a compressible fluid are
+\begin{equation}
+\begin{split}
+    \rho \frac{d \bm{u}}{dt} = \rho \left[ \partial_t \bm{u}+\left(\bm{u}\cdot \nabla \right)\bm{u} \right] & =-\nabla p + \eta \nabla \left(\nabla \cdot \bm{u} \right)+\mu \nabla^2 \bm{u} + \bm{f}_B \\
+     \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
+     \label{eq:NS_equations}
+\end{split}
+\end{equation}
+
+Here, $\bm{u}(\bm{x},t)$ is the velocity of a fluid parcel at position $\bm{x}$ at time $t$, $\rho (\bm{x},t)$ is the fluid's density, $p$ is the pressure, $\eta \equiv \left(\xi + \frac{2}{3}\mu \right)$ is the \emph{bulk} viscosity with $\mu$ the \emph{shear} viscosity, and $\bm{f}_B$ are \emph{body} forces (e.g.
+gravity, electrostatic forces) which are usually from external sources and will be assumed negligible from here on.
+The subscrits on $\partial$ denote what deriviative is being taken and $\nabla$ and $\nabla^2$ have thier usual meanings.
+The first equation is \emph{conservation of momentum} for the fluid, i.e.
+Newton's $2^{nd}$ law; the second equation is \emph{mass convervation}.
+The total-derivative of a moving fluid $\frac{d}{dt}\equiv D_t$ is sometimes called the \emph{material derivative}.
+In the case of velocity $D_t \bm{u}$ is the \emph{acceleration} of the fluid.
+It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies \emph{in space} as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
+The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
+On the other hand, the terms $\sim \bm{u}$ are viscous (damping) terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
+The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
+dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
+
+Solving eqs.
+\ref{eq:NS_equations} is an intractable problem.
+To make progress, we will make some drastic approximations that are more-or-less valid for our problem.
+Firslty, we will assume irrotational flow (i.e.
+$\nabla\times\bm{u}\equiv 0$) so that there is only radial motion in the liquid (i.e.
+$\bm{u}=u\bm{r}$).
+Note that, since the flow is assumed irrotational, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
+This amounts to assuming that the bubble is always spherical which seems like a rather drastic approximation but has been validated experimentally in many cases \cite{}.
+The fact that the bubble tends to remain spherical can be understood by accounting for surface-tension at the liquid-bubble interface \cite{prosperetti1999old}.
+
+Nextly, we assume that the viscous terms are negligible in the bulk dyanimcs of the liquid.
+For (relatively) low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation and is widely used \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
+We will acount, to some extent, for viscosity later when looking at the bubble-liquid interface.
+Moreover, we are assuming that the flow is \emph{isentropic}, i.e.
+that it is reversible (no damping) and that no heat is exchanged between fluid parcels.
+We will further assume that the liquid is isothermal, i.e.
+its temperature is constant (in space).
+Then pressure $p$ is determined from an instantaneous equation-of-state $p=p(\rho,T)$.
+In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and, as we will later see, heat-transport across the liquid-bubble interfacing is usually neglected anyway (i.e.
+the bubble is compressed adiabatically in the thermodynamic sense \footnote{According to Wikipedia, \emph{adiabatic} means \emph{fast} in thermodynamics lingo.
+This is relevant to us since we are claiming the bubble wall moves so quickly that it is compressed to its maximum pressure before any heat can flow out.
+On the otherhand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
+This is, e.g.
+the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its groundstate at all times.})
+
+Lastly, related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
+Of course the bubble-liquid interface enters both systems as a boundary condition \cite{prosperetti1986bubble}.
+
+With these simplifying assumptions (and a vector identity \footnote{$\left( \nabla \cdot \bm{u} \right) \bm{u} = \frac{1}{2} \nabla (u^2) - \bm{u}\times \left( \nabla \times \bm{u} \right)$}) we rewrite eqs.
+\ref{eq:NS_equations} as a system of $1+1$ dimensional partial differential equations:
+\begin{equation}
+\begin{split}
+     \partial_r \left[ \partial_t \phi +\frac{1}{2} \left( \partial_r \phi \right)^2 \right] & = - \frac{1}{\rho} \partial_r p  \\ 
+     \partial_t \rho+ \partial_r \phi \partial_r \rho + \rho \partial^2_r \phi & = 0.
+     \label{eq:NS_1D}
+\end{split}
+\end{equation}
+We can integrate the first equation above using the fact that the compressibility is negligible to move the density through the integration \cite{leighton2007derivation}
+\begin{equation}
+    \int_{r_\infty}^{r} \frac{\partial}{\partial r'} \left[ \frac{\partial \phi}{\partial t} +\frac{1}{2} \left( \frac{\partial \phi}{\partial r'} \right)^2 \right] dr' = - \frac{1}{\rho} \int_{r_\infty}^{r} \frac{\partial p}{\partial r'} dr'.
+\end{equation}
+On the left side, we take the velocity to vanish at infinity.
+On the right side, we assume that only the static and unperturbed pressures, $p_\infty$ and $P(t)$ respectively, are relevant so that \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}
+\begin{equation}
+    \partial_t \phi + \frac{1}{2}\left( \partial_r \phi \right)^2 = \frac{(p_\infty+P(t))-p(t,r)}{\rho}.
+    \label{eq:press_vel}
+\end{equation}
+Eq. \ref{eq:press_vel} gives the pressure at cordinate $(r,t)$ in terms of the velocity of the fluid, the density, and the appied pressure.
+
+Now, let us assume that the velocity potential satisfies a wave equation $\nabla^2 \phi - \frac{1}{c^2}\partial_t^2 \phi =0 $ \footnote{This doesn't come from thin air.
+We could write the r.h.s.
+of eq.
+\ref{eq:press_vel} as the enthalpy, $h\equiv \int \frac{dp}{\rho}$, and similarly introduce $dh=\frac{dp}{\rho}$ in the first line of eq.
+\ref{eq:NS_1D}.
+Eliminating $h$ between the two and dropping small terms $\sim \frac{u}{c}$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.} with $c$ the speed of sound in the fluid.
+Again, we use incompressibilty by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$ \footnote{For our problem, it is true that $\left( \frac{1}{c^2} \partial_t \phi \right) / \nabla^2 \phi \sim \left( \frac{R}{\lambda} \right)^2 \ll 1$, i.e.
+the long wave-length approximation is accurate \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}.} .
+In this case, we can ignore retardation effects and, remembering that we have spherical symmetry, write the solution of $\nabla^2 \phi$ as
+\begin{equation}
+    \phi = \frac{\psi(t)}{r}
+\end{equation}
+with $\psi$ the time-dependent coefficient. We now use the boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, to determine $\psi$:
+\begin{equation}
+    \dot{R} = -\frac{\psi}{R^2} \Rightarrow \psi = -R^2 \dot{R}.
+\end{equation}  
+Thus, we find $\phi(r,t) = -\frac{R^2 \dot{R}}{r}$. Plugging this into the left hand side of eq. \ref{eq:press_vel} and evaluating at the bubble wall, $r\equiv R$
+\begin{equation}
+\begin{split}
+    \partial_t \phi \vert_{r=R} & = -2\dot{R}^2-R\ddot{R} \\
+    \partial_r \phi \vert_{r=R} & = \dot{R} \equiv u(R)
+\end{split}
+\end{equation}
+
+Finally, sticking all this together, we arrive at the Rayleigh-Plesset equation \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,plesset1949dynamics,plesset1977bubble}
+\begin{equation}
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{p_B(t)-(p_\infty+P(t))}{\rho}
+    \label{eq:RP_1}
+\end{equation}
+with $p_B \equiv p(r=R,t)$ the pressure in the fluid at the bubble wall.
+Evaluated beyond the bubble wall, $r > R$, eq.
+\ref{eq:RP_1} gives the pressure radiated by the fluctuating bubble.
+Some comments are in order.
+The combination $p_\infty+P(t)$ is the \emph{ambient} pressure \cite{prosperetti1999old,prosperetti1986bubble} which gives the pressure in the fluid in the absence of the bubble.
+The left hand side of eq.
+\ref{eq:RP_1} may be regarded as the kinetic energy (density).
+The right hand side is the change in enthalpy (density) due to the fluctuating bubble, i.e.
+the dyanmics of the bubble given on the left hand side are determined by the enthalpy in the fluid due to the bubble's motion.
+In deriving this equation, we have assumed that the speed of propogation in the fluid is infinite.
+In reality, this is not true but this approximation will be very good close to the bubble (in the \emph{near-field}) where retardation effects are negligible anyway.
+In a similar way, we have assumed that the energy in the fluid due to distorting its volume is negligible.
+This approximation will be accurate near the bubble as well since the kinetic energy in this region will dominate \cite{prosperetti1999old}.
+
+We can refine eq.
+\ref{eq:RP_1} a little further.
+Since the motion in the fluid is purely radial, we expect the only relevant stresses to be normal to the bubble wall.
+Let's denote the radial stress in the liquid at some point $r$ by $s_r(r)=-p(r,t)+2\mu \partial_r u$.
+Recall, $\mu$ is the fluid viscosity from earlier.
+Using conservation of mass and momentum, and our expression above for the velocity, we can equate the normal stresses in the fluid to those in the bubble:
+\begin{equation}
+\begin{split}
+    s_r(R)=-p_B-4\frac{\mu \dot{R}}{R}=-p_g+2\frac{\sigma}{R}
+\end{split}
+\end{equation}
+where $p_g$ is the pressure in the gas and $\sigma$ is the surface tension. Solving for $p_B(t)$, we have the relation \cite{brenner2002single,prosperetti1999old,prosperetti1986bubble}
+\begin{equation}
+    p_B(t)=p_g(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)
+    \label{eq:p_B}.
+\end{equation}
+Note that in deriving the equations \ref{eq:NS_1D} we neglected the viscosity terms.
+This was valid since only combinations of viscosity and compressibility $\nabla \cdot \bm{u}$ appeared, both of which are assumed small.
+Here, however, we opt to keep the viscosity terms coupling the bubble wall to the fluid.
+With this, we can rewrite the RP equation as
+\begin{equation}
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) \right]
+    \label{eq:RP_2}
+\end{equation}
+with $\Delta p = p_g(t)-p_\infty$ the deviation of the pressure inside the bubble from the static pressure.
+With the above assumption that the compression is adiabatic, the conditions inside the bubble are decoupled from the liquid (besides the coupling through the RP equation) and we may regard $p_g(t)$ in the bubble as a given quantity to be determined from an equation of state.
+Recall $P(t)$ is the time-dependent part of the pressure in the absence of the bubble: we may regard this as an external pressure from e.g.
+a driving stress.
+This will be from the transducers driving the acoustic levitation cell.
+
+Eq.
+\ref{eq:RP_2} is the equation of motion for a driven spherical bubble in an infinite, incompressible fluid.
+It is worth breifly summarizing this all again.
+On the left hand side, $R$ is the radius of the bubble, $\dot{R}$ is the velocity of the bubble's radius i.e.
+the velocity of the \emph{bubble wall}, and $\ddot{R}$ is the bubble wall's acceleration.
+On the right hand side, $\rho$ is the (constant) density of the fluid.
+$\Delta p(t) = p_g(t)-p_\infty$ is the deviation of the gas pressure in the bubble, $p_g(t)$, from the static fluid pressure at infinite distance, $p_\infty$.
+$p_g(t)$ is evaluated at the bubble wall.
+In the second term, the $\sim \sigma$ part gives the restoring force from surface tension stresses while the $\sim \mu$ term is due to the fluid viscos stresses: $\sigma$ and $\mu$ are the surface tension and viscosity coefficients respectively.
+The last term on the right, $P(t)$, is the \emph{external pressure}.
+For problems relevant to SBSL, the driving pressure, $P(t)$ in the RP equations, is an acoustic plane wave.
+Keller and coworkers showed that the modification required to the RP equations result in the replacement $P(t) \rightarrow P_0 \sin(\omega t)$ \cite{keller1980bubble}.
+If we linearize in the bubble's radius $R(t)\approx R_0 \left( 1+x(t) \right)$ with $x(t) \ll 1$, we can derive a forced harmonic oscillator equation for $x(t)$ \cite{brennen2014cavitation,yasui2018acoustic}.
+However, the low driving amplitude regime where this method is valid is irrelevant to SBSL so we won't pursue it here.
+
+Let us breifly digress to discuss retardation effects.
+If we had not assumed incompressibility, our solution would include a term arising from the sound radiation from the bubble itself.
+There are practical issues with this solution however; it is \emph{third} order in time and has unphysical, exponentially diverging solutions \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,lezzi1987bubble}.
+Both issues are solved by imposing boundary conditions on the $\ddot{R}$ term.
+The method for doing so is also challenging and apparently no unique way to do it exists \cite{prosperetti1986bubble,prosperetti1988nonlinear,keller1956damping}.
+Instead, there is a whole family of solutions that can be derived; different variations of the improved RP equation, which are collectively referred to as ``Rayleigh-Plesset equations" are derived by neglecting different terms in the full compressible solution.
+We will discuss this again below.
+
+The RP equation (eq.
+\ref{eq:RP_2}) is pretty but its analytical solution intractable.
+Direct solutions of the RP equation or its variants are done numerically using e.g.
+the Euler method \cite{yasui2018acoustic,yasui2015dynamics}.
+It is no great challenge computationally.
+For now, let's see what we else we can understand from this analytically by making some more approximations.
+If we take $p_B-(p_\infty+P) \equiv p_0$ with $p_0$ constant, we recover Rayleigh's equation discussed above \cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
+This is valid during the part of the bubble's cycle called \emph{The Rayleigh collapse}: when the bubble's wall is very fast, the $\dot{R}^2$ term will be much larger than the right hand side of eq.
+\ref{eq:RP_2} and we may neglect it.
+Rayleigh looked at the same problem, i.e.
+a cavitating \emph{void} in water, by replacing $p_0\rightarrow 0$ \cite{brenner2002single,rayleigh1917pressure}:
+\begin{equation}
+    R\ddot{R}+\frac{3}{2}\dot{R}^2=0
+    \label{eq:Rayleigh}
+\end{equation}
+which has a solution that diverges as $R(t)\sim\left[(t_*-t)/t_* \right]^{2/5}$ \cite{brenner2002single}.
+This observation is what lead to Rayleigh's argument for the origin of cavitation damage to ship's propellers.
+Yasui gave a rather elegant heurestic explanation for the origin of \emph{caviation} in a fluid that is worth stating here \cite{yasui2018acoustic}.
+Condider the mass flowing through contentric shells around a bubble, $4\pi R_i^2 v_i \rho$.
+For a source free field, we require that neighboring shells satisfy conservation of mass: $4\pi R_1^2 v_1 = 4\pi R_2^2 v_2$.
+For $R_2 > R_1$, we see that $v_1/v_2 = \left(R_2/R_1\right)^2$.
+The velocity for a smaller radius is always larger, diverging towards the origin.
+
+In reality, the bubble's velocity does not become infinite and, in the case of \emph{stable} SBSL, the radius remains finite at all times.
+In the full RP equation (eq.
+\ref{eq:RP_1}) the only term capable of compensating the bubble wall's motion is the divering pressure inside the bubble itself \cite{brenner2002single}.
+Applying the same analysis to the family of solutions derived when including the bubble's sound radiation, it turns out that the most important term for slowing the bubble wall's diverging velocity is $\propto\dot{p}_g$.
+Keeping only this term, we arrive at the most popular variant of the RP equation \cite{lofstedt1995sonoluminescing,barber1997defining}:
+\begin{equation}
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) +\frac{R}{c} \dot{p}_g \right].
+    \label{eq:RP_3}
+\end{equation}
+This equation is only slightly more complicated than the RP equation (eq.
+\ref{eq:RP_2}), the only chage being the added term $\propto \dot{p}_g$ which is determined from solving the decoupled problem of the gas inside the bubble.
+Eq.
+\ref{eq:RP_3} is usually solved numerically; we will look at results of numerical solutions later when talking about experiments.
+For now, we note that the observed error between this equation and experiment is only significant in the interval during bubble collapse and even still, results of solving this equation are quantitavely in good agreement with measurements of the bubble's radius \cite{brenner2002single}.
+
+\subsection{The Bubble's Interior}
+The dynamics of the bubble's wall, determined by the RP equations (eq.
+\ref{eq:RP_2} and eq.
+\ref{eq:RP_3}), require the pressure inside the bubble as input.
+To that end, we now examine the physics of the \emph{bubble's interior}.
+Our goal for this section is to find a suitable form for $\Delta p(t)$ in the RP equations.
 
 \subsection{Experimental Results}
-what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}) (Ar content, driving frequency, P$_{max}$)
 
-\section{What don't we know}
-\subsection{The Bubble Wall (Again)}
-\subsection{Experimental Anomalies}
+Creating stable, single bubbles in a laboratory experiment is not particulary challenging and can be done with standard and low-cost materials \footnote{Actually, the phase-space for SBSL is ``small".
+The coordinates being driving amplitude, driving frequency, fluid composition, temperature, ambient pressure, gas content and composition, etc...
+What we are claiming is simple is the experimental setup once the correct parameters are identified.}.
+A typical experimental setup is shown in Fig.
+\ref{fig:flask}.
+Since we are mainly concerned with the light emission process itsself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}.
+A sample of degassed liquid is placed into a flask \footnote{Actually, the degassing part is quite important.
+We focus on it elsewhere.}.
+Coupled to the outside of the flask are piezoelectric transducers of some sort.
+The piezos are driven by a sinusoidal function generator to excite an acoustic resonance of the flask: a typical flask is a few cm across with a resonance at $\approx 20$ kHz \cite{brenner2002single}.
+The frequency is chosen to form a pressure \emph{antinode} at the center of the flask.
+
+Let us take a brief aside: in deriving the RP equations, we assumed the center of the bubble was stationary.
+Actually, this is what we want in an SBSL experiment and it is worth discussing under what conditions this is true.
+In an ideal, incompressible fluid the net force acting on a bubble may be written as
+\begin{equation}
+    \bm{f}=-\int_{\partial \Omega} p \bm{n} dA = -\int_\Omega \nabla p dV
+\end{equation}
+Here, $\bm{f}$ is the net force, $p$ is the pressure at the bubble's surface, $\bm{n}$ is the unit normal vector point out of the bubbles surface, and $\partial \Omega$ is the bubble's boundary with $\Omega$ the bubble's domain.
+The last term is due to an application of Gauss's theorem.
+Bjerknes \cite{bjerknes1909kraftfelder} took $\nabla p \equiv \nabla p(r=0,t)$ to be constant and averaged over one period of oscillation to find
+\begin{equation}
+    \bm{f} = -\frac{4}{3}\pi \langle R^3 \nabla p \rangle.
+    \label{eq:Bj_force}
+\end{equation}
+This is the so called \emph{Bjerknes force}.
+The point is that if the bubble is located at a pressure extrema, e.g.
+the center of the flask, then this force should vanish and the bubble will remain stationary, an essential result for SBSL experiments.
+
+Some authors have looked at corrections to the bubble's dynamics in the presence of bouyancy, which causes a net translation other than the Bjerknes force.
+The result is the bubble oscillates around an equilibrium position near but away from the center of the flask \cite{matula2000single,matula1997bjerknes,matula1999inertial}.
+The translational motion of the bubble leads to shape distortion which is beleived to supress SL; this is confirmed by SBSL experiments in microgravity resulting in larger intensity light production \cite{matula2000single}.
+
+Back to the experimental setup.
+The typical driving pressure relevant to SBSL ($\sim 1.5$ atm) is too small to cause bubbles to form spontaneously in the flask \cite{}.
+Instead, the bubble is usually seeded somehow.
+In the original work of Gaitan, an air bubble was injected using a needle and syringe through a small hole in the flask's seal \cite{gaitan1992sonoluminescence}.
+More recently, seeding methods involve shooting a small jet of water into the flask (like our shrimp friends!) or blasting the water with a laser to boil a small volume \cite{}.
+The later method is preferred as it enables more precise control over the initial bubble's radius.
+The small bubbles formed this way coalesce at the center of the flask into a single, stable bubble \cite{}.
+
+Now that we know about how single, stable bubbles are formed in the lab, we can discuss a number of quantities of experimental interest.
+The bubble's radius may be measured by \emph{Mie scattering} \cite{}.
+We may also analyze light spectrum which contains information on the bubble's contents: in principle, we can learn about the temperature inside the bubble and spectral lines contain information on the matter \cite{}.
+As we will see, it is actually not so simple to understand the light spectrum \cite{}.
+...
+
+{\color{red} what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}) (Ar content, driving frequency, P$_{max}$)
+
+\cite{gaitan1992sonoluminescence} has good discussion of early experimental errors.
+Measuring phase of SL relative to the acoustic field was an early attempt to explain SL.
+Phase here means how far in time is the SL light emission offset from the driven pressure.
+In the case of periodic oscillations, this is a phase.
+A handful of early models assumed that the light emission occured during expansion of the bubble.
+Others assumed that it was at some point during.
+Most of these were ruled out when it was observed that the light emission is for a very short duration near maximum compression, though the phase depends quite a bit on the initial bubble radius, the driving frequency, the resononances of the flask, the physical properties of the liquid, etc \cite{brenner2002single,gaitan1992sonoluminescence}.
+These complicated relationships were unkown to the early investigators of SL and were quite difficult to measure in MBSL.
+For this reason, it wasn't until SBSL was discovered that many of these models were dismissed \cite{gaitan1992sonoluminescence}.
+In this paper, they present the discovery of SBSL...
+}
 
@@
 
-\section{The Rayleigh-Plesset Equation}
-The Rayleigh-Plesset (RP) equation can be deduced from the compressible Navier-Stokes equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
-To that end, we first remind the reader of the Navier-Stokes equations.
+\section{The Navier-Stokes Equations}
+The Rayleigh-Plesset (RP) equation can be deduced from the Navier-Stokes equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
+To that end, we remind ourselves of the Navier-Stokes equations here.
 
@@
 
-If instead the fluid is real, we can't neglect viscous forces (i.e.
-dissipation).
+If instead the fluid is real, we can't neglect viscous forces.
 We suppose that the viscous surface term can be represented by a tensor, $\hat{\bm{\tau}}$ such that $\bm{F}_v=\int_{\partial \Omega_t} \hat{\bm{\tau}}\cdot \bm{n} dA = \int_{\Omega_t} \nabla \cdot \hat{\bm{\tau}} dV$.
@@
 In general, $\hat{\bm{\mu}}$ is a symmetric rank-4 tensor.
-However, if we assume an isotropic fluid \footnote{Definitely true for water which is usually the fluid used in sonoluminescence experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
+However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in sonoluminescence experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
 \begin{equation}
@@
 It is usual to combine the pressure (i.e.
-elastic stresses) and the viscous-stresses into the stress-tensor $\hat{\bm{\sigma}}=-p\hat{\bm{I}}+\xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)}$.
+elastic-stresses) and the viscous-stresses into the stress-tensor $\hat{\bm{\sigma}}=-p\hat{\bm{I}}+\xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)}$.
 Combined with this, Newton's second law becomes
@@
      \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
-     \label{eq:ns_equations}
+     \label{eq:ns_equations_2}
 \end{split}
@@
 The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
-On the other hand, the terms $\sim \bm{u}$ are viscous (damping) terms that tend to make the velocity field spatially uniform: the vanish when the (spatial) derivatives of the velocity vanish.
+On the other hand, the terms $\sim \bm{u}$ are viscous terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
 The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
 dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
-The body term $\bm{f}_B$ is from external sources: from here on, we will assume it is zero.
-
-Even with these simplifying assumptions, solving eqs.
-\ref{eq:ns_equations} is a formidable task \footnote{In fact, it's not even clear that the NS equations \emph{can} be solved even for incompressible fluids.
-Proving that smooth, sensible solutions exist earns \$1,000,000 \cite{fefferman2006existence}.
+The body term $\bm{f}_B$ is from external sources.
+It is worth noting that solving eqs.
+\ref{eq:ns_equations_2} is a formidable task \footnote{In fact, it's not even clear that the NS equations \emph{can} be solved even for incompressible fluids.
+In 2000, it was announced that proving smooth, sensible solutions exist earns \$1,000,000 \cite{fefferman2006existence}.
 Better hurry though...
 inflation was 6.8\% in Nov.
-2021} To that end, several more-drastic approximations have to be made that allow us to make progess.
-\begin{enumerate}
-    \item We will assume irrotational flow (i.e.
-$\nabla\times\bm{u}\equiv 0$) and only radial motion in the liquid (i.e.
-$\bm{u}=u\bm{r}$).
-Note that, since the flow is assumed irrotation, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
-This amounts to assuming that the bubble is always spherical.
-This seems like a rather drastic approximation but has been validated experimentally in many cases \cite{}.
-This can be understood by accounting for surface-tension at the liquid-bubble interface \cite{}.
-    \item Next, we assume that the viscous terms are negligible in the bulk dyanimcs of the liquid.
-For (relatively) low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation and is widely used \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single}.
-We will acount, to some extent, for viscosity later when looking at the bubble-liquid interface.
-Moreover, we are assuming that the flow is \emph{isentropic}, i.e.
-that it is reversible (no damping) and that no heat is exchanged between fluid parcels.
-We will further assume that the liquid is isothermal, i.e.
-its temperature is constant (in space).
-Then pressure $p$ is determined from an instantaneous equation-of-state $p=p(\rho,T)$.
-In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and, as we will later see, heat-transport across the liquid-bubble interfacing is usually neglected anyway (i.e.
-the bubble is compressed adiabatically in the thermodynamic sense \footnote{According to Wikipedia, \emph{adiabatic} means \emph{fast} in thermodynamics lingo.
-This is relevant to us since we are claiming the bubble wall moves so quickly that it is compressed to its maximum pressure before any heat can flow out.
-On the otherhand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
-This is, e.g.
-the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its groundstate at all times.}
-    \item Related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
-Of course the bubble-liquid interface enters both systems as a boundary condition.
-\end{enumerate}
-
-With these simplifying assumptions (and a vector identity \footnote{$\left( \nabla \cdot \bm{u} \right) \bm{u} = \frac{1}{2} \nabla (u^2) - \bm{u}\times \left( \nabla \times \bm{u} \right)$}) we rewrite eqs.
-\ref{eq:ns_equations} as a system of $1+1$ dimensional partial differential equations:
-\begin{equation}
-\begin{split}
-     \partial_r \left[ \partial_t \phi +\frac{1}{2} \left( \partial_r \phi \right)^2 \right] & = - \frac{1}{\rho} \partial_r p  \\ 
-     \partial_t \rho+ \partial_r \phi \partial_r \rho + \rho \partial^2_r \phi & = 0
-%     \label{eq:ns_equations}
-\end{split}
-\end{equation}
+2021.} and is usually done using computers.
 
@@
 
-\section{random notes}
+%\section{random notes}
 
-argon flash: a method of 'imaging explosions' similar to why argon is more useful for SBSL.
+%argon flash: a method of 'imaging explosions' similar to why argon is more useful for SBSL.
 argon has no internal DOF whereas air (i.e.
@@
 
-Brennen defines caviation as (bubble) nucleation when the pressure drops below the vapor pressure...
+%Brennen defines caviation as (bubble) nucleation when the pressure drops below the vapor pressure...
 as compared to boiling in which the temperature goes above the saturated liquid/vapor temp.
@@
 
-Yasui defines \emph{acoustic cavitation} as the formation and collapse of a bubble driven by a powerful ultrasonic wave \cite{yasui2018acoustic}.
+%Yasui defines \emph{acoustic cavitation} as the formation and collapse of a bubble driven by a powerful ultrasonic wave \cite{yasui2018acoustic}.
 He defines bubble dynamics as bubble pulsation under the driving sound.
@@
 
-some dickheads 3d printed a shrimp claw and it also shrimpoluminesces \cite{tang2019bioinspired}
+%some dickheads 3d printed a shrimp claw and it also shrimpoluminesces \cite{tang2019bioinspired}
 
-apparently a shitty movie called 'Chain Reaction' with Keanu Reeves was centered around the premise of MBSL as a means of clean energy production.
+%apparently a shitty movie called 'Chain Reaction' with Keanu Reeves was centered around the premise of MBSL as a means of clean energy production.
 it was based on the speculation that temperatures inside the bubbles ciould reach up to millions of Kelvin and lead to fusion.
@@
 
-for fusions, also see wikipedia article on 'bubble fusion' and 'Rusi P Taleyarkhan'.
+%for fusions, also see wikipedia article on 'bubble fusion' and 'Rusi P Taleyarkhan'.
 Ruse was some asshole prof at purdue that claimed fusion occured inside cavitating bubbles, but it was later found out that the faked data and interfered with peer reviews leading to him being fucking fired.
 
-\emph{heurestic explanation of bubble trapping given by Brenner in \cite{chainreaction}.}: the bonds between molecules in a fluid are very strong, so instead of 'tearing', the fluid stretches.
+%\emph{heurestic explanation of bubble trapping given by Brenner in \cite{chainreaction}.}: the bonds between molecules in a fluid are very strong, so instead of 'tearing', the fluid stretches.
 however, if there is a bubble present, the local dialation of the fluid can change the bubbles diameter by $\sim$ 1000 times to avoid breaking bonds at the interface between bubble and fluid.
@@
 
-\section{1st discovery}
-MBSL was discovered first.
+%\section{1st discovery}
+%MBSL was discovered first.
 its (obviously) hard to create and control single bubbles so this wasnt feasible until the 1990s, attributed to Gaitan (see below).
@@
 
-Update: Brenner \cite{brenner2002single} consisely summarizes the history: Marinesco notes the photographic plates fogs in an ultrasonic bath \cite{marinesco1933actions} and Frenzel et al \cite{frenzel1934luminescenz} deliberately studied the light emission.
+%Update: Brenner \cite{brenner2002single} consisely summarizes the history: Marinesco notes the photographic plates fogs in an ultrasonic bath \cite{marinesco1933actions} and Frenzel et al \cite{frenzel1934luminescenz} deliberately studied the light emission.
 According to brenner, no one was surprised by this since it was already known that cavitating bubbles fuck shit up.
 
-SBSL was discovered by Gaitan \cite{gaitan1990experimental,gaitan1992sonoluminescence} etc.
+%SBSL was discovered by Gaitan \cite{gaitan1990experimental,gaitan1992sonoluminescence} etc.
 who trapped a single air bubble in a water filled flask using piezoacoutics stuck to the outside.
@@
 
-\section{notes from \cite{crum1994sonoluminescence}}
-pop sci type of magazine article.
+%\section{notes from \cite{crum1994sonoluminescence}}
+%pop sci type of magazine article.
 Gives heurestic expanation: When an intense sound field is produced within a liquid, microscopic cavities of gas or vapor can be generated when the liquid fails under tensile stress.
@@
 
-\section{notes from ref \cite{lohse2018bubble}} 
-SBSL was discovered by Giatan et.
+%\section{notes from ref \cite{lohse2018bubble}} 
+%SBSL was discovered by Giatan et.
 al.
@@
 
-This paper refs the review by Propsperetti: \cite{plesset1977bubble}.
+%This paper refs the review by Propsperetti: \cite{plesset1977bubble}.
 This paper gives EOM for the bubbles radius, the so called \emph{Plesset-Rayleight} equation which I have seen everywhere related to bubbles.
@@
 
-Later in the paper, also revisists snapping shrimp.
+%Later in the paper, also revisists snapping shrimp.
 
-\section{notes from ref \cite{prosperetti2004bubbles}}
-has shit tons of relevant math. take a look. a pretty paper overall
+%\section{notes from ref \cite{prosperetti2004bubbles}}
+%has shit tons of relevant math. take a look. a pretty paper overall
 
-\section{notes from \cite{gaitan1990experimental}}
-disseration by Giatan who is credited with discovering SBSL.
+%\section{notes from \cite{gaitan1990experimental}}
+%disseration by Giatan who is credited with discovering SBSL.
 says apparently SL was discovered in 1933.
@@
 
-also says caviation was predicted by 'Euler' in 1754: \emph{if the velocity of a fluid is high enouhg, negative pressures can form and the liquid can 'break'.} the 'breaking' was called cavitation in 895 by R.E.
+%also says caviation was predicted by 'Euler' in 1754: \emph{if the velocity of a fluid is high enouhg, negative pressures can form and the liquid can 'break'.} the 'breaking' was called cavitation in 895 by R.E.
 Froude, whoever the fuck this is, when studying bubbles near propellers.
@@
 
-\section{notes from \cite{brenner2002single}}
-defines SBSL as ``Single-bubble sonoluminescence occurs when an acoustically trapped and periodically driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission".
+%\section{notes from \cite{brenner2002single}}
+%defines SBSL as ``Single-bubble sonoluminescence occurs when an acoustically trapped and periodically driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission".
 
-\section{QED}
-the ref \cite{liberati2000sonoluminescence} has lots of good refs.
+%\section{QED}
+%the ref \cite{liberati2000sonoluminescence} has lots of good refs.
 they have follow up papers extending Schwingers papers (which are refd therein) and referencing a bunch of other relevant literature.
@@
 
-\section{Modelling}
-in the ref \cite{schanz2012molecular}, they use MD to modelling the equation of state for a collapsing bubble. they calculate the temperature and study the conditions for emitting light.
+%\section{Modelling}
+%in the ref \cite{schanz2012molecular}, they use MD to modelling the equation of state for a collapsing bubble. they calculate the temperature and study the conditions for emitting light.
 
-\section{Refs. to different arguments}
-ref \cite{didenko2000molecular} gives examples of other origin theories (molecular excited states, black-body radiation, bremsstrahlung, ion-electron recombination, confined electrons)
+%\section{Refs. to different arguments}
+%ref \cite{didenko2000molecular} gives examples of other origin theories (molecular excited states, black-body radiation, bremsstrahlung, ion-electron recombination, confined electrons)
 

### 2022-01-05 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 (b) Photograph of the same.
-Here, the fluid is H$_2$SO$_4$ instead of water.
+Here, the fluid is aqueous H$_2$SO$_4$ instead of water.
 A sonoluminescing bubble is clearly visible in the center of the flask.
@@
 \subsection{The Bubble's Interior}
-The dynamics of the bubble's wall, determined by the RP equations (eq.
-\ref{eq:RP_2} and eq.
-\ref{eq:RP_3}), require the pressure inside the bubble as input.
-To that end, we now examine the physics of the \emph{bubble's interior}.
-Our goal for this section is to find a suitable form for $\Delta p(t)$ in the RP equations.
+
+The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (i.e.
+the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
+(ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{}.
+Obviously these are related topics, but the distinction in the context of SBSL is important as we will now see.
+
+Most early progress on the bubble's interior depended on the former method: if we predict the right dynamics, then we might know the correct pressure and temperature in the bubble.
+These data are then used to try to describe the light emission \cite{}.
+This was preferred to studying the emitted spectra directly as careful measurements resulted in an almost featureless spectrum that could not be explained \cite{}.
+Calculating the dynamics of the bubble's wall from the RP equations (eq.
+\ref{eq:RP_2} or eq.
+\ref{eq:RP_3}) requires the pressure inside the bubble as input.
+To that end, we now examine the dynamics of the \emph{bubble's interior}.
+We would like to find a suitable form for $\Delta p(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating buble.
+It turns out that this is a very hard problem for a number of reasons: the conditions inside the bubble depend on gas diffusion and rectification (one-directional diffusion) between the liquid and the bubble, water-vapor conensation and evaporation inside the bubble, and chemical reactions between species in the bubble \cite{brenner2002single}.
+Brenner et.
+al elegantly summarized the impact of the problem \cite{brenner2002single}: ``one of the exciting features of modern research on SBSL is that it is a testing ground for how well mathematical models can deal with such a complicated situation"
+
+Now let us breifly discuss the bubble's light spectrum.
+For the first decade or so, SBSL experiments were done using partially degased water as the liquid medium \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}.
+Careful measurements of the emission from SBSL in water resulted in an almost featureless spectrum \cite{}.
+Early attempts were made to fit this as black-body radiation resulting in \emph{huge} temperature estimates, e.g.
+$10^8$ K \cite{}.
+It was later discovered that SBSL is volume emission, indicating that black-body radiation is not valid \cite{}.
+Little progress was made on this front for quite some time \cite{brenner2002single}.
+The utility of analyzing the light spectrum of SBSL in water is succinctly described by Suslick \cite{suslick2008inside}: ``because of the inherent ambiguity associated with the analysis of featureless spectra of unknown origin, a more rigorous explanation is unlikely to be generated".
+
+It was later realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing much better measurments \cite{}.
+New measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressures than water.
+Careful experiments revealed that the spectrum depended critically on the fluid composition, gas content, and driving pressure.
+Suslick et.
+al measured the different emission spectra from Ar, Xe, and Kr bubbles in aqueous H$_2$SO$_4$ \cite{flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
+Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
+Perhaps equally as crucially, they found the \emph{absence} of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas.
+This observation greatly helped to refine models for the gas dynamics in the bubble and clarified several unexplained experimental observations \cite{}.
+With this in mind, we choose only to focus on presenting the essential results needed to discuss the dynamics of a bubble undergoing SBSL.
+Excellent books and modern reviews on modelling the gas dynamics exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.
+We will hold off on discussing the dynamics in the plasma until a later section.
+
+The most straighforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
+With simplifying assumptions, equations of motion with varying degrees of sophistication are derived for the gas and solved numerically, consistent with the RP equations.
+The discussion proceeds in much the same way as deriving the RP equations above \cite{}, so we will omit it here.
+Instead, we will highlight the salient features before introducing a simple model for the gas that will enable to gain some physical understanding analytically.
+
+The important results of direct numerical solution of the gas dynamics are as follows: chemical reactions in the bubble are important but only in the immediate vacinity of SBSL.
+Diffusion into/out of the bubble of dissociated species is also important but also mainly occurs in the vacinity of SBSL.
+In model calculations, these things are then usually neglected.
+{\color{red} Go on to talk about heat transfer, mass transfer, etc.
+following \cite{brenner2002single}.
+} To summarize, careful solution of the equations of motion show that the contents of the bubble are, to a good approximation, spatially uniform \cite{}.
+This fact allows us to simplify a model for the gas dynamics greatly \cite{}.
+Before we turn to discuss this, let us pause here to mention that another theoretically appealing method for modelling the gas dynamics is \emph{molecular dynamics} \cite{}.
+While this method is attractive for a number of reasons, practical calculations containing realistic numbers of particles aren't possible \cite{brenner2002single}.
+
+The simplest model for the gas dyanmics in the bubble is {}:
 
 \subsection{Experimental Results}
-
 Creating stable, single bubbles in a laboratory experiment is not particulary challenging and can be done with standard and low-cost materials \footnote{Actually, the phase-space for SBSL is ``small".
@@
 \section{Let there be light!}
-\subsection{Thermal Arguments}
-\subsection{Electric Arguments}
-
-\subsection{The Casimir Effect}
-\subsection{The Casimir Argument}
+\subsection{The Plasma}
+\subsection{Brehmasstunglingnls}
+\subsection{Recombination}
+\subsection{Electronic transitions}
 
@@
 2021.} and is usually done using computers.
+
+\section{Electrical Arguments}
+
+\section{The Casimir Argument}
 

### 2022-01-06 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 
-\documentclass[rmp,aps,nofootinbib,superscriptaddress,floatfix,10pt]{revtex4-2} 
+\documentclass[rmp,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
 \usepackage{graphicx}
@@
 \ref{eq:RP_3}) requires the pressure inside the bubble as input.
-To that end, we now examine the dynamics of the \emph{bubble's interior}.
-We would like to find a suitable form for $\Delta p(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating buble.
+We would like to find a suitable form for $\Delta p(t) \sim p_g(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating bubble.
 It turns out that this is a very hard problem for a number of reasons: the conditions inside the bubble depend on gas diffusion and rectification (one-directional diffusion) between the liquid and the bubble, water-vapor conensation and evaporation inside the bubble, and chemical reactions between species in the bubble \cite{brenner2002single}.
@@
 
-It was later realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing much better measurments \cite{}.
-New measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressures than water.
-Careful experiments revealed that the spectrum depended critically on the fluid composition, gas content, and driving pressure.
+SBSL was first measured in partially degassed water \cite{gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single} \footnote{Degassing lowers the ambient pressure, making it easier for bubbles to form.} Early attempts to understand the physics required exploring the phase-space over which SBSL occurs.
+Some attempts were made using non-aqueous host liquids e.g.
+alcohols, silicone oils \cite{weninger1995sonoluminescence,barber1997defining}, but experiments with air bubbles were not very sucessful \cite{barber1997defining}.
+Instead, different gases were tried in water.
+Recall, air is $\sim 80\%$ N$_2$, $\sim 20\%$ O$_2$, and $\sim 1\% $ Ar.
+Considering the relatively large content of O$_2$ and N$_2$, degassed water \emph{regassed} with N$_2$ or O$_2$ or a mixture were checked first and found not to produce SBSL \cite{hiller1994effect}.
+What was discovered was that a small amount of \emph{noble gas} was required for SBSL to occur \cite{barber1997defining,hiller1994effect,brenner2002single}.
+A result of this work was the \emph{argon rectification hypothesis} \cite{lohse1997sonoluminescing,brenner2002single,yasui2018acoustic,suslick2008inside}.
+The hypothesis claims that all species (in air) in the bubble besides argon are gradually ejected from bubble's interior until all that remains is pure argon.
+The theory is based on the fact that, at the elevated temperatures and pressures inside the bubble, dissociation of O$_2$ and N$_2$ into radicals is possible.
+These species then react with the water vapor radicals to form species that are soluble in the host fluid.
+As the pressure becomes very large during the compression stage of SBSL, the soluble materials leave the bubble and do not re-enter since thier solubility in water is enormous compared to the Ar content of the bubble \cite{lohse1997sonoluminescing}.
+Over many cycles, the contents of the bubble become nearly pure argon.
+Moreover, it was realized that the SBSL was much more intense when the contents of the bubble are a pure inert gas: if the contents were e.g.
+molecules, bond breaking/formation would alter the pressure of the gas, ultimately reducing the temperature and reducing the light production \cite{brenner2002single,yasui2018acoustic,lohse2018bubble,suslick2008inside}.
+This mechanism has been used to explain the relatively weak light, i.e.
+the low temperature, in MBSL: the bubbles are transient and cannot remove a significant amount of O$_2$ or N$_2$ over a single cycle.
+The current belief is that the contents of a bubble in SBSL are pure argon after $\sim 10^3$ cycles \cite{brenner2002single,yasui2018acoustic}.
+
+Most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurments of the light spectum \cite{}.
+More importantly, new measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressures than water.
+Careful experiments revealed that the spectrum depended critically on the noble gas content and driving pressure.
 Suslick et.
@@
 Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
-Perhaps equally as crucially, they found the \emph{absence} of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas.
-This observation greatly helped to refine models for the gas dynamics in the bubble and clarified several unexplained experimental observations \cite{}.
-With this in mind, we choose only to focus on presenting the essential results needed to discuss the dynamics of a bubble undergoing SBSL.
+Perhaps equally as crucially, they found the \emph{absence} of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas, experimentally confirming the argon rectification hypothesis.
+Moreover, the spectra at different pressures were in very good agreement with detailed calcutions of emission from noble gas plasmas \cite{an2009diagnosing,an2008spectral}.
+Increased pressure results in increased collision rate between particles, broadening the excited state transition spectral lines \cite{an2008spectral,suslick2008inside,flannigan2005plasma,flannigan2006measurement}.
+Further calculations showed that the greatly increased viscosity and density of aqueous H$_2$SO$_4$, $\sim 10 \times$ and $\sim 1.3 \times$ larger than that that of water, lead to reduced (by orders of magnitude) pressure in the bubble at the same driving pressures \cite{an2009diagnosing}.
+The larger pressure in the bubbles in water was found to be consistent with the featureless emission spectrum \cite{suslick2008inside,yasui2018acoustic}.
+
+This argon rectification hypothesis greatly helped to refine models for the gas dynamics in the bubble and clarified several unexplained experimental observations \cite{suslick2008inside,brenner2002single}.
+With this in mind, we choose only to focus on results known to be relevant to the dynamics of a bubble undergoing SBSL.
 Excellent books and modern reviews on modelling the gas dynamics exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.
@@
 
+\subsection{Gas Dynamics}
 The most straighforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
-With simplifying assumptions, equations of motion with varying degrees of sophistication are derived for the gas and solved numerically, consistent with the RP equations.
-The discussion proceeds in much the same way as deriving the RP equations above \cite{}, so we will omit it here.
-Instead, we will highlight the salient features before introducing a simple model for the gas that will enable to gain some physical understanding analytically.
+In fact, serious quantitative predictions of the conditions of the bubble's interior take this path \cite{an2006mechanism,an2008spectral,an2009diagnosing,flannigan2005plasma,flannigan2006measurement}.
+With simplifying assumptions, equations of motion with varying degrees of sophistication can be derived for the gas and the system of gas dynamical equations and the RP equations are solved numerically.
+The discussion proceeds in much the same way as deriving the RP equations above but is considerally more complicated \cite{brenner2002single,yasui2018acoustic}, so we will omit it here.
+An important aspect of these studies is the understanding of why relatively simple models work quite well.
+We will highlight the salient features before introducing a simple model for the gas that will enable us to gain some physical understanding analytically.
+An especially important result of direct numerical solution of the gas dynamics was (theoretical) confirmation of the argon rectification hypothesis \cite{}, reducing the possible gas dynamical models to those applicable to a monatomic gas.
+Various other models fully including dissipation and heat transport into/out of the bubble showed that the bubble's compression is, to a good approximation, adiabatic and quasistatic \cite{}.
+Before we turn to discuss a model for the gas dynamics, let us pause to mention that another theoretically appealing method for modelling the gas dynamics is \emph{molecular dynamics} \cite{}.
+While this method is attractive for a number of reasons, practical calculations containing realistic numbers of particles aren't possible, limiting the applicability \cite{brenner2002single}.
 
-The important results of direct numerical solution of the gas dynamics are as follows: chemical reactions in the bubble are important but only in the immediate vacinity of SBSL.
-Diffusion into/out of the bubble of dissociated species is also important but also mainly occurs in the vacinity of SBSL.
-In model calculations, these things are then usually neglected.
-{\color{red} Go on to talk about heat transfer, mass transfer, etc.
-following \cite{brenner2002single}.
-} To summarize, careful solution of the equations of motion show that the contents of the bubble are, to a good approximation, spatially uniform \cite{}.
-This fact allows us to simplify a model for the gas dynamics greatly \cite{}.
-Before we turn to discuss this, let us pause here to mention that another theoretically appealing method for modelling the gas dynamics is \emph{molecular dynamics} \cite{}.
-While this method is attractive for a number of reasons, practical calculations containing realistic numbers of particles aren't possible \cite{brenner2002single}.
+A simple but reasonable model for the gas dyanmics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.}.
+Let us remind ourselves what the pressure is in this case.
+Suppose that we have a volume, $V$, of an ideal gas containing $N$ particles at a temperature $T$.
+The ideal gas law tells us the equation of state is $PV=Nk_BT$ with $k_B$ \emph{Boltzmann's constant}.
+Since we are assuming adiabatic compression, the second law of thermodynamics tells us that $dU=-PdV$ with $U$ the internal energy in the gas.
+Moreover, we may use the equipartition theorem to write $dU=N f k_B dT/2$ with $f$ the number of \emph{degrees of freedom} per particle \cite{}.
+Recall that for an ideal gas \cite{}:
+\begin{equation}
+\begin{split}
+    C_V &= \frac{N f k_B}{2} \\
+    C_P &= C_V+N k_B \\
+    \frac{C_P}{C_V} &= 1+\frac{f}{2} \equiv \Gamma 
+\end{split}
+\end{equation}
+with $C_V$ and $C_P$ the heat capcities at constant volume and pressure respectively.
+For a monatomic ideal gas, $f=3$ and $\Gamma=5/2$.
+Now we can use the second law of thermodynamics to write $C_V dT = -P dV$.
+From the equation of state, we have
+\begin{equation}
+    dT = \frac{PdV+VdP}{Nk_B}=\frac{PdV+VdP}{C_P-C_V}.
+\end{equation}
+Eliminating $dT$ and simplifying, 
+\begin{equation}
+    \frac{dP}{P}=-\left(\frac{C_P}{C_V}\right) \frac{dV}{V} = -\Gamma \frac{dV}{V}.
+    \label{eq:dP}
+\end{equation}
+Integrating eq. \ref{eq:dP}, we find 
+\begin{equation}
+    P_2 = P_1 \left(\frac{V_1}{V_2}\right)^\Gamma
+\label{eq:P}
+\end{equation}
+where $P_2$ is the pressure in a gas at volume $V_2$ that was initially contained in a volume $V_1$ at pressure $P_1$. Importantly we have not made any assumptions about the shape of the container.
 
-The simplest model for the gas dyanmics in the bubble is {}:
+Let us now connect this to the problem of SBSL.
+In our case, we assume that our bubble has an \emph{ambient} radius $R_0$, leading to volume $V_0 =  4/3 \pi R_0^3$.
+Similarly, we will assume the pressure in the bubble at ambient radius $R_0$ is known and we will call it $p_0$.
+What we want to calculate is the pressure in the gas as a function of the bubble's radius $R(t)$ determined from the RP equations (eq.
+\ref{eq:RP_2} or eq.
+\ref{eq:RP_3}).
+Earlier we called the pressure in the gas $p_g(t)$ and we will stick with this here.
+Plugging these into eq.
+\ref{eq:P}, (and sneaking in a modification) we arrive at a very commonly used equation in the context of SBSL \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward}:
+\begin{equation}
+    p_g(t) = p_0 \left( \frac{R_0^3}{R^3(t)-h^3} \right)^ \Gamma
+    \label{eq:p_g}
+\end{equation}
+The new variable $h$ that was added willy-nilly is called the \emph{Van der Walls hard-core radius} and its purpose is to make sure that, if the bubble collapses so strongly that its contents become incompressible (i.e.
+$R(t)\rightarrow h$), the pressure diverges \cite{lofstedt1993toward,brenner2002single}.
+Eq.
+\ref{eq:p_g} can be inserted into an RP equation which can then be integrated numerically to determine the bubble's dynamics.
+Now recall that one of main goals of modelling the bubble dynamics in SBSL was to calculate the temperature inside the bubble in order to model the light emission.
+Following similar steps to deriving $p_g(t)$, we can write down an equation for the temperature $T$ \cite{barber1997defining,brenner2002single}:
+\begin{equation}
+    T(t) = T_0 \left( \frac{R_0^3}{R^3(t)-h^3} \right)^ {\Gamma-1}.
+    \label{eq:T(t)}
+\end{equation}
+Of course we still have to assign a reasonable guess to $T_0$, the temperature in the bubble at ambient radius $R_0$.
+
+Our physical intuition tells us why eq.
+\ref{eq:p_g} is sensible during the bubble's collapse: the bubble wall's velocity is fast and very little heat can flow out during compression.
+Detailed calculations starting from the full NS equations further tell us that the gas dynamics in the bubble can be regarded as \emph{quasistatic} \cite{}.
+During the expansion and afterbounce stages of the bubbles cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal}.
+This is true because, when the bubble's velocity is on the order of the \emph{heat diffusion} timescale, the temperature throughout the bubble is nearly equal to that in the liquid \cite{prosperetti1999old,brenner2002single,yasui2018acoustic}.
+The relevant modifications to eq.
+\ref{eq:p_g} turns out to be replacing $\Gamma\rightarrow 1$.
+A neat way of including both isothermal and adiabatic gas dynamics is continously varying the exponent, $\Gamma$ in eq.
+\ref{eq:p_g}, between its adiabatic and isothermal values, $C_P/C_V$ and $1$, respectively.
+A summary of this work is given in refs.
+\cite{brenner2002single} and \cite{prosperetti1999old}.
+It turns out that this process doesn't result in better quantitative predictions since, during the assumed isothermal part of the process, dissipation becomes imporant \cite{brenner2002single}.
+It is also possible to include heat and mass diffusion in simple model forms and these do give slighlty improved quantitative agreement with experiment; however we choose to discuss solutions of the RP equations using the simpler form given in eq.
+\ref{eq:p_g} as this will enable to understand the important features of the bubble's cycle.
+More advanced solutions using the full gas dynamical equations are qualitatively similar \cite{brenner2002single,yasui2018acoustic}.
 

### 2022-01-06 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 
-A simple but reasonable model for the gas dyanmics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.}.
+A simple but reasonable model for the gas dyanmics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.
+The modifications to the result for an ideal gas are simple.}.
 Let us remind ourselves what the pressure is in this case.
@@
 \begin{equation}
-    p_g(t) = p_0 \left( \frac{R_0^3}{R^3(t)-h^3} \right)^ \Gamma
+    p_g(t) = p_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ \Gamma
     \label{eq:p_g}
@@
 \begin{equation}
-    T(t) = T_0 \left( \frac{R_0^3}{R^3(t)-h^3} \right)^ {\Gamma-1}.
+    T(t) = T_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ {\Gamma-1}.
     \label{eq:T(t)}

### 2022-01-06 sonoluminescence/paper/bak/v1/paper.tex
sdfadsfa

--- 2022-01-05 sonoluminescence/paper/paper.tex
@@
 With unprecidented experimental control, many discrepencies in previous assumptions about SL were discovered.
-Experimental Results measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
+Experiments measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
 This discovery implied that SL was nearly decoupled from the bubble's dynamics, contradicting models based on Rayleigh's equation.
@@
 
-History has shown that the most important part of understanding SBSL is understanding the dyanmics of a single acoustically driven bubble.
-We will see that the dynamics of the gas \emph{inside} the bubble is nearly decoupled from the dyanmics of the \emph{fluid} containing the bubble so that we may look at these two problems seperately.
-Firstly, we look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
-Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
-After that, we will combine these two and discuss solutions of the full problem.
-The ultimate result of this will turn out to be an estimate of the \emph{temperature} in the bubble.
-We will then look at theories presenting the origin of light from the very hot contents of the bubble.
-Finally, we will wrap this whole thing up with a summary of unsolved problems and an outlook to the future.
-
 \begin{figure}
@@
 (b) Photograph of the same.
-Here, the fluid is aqueous H$_2$SO$_4$ instead of water.
+Here, the fluid is H$_2$SO$_4$ instead of water.
 A sonoluminescing bubble is clearly visible in the center of the flask.
@@
 
-\section{Bubble Dyanmics}
-The conditions in the bubble's \emph{interior} are still not entirely understood and will be the subject of an entire subsection of this paper.
-Before we get to this, we will at an older problem: the dynamics of a periodically driven bubble in a liquid.
-An important aspect of the bubble's dyanmics is that they are experimentally accessible.
-Moreover, as we shall see, a (relatively) simple theory describles the dynamics of the bubble quite well.
-On the other hand, measuring what is going on in the bubble's interior is not at all simple though much progress has been made theoretically.
-
-%Before we look at what we don't know, it is useful to summarize what we \emph{do} know about SBSL.
+\section{What do we know?}
+The conditions in the bubble's \emph{interior} are still not entirely understood and will be the subject of an entire section of this paper.
+Before we look at what we don't know, it is useful to summarize what we \emph{do} know about SBSL.
 This has two parts: on the one hand, the dynamics of the bubble are experimentally accessible and have been well studied by \emph{Mie scattering} experiments.
@@
 As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
-This isn't particularly surprising since the light emission and bubble collapse occur at different timescales (the duration of maximum compressions is $\sim 0.003\%$ of the bubble's period) \cite{suslick2008inside,prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
-The starting point for any theory of the bubble dynamics is Rayleigh's original work \cite{rayleigh1917pressure}, now expounded upon by many others.
-The most widely known modern work is that of Plesset \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}, resulting in the so called ``the Rayleigh-Plesset" (RP) equation.
-The derivation assumes the host liquid is incompressible, ultimately resulting in us neglecting the effects of sound radiated by the bubble.
-Corrections exist that improve upon the RP equation in various ways and we will discuss these below \cite{}.
-However, for brevity's sake, we will only derive the RP equation here and merely quote the others as they are very similar in form to the RP equation.
-The field of \emph{bubble dynamics} is quite mature and attempting to review it here would be out of scope of this paper; instead, the reader is referred to numerous books and reviews \cite{}.
-
-The RP equation can be deduced from the \emph{Navier-Stokes} (NS) equations \footnote{The NS equations are derived in the appendix} \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
-The NS equations for a compressible fluid are
+This isn't particularly surprising since the light emission and bubble collapse occur at different timescales \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
+The starting point for any theory of the bubble dynamics is Rayleigh's original work expounded upon by many others \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}: the so called ``the Rayleigh-Plesset" (RP) equation \footnote{The RP equation is derived in the appendix}.
 \begin{equation}
-\begin{split}
-    \rho \frac{d \bm{u}}{dt} = \rho \left[ \partial_t \bm{u}+\left(\bm{u}\cdot \nabla \right)\bm{u} \right] & =-\nabla p + \eta \nabla \left(\nabla \cdot \bm{u} \right)+\mu \nabla^2 \bm{u} + \bm{f}_B \\
-     \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
-     \label{eq:NS_equations}
-\end{split}
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) \right].
+    \label{eq:RP}
 \end{equation}
-
-Here, $\bm{u}(\bm{x},t)$ is the velocity of a fluid parcel at position $\bm{x}$ at time $t$, $\rho (\bm{x},t)$ is the fluid's density, $p$ is the pressure, $\eta \equiv \left(\xi + \frac{2}{3}\mu \right)$ is the \emph{bulk} viscosity with $\mu$ the \emph{shear} viscosity, and $\bm{f}_B$ are \emph{body} forces (e.g.
-gravity, electrostatic forces) which are usually from external sources and will be assumed negligible from here on.
-The subscrits on $\partial$ denote what deriviative is being taken and $\nabla$ and $\nabla^2$ have thier usual meanings.
-The first equation is \emph{conservation of momentum} for the fluid, i.e.
-Newton's $2^{nd}$ law; the second equation is \emph{mass convervation}.
-The total-derivative of a moving fluid $\frac{d}{dt}\equiv D_t$ is sometimes called the \emph{material derivative}.
-In the case of velocity $D_t \bm{u}$ is the \emph{acceleration} of the fluid.
-It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies \emph{in space} as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
-The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
-On the other hand, the terms $\sim \bm{u}$ are viscous (damping) terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
-The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
-dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
-
-Solving eqs.
-\ref{eq:NS_equations} is an intractable problem.
-To make progress, we will make some drastic approximations that are more-or-less valid for our problem.
-Firslty, we will assume irrotational flow (i.e.
-$\nabla\times\bm{u}\equiv 0$) so that there is only radial motion in the liquid (i.e.
-$\bm{u}=u\bm{r}$).
-Note that, since the flow is assumed irrotational, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
-This amounts to assuming that the bubble is always spherical which seems like a rather drastic approximation but has been validated experimentally in many cases \cite{}.
-The fact that the bubble tends to remain spherical can be understood by accounting for surface-tension at the liquid-bubble interface \cite{prosperetti1999old}.
-
-Nextly, we assume that the viscous terms are negligible in the bulk dyanimcs of the liquid.
-For (relatively) low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation and is widely used \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
-We will acount, to some extent, for viscosity later when looking at the bubble-liquid interface.
-Moreover, we are assuming that the flow is \emph{isentropic}, i.e.
-that it is reversible (no damping) and that no heat is exchanged between fluid parcels.
-We will further assume that the liquid is isothermal, i.e.
-its temperature is constant (in space).
-Then pressure $p$ is determined from an instantaneous equation-of-state $p=p(\rho,T)$.
-In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and, as we will later see, heat-transport across the liquid-bubble interfacing is usually neglected anyway (i.e.
-the bubble is compressed adiabatically in the thermodynamic sense \footnote{According to Wikipedia, \emph{adiabatic} means \emph{fast} in thermodynamics lingo.
-This is relevant to us since we are claiming the bubble wall moves so quickly that it is compressed to its maximum pressure before any heat can flow out.
-On the otherhand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
-This is, e.g.
-the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its groundstate at all times.})
-
-Lastly, related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
-Of course the bubble-liquid interface enters both systems as a boundary condition \cite{prosperetti1986bubble}.
-
-With these simplifying assumptions (and a vector identity \footnote{$\left( \nabla \cdot \bm{u} \right) \bm{u} = \frac{1}{2} \nabla (u^2) - \bm{u}\times \left( \nabla \times \bm{u} \right)$}) we rewrite eqs.
-\ref{eq:NS_equations} as a system of $1+1$ dimensional partial differential equations:
-\begin{equation}
-\begin{split}
-     \partial_r \left[ \partial_t \phi +\frac{1}{2} \left( \partial_r \phi \right)^2 \right] & = - \frac{1}{\rho} \partial_r p  \\ 
-     \partial_t \rho+ \partial_r \phi \partial_r \rho + \rho \partial^2_r \phi & = 0.
-     \label{eq:NS_1D}
-\end{split}
-\end{equation}
-We can integrate the first equation above using the fact that the compressibility is negligible to move the density through the integration \cite{leighton2007derivation}
-\begin{equation}
-    \int_{r_\infty}^{r} \frac{\partial}{\partial r'} \left[ \frac{\partial \phi}{\partial t} +\frac{1}{2} \left( \frac{\partial \phi}{\partial r'} \right)^2 \right] dr' = - \frac{1}{\rho} \int_{r_\infty}^{r} \frac{\partial p}{\partial r'} dr'.
-\end{equation}
-On the left side, we take the velocity to vanish at infinity.
-On the right side, we assume that only the static and unperturbed pressures, $p_\infty$ and $P(t)$ respectively, are relevant so that \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}
-\begin{equation}
-    \partial_t \phi + \frac{1}{2}\left( \partial_r \phi \right)^2 = \frac{(p_\infty+P(t))-p(t,r)}{\rho}.
-    \label{eq:press_vel}
-\end{equation}
-Eq. \ref{eq:press_vel} gives the pressure at cordinate $(r,t)$ in terms of the velocity of the fluid, the density, and the appied pressure.
-
-Now, let us assume that the velocity potential satisfies a wave equation $\nabla^2 \phi - \frac{1}{c^2}\partial_t^2 \phi =0 $ \footnote{This doesn't come from thin air.
-We could write the r.h.s.
-of eq.
-\ref{eq:press_vel} as the enthalpy, $h\equiv \int \frac{dp}{\rho}$, and similarly introduce $dh=\frac{dp}{\rho}$ in the first line of eq.
-\ref{eq:NS_1D}.
-Eliminating $h$ between the two and dropping small terms $\sim \frac{u}{c}$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.} with $c$ the speed of sound in the fluid.
-Again, we use incompressibilty by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$ \footnote{For our problem, it is true that $\left( \frac{1}{c^2} \partial_t \phi \right) / \nabla^2 \phi \sim \left( \frac{R}{\lambda} \right)^2 \ll 1$, i.e.
-the long wave-length approximation is accurate \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}.} .
-In this case, we can ignore retardation effects and, remembering that we have spherical symmetry, write the solution of $\nabla^2 \phi$ as
-\begin{equation}
-    \phi = \frac{\psi(t)}{r}
-\end{equation}
-with $\psi$ the time-dependent coefficient. We now use the boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, to determine $\psi$:
-\begin{equation}
-    \dot{R} = -\frac{\psi}{R^2} \Rightarrow \psi = -R^2 \dot{R}.
-\end{equation}  
-Thus, we find $\phi(r,t) = -\frac{R^2 \dot{R}}{r}$. Plugging this into the left hand side of eq. \ref{eq:press_vel} and evaluating at the bubble wall, $r\equiv R$
-\begin{equation}
-\begin{split}
-    \partial_t \phi \vert_{r=R} & = -2\dot{R}^2-R\ddot{R} \\
-    \partial_r \phi \vert_{r=R} & = \dot{R} \equiv u(R)
-\end{split}
-\end{equation}
-
-Finally, sticking all this together, we arrive at the Rayleigh-Plesset equation \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,plesset1949dynamics,plesset1977bubble}
-\begin{equation}
-    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{p_B(t)-(p_\infty+P(t))}{\rho}
-    \label{eq:RP_1}
-\end{equation}
-with $p_B \equiv p(r=R,t)$ the pressure in the fluid at the bubble wall.
-Evaluated beyond the bubble wall, $r > R$, eq.
-\ref{eq:RP_1} gives the pressure radiated by the fluctuating bubble.
-Some comments are in order.
-The combination $p_\infty+P(t)$ is the \emph{ambient} pressure \cite{prosperetti1999old,prosperetti1986bubble} which gives the pressure in the fluid in the absence of the bubble.
-The left hand side of eq.
-\ref{eq:RP_1} may be regarded as the kinetic energy (density).
-The right hand side is the change in enthalpy (density) due to the fluctuating bubble, i.e.
-the dyanmics of the bubble given on the left hand side are determined by the enthalpy in the fluid due to the bubble's motion.
-In deriving this equation, we have assumed that the speed of propogation in the fluid is infinite.
-In reality, this is not true but this approximation will be very good close to the bubble (in the \emph{near-field}) where retardation effects are negligible anyway.
-In a similar way, we have assumed that the energy in the fluid due to distorting its volume is negligible.
-This approximation will be accurate near the bubble as well since the kinetic energy in this region will dominate \cite{prosperetti1999old}.
-
-We can refine eq.
-\ref{eq:RP_1} a little further.
-Since the motion in the fluid is purely radial, we expect the only relevant stresses to be normal to the bubble wall.
-Let's denote the radial stress in the liquid at some point $r$ by $s_r(r)=-p(r,t)+2\mu \partial_r u$.
-Recall, $\mu$ is the fluid viscosity from earlier.
-Using conservation of mass and momentum, and our expression above for the velocity, we can equate the normal stresses in the fluid to those in the bubble:
-\begin{equation}
-\begin{split}
-    s_r(R)=-p_B-4\frac{\mu \dot{R}}{R}=-p_g+2\frac{\sigma}{R}
-\end{split}
-\end{equation}
-where $p_g$ is the pressure in the gas and $\sigma$ is the surface tension. Solving for $p_B(t)$, we have the relation \cite{brenner2002single,prosperetti1999old,prosperetti1986bubble}
-\begin{equation}
-    p_B(t)=p_g(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)
-    \label{eq:p_B}.
-\end{equation}
-Note that in deriving the equations \ref{eq:NS_1D} we neglected the viscosity terms.
-This was valid since only combinations of viscosity and compressibility $\nabla \cdot \bm{u}$ appeared, both of which are assumed small.
-Here, however, we opt to keep the viscosity terms coupling the bubble wall to the fluid.
-With this, we can rewrite the RP equation as
-\begin{equation}
-    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) \right]
-    \label{eq:RP_2}
-\end{equation}
-with $\Delta p = p_g(t)-p_\infty$ the deviation of the pressure inside the bubble from the static pressure.
-With the above assumption that the compression is adiabatic, the conditions inside the bubble are decoupled from the liquid (besides the coupling through the RP equation) and we may regard $p_g(t)$ in the bubble as a given quantity to be determined from an equation of state.
-Recall $P(t)$ is the time-dependent part of the pressure in the absence of the bubble: we may regard this as an external pressure from e.g.
-a driving stress.
-This will be from the transducers driving the acoustic levitation cell.
-
-Eq.
-\ref{eq:RP_2} is the equation of motion for a driven spherical bubble in an infinite, incompressible fluid.
-It is worth breifly summarizing this all again.
+This is the equation of motion for a driven spherical bubble in an infinite, incompressible fluid.
 On the left hand side, $R$ is the radius of the bubble, $\dot{R}$ is the velocity of the bubble's radius i.e.
@@
 The last term on the right, $P(t)$, is the \emph{external pressure}.
-For problems relevant to SBSL, the driving pressure, $P(t)$ in the RP equations, is an acoustic plane wave.
-Keller and coworkers showed that the modification required to the RP equations result in the replacement $P(t) \rightarrow P_0 \sin(\omega t)$ \cite{keller1980bubble}.
-If we linearize in the bubble's radius $R(t)\approx R_0 \left( 1+x(t) \right)$ with $x(t) \ll 1$, we can derive a forced harmonic oscillator equation for $x(t)$ \cite{brennen2014cavitation,yasui2018acoustic}.
-However, the low driving amplitude regime where this method is valid is irrelevant to SBSL so we won't pursue it here.
+In the context of SBSL, this is a sinusoidal function driven by the transducers connected to the flask containing the bubble (see the next section).
 
-Let us breifly digress to discuss retardation effects.
-If we had not assumed incompressibility, our solution would include a term arising from the sound radiation from the bubble itself.
-There are practical issues with this solution however; it is \emph{third} order in time and has unphysical, exponentially diverging solutions \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,lezzi1987bubble}.
-Both issues are solved by imposing boundary conditions on the $\ddot{R}$ term.
-The method for doing so is also challenging and apparently no unique way to do it exists \cite{prosperetti1986bubble,prosperetti1988nonlinear,keller1956damping}.
-Instead, there is a whole family of solutions that can be derived; different variations of the improved RP equation, which are collectively referred to as ``Rayleigh-Plesset equations" are derived by neglecting different terms in the full compressible solution.
-We will discuss this again below.
+The Rayleigh equation, as opposed to the RP equation, is equivalent to eq.
+\ref{eq:RP} with the pressure term at a constant value.
+{\color{red} Write down Rayleighs equation, show that it is recovered from the RP equation, discuss its solution, show that its wall velocity diverges.
+\cite{brenner2002single} discusses these and how it is necessary for something to diverege with the same power to oppose the collapse.
+They call this the Rayleigh collapse.}
 
-The RP equation (eq.
-\ref{eq:RP_2}) is pretty but its analytical solution intractable.
-Direct solutions of the RP equation or its variants are done numerically using e.g.
-the Euler method \cite{yasui2018acoustic,yasui2015dynamics}.
-It is no great challenge computationally.
-For now, let's see what we else we can understand from this analytically by making some more approximations.
-If we take $p_B-(p_\infty+P) \equiv p_0$ with $p_0$ constant, we recover Rayleigh's equation discussed above \cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
-This is valid during the part of the bubble's cycle called \emph{The Rayleigh collapse}: when the bubble's wall is very fast, the $\dot{R}^2$ term will be much larger than the right hand side of eq.
-\ref{eq:RP_2} and we may neglect it.
-Rayleigh looked at the same problem, i.e.
-a cavitating \emph{void} in water, by replacing $p_0\rightarrow 0$ \cite{brenner2002single,rayleigh1917pressure}:
-\begin{equation}
-    R\ddot{R}+\frac{3}{2}\dot{R}^2=0
-    \label{eq:Rayleigh}
-\end{equation}
-which has a solution that diverges as $R(t)\sim\left[(t_*-t)/t_* \right]^{2/5}$ \cite{brenner2002single}.
-This observation is what lead to Rayleigh's argument for the origin of cavitation damage to ship's propellers.
-Yasui gave a rather elegant heurestic explanation for the origin of \emph{caviation} in a fluid that is worth stating here \cite{yasui2018acoustic}.
-Condider the mass flowing through contentric shells around a bubble, $4\pi R_i^2 v_i \rho$.
-For a source free field, we require that neighboring shells satisfy conservation of mass: $4\pi R_1^2 v_1 = 4\pi R_2^2 v_2$.
-For $R_2 > R_1$, we see that $v_1/v_2 = \left(R_2/R_1\right)^2$.
-The velocity for a smaller radius is always larger, diverging towards the origin.
-
-In reality, the bubble's velocity does not become infinite and, in the case of \emph{stable} SBSL, the radius remains finite at all times.
-In the full RP equation (eq.
-\ref{eq:RP_1}) the only term capable of compensating the bubble wall's motion is the divering pressure inside the bubble itself \cite{brenner2002single}.
-Applying the same analysis to the family of solutions derived when including the bubble's sound radiation, it turns out that the most important term for slowing the bubble wall's diverging velocity is $\propto\dot{p}_g$.
-Keeping only this term, we arrive at the most popular variant of the RP equation \cite{lofstedt1995sonoluminescing,barber1997defining}:
-\begin{equation}
-    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) +\frac{R}{c} \dot{p}_g \right].
-    \label{eq:RP_3}
-\end{equation}
-This equation is only slightly more complicated than the RP equation (eq.
-\ref{eq:RP_2}), the only chage being the added term $\propto \dot{p}_g$ which is determined from solving the decoupled problem of the gas inside the bubble.
-Eq.
-\ref{eq:RP_3} is usually solved numerically; we will look at results of numerical solutions later when talking about experiments.
-For now, we note that the observed error between this equation and experiment is only significant in the interval during bubble collapse and even still, results of solving this equation are quantitavely in good agreement with measurements of the bubble's radius \cite{brenner2002single}.
-
-\subsection{The Bubble's Interior}
-
-The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (i.e.
-the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
-(ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{}.
-Obviously these are related topics, but the distinction in the context of SBSL is important as we will now see.
-
-Most early progress on the bubble's interior depended on the former method: if we predict the right dynamics, then we might know the correct pressure and temperature in the bubble.
-These data are then used to try to describe the light emission \cite{}.
-This was preferred to studying the emitted spectra directly as careful measurements resulted in an almost featureless spectrum that could not be explained \cite{}.
-Calculating the dynamics of the bubble's wall from the RP equations (eq.
-\ref{eq:RP_2} or eq.
-\ref{eq:RP_3}) requires the pressure inside the bubble as input.
-To that end, we now examine the dynamics of the \emph{bubble's interior}.
-We would like to find a suitable form for $\Delta p(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating buble.
-It turns out that this is a very hard problem for a number of reasons: the conditions inside the bubble depend on gas diffusion and rectification (one-directional diffusion) between the liquid and the bubble, water-vapor conensation and evaporation inside the bubble, and chemical reactions between species in the bubble \cite{brenner2002single}.
-Brenner et.
-al elegantly summarized the impact of the problem \cite{brenner2002single}: ``one of the exciting features of modern research on SBSL is that it is a testing ground for how well mathematical models can deal with such a complicated situation"
-
-Now let us breifly discuss the bubble's light spectrum.
-For the first decade or so, SBSL experiments were done using partially degased water as the liquid medium \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}.
-Careful measurements of the emission from SBSL in water resulted in an almost featureless spectrum \cite{}.
-Early attempts were made to fit this as black-body radiation resulting in \emph{huge} temperature estimates, e.g.
-$10^8$ K \cite{}.
-It was later discovered that SBSL is volume emission, indicating that black-body radiation is not valid \cite{}.
-Little progress was made on this front for quite some time \cite{brenner2002single}.
-The utility of analyzing the light spectrum of SBSL in water is succinctly described by Suslick \cite{suslick2008inside}: ``because of the inherent ambiguity associated with the analysis of featureless spectra of unknown origin, a more rigorous explanation is unlikely to be generated".
-
-It was later realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing much better measurments \cite{}.
-New measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressures than water.
-Careful experiments revealed that the spectrum depended critically on the fluid composition, gas content, and driving pressure.
-Suslick et.
-al measured the different emission spectra from Ar, Xe, and Kr bubbles in aqueous H$_2$SO$_4$ \cite{flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
-Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
-Perhaps equally as crucially, they found the \emph{absence} of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas.
-This observation greatly helped to refine models for the gas dynamics in the bubble and clarified several unexplained experimental observations \cite{}.
-With this in mind, we choose only to focus on presenting the essential results needed to discuss the dynamics of a bubble undergoing SBSL.
-Excellent books and modern reviews on modelling the gas dynamics exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.
-We will hold off on discussing the dynamics in the plasma until a later section.
-
-The most straighforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
-With simplifying assumptions, equations of motion with varying degrees of sophistication are derived for the gas and solved numerically, consistent with the RP equations.
-The discussion proceeds in much the same way as deriving the RP equations above \cite{}, so we will omit it here.
-Instead, we will highlight the salient features before introducing a simple model for the gas that will enable to gain some physical understanding analytically.
-
-The important results of direct numerical solution of the gas dynamics are as follows: chemical reactions in the bubble are important but only in the immediate vacinity of SBSL.
-Diffusion into/out of the bubble of dissociated species is also important but also mainly occurs in the vacinity of SBSL.
-In model calculations, these things are then usually neglected.
-{\color{red} Go on to talk about heat transfer, mass transfer, etc.
-following \cite{brenner2002single}.
-} To summarize, careful solution of the equations of motion show that the contents of the bubble are, to a good approximation, spatially uniform \cite{}.
-This fact allows us to simplify a model for the gas dynamics greatly \cite{}.
-Before we turn to discuss this, let us pause here to mention that another theoretically appealing method for modelling the gas dynamics is \emph{molecular dynamics} \cite{}.
-While this method is attractive for a number of reasons, practical calculations containing realistic numbers of particles aren't possible \cite{brenner2002single}.
-
-The simplest model for the gas dyanmics in the bubble is {}:
+While the RP equation is pretty, it is non-linear in $R$, making its analytical solution intractable.
+In some cases, linearizating in $R$ is reasonable and makes an analytical solution possible but in general, this equation is solved numerically.
+We will explore this below.
+{\color{red} Go thru \cite{brenner2002single} and maybe \cite{gaitan1990experimental}.
+Different models include, to different extents, mass and heat transfer across the interface, with good results.
+These are included by integegrating the Navier-Stokes equations assuming energy is not conserved, mass is not conserved, etc.
+Just state the different equations and heurestically describe how they come about.
+The point of all of them is the dynamics are modified somewhat.
+In \cite{prosperetti1999old}, these different formulations are all reduced to a particular form of \emph{one parameter} equations with different numerical results.
+Discuss these and show which one is the most accurate so far.}
 
 \subsection{Experimental Results}
+
 Creating stable, single bubbles in a laboratory experiment is not particulary challenging and can be done with standard and low-cost materials \footnote{Actually, the phase-space for SBSL is ``small".
-The coordinates being driving amplitude, driving frequency, fluid composition, temperature, ambient pressure, gas content and composition, etc...
-What we are claiming is simple is the experimental setup once the correct parameters are identified.}.
+The phase-space being over the variables driving pressure, driving frequency, fluid composition, temperature, ambient pressure, gas content, gas composition, etc...
+What we are claiming is simple is the experimental setup once the correct parameters are identified is relatively simple.}.
 A typical experimental setup is shown in Fig.
-\ref{fig:flask}.
-Since we are mainly concerned with the light emission process itsself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}.
-A sample of degassed liquid is placed into a flask \footnote{Actually, the degassing part is quite important.
-We focus on it elsewhere.}.
+\ref{fig:flask}; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation}.
+A sample of degassed liquid is placed into a flask.
 Coupled to the outside of the flask are piezoelectric transducers of some sort.
-The piezos are driven by a sinusoidal function generator to excite an acoustic resonance of the flask: a typical flask is a few cm across with a resonance at $\approx 20$ kHz \cite{brenner2002single}.
-The frequency is chosen to form a pressure \emph{antinode} at the center of the flask.
 
-Let us take a brief aside: in deriving the RP equations, we assumed the center of the bubble was stationary.
-Actually, this is what we want in an SBSL experiment and it is worth discussing under what conditions this is true.
-In an ideal, incompressible fluid the net force acting on a bubble may be written as
-\begin{equation}
-    \bm{f}=-\int_{\partial \Omega} p \bm{n} dA = -\int_\Omega \nabla p dV
-\end{equation}
-Here, $\bm{f}$ is the net force, $p$ is the pressure at the bubble's surface, $\bm{n}$ is the unit normal vector point out of the bubbles surface, and $\partial \Omega$ is the bubble's boundary with $\Omega$ the bubble's domain.
-The last term is due to an application of Gauss's theorem.
-Bjerknes \cite{bjerknes1909kraftfelder} took $\nabla p \equiv \nabla p(r=0,t)$ to be constant and averaged over one period of oscillation to find
-\begin{equation}
-    \bm{f} = -\frac{4}{3}\pi \langle R^3 \nabla p \rangle.
-    \label{eq:Bj_force}
-\end{equation}
-This is the so called \emph{Bjerknes force}.
-The point is that if the bubble is located at a pressure extrema, e.g.
-the center of the flask, then this force should vanish and the bubble will remain stationary, an essential result for SBSL experiments.
-
-Some authors have looked at corrections to the bubble's dynamics in the presence of bouyancy, which causes a net translation other than the Bjerknes force.
-The result is the bubble oscillates around an equilibrium position near but away from the center of the flask \cite{matula2000single,matula1997bjerknes,matula1999inertial}.
-The translational motion of the bubble leads to shape distortion which is beleived to supress SL; this is confirmed by SBSL experiments in microgravity resulting in larger intensity light production \cite{matula2000single}.
-
-Back to the experimental setup.
-The typical driving pressure relevant to SBSL ($\sim 1.5$ atm) is too small to cause bubbles to form spontaneously in the flask \cite{}.
-Instead, the bubble is usually seeded somehow.
-In the original work of Gaitan, an air bubble was injected using a needle and syringe through a small hole in the flask's seal \cite{gaitan1992sonoluminescence}.
-More recently, seeding methods involve shooting a small jet of water into the flask (like our shrimp friends!) or blasting the water with a laser to boil a small volume \cite{}.
-The later method is preferred as it enables more precise control over the initial bubble's radius.
-The small bubbles formed this way coalesce at the center of the flask into a single, stable bubble \cite{}.
-
-Now that we know about how single, stable bubbles are formed in the lab, we can discuss a number of quantities of experimental interest.
-The bubble's radius may be measured by \emph{Mie scattering} \cite{}.
-We may also analyze light spectrum which contains information on the bubble's contents: in principle, we can learn about the temperature inside the bubble and spectral lines contain information on the matter \cite{}.
-As we will see, it is actually not so simple to understand the light spectrum \cite{}.
-...
-
-{\color{red} what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}) (Ar content, driving frequency, P$_{max}$)
+what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}) (Ar content, driving frequency, P$_{max}$)
 
@@
 In this paper, they present the discovery of SBSL...
-}
+
+\section{What don't we know}
+\subsection{The Bubble Wall (Again)}
+\subsection{Experimental Anomalies}
 
 \section{Let there be light!}
-\subsection{The Plasma}
-\subsection{Brehmasstunglingnls}
-\subsection{Recombination}
-\subsection{Electronic transitions}
+\subsection{Thermal Arguments}
+\subsection{Electric Arguments}
+
+\subsection{The Casimir Effect}
+\subsection{The Casimir Argument}
 
@@
 
-\section{The Navier-Stokes Equations}
+\section{The Rayleigh-Plesset Equation}
 The Rayleigh-Plesset (RP) equation can be deduced from the Navier-Stokes equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
-To that end, we remind ourselves of the Navier-Stokes equations here.
+To that end, we first remind ourselves of the Navier-Stokes equations.
 
@@
 
-If instead the fluid is real, we can't neglect viscous forces.
+If instead the fluid is real, we can't neglect viscous forces (i.e.
+dissipation).
 We suppose that the viscous surface term can be represented by a tensor, $\hat{\bm{\tau}}$ such that $\bm{F}_v=\int_{\partial \Omega_t} \hat{\bm{\tau}}\cdot \bm{n} dA = \int_{\Omega_t} \nabla \cdot \hat{\bm{\tau}} dV$.
@@
 In general, $\hat{\bm{\mu}}$ is a symmetric rank-4 tensor.
-However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in sonoluminescence experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
+However, if we assume an isotropic fluid \footnote{Definitely true for water which is usually the fluid used in sonoluminescence experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
 \begin{equation}
@@
 It is usual to combine the pressure (i.e.
-elastic-stresses) and the viscous-stresses into the stress-tensor $\hat{\bm{\sigma}}=-p\hat{\bm{I}}+\xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)}$.
+elastic stresses) and the viscous-stresses into the stress-tensor $\hat{\bm{\sigma}}=-p\hat{\bm{I}}+\xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)}$.
 Combined with this, Newton's second law becomes
@@
      \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
-     \label{eq:ns_equations_2}
+     \label{eq:ns_equations}
 \end{split}
@@
 The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
-On the other hand, the terms $\sim \bm{u}$ are viscous terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
+On the other hand, the terms $\sim \bm{u}$ are viscous (damping) terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
 The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
 dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
-The body term $\bm{f}_B$ is from external sources.
-It is worth noting that solving eqs.
-\ref{eq:ns_equations_2} is a formidable task \footnote{In fact, it's not even clear that the NS equations \emph{can} be solved even for incompressible fluids.
+The body term $\bm{f}_B$ is from external sources: from here on, we will assume it is zero.
+
+Even with these simplifying assumptions, solving eqs.
+\ref{eq:ns_equations} is a formidable task \footnote{In fact, it's not even clear that the NS equations \emph{can} be solved even for incompressible fluids.
 In 2000, it was announced that proving smooth, sensible solutions exist earns \$1,000,000 \cite{fefferman2006existence}.
@@
 inflation was 6.8\% in Nov.
-2021.} and is usually done using computers.
+2021.} To that end, several more-drastic approximations, valid for our problem, have to be made that allow us to make progess.
+\begin{enumerate}
+    \item We will assume irrotational flow (i.e.
+$\nabla\times\bm{u}\equiv 0$) and only radial motion in the liquid (i.e.
+$\bm{u}=u\bm{r}$).
+Note that, since the flow is assumed irrotational, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
+This amounts to assuming that the bubble is always spherical.
+This seems like a rather drastic approximation but has been validated experimentally in many cases \cite{}.
+The fact that the bubble tends to remain spherical can be understood by accounting for surface-tension at the liquid-bubble interface \cite{}.
+    \item Next, we assume that the viscous terms are negligible in the bulk dyanimcs of the liquid.
+For (relatively) low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation and is widely used \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
+We will acount, to some extent, for viscosity later when looking at the bubble-liquid interface.
+Moreover, we are assuming that the flow is \emph{isentropic}, i.e.
+that it is reversible (no damping) and that no heat is exchanged between fluid parcels.
+We will further assume that the liquid is isothermal, i.e.
+its temperature is constant (in space).
+Then pressure $p$ is determined from an instantaneous equation-of-state $p=p(\rho,T)$.
+In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and, as we will later see, heat-transport across the liquid-bubble interfacing is usually neglected anyway (i.e.
+the bubble is compressed adiabatically in the thermodynamic sense \footnote{According to Wikipedia, \emph{adiabatic} means \emph{fast} in thermodynamics lingo.
+This is relevant to us since we are claiming the bubble wall moves so quickly that it is compressed to its maximum pressure before any heat can flow out.
+On the otherhand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
+This is, e.g.
+the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its groundstate at all times.})
+    \item Related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
+Of course the bubble-liquid interface enters both systems as a boundary condition \cite{prosperetti1986bubble}.
+\end{enumerate}
 
-\section{Electrical Arguments}
+With these simplifying assumptions (and a vector identity \footnote{$\left( \nabla \cdot \bm{u} \right) \bm{u} = \frac{1}{2} \nabla (u^2) - \bm{u}\times \left( \nabla \times \bm{u} \right)$}) we rewrite eqs.
+\ref{eq:ns_equations} as a system of $1+1$ dimensional partial differential equations:
+\begin{equation}
+\begin{split}
+     \partial_r \left[ \partial_t \phi +\frac{1}{2} \left( \partial_r \phi \right)^2 \right] & = - \frac{1}{\rho} \partial_r p  \\ 
+     \partial_t \rho+ \partial_r \phi \partial_r \rho + \rho \partial^2_r \phi & = 0.
+     \label{eq:NS_1D}
+\end{split}
+\end{equation}
+We can integrate the first equation above using the fact that the compressibility is negligible to move the density through the integration \cite{leighton2007derivation}
+\begin{equation}
+    \int_{r_\infty}^{r} \frac{\partial}{\partial r'} \left[ \frac{\partial \phi}{\partial t} +\frac{1}{2} \left( \frac{\partial \phi}{\partial r'} \right)^2 \right] dr' = - \frac{1}{\rho} \int_{r_\infty}^{r} \frac{\partial p}{\partial r'} dr'.
+\end{equation}
+On the left side, we take the velocity to vanish at infinity.
+On the right side, we assume that only the static and unperturbed pressures, $p_\infty$ and $P(t)$ respectively, are relevant so that \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}
+\begin{equation}
+    \partial_t \phi + \frac{1}{2}\left( \partial_r \phi \right)^2 = \frac{(p_\infty+P(t))-p(t,r)}{\rho}.
+    \label{eq:press_vel}
+\end{equation}
+Eq. \ref{eq:press_vel} gives the pressure at cordinate $(r,t)$ in terms of the velocity of the fluid, the density, and the appied pressure.
 
-\section{The Casimir Argument}
+Now, let us assume that the velocity potential satisfies a wave equation $\nabla^2 \phi - \frac{1}{c^2}\partial_t^2 \phi =0 $ \footnote{This doesn't come from thin air.
+We could write the r.h.s.
+of eq.
+\ref{eq:press_vel} as the enthalpy, $h\equiv \int \frac{dp}{\rho}$, and similarly introduce $dh=\frac{dp}{\rho}$ in the first line of eq.
+\ref{eq:NS_1D}.
+Eliminating $h$ between the two and dropping small terms $\sim \frac{u}{c}$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.} with $c$ the speed of sound in the fluid.
+We can again include incompressibilty by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$.
+\footnote{For our problem, it is true that $\left( \frac{1}{c^2} \partial_t \phi \right) / \nabla^2 \phi \sim \left( \frac{R}{\lambda} \right)^2 \ll 1$, i.e.
+the long wave-length approximation is accurate \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}.} In this case, we can ignore retardation effects and, remembering that we have spherical symmetry, write the solution of $\nabla^2 \phi$ as
+\begin{equation}
+    \phi = \frac{\psi(t)}{r}
+\end{equation}
+with $\psi$ the time-dependent coefficient. We now use the boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, to determine $\psi$:
+\begin{equation}
+    \dot{R} = -\frac{\psi}{R^2} \Rightarrow \psi = -R^2 \dot{R}.
+\end{equation}  
+Thus, we find $\phi(r,t) = -\frac{R^2 \dot{R}}{r}$. Plugging this into the left hand side of eq. \ref{eq:press_vel} and evaluating at the bubble wall, $r\equiv R$
+\begin{equation}
+\begin{split}
+    \partial_t \phi \vert_{r=R} & = -2\dot{R}^2-R\ddot{R} \\
+    \partial_r \phi \vert_{r=R} & = \dot{R} \equiv u(R)
+\end{split}
+\end{equation}
+
+Finally, sticking all this together, we arrive at the Rayleigh-Plesset equation \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,plesset1949dynamics,plesset1977bubble}
+\begin{equation}
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{p_B(t)-(p_\infty+P(t))}{\rho}
+    \label{eq:RP_1}
+\end{equation}
+with $p_B \equiv p(r=R,t)$ the pressure in the fluid at the bubble wall.
+Evaluated beyond the bubble wall, $r > R$, eq.
+\ref{eq:RP_1} gives the pressure radiated by the fluctuating bubble.
+Some comments are in order.
+The combination $p_\infty+P(t)$ is the \emph{ambient} pressure \cite{prosperetti1999old,prosperetti1986bubble} which gives the pressure in the fluid in the absence of the bubble.
+The left hand side of eq.
+\ref{eq:RP_1} may be regarded as the kinetic energy (density).
+The right hand side is the change in enthalpy (density) due to the fluctuating bubble, i.e.
+the dyanmics of the bubble given on the left hand side are determined by the enthalpy in the fluid due to the bubbles motion.
+In deriving this equation, we have assumed that the speed of propogation in the fluid is infinite.
+In reality, this is not true but this approximation will be very good close to the bubble (in the \emph{near field}) where retardation effects are negligible anyway.
+In a similar way, we have assumed that the energy in the fluid due to distorting its volume is negligible.
+This approximation will be accurate near the bubble as well since the kinetic energy in this region will dominate \cite{prosperetti1999old}.
+
+We can refine eq.
+\ref{eq:RP_1} a little further.
+Since the motion in the fluid is purely radial, we expect the only relevant stresses to be normal to the bubble wall.
+Let's denote the radial stress in the liquid at some point $r$ by $s_r(r)=-p(r,t)+2\mu \partial_r u$.
+Recall, $\mu$ is the fluid viscosity from earlier.
+Using conservation of mass and momentum, and our expression above for the velocity, we can equate the normal stresses in the fluid to those in the bubble:
+\begin{equation}
+\begin{split}
+    s_r(R)=-p_B-4\frac{\mu \dot{R}}{R}=-p_g+2\frac{\sigma}{R}
+\end{split}
+\end{equation}
+where $p_g$ is the pressure in the gas and $\sigma$ is the surface tension. Solving for $p_B(t)$, we have the relation \cite{brenner2002single,prosperetti1999old,prosperetti1986bubble}
+\begin{equation}
+    p_B(t)=p_g(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)
+    \label{eq:p_B}.
+\end{equation}
+Note that in deriving the equations \ref{eq:NS_1D} we neglected the viscosity terms.
+This was valid since only combinations of viscosity and compressibility $\nabla \cdot \bm{u}$ appeared, both of which are assumed small.
+Here, however, we opt to keep the viscosity terms coupling the bubble wall to the fluid.
+With this, we can rewrite the RP equation as
+\begin{equation}
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) \right]
+    \label{eq:RP_2}
+\end{equation}
+with $\Delta p = p_g(t)-p_\infty$ the deviation of the pressure inside the bubble from the static pressure.
+With the above assumption that the compression is adiabatic, the conditions inside the bubble are decoupled from the liquid (besides the coupling through the RP equation) and we may regard $p_g(t)$ in the bubble as a given quantity to be determined from an equation of state.
+Recall $P(t)$ is the time-dependent part of the pressure in the absence of the bubble: we may regard this as an external pressure from e.g.
+a driving stress (this will be important in the main text).
+
+To summarize, the Rayleigh-Plesset equation eq.
+\ref{eq:RP_2} gives the equation of motion for a spherical bubble in an incompressible liquid driven by a time dependent pressure $P(t)$.
+This is one of the more important results for the subject of this paper.
 

### 2022-01-06 sonoluminescence/paper/bak/v2/paper.tex
sdfadsfa

--- 2022-01-06 sonoluminescence/paper/paper.tex
@@
 
-The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (i.e.
-the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
+The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (from which we may calculate the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
 (ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{}.
@@
 A result of this work was the \emph{argon rectification hypothesis} \cite{lohse1997sonoluminescing,brenner2002single,yasui2018acoustic,suslick2008inside}.
-The hypothesis claims that all species (in air) in the bubble besides argon are gradually ejected from bubble's interior until all that remains is pure argon.
+The hypothesis claims that all species in the air inside the bubble besides argon are gradually ejected from bubble's interior until all that remains is pure argon.
 The theory is based on the fact that, at the elevated temperatures and pressures inside the bubble, dissociation of O$_2$ and N$_2$ into radicals is possible.
@@
 Over many cycles, the contents of the bubble become nearly pure argon.
-Moreover, it was realized that the SBSL was much more intense when the contents of the bubble are a pure inert gas: if the contents were e.g.
-molecules, bond breaking/formation would alter the pressure of the gas, ultimately reducing the temperature and reducing the light production \cite{brenner2002single,yasui2018acoustic,lohse2018bubble,suslick2008inside}.
+Moreover, it was hypothesized that the SL would much more intense when the contents of the bubble are a pure inert gas: if the contents were e.g.
+molecules, bond breaking/formation would alter the pressure of the gas, ultimately reducing the temperature and reducing the amount of light production \cite{brenner2002single,yasui2018acoustic,lohse2018bubble,suslick2008inside}.
 This mechanism has been used to explain the relatively weak light, i.e.
 the low temperature, in MBSL: the bubbles are transient and cannot remove a significant amount of O$_2$ or N$_2$ over a single cycle.
-The current belief is that the contents of a bubble in SBSL are pure argon after $\sim 10^3$ cycles \cite{brenner2002single,yasui2018acoustic}.
+The current belief is that the contents of a bubble undergoing SBSL in water are pure argon after $\sim 10^3$ cycles \cite{brenner2002single,yasui2018acoustic}.
 
@@
 A simple but reasonable model for the gas dyanmics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.
-The modifications to the result for an ideal gas are simple.}.
+The modifications to the result for an ideal gas are simple \cite{sivasubramanian2002temperature}.}.
 Let us remind ourselves what the pressure is in this case.
@@
 The point is that if the bubble is located at a pressure extrema, e.g.
-the center of the flask, then this force should vanish and the bubble will remain stationary, an essential result for SBSL experiments.
+the antinode at the center of the flask, then this force should vanish and the bubble will remain stationary, an essential result for SBSL experiments.
 

### 2022-01-07 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 Snapping shrimp, like our cute friend in Fig.
-\ref{fig:shrimp_claw} (left), have long been known to produce cavitating bubbles by \emph{snapping} their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
+\ref{fig:shrimp} (left), have long been known to produce cavitating bubbles by \emph{snapping} their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
 They have a strong appendage (the \emph{dactyl}) in thier claw [Fig.
-\ref{fig:shrimp_claw} (center)] that is used to create a very high-velocity jet of water.
+\ref{fig:shrimp} (center)] that is used to create a very high-velocity jet of water.
 The low pressure region in the jet's wake forms a bubble [Fig.
-\ref{fig:shrimp_claw} (t=0 ms)] that, among other things, produces an exceptionally loud noise when it collapses.
+\ref{fig:shrimp} (right)] that, among other things, produces an exceptionally loud noise when it collapses.
 The sound (i.e.
@@
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{figs/shrimp_claw.pdf}
+\includegraphics[width=0.95\linewidth]{figs/shrimp.pdf}
     \caption{(left) Snapping shrimp (\emph{Alpheus heterochaelis}).
@@
 The \emph{plunger} (pl) on the \emph{dactyl} (d) rapidly enters the \emph{socket} (s), ejecting a high-velocity jet of water.
-The water ejection and subsequent bubble formation and, finally, bubble collape (at t=0) are shown on the right.
-The time-offset are shown in the panels in ms.
-Adapted from ref.
-\cite{versluis2000snapping}}
-\label{fig:shrimp_claw}
+A time near the bubble collape is shown on the right.
+The light emission is too dim to be seen by the naked eye.
+Adapted from refs.
+\cite{versluis2000snapping} and \cite{lohse2001snapping}.}
+\label{fig:shrimp}
 \end{figure}
@@
 The shock-wave produced by the cavitating bubble is used to stun and even kill prey \cite{versluis2000snapping}.
-If the shrimp's prey had very sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.
+If the shrimp's prey had very\footnote{We stress \emph{very}: shrimpoluminescence is too weak to be seen by the naked-eye.} sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.
 
 Sonoluminescence (SL) is more precisely defined as the process by which a ``driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission" \cite{brenner2002single}.
-Sonoluminescence comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonolumiscence.
+SL comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonoluminescence.
 The distinction is self-explanatory: multi-bubble sonolumiscence (MBSL) consists of  ``the simultaneous creation and destruction of many seperate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
 The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
-We will discuss MBSL breifly throughout this paper, but for now it suffices to say that nearly all theoretical and experimental progress to explain sonoluminescence has been made using SBSL, with some authors even calling it ``the hydrogen atom of sonoluminescence" \cite{lohse2018bubble,crum1994sonoluminescence}.
+In SBSL, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volumetric-light emission from the bubble cloud, etc.
+Similiarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls.
+As a result, the theory is greatly simplified compared to MBSL.
+It is not an exaggeration to say that nearly all theoretical and experimental progress to explain SL has been made using SBSL, with some authors even calling it ``the hydrogen atom of SL" \cite{lohse2018bubble,crum1994sonoluminescence}.
 
-In SBSL, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volumetric-light emission from the bubble cloud, etc.
-Similiarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls and the theory is greatly simplified compared to MBSL.
-SBSL, while offering advantages, also comes with complications.
+Studying SBSL, while offering advantages, also comes with complications.
 Due to the tiny volume of the bubble, directly measuring the temperature or pressure \emph{inside} the bubble is practically impossible \cite{suslick2008inside}.
@@
 \cite{brenner2002single}.} to more exotic explanations based of quantum field theory \cite{schwinger1993casimir,eberlein1996sonoluminescence,liberati2000sonoluminescence}.
+Over time, the \emph{phase-space} of SBSL has been refined: SBSL only occurs for a narrow range of driving pressures and for a particular combination of host-liquid and gas contnent in the bubble.
+The mapping of the phase-space of SBSL is a very rich subject and merits a review in itself (several exist \cite{brenner2002single,yasui2018acoustic}).
+We will not pursue this; our goal in this paper is to explain \emph{shrimpoluminescence} so we will mainly focus on SBSL as it occurs in \emph{water} containing \emph{air bubbles}, though occasionally we will comment on SBSL in other fluids.
 
-But besides the obvious case of shrimpoluminescence, why is SL interesting?
-Well, at a glance, it is not obvious why SL occurs.
-The acoustic wave in the liquid displaces molecules on the order of nanometers, costing an elastic energy of  $\sim1\times 10^{-12}$ eV/molecule while emitting visible light (through e.g.
-electronic transitions) costs an energy $\sim1$ eV/molecule \cite{lohse2018bubble,crum1994sonoluminescence}...
-there is a 12-orders-of-magnitude concentration of energy.
-That is huge!
-Estimates of the temperature at the center bubble are ...
-
-%After a moments thought, however, the massive energy focusing is not that surprising.
-For example, the energy in the shock wave produced by a cavitating bubble, regardless of the light production, has been known to damage ship propellers since atleast 1917 when Lord Rayleigh's gave the first rigorous treatment of caviating bubbles \cite{gaitan1990experimental}.
-
-%So again, why is the light emmission interesting? In short, after 3 decades of studying SBSL, the mechanism of light production is still not understood.
-
-%The excitment wasn't restricted to academics however, with SBSL catching Hollywood's attention leading to an unpopular movie called \emph{Chain Reaction} where t
-
-After 3 decades of studying SBSL, the physical mechanism of light production in SL is still not understood.
-This can be seen by glancing at the literature and noticing that there are many recent papers with different arguments for \emph{where} \footnote{We emphasive ``where" since it's not even certain if the light is \emph{surface} or \emph{volumetric} emission, i.e.
-it is literally not known \emph{where} the light comes from \cite{}.} the light comes from \cite{borisenok2020mechanisms,flannigan2013non,flannigan2012temperature,tatartchenko2017sonoluminescence}.
-Numerous aspects of the process are conveniently accessible to the experimentalist, while at the same time the theory of the bubble's interior is quite mature.
-Many theories require experimental inputs as parameters, while in other cases experimental results are indirectly \emph{inferred} by fitting to a theory.
-Put simply, the history of studying SBSL is a rather beautiful example of the scientific process of explaining nature.
-Accorind to Brenner ``SBSL has become a rather sophisticated testing ground for the ability of mathematical models and numerical simulations to explain detailed experimental data from a complicated physical process" \cite{brenner2002single}.
-
-The goal of this paper is to catch the reader up on recent efforts to explain SBSL.
-While the author of this paper is particularly excited about shrimpoluminescence, it is important to stress that this work will focus on SBSL in general.
-It is apparently simpler to create and characterize SBSL in the laboratory without involving shrimp (for example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}), so practically all work to study SL has not involved shrimp \footnote{ With notable exceptions \cite{tang2019bioinspired}}.
-Also notably absent will be any detailed discussion of MBSL since this will take us too far afield.
-Instead, we will summarize the history of SBSL research up to now starting at the discovery of MBSL leading to the subsequent discovery of SBSL and a rush of effort to explain it.
-We will explore both the experimental and theoretical discoveries along the way.
-Finally, we will review the current state of agreement between what is known experimentally and theoretically about SBSL.
-
-\section{Historical Overview}
-
-SL was discovered by accident in 1933 (in the form of MBSL) by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz} \footnote{Neither the paper due to Marinesco et al.
-nor the one due to Frenzel et al.
-can be found by the author; in any case he can't read French or German so having the papers wouldn't be much use.
-As such, the story of how the effect was discovered is taken from more recent sources whose authors hopefully could read French and German \cite{brenner2002single,gaitan1990experimental,crum1994sonoluminescence}} Marinesco et al.
-were trying to accelerate photo development by \emph{insonating} developing fluid.
-They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy" which they attributed to exposure to light.
-Shortly after, Frenzel et al.
-repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
-
-This result was not particularly surprising to the community since it had been known for a while that cavitating bubbles could do tremendous damage to e.g.
-ship's propellers \footnote{A more modern example could be my van's oil pump, which is in need of replacing.}.
-The discovery's impact is consicesly summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
-In fact, cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
-The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures could become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence}.
-Cavitation was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
-Shortly after, Lord Rayliegh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigoruous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
-Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
-Still, very little information was accessible about the light emission until the 1990's when stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing caviation.
-A stable bubble persists through multiple cycles of cavitation, oscillating nonlinearly around an equilibrium size.
-Transient bubbles appear and then collapse to dissapear}, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
-With this discovery, serious interest took root.
-
-Details of a typical SBSL experiment will be given later, but for now it suffices to note that SBSL, unlike previous studies on MBSL, allowed very precise control and measurment of the SL process.
-With unprecidented experimental control, many discrepencies in previous assumptions about SL were discovered.
-Experimental Results measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
-This discovery implied that SL was nearly decoupled from the bubble's dynamics, contradicting models based on Rayleigh's equation.
-New models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubbles $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
-At the same time, experimentalists fit the bubble's light emission spectra as a \emph{black body emitter} and concluded that the temperature in the bubble was atleast 25,000 K \cite{hiller1992spectrum}.
-
-These very high estimates for the temperature in the bubbles lead to a spark of interest more broadly: if the temerature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
-An action movie featuring Keanu Reeves was even made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
-In the early 2000's, several papers were even published claiming that stable nuclear fusion was possible in the lab using a method based on SBSL \cite{}.
-Unfortunately these results turned out to be fraudulent \cite{}.
-Shortly after this fiasco, evidenced emerged that there are shape instabilities in the bubbles dynamics indicating that the converging shock-wave hypothesis is probably not right.
-This severely reduced the estimates for the temperature to below $10,000$ K, ruling out the possibility of fusion occuring at all \cite{}.
-Moreover, it was shown that the light is most likely emitted from the \emph{bulk} of the bubble as opposed to its surface, making fits to a black body specturm (i.e.
-surface emission) incorrect \cite{}.
-Current estimates for the temperature at the center of a sonoluminescing bubble are $...$ \cite{}
-
-Still, the very high pressures and temperatures in the bubble during sonoluminescence see technological applications in chemistry and biosciences. ...
-
-History has shown that the most important part of understanding SBSL is understanding the dyanmics of a single acoustically driven bubble.
-We will see that the dynamics of the gas \emph{inside} the bubble is nearly decoupled from the dyanmics of the \emph{fluid} containing the bubble so that we may look at these two problems seperately.
-Firstly, we look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
-Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
-After that, we will combine these two and discuss solutions of the full problem.
-The ultimate result of this will turn out to be an estimate of the \emph{temperature} in the bubble.
-We will then look at theories presenting the origin of light from the very hot contents of the bubble.
-Finally, we will wrap this whole thing up with a summary of unsolved problems and an outlook to the future.
+The goal of this paper is to explain the basic ingredients needed to understand SBSL.
+While the author of this paper is particularly excited about shrimpoluminescence, this work will focus on SBSL in general as it is apparently simpler to create and characterize SBSL in the laboratory without involving shrimp (for example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}).
+Besides notable exceptions \cite{tang2019bioinspired}, practically all work to study SL has not involved shrimp.
+Also absent will be any detailed discussion of MBSL.
+Instead, we will summarize the history of SBSL research up to now starting at the discovery of MBSL, leading to the subsequent discovery of SBSL and a rush of effort to explain it.
+We will then present a sketch of the relevant physics before visiting each topic in detail, finally combining everything together to give a complete picture of how shrimpoluminsence occurs.
 
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{figs/flask.pdf}
+\includegraphics[width=0.95\linewidth]{figs/flask.pdf}
     \caption{(a) Schematic of a spherical acoustic levitation cell.
@@
 
+\subsection{Historical Overview}
+
+MBSL was discovered by accident in 1933 by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz} \footnote{Neither the paper due to Marinesco et al.
+nor the one due to Frenzel et al.
+can be found by the author; in any case he can't read French or German so having the papers wouldn't be much use.
+As such, the story of how the effect was discovered is taken from more recent sources whose authors hopefully could read French and German \cite{brenner2002single,gaitan1990experimental,crum1994sonoluminescence}} Marinesco et al.
+were trying to accelerate photo development by \emph{insonating} developing fluid.
+They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy," which they attributed to exposure to light.
+Shortly after, Frenzel et al.
+repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
+
+This result was not particularly surprising to the community since it had been known for a while that cavitating bubbles, e.g.
+those generated by the sound field, could do tremendous damage to ship's propellers \footnote{A more modern example could be my van's oil pump, which is in need of replacing.}.
+The discovery's impact is consicesly summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
+In fact, cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
+The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures could become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence}.
+Cavitation was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
+Shortly after, Lord Rayliegh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigoruous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
+Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
+The theory has been refined considerably since then \cite{prosperetti1999old,plesset1977bubble,plesset1977bubble,brenner2002single,lofstedt1995sonoluminescing,barber1992resolving} and we will devote a large portion of this paper to this subject.
+Still, very little information was accessible about the light emission until the 1990's when stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing caviation.
+A stable bubble persists through multiple cycles of cavitation, oscillating nonlinearly around an equilibrium size.
+Transient bubbles appear and then collapse to dissapear}, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
+With this discovery, serious interest took root.
+
+SBSL, unlike previous studies on MBSL, allowed very precise control and measurment of the SL process.
+With unprecidented experimental control, many discrepencies in previous assumptions about SL were discovered.
+Experiments measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
+This discovery implied that SL was nearly decoupled from the bubble's dynamics.
+New models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubbles $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
+At the same time, experimentalists fit the bubble's light emission spectra as a \emph{black body emitter} and concluded that the temperature in the bubble was atleast 25,000 K \cite{hiller1992spectrum}.
+
+These very high estimates for the temperature in the bubbles led to a spark of interest more broadly: if the temerature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
+An action movie featuring Keanu Reeves was even made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
+In the early 2000's, several papers were even published claiming that stable nuclear fusion was possible in the lab using a method based on SBSL \cite{}.
+Unfortunately these results turned out to be fraudulent \cite{}.
+Shortly after this fiasco, evidenced emerged that there are shape instabilities in the bubbles dynamics indicating that the converging shock-wave hypothesis is probably not right.
+This severely reduced the estimates for the temperature to $\sim 10,000$ K, ruling out the possibility of fusion occuring at all \cite{}.
+Moreover, it was shown that the light is most likely emitted from the \emph{bulk} of the bubble as opposed to its surface, making fits to a black body specturm (i.e.
+surface emission) incorrect \cite{}.
+Current estimates for the temperature at the center of a sonoluminescing bubble are $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
+
+\subsection{Overview of Single-Bubble Sonoluminescence}
+
+\begin{figure}
+\includegraphics[width=0.6\linewidth]{figs/pulses}
+    \caption{\cite{brenner2002single,crum1994sonoluminescence}}
+\label{fig:pulses}
+\end{figure}
+
+A sketch of SBSL is as follows: an air bubble is trapped at the center of an acoustically driven flask containing a fluid.
+The pressure wave in the fluid drives the expansion and contraction of a bubble and, in turn, the dyanmics of the gas in the bubble respond to the bubble's oscillations.
+The result is a very complicated, non-linear behavior that we will spend a good deal of this paper explaining.
+We will see that the dynamics of the gas \emph{inside} the bubble is nearly decoupled from the dyanmics of the \emph{fluid} containing the bubble so that we may look at these two problems seperately.
+What is now known is that the conditions required for a cavitating bubble to emit light depend sensitively on both the mechanical and chemical properties of the host-liquid and gas in the bubble, as well as the driving pressure and frequency \cite{brenner2002single,suslick2008inside}.
+The mechanical properties determine the pressure and temperature inside the bubble which in turn control chemical reactions: all of these data are used to explain the light emission.
+For extreme enough conditions in the bubble molecular dissociation, atomic/excited-state transitions, and electron-ion/electron-neutral atom bremsstrahlung all occur \cite{}.
+These are the processes which emit light \cite{}.
+
+In very viscous fluids, such as 85\% aqueous H$_2$SO$_4$, the pressure and temperature are lower and most light emission occurs from molecular dissociation and ...
+\cite{}.
+In water, the pressure in the bubble is considerably higher and most light emission is due the electron-ion and electron-neutral atom bremsstrahlung \cite{}.
+Since our goal in this paper is to explain \emph{shrimpoluminescence}, which occurs in water, we will mainly focus on the later emission processes.
+
+Firstly, we will look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
+Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
+After that, we will combine these two and discuss solutions of the full problem.
+The ultimate result of this will turn out to be an estimate of the \emph{temperature} and \emph{pressure} in the bubble.
+We will then look at theories presenting the origin of light from the very hot contents of the bubble.
+Throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
+Connections to more advanced treatments, and thier implications, are provided for completeness.
 \section{Bubble Dyanmics}
-The conditions in the bubble's \emph{interior} are still not entirely understood and will be the subject of an entire subsection of this paper.
-Before we get to this, we will at an older problem: the dynamics of a periodically driven bubble in a liquid.
-An important aspect of the bubble's dyanmics is that they are experimentally accessible.
-Moreover, as we shall see, a (relatively) simple theory describles the dynamics of the bubble quite well.
-On the other hand, measuring what is going on in the bubble's interior is not at all simple though much progress has been made theoretically.
-
-%Before we look at what we don't know, it is useful to summarize what we \emph{do} know about SBSL.
-This has two parts: on the one hand, the dynamics of the bubble are experimentally accessible and have been well studied by \emph{Mie scattering} experiments.
-As we shall see, a (relatively) simple theory describles the dynamics of the bubble quite well.
-On the other hand, a great deal of effort has been invested in measuring what is going on in the bubble's interior.
-We will look at what we have been able to figure out.
-
-\subsection{The Bubble Wall}
 As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
@@
 
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{figs/bubble_radius_2.pdf}
+    \caption{\cite{brenner2002single}}
+\label{fig:bubble_radius_2}
+\end{figure}
+
 Let us breifly digress to discuss retardation effects.
@@
 
-\subsection{The Bubble's Interior}
+\section{The Bubble's Interior}
 
-The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (i.e.
-the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
+The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (from which we may calculate the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
 (ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{}.
@@
 Plugging these into eq.
-\ref{eq:P}, (and sneaking in a modification) we arrive at a very commonly used equation in the context of SBSL \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward}:
+\ref{eq:P}, (and sneaking in a modification) we arrive at a very commonly used equation in the context of SBSL \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward,hilgenfeldt1999simple}:
 \begin{equation}
@@
 It turns out that this process doesn't result in better quantitative predictions since, during the assumed isothermal part of the process, dissipation becomes imporant \cite{brenner2002single}.
+A simple alternative is choosing a constant exponent somewhere between the adiabatic $\Gamma=5/2$ and isothermal $\Gamma=1$ values \cite{hilgenfeldt1999simple}.
 It is also possible to include heat and mass diffusion in simple model forms and these do give slighlty improved quantitative agreement with experiment; however we choose to discuss solutions of the RP equations using the simpler form given in eq.
@@
 \subsection{Experimental Results}
+
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figs/bubble_radius_1.pdf}
+    \caption{\cite{barber1997defining,barber1992light}}
+\label{fig:bubble_radius_1}
+\end{figure}
+
 Creating stable, single bubbles in a laboratory experiment is not particulary challenging and can be done with standard and low-cost materials \footnote{Actually, the phase-space for SBSL is ``small".
@@
 
-Now that we know about how single, stable bubbles are formed in the lab, we can discuss a number of quantities of experimental interest.
-The bubble's radius may be measured by \emph{Mie scattering} \cite{}.
-We may also analyze light spectrum which contains information on the bubble's contents: in principle, we can learn about the temperature inside the bubble and spectral lines contain information on the matter \cite{}.
-As we will see, it is actually not so simple to understand the light spectrum \cite{}.
-...
+Now that we know about how single, stable bubbles are formed in the lab, we can discuss two quantities of experimental quantities that have turned out the be the most important: (i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations \cite{}.
+This can tell us about the temperature and pressure in the bubble.
+(ii) The other quantity of interest is the light spectrum which contains information on the bubble's contents: in principle, we can learn about temperature and pressure inside the bubble, with spectral lines telling us about the matter content \cite{}.
+As we will see, it is not simple to understand the light spectrum \cite{} and we will devote a section of this paper to this topic.
+For now, let us see what measurements of the bubble's dynamics have told us.
 
-{\color{red} what conditions are necessary for light emission (i.e. what is the phase space \cite{lohse2018bubble}) (Ar content, driving frequency, P$_{max}$)
+Mie scattering measurements scatter a laser off a bubble assuming the bubble is a homogenous dielectric sphere \cite{}.
+The angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measurethe bubble's radius as a function of time \cite{}.
+A typical Mie scattering experiment is shown in fig.
+\ref{fig:flask} (c).
+Two important sources of experimental error are changes to the dielectric properties of the bubble during its cycle \cite{} and the fact that cavitating bubbles can emit an outward going shock \cite{}: some light scatters off the outward-going spherical wave too \cite{}.
+Both effects are largest during the SL period of the bubble's cycle, leading to an important source of experimental error \cite{}.
 
-\cite{gaitan1992sonoluminescence} has good discussion of early experimental errors.
-Measuring phase of SL relative to the acoustic field was an early attempt to explain SL.
-Phase here means how far in time is the SL light emission offset from the driven pressure.
-In the case of periodic oscillations, this is a phase.
-A handful of early models assumed that the light emission occured during expansion of the bubble.
-Others assumed that it was at some point during.
-Most of these were ruled out when it was observed that the light emission is for a very short duration near maximum compression, though the phase depends quite a bit on the initial bubble radius, the driving frequency, the resononances of the flask, the physical properties of the liquid, etc \cite{brenner2002single,gaitan1992sonoluminescence}.
-These complicated relationships were unkown to the early investigators of SL and were quite difficult to measure in MBSL.
-For this reason, it wasn't until SBSL was discovered that many of these models were dismissed \cite{gaitan1992sonoluminescence}.
-In this paper, they present the discovery of SBSL...
-}
+Parameters used to calculate the gas dynamics in the bubble can then be fit to the radius to build a full model of the dyanmics.
+In the case of eq.
+\ref{eq:p_g}, the quantities are $P_0$ and $R_0$.
+The results of a calculation using eq.
+\ref{eq:p_g} combined with the eq.
+\ref{eq:RP_3} RP variant fit to experiment are shown in fig.
+\ref{fig:bubble_radius_1} (a).
+The ambient ambient radius and pressure were $R_0=1.45$ $\mu$m and $P_0=1.45$ atm \cite{barber1997defining,barber1992light}.
 
 \section{Let there be light!}
-\subsection{The Plasma}
-\subsection{Brehmasstunglingnls}
-\subsection{Recombination}
-\subsection{Electronic transitions}
+The radiative processes occuring during \cite{hilgenfeldt1999simple}
 
@@
 In general, $\hat{\bm{\mu}}$ is a symmetric rank-4 tensor.
-However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in sonoluminescence experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
+However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in SL experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
 \begin{equation}
@@
 
-\newpage
-
-%\section{random notes}
-
-%argon flash: a method of 'imaging explosions' similar to why argon is more useful for SBSL.
-argon has no internal DOF whereas air (i.e.
-N2, O2, water vapor) does and energy from explosions is absorbed dissociating bonds.
-argon, instead, becomes 'hotter' and emits more light, making photographing explosions easier.
-see wikipedia article "argon flash"
-
-%Brennen defines caviation as (bubble) nucleation when the pressure drops below the vapor pressure...
-as compared to boiling in which the temperature goes above the saturated liquid/vapor temp.
-\cite{brennen2014cavitation}.
-
-%Yasui defines \emph{acoustic cavitation} as the formation and collapse of a bubble driven by a powerful ultrasonic wave \cite{yasui2018acoustic}.
-He defines bubble dynamics as bubble pulsation under the driving sound.
-\textcolor{red}{read this book}
-
-%some dickheads 3d printed a shrimp claw and it also shrimpoluminesces \cite{tang2019bioinspired}
-
-%apparently a shitty movie called 'Chain Reaction' with Keanu Reeves was centered around the premise of MBSL as a means of clean energy production.
-it was based on the speculation that temperatures inside the bubbles ciould reach up to millions of Kelvin and lead to fusion.
-this was debunked by Prosperetti (see the pop sci article \cite{chainreaction}).
-He points out that the temp is speculative and is based on assumptions of a perfect sphere collapsing.
-he notes that its probably not spherical when light is produced and even points out that groups have tried to produce fusion power by intentionally collapsing gas filled micro baloons with lasers...
-and failed.
-Tatartcheceko also shows that this is horsehit \cite{tatartchenko2017sonoluminescence}.
-in this article, Prosperetti also explains how a 'jet' of water whooshing across the bubble could be going faster that 4000 mph, fast enough to cause non-netwownian fracturing of the water leading to light producuction through (apparently) the same mechanism of light emession when ice or wint-o-green mints are snapped in half...
-huh.
-this is good introduction fodder.
-need to find the primary lit.
-
-%for fusions, also see wikipedia article on 'bubble fusion' and 'Rusi P Taleyarkhan'.
-Ruse was some asshole prof at purdue that claimed fusion occured inside cavitating bubbles, but it was later found out that the faked data and interfered with peer reviews leading to him being fucking fired.
-
-%\emph{heurestic explanation of bubble trapping given by Brenner in \cite{chainreaction}.}: the bonds between molecules in a fluid are very strong, so instead of 'tearing', the fluid stretches.
-however, if there is a bubble present, the local dialation of the fluid can change the bubbles diameter by $\sim$ 1000 times to avoid breaking bonds at the interface between bubble and fluid.
-But when the compression waves comes by, the 'gas' in the bubble, which has boiled due to low pressure, offers very little resistance to compression.
-the wave compresses the bubble rapidly, which spikes the temp in the gas to incandescent temperatueres.
-this heuristic explanation is on-track with Crum \cite{crum1994sonoluminescence} below.
-
-%\section{1st discovery}
-%MBSL was discovered first.
-its (obviously) hard to create and control single bubbles so this wasnt feasible until the 1990s, attributed to Gaitan (see below).
-allegedly, it was Marinesco et al \cite{marinesco1933actions} that discovered MBSL in 1933 (according to \cite{gaitan1990experimental} below and to \cite{tatartchenko2017sonoluminescence}) but I cant find this paper (or understand it since its in french).
-Other sources (e.g..
-wikipedia, \cite{tatartchenko2017sonoluminescence,crum1994sonoluminescence}) credit Frenzel et al \cite{frenzel1934luminescenz} as the discoverers.
-I cant German or find this paper either, but in anycase, Tatartchenko summarized thier discovery: Freznel and coworkers were trying to speed up photo development by 'insonating' i.e.
-subjected developer-fluid to sound.
-they found that when they immersed photo-plates or whatever in the fluid, it became 'foggy' and overexposed, as if subjected to diffuse light.
-A bunch of other papers followed up on this.
-get the deets from \cite{gaitan1990experimental} below.
-
-%Update: Brenner \cite{brenner2002single} consisely summarizes the history: Marinesco notes the photographic plates fogs in an ultrasonic bath \cite{marinesco1933actions} and Frenzel et al \cite{frenzel1934luminescenz} deliberately studied the light emission.
-According to brenner, no one was surprised by this since it was already known that cavitating bubbles fuck shit up.
-
-%SBSL was discovered by Gaitan \cite{gaitan1990experimental,gaitan1992sonoluminescence} etc.
-who trapped a single air bubble in a water filled flask using piezoacoutics stuck to the outside.
-(note: the trapping force is called \emph{Bjerknes force}, refs in \cite{lohse2018bubble}).
-
-%\section{notes from \cite{crum1994sonoluminescence}}
-%pop sci type of magazine article.
-Gives heurestic expanation: When an intense sound field is produced within a liquid, microscopic cavities of gas or vapor can be generated when the liquid fails under tensile stress.
-The subsequent acoustic compression cycle forces these cavities to collapse violently, which results in a remarkable concentration of mechanical energy, estimated to be as high as 12orders of magnitude (refs are given too).
-Also notes that MBSL is very hard to study since the bubbles occur randomly and transiently.
-Describes SBSL as the \emph{hydrogen atom of sonoluminescence}.also discusses that attempts to measure the (time) interval over which light is produced have found it be less than $\sim$50 ps which is 1/1000 times smaller than theory predicts.
-They also found the light to be more 'stable' than the acoustic generator, i.e.
-more regular with more controlled intervals.
-an old but probably still relevant result.
-mentions a neat claim that the conditions of the gas at the center of the bubble are probably more like a metal than a gas.
-\textcolor{red}{hell, this article basically summarizes all of the anomalies present in 1994.
-a very good resource.}
-
-%\section{notes from ref \cite{lohse2018bubble}} 
-%SBSL was discovered by Giatan et.
-al.
-in 1990s at U.
-Mississippi \cite{gaitan1990experimental,crum1994sonoluminescence,gaitan1992sonoluminescence} (note, \cite{gaitan1992sonoluminescence} has $\sim$1000 citations..).
-The SBSL experiments were done by sticking peizoacoustic transducers to the outside of a water filled glass.
-the 'Bjerknes' force trapps the bubble and the acoustic wave drives its expansion/collapse.
-What Gaitan saw was that even the single bubble emits light.
-Wow!
-This was baffling since 'acoustic energy' \textcolor{red}{(i dont understand this)} is on the order of $\sim10^{-12}$ eV/molecule while visible light is $\sim1$ eV.
-NVM: I do get this.
-Typical acoustic waves, e.g.
-in air, only displace molecules on the order of nanometers, which aquires an elastic energy gaine between molecules of $\sim10^{-12}$ eV/molecule: certainly not enough to excite an electronic transition.
-
-%This paper refs the review by Propsperetti: \cite{plesset1977bubble}.
-This paper gives EOM for the bubbles radius, the so called \emph{Plesset-Rayleight} equation which I have seen everywhere related to bubbles.
-Talks about solving it, some other crap related to bubble dynamics.
-this requires a detailed re-read.
-
-%Later in the paper, also revisists snapping shrimp.
-
-%\section{notes from ref \cite{prosperetti2004bubbles}}
-%has shit tons of relevant math. take a look. a pretty paper overall
-
-%\section{notes from \cite{gaitan1990experimental}}
-%disseration by Giatan who is credited with discovering SBSL.
-says apparently SL was discovered in 1933.
-need to go read this crap.
-Apparently this was by Marinesco et.
-al.
-in 1933.
-Zimakov tried to explain it in 1934 and concluded that the light was due to electric discharge between vapor cavities and the glass container.
-Frenzel and Schultes in 1935 instead claimed it was from friction between cavitating bubbles and the water.
-Chamber in 1936 studied SL in other nonwater liquids.
-Chamber also postulated the light was from 'tribolumiscence' which occurs when the crystalline structure of the liquid-bubble lattice collapses?
-Apparently somthing similar occurs in real crystals.
-...
-shit, a bunch more models are given.
-give this dis.
-a lot of attention.
-
-%also says caviation was predicted by 'Euler' in 1754: \emph{if the velocity of a fluid is high enouhg, negative pressures can form and the liquid can 'break'.} the 'breaking' was called cavitation in 895 by R.E.
-Froude, whoever the fuck this is, when studying bubbles near propellers.
-Rayleigh, unsurpisingly, was the first to give a rigorous theoretical treatment of cavitation in 1917.
-
-%\section{notes from \cite{brenner2002single}}
-%defines SBSL as ``Single-bubble sonoluminescence occurs when an acoustically trapped and periodically driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission".
-
-%\section{QED}
-%the ref \cite{liberati2000sonoluminescence} has lots of good refs.
-they have follow up papers extending Schwingers papers (which are refd therein) and referencing a bunch of other relevant literature.
-I should look for refs citing Schwingers paper to try to figure out the modern state of the art of this thin.
-
-%\section{Modelling}
-%in the ref \cite{schanz2012molecular}, they use MD to modelling the equation of state for a collapsing bubble. they calculate the temperature and study the conditions for emitting light.
-
-%\section{Refs. to different arguments}
-%ref \cite{didenko2000molecular} gives examples of other origin theories (molecular excited states, black-body radiation, bremsstrahlung, ion-electron recombination, confined electrons)
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-01-08 sonoluminescence/paper/bak/v3/paper.tex
sdfadsfa

--- 2022-01-07 sonoluminescence/paper/paper.tex
@@
 
+\begin{figure}
+\includegraphics[width=0.6\linewidth]{figs/pulses}
+    \caption{\cite{brenner2002single,crum1994sonoluminescence}}
+\label{fig:pulses}
+\end{figure}
+
 Studying SBSL, while offering advantages, also comes with complications.
@@
 \cite{brenner2002single}.} to more exotic explanations based of quantum field theory \cite{schwinger1993casimir,eberlein1996sonoluminescence,liberati2000sonoluminescence}.
-Over time, the \emph{phase-space} of SBSL has been refined: SBSL only occurs for a narrow range of driving pressures and for a particular combination of host-liquid and gas contnent in the bubble.
+Over time, the \emph{phase-space} of SBSL has been refined: SBSL only occurs for a narrow range of driving pressures and for a particular combination of host-liquid and gas content in the bubble.
 The mapping of the phase-space of SBSL is a very rich subject and merits a review in itself (several exist \cite{brenner2002single,yasui2018acoustic}).
@@
 
-The goal of this paper is to explain the basic ingredients needed to understand SBSL.
-While the author of this paper is particularly excited about shrimpoluminescence, this work will focus on SBSL in general as it is apparently simpler to create and characterize SBSL in the laboratory without involving shrimp (for example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}).
-Besides notable exceptions \cite{tang2019bioinspired}, practically all work to study SL has not involved shrimp.
-Also absent will be any detailed discussion of MBSL.
-Instead, we will summarize the history of SBSL research up to now starting at the discovery of MBSL, leading to the subsequent discovery of SBSL and a rush of effort to explain it.
-We will then present a sketch of the relevant physics before visiting each topic in detail, finally combining everything together to give a complete picture of how shrimpoluminsence occurs.
+Let us now sketch the SBSL process a little more precisely.
+An air bubble is trapped at the center of an acoustically driven flask containing a fluid.
+The pressure wave in the fluid, shown as a function of time in Fig.
+\ref{fig:pulses}, drives the expansion and contraction of the bubble and, in turn, the dyanmics of the gas in the bubble responds to the bubble's oscillations.
+Within the phase-space relevant to SBSL, we get the very complicated non-linear response of the bubble's radius as shown in the figure.
+These dynamics lead to SBSL.
+
+First, as the pressure decreases and becomes negative, the bubble expands gradually: surface tension and the negative pressure in the bubble tend to slow expansion.
+Eventually the pressure is increasing and.
+Once it becomes positive, the bubble violently collapses to its minimum radius: during compression, the negative pressure in the bubble is very large and works with the driving pressure to rapidly accelerate the bubble's wall.
+Since the collapse is so quick, the bubble's contents are adiabatically compressed, resulting in very large temperature and pressure in the bubble.
+If everything is just right, the temperature and pressure are large enough for a burst of light to produced as shown by the bottom curve in Fig.
+\ref{fig:pulses}.
+After the collapse, there is a period of \emph{after-bouncing} that continues until the pressure is negative again.
+
+It turns out that the light pulse is very short: only $\sim0.003\%$ of the bubble's cycle \cite{suslick2008inside}, and its spectrum depends sensitively on both the mechanical and chemical properties of the host-liquid and gas in the bubble, as well as the driving pressure and frequency \cite{brenner2002single,suslick2008inside}.
+The mechanical properties determine the pressure and temperature inside the bubble which in turn control chemical reactions: all of these data are used to explain the light emission.
+For extreme enough conditions in the bubble, which occur only breifly, molecular dissociation and recombination, atomic/excited-state transitions, and electron-ion/electron-neutral atom bremsstrahlung are possible \cite{an2009diagnosing,an2008spectral,an2006mechanism,flannigan2005plasma,suslick2008inside,flannigan2006measurement}.
+These are now known to be the processes which emit light \cite{lohse2018bubble,yasui2018acoustic}.
+
+Our goal in this paper is to present the bare minimum theory needed to understand \emph{shrimpoluminescence}.
+To keep things simple, we will concentrate on the dynamics of a bubble undergoing SBSL in the lab as it is apparently simpler to create and characterize SBSL without involving shrimp \footnote{ For example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}.
+Besides notable exceptions \cite{tang2019bioinspired}, practically all work to study SL has not involved shrimp.}.
+Firstly, we will look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
+Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
+The result of this will turn out to be an estimate of the \emph{temperature} and \emph{pressure} in the bubble.
+We will then look at theories presenting the origin of light from the very hot contents of the bubble.
+For relatively mild conditions inside the bubble, most light emission occurs from molecular dissociation processes \cite{}.
+These are the conditions in e.g.
+85\% H$_2$SO$_4$.
+In water, the pressure in the bubble is considerably higher and most light emission is due the electron-ion and electron-neutral atom bremsstrahlung which occur at higher temperature \cite{}.
+Since we want to understand, shrimpoluminescence which occurs in water, we will mainly focus on the later processes.
+Moreover, throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
+Connections to more advanced treatments, and thier implications, are provided for completeness.
+
+\subsection{Historical Overview}
+
+MBSL was discovered by accident in 1933 by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz} \footnote{Neither the paper due to Marinesco et al.
+nor the one due to Frenzel et al.
+can be found by the author; in any case he can't read French or German so having the papers wouldn't be much use.
+As such, the story of how the effect was discovered is taken from more recent sources whose authors hopefully could read French and German \cite{brenner2002single,gaitan1990experimental,crum1994sonoluminescence}} Marinesco et al.
+were trying to accelerate photo development by \emph{insonating} developing fluid.
+They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy," which they attributed to exposure to light.
+Shortly after, Frenzel et al.
+repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
 
@@
 \end{figure}
-
-\subsection{Historical Overview}
-
-MBSL was discovered by accident in 1933 by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz} \footnote{Neither the paper due to Marinesco et al.
-nor the one due to Frenzel et al.
-can be found by the author; in any case he can't read French or German so having the papers wouldn't be much use.
-As such, the story of how the effect was discovered is taken from more recent sources whose authors hopefully could read French and German \cite{brenner2002single,gaitan1990experimental,crum1994sonoluminescence}} Marinesco et al.
-were trying to accelerate photo development by \emph{insonating} developing fluid.
-They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy," which they attributed to exposure to light.
-Shortly after, Frenzel et al.
-repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
 
@@
 surface emission) incorrect \cite{}.
-Current estimates for the temperature at the center of a sonoluminescing bubble are $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
+Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble at $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
 
-\subsection{Overview of Single-Bubble Sonoluminescence}
+\section{Bubble Dyanmics}
+In this section, we will present the theory needed to model and understand the dynamics of a bubble undergoing SBSL.
+We will split this into parts.
+Firstly, we will derive the equations of motion for the bubble's wall.
+Nextly, we will review some important experimental developments that we will then use to argue for a simple model for the dynamics of the gas in the bubble.
+Lastly, we will compare results of a simulated bubble to experimental data.
+The connection to experiment will be what allows us to calculate the temperature in real bubbles.
 
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{figs/pulses}
-    \caption{\cite{brenner2002single,crum1994sonoluminescence}}
-\label{fig:pulses}
-\end{figure}
-
-A sketch of SBSL is as follows: an air bubble is trapped at the center of an acoustically driven flask containing a fluid.
-The pressure wave in the fluid drives the expansion and contraction of a bubble and, in turn, the dyanmics of the gas in the bubble respond to the bubble's oscillations.
-The result is a very complicated, non-linear behavior that we will spend a good deal of this paper explaining.
-We will see that the dynamics of the gas \emph{inside} the bubble is nearly decoupled from the dyanmics of the \emph{fluid} containing the bubble so that we may look at these two problems seperately.
-What is now known is that the conditions required for a cavitating bubble to emit light depend sensitively on both the mechanical and chemical properties of the host-liquid and gas in the bubble, as well as the driving pressure and frequency \cite{brenner2002single,suslick2008inside}.
-The mechanical properties determine the pressure and temperature inside the bubble which in turn control chemical reactions: all of these data are used to explain the light emission.
-For extreme enough conditions in the bubble molecular dissociation, atomic/excited-state transitions, and electron-ion/electron-neutral atom bremsstrahlung all occur \cite{}.
-These are the processes which emit light \cite{}.
-
-In very viscous fluids, such as 85\% aqueous H$_2$SO$_4$, the pressure and temperature are lower and most light emission occurs from molecular dissociation and ...
-\cite{}.
-In water, the pressure in the bubble is considerably higher and most light emission is due the electron-ion and electron-neutral atom bremsstrahlung \cite{}.
-Since our goal in this paper is to explain \emph{shrimpoluminescence}, which occurs in water, we will mainly focus on the later emission processes.
-
-Firstly, we will look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
-Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
-After that, we will combine these two and discuss solutions of the full problem.
-The ultimate result of this will turn out to be an estimate of the \emph{temperature} and \emph{pressure} in the bubble.
-We will then look at theories presenting the origin of light from the very hot contents of the bubble.
-Throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
-Connections to more advanced treatments, and thier implications, are provided for completeness.
-\section{Bubble Dyanmics}
+\subsection{The Bubble Wall}
 As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
@@
 
-\section{The Bubble's Interior}
-
+\subsection{The Bubble's Interior}
 The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (from which we may calculate the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
@@
 We would like to find a suitable form for $\Delta p(t) \sim p_g(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating bubble.
-It turns out that this is a very hard problem for a number of reasons: the conditions inside the bubble depend on gas diffusion and rectification (one-directional diffusion) between the liquid and the bubble, water-vapor conensation and evaporation inside the bubble, and chemical reactions between species in the bubble \cite{brenner2002single}.
+It turns out that this is a very complicated problem: gas diffusion and rectification (one-directional diffusion) between the bubble and liquid vary the number of particles present and, to make things worse, the conditions inside the bubble facilitate chemical reactions between the air and water-vapor changing the properties of gas dynamically \cite{brenner2002single}.
 Brenner et.
-al elegantly summarized the impact of the problem \cite{brenner2002single}: ``one of the exciting features of modern research on SBSL is that it is a testing ground for how well mathematical models can deal with such a complicated situation"
+al elegantly summarized the significance of this problem \cite{brenner2002single}: ``one of the exciting features of modern research on SBSL is that it is a testing ground for how well mathematical models can deal with such a complicated situation"
 
-Now let us breifly discuss the bubble's light spectrum.
-For the first decade or so, SBSL experiments were done using partially degased water as the liquid medium \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}.
+The other method to understand the bubble's interior is its light spectrum.
+For the first decade or so, SBSL experiments were done using partially degassed\footnote{Degassing lowers the ambient pressure, making it easier for bubbles to form.} water as the liquid medium \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}.
 Careful measurements of the emission from SBSL in water resulted in an almost featureless spectrum \cite{}.
@@
 
-SBSL was first measured in partially degassed water \cite{gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single} \footnote{Degassing lowers the ambient pressure, making it easier for bubbles to form.} Early attempts to understand the physics required exploring the phase-space over which SBSL occurs.
-Some attempts were made using non-aqueous host liquids e.g.
+Some attempts were made to measure SBSL using non-aqueous host liquids e.g.
 alcohols, silicone oils \cite{weninger1995sonoluminescence,barber1997defining}, but experiments with air bubbles were not very sucessful \cite{barber1997defining}.
-Instead, different gases were tried in water.
+Eventually, different gases were tried in water.
 Recall, air is $\sim 80\%$ N$_2$, $\sim 20\%$ O$_2$, and $\sim 1\% $ Ar.
@@
 As the pressure becomes very large during the compression stage of SBSL, the soluble materials leave the bubble and do not re-enter since thier solubility in water is enormous compared to the Ar content of the bubble \cite{lohse1997sonoluminescing}.
-Over many cycles, the contents of the bubble become nearly pure argon.
+Over many cycles, the contents of an air bubble in water become nearly pure argon.
 Moreover, it was hypothesized that the SL would much more intense when the contents of the bubble are a pure inert gas: if the contents were e.g.
@@
 
-Most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurments of the light spectum \cite{}.
-More importantly, new measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressures than water.
+Still, most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurments of the light spectum \cite{}.
+More importantly, new measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressure than water.
 Careful experiments revealed that the spectrum depended critically on the noble gas content and driving pressure.
 Suslick et.
-al measured the different emission spectra from Ar, Xe, and Kr bubbles in aqueous H$_2$SO$_4$ \cite{flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
+al measured the different emission spectra from Ar, Xe, and Kr bubbles in aqueous H$_2$SO$_4$, driven to have similar conditions to those in an air bubble in water \cite{flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
 Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
@@
 Let us now connect this to the problem of SBSL.
-In our case, we assume that our bubble has an \emph{ambient} radius $R_0$, leading to volume $V_0 =  4/3 \pi R_0^3$.
-Similarly, we will assume the pressure in the bubble at ambient radius $R_0$ is known and we will call it $p_0$.
-What we want to calculate is the pressure in the gas as a function of the bubble's radius $R(t)$ determined from the RP equations (eq.
+In our case, we assume that our bubble has an \emph{ambient} radius $R_0$ where the pressure is at its equilibrium value, $p_0$.
+The volume occupied by the gas is $V_0 =  4/3 \pi R_0^3$.
+We know that $p_0=p_\infty+2\sigma /R$ from earlier.
+What we want to calculate is the pressure in the gas as a function of the bubble's instantaneous radius $R(t)$ determined from the RP equations (eq.
 \ref{eq:RP_2} or eq.
@@
 \begin{equation}
-    p_g(t) = p_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ \Gamma
+    p_g(t) = \left( p_\infty+2\frac{\sigma}{R_0} \right) \left[ \frac{R_0^3-h^3}{R^3(t)-h^3} \right] ^ \Gamma
     \label{eq:p_g}
@@
 \ref{eq:p_g} can be inserted into an RP equation which can then be integrated numerically to determine the bubble's dynamics.
-Now recall that one of main goals of modelling the bubble dynamics in SBSL was to calculate the temperature inside the bubble in order to model the light emission.
-Following similar steps to deriving $p_g(t)$, we can write down an equation for the temperature $T$ \cite{barber1997defining,brenner2002single}:
-\begin{equation}
-    T(t) = T_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ {\Gamma-1}.
-    \label{eq:T(t)}
-\end{equation}
-Of course we still have to assign a reasonable guess to $T_0$, the temperature in the bubble at ambient radius $R_0$.
 
@@
 
-\subsection{Experimental Results}
+Now recall that one of main goals of modelling the bubble dynamics in SBSL was to calculate the temperature inside the bubble in order to model the light emission.
+Following similar steps to deriving $p_g(t)$, we can write down an equation for the temperature $T$ \cite{barber1997defining,brenner2002single}:
+\begin{equation}
+    T(t) = T_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ {\Gamma-1}.
+    \label{eq:T(t)}
+\end{equation}
+Earlier, we assumed that during the bubble's afterbounces and expansion, the gas dynamics were isothermal.
+Then we take $T_0$ to be the temperature of the liquid.
+The RP equation, eq.
+\ref{eq:RP_3}, combined with the gas dynamical equation, eq.
+\ref{eq:p_g}, and the known liquid and gas properties $\mu$, $\sigma$, and $h$, can be used to calculate the dynamics of an acoustically driven bubble, determining the time-dependence of the pressure and temperature in the bubble along the way.
 
+\subsection{Bubbles in the Lab}
 \begin{figure}
@@
 \ref{fig:flask}.
-Since we are mainly concerned with the light emission process itsself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}.
+Since we are mainly concerned with the explaining the light emission process itself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}.
 A sample of degassed liquid is placed into a flask \footnote{Actually, the degassing part is quite important.
@@
 
-Now that we know about how single, stable bubbles are formed in the lab, we can discuss two quantities of experimental quantities that have turned out the be the most important: (i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations \cite{}.
+Now that we know about how single, stable bubbles are formed in the lab, we can discuss two experimentally accessible quantities that have turned out the be the most important: (i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations \cite{}.
 This can tell us about the temperature and pressure in the bubble.
@@
 
-Mie scattering measurements scatter a laser off a bubble assuming the bubble is a homogenous dielectric sphere \cite{}.
-The angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measurethe bubble's radius as a function of time \cite{}.
+Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
+Assuming the bubble is a homogenous dielectric sphere \cite{}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measurethe bubble's radius as a function of time \cite{}.
 A typical Mie scattering experiment is shown in fig.
@@
 Two important sources of experimental error are changes to the dielectric properties of the bubble during its cycle \cite{} and the fact that cavitating bubbles can emit an outward going shock \cite{}: some light scatters off the outward-going spherical wave too \cite{}.
-Both effects are largest during the SL period of the bubble's cycle, leading to an important source of experimental error \cite{}.
+Both effects are largest during the SL period of the bubble's cycle, leading to an important source of error when fitting the calculated $R(t)$ curve \cite{}.
+Parameters used to calculate the gas dynamics in the bubble can then be fit to the measured radius to build a full model of the dyanmics.
+In the case of eq.
+\ref{eq:p_g}, the quantity to fit is $R_0$.
+The results of a calculation using eq.
+\ref{eq:p_g} combined with the RP variant eq.
+\ref{eq:RP_3} and fit to experiment are shown in fig.
+\ref{fig:bubble_radius_1} (a).
+The ambient ambient radius is $R_0=1.45$ $\mu$m for driving pressure $P_0=1.45$ atm \cite{barber1997defining,barber1992light}.
+Note the large change in the calculated dynamics for relatively small changes in $R_0$ and $P_0$ in Fig.
+\ref{fig:bubble_radius_1} (b) and (c).
 
-Parameters used to calculate the gas dynamics in the bubble can then be fit to the radius to build a full model of the dyanmics.
-In the case of eq.
-\ref{eq:p_g}, the quantities are $P_0$ and $R_0$.
-The results of a calculation using eq.
-\ref{eq:p_g} combined with the eq.
-\ref{eq:RP_3} RP variant fit to experiment are shown in fig.
-\ref{fig:bubble_radius_1} (a).
-The ambient ambient radius and pressure were $R_0=1.45$ $\mu$m and $P_0=1.45$ atm \cite{barber1997defining,barber1992light}.
+More accurate treatments of the gas dynamics from numerically solving the gas's NS equations are qualitatively in good agreement with these simpler model calculations \cite{brenner2002single,yasui2018acoustic}.
+They are, however, in better quantitative agreement when their results are used to calculate the light spectrum \cite{an2009diagnosing,an2008spectral,an2006mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
+Let us turn to discuss the light's spectrum now.
 
@@
 
-\section{The Navier-Stokes Equations}
+\section{The Navier-Stokes Equation}
 The Rayleigh-Plesset (RP) equation can be deduced from the Navier-Stokes equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.

### 2022-01-13 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 \ref{fig:shrimp} (left), have long been known to produce cavitating bubbles by \emph{snapping} their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
-They have a strong appendage (the \emph{dactyl}) in thier claw [Fig.
+They have a strong appendage (the \emph{dactyl}) in their claw [Fig.
 \ref{fig:shrimp} (center)] that is used to create a very high-velocity jet of water.
@@
 The noise produced by groups of shrimp is so intense that the U.S.
-Navy used them as ``sonar-camoflauge" in the Pacific ocean during World War II \cite{versluis2000snapping}.
+Navy used them as ``sonar-camouflage" in the Pacific ocean during World War II \cite{versluis2000snapping}.
 
@@
 The \emph{plunger} (pl) on the \emph{dactyl} (d) rapidly enters the \emph{socket} (s), ejecting a high-velocity jet of water.
-A time near the bubble collape is shown on the right.
+A time near the bubble collapse is shown on the right.
 The light emission is too dim to be seen by the naked eye.
@@
 SL comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonoluminescence.
-The distinction is self-explanatory: multi-bubble sonolumiscence (MBSL) consists of  ``the simultaneous creation and destruction of many seperate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
+The distinction is self-explanatory: multi-bubble sonoluminescence (MBSL) consists of  ``the simultaneous creation and destruction of many separate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
 The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
 In SBSL, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volumetric-light emission from the bubble cloud, etc.
-Similiarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls.
+Similarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls.
 As a result, the theory is greatly simplified compared to MBSL.
 It is not an exaggeration to say that nearly all theoretical and experimental progress to explain SL has been made using SBSL, with some authors even calling it ``the hydrogen atom of SL" \cite{lohse2018bubble,crum1994sonoluminescence}.
+
+\begin{figure}
+\includegraphics[width=0.6\linewidth]{figs/pulses}
+    \caption{\cite{brenner2002single,crum1994sonoluminescence}}
+\label{fig:pulses}
+\end{figure}
 
@@
 
-The goal of this paper is to explain the basic ingredients needed to understand SBSL.
-While the author of this paper is particularly excited about shrimpoluminescence, this work will focus on SBSL in general as it is apparently simpler to create and characterize SBSL in the laboratory without involving shrimp (for example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}).
-Besides notable exceptions \cite{tang2019bioinspired}, practically all work to study SL has not involved shrimp.
-Also absent will be any detailed discussion of MBSL.
-Instead, we will summarize the history of SBSL research up to now starting at the discovery of MBSL, leading to the subsequent discovery of SBSL and a rush of effort to explain it.
-We will then present a sketch of the relevant physics before visiting each topic in detail, finally combining everything together to give a complete picture of how shrimpoluminsence occurs.
+Let us now sketch the SBSL process a little more precisely.
+An air bubble is trapped at the center of an acoustically driven flask containing a fluid.
+The pressure wave in the fluid, shown as a function of time in Fig.
+\ref{fig:pulses}, drives the expansion and contraction of the bubble and, in turn, the dynamics of the gas in the bubble responds to the bubble's oscillations.
+Within the phase-space relevant to SBSL, we get the very complicated non-linear response of the bubble's radius as shown in Fig.
+\ref{fig:pulses}.
+Let's see how these dynamics lead to SBSL.
 
-\begin{figure}
-\includegraphics[width=0.95\linewidth]{figs/flask.pdf}
-    \caption{(a) Schematic of a spherical acoustic levitation cell.
-The piezoelectirc transducers that drive the bubble trapped in the flask are labeled in the diagram.
-(b) Photograph of the same.
-Here, the fluid is aqueous H$_2$SO$_4$ instead of water.
-A sonoluminescing bubble is clearly visible in the center of the flask.
-The photo was taken in a fully lit room with exposure time 2 s.
-(c) Schematic of a Mie scattering experimental setup.
-A similar acoustic levitation cell as (a) is shown with an added laser source and photomultiplier tubes (PMT) to analyze the scattered and emitted light.
-(a), (b), and (c) are from refs.
-\cite{brenner2002single}, \cite{suslick2008inside}, and \cite{gompf2000mie} respectively.}
-\label{fig:flask}
-\end{figure}
+First, as the pressure decreases and becomes negative, the bubble expands gradually: surface tension and the negative pressure in the bubble tend to slow this down.
+Eventually the pressure is increasing again.
+Once it becomes positive, the bubble violently collapses to its minimum radius: during compression, the negative pressure in the bubble is very large and works with the driving pressure to rapidly accelerate the bubble's wall.
+Since the collapse is so quick, the bubble's contents are adiabatically compressed, resulting in very large temperature and pressure in the bubble.
+If everything is just right, the temperature and pressure are large enough for a burst of light to be produced as shown by the bottom curve in Fig.
+\ref{fig:pulses}.
+After the collapse, there is a period of \emph{after-bouncing} that continues until the pressure is negative again.
+
+It turns out that the light pulse is very short (only $\sim0.003\%$ of the bubble's cycle \cite{suslick2008inside}) and its spectrum depends sensitively on both the mechanical and chemical properties of the host-liquid and gas in the bubble, as well as the driving pressure and frequency \cite{brenner2002single,suslick2008inside}.
+The mechanical properties determine the pressure and temperature inside the bubble which in turn control chemical reactions: all of these data are used to explain the light emission.
+For extreme enough conditions in the bubble, which occur only briefly, molecular dissociation and recombination, atomic/excited-state transitions, and electron-ion/electron-neutral atom bremsstrahlung are possible \cite{an2009diagnosing,an2008spectral,an2006mechanism,flannigan2005plasma,suslick2008inside,flannigan2006measurement}.
+These are now known to be the processes which emit light \cite{lohse2018bubble,yasui2018acoustic}.
+
+Our goal in this paper is to present the simplest theory needed to understand \emph{shrimpoluminescence}, though we will actually concentrate on the related physics of a bubble undergoing SBSL in the lab as it is apparently simpler to create and characterize SBSL without involving shrimp \footnote{ For example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}.
+Besides notable exceptions \cite{tang2019bioinspired}, practically all work to study SL has not involved shrimp.}.
+Firstly, we will look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
+Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
+The result of this will turn out to be an estimate of the \emph{temperature} and \emph{pressure} in the bubble.
+We will then look at theories presenting the origin of light from the very hot contents of the bubble.
+For relatively mild conditions inside the bubble, most light emission occurs during molecular dissociation processes \cite{}.
+These are the conditions in e.g.
+weakly driven 85\% H$_2$SO$_4$.
+In water, the pressure in the bubble is considerably higher and most light emission is due the electron-ion and electron-neutral atom bremsstrahlung which occur at higher temperatures \cite{flannigan2005plasma,an2009diagnosing,suslick2008inside}.
+Recalling that we want to understand SBSL in water, we will mainly focus on these latter processes.
+Moreover, throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
+Connections to more advanced treatments, and their implications, are provided for completeness.
 
@@
 
+\begin{figure}
+\includegraphics[width=0.95\linewidth]{figs/flask.pdf}
+    \caption{(a) Schematic of a spherical acoustic levitation cell.
+The piezoelectric transducers that drive the bubble trapped in the flask are labeled in the diagram.
+(b) Photograph of the same.
+Here, the fluid is aqueous H$_2$SO$_4$ instead of water.
+A sonoluminescing bubble is clearly visible in the center of the flask.
+The photo was taken in a fully lit room with exposure time 2 s.
+(c) Schematic of a Mie scattering experimental setup.
+A similar acoustic levitation cell as (a) is shown with an added laser source and photomultiplier tubes (PMT) to analyze the scattered and emitted light.
+(a), (b), and (c) are from refs.
+\cite{brenner2002single}, \cite{suslick2008inside}, and \cite{gompf2000mie} respectively.}
+\label{fig:flask}
+\end{figure}
+
 This result was not particularly surprising to the community since it had been known for a while that cavitating bubbles, e.g.
 those generated by the sound field, could do tremendous damage to ship's propellers \footnote{A more modern example could be my van's oil pump, which is in need of replacing.}.
-The discovery's impact is consicesly summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
+The discovery's impact is concisely summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
 In fact, cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
@@
 Cavitation was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
-Shortly after, Lord Rayliegh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigoruous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
+Shortly after, Lord Rayleigh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigorous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
 Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
 The theory has been refined considerably since then \cite{prosperetti1999old,plesset1977bubble,plesset1977bubble,brenner2002single,lofstedt1995sonoluminescing,barber1992resolving} and we will devote a large portion of this paper to this subject.
-Still, very little information was accessible about the light emission until the 1990's when stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing caviation.
-A stable bubble persists through multiple cycles of cavitation, oscillating nonlinearly around an equilibrium size.
-Transient bubbles appear and then collapse to dissapear}, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
+Still, very little information was accessible about the light emission until the 1990's when stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing cavitation.
+A stable bubble persists through multiple cycles of cavitation, oscillating non-linearly around an equilibrium size.
+Transient bubbles appear and then collapse to disappear}, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
 With this discovery, serious interest took root.
 
-SBSL, unlike previous studies on MBSL, allowed very precise control and measurment of the SL process.
-With unprecidented experimental control, many discrepencies in previous assumptions about SL were discovered.
+SBSL, unlike previous studies on MBSL, allowed very precise control and measurement of the SL process.
+With unprecedented experimental control, many discrepancies in previous assumptions about SL were discovered.
 Experiments measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
@@
 New models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubbles $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
-At the same time, experimentalists fit the bubble's light emission spectra as a \emph{black body emitter} and concluded that the temperature in the bubble was atleast 25,000 K \cite{hiller1992spectrum}.
+At the same time, experimentalists fit the bubble's light emission spectra as a \emph{black body emitter} and concluded that the temperature in the bubble was at least 25,000 K \cite{hiller1992spectrum}.
 
-These very high estimates for the temperature in the bubbles led to a spark of interest more broadly: if the temerature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
+These very high estimates for the temperature in the bubbles led to a spark of interest more broadly: if the temperature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
 An action movie featuring Keanu Reeves was even made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
@@
 Shortly after this fiasco, evidenced emerged that there are shape instabilities in the bubbles dynamics indicating that the converging shock-wave hypothesis is probably not right.
-This severely reduced the estimates for the temperature to $\sim 10,000$ K, ruling out the possibility of fusion occuring at all \cite{}.
-Moreover, it was shown that the light is most likely emitted from the \emph{bulk} of the bubble as opposed to its surface, making fits to a black body specturm (i.e.
+This severely reduced the estimates for the temperature to $\sim 10,000$ K, ruling out the possibility of fusion occurring at all \cite{}.
+Moreover, it was shown that the light is most likely emitted from the \emph{bulk} of the bubble as opposed to its surface, making fits to a black body spectrum (i.e.
 surface emission) incorrect \cite{}.
-Current estimates for the temperature at the center of a sonoluminescing bubble are $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
+Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble at $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
 
-\subsection{Overview of Single-Bubble Sonoluminescence}
+\section{Bubble Dynamics}
+In this section, we will present the theory needed to model and understand the dynamics of a bubble undergoing SBSL.
+We will split this into parts.
+Firstly, we will derive the equations of motion for the bubble's wall.
+Nextly, we will review some important experimental developments that we will then use to argue for a simple model for the dynamics of the gas in the bubble.
+Lastly, we will compare results of a simulated bubble to experimental data.
+The connection to experiment will be what allows us to calculate the temperature in real bubbles.
 
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{figs/pulses}
-    \caption{\cite{brenner2002single,crum1994sonoluminescence}}
-\label{fig:pulses}
-\end{figure}
-
-A sketch of SBSL is as follows: an air bubble is trapped at the center of an acoustically driven flask containing a fluid.
-The pressure wave in the fluid drives the expansion and contraction of a bubble and, in turn, the dyanmics of the gas in the bubble respond to the bubble's oscillations.
-The result is a very complicated, non-linear behavior that we will spend a good deal of this paper explaining.
-We will see that the dynamics of the gas \emph{inside} the bubble is nearly decoupled from the dyanmics of the \emph{fluid} containing the bubble so that we may look at these two problems seperately.
-What is now known is that the conditions required for a cavitating bubble to emit light depend sensitively on both the mechanical and chemical properties of the host-liquid and gas in the bubble, as well as the driving pressure and frequency \cite{brenner2002single,suslick2008inside}.
-The mechanical properties determine the pressure and temperature inside the bubble which in turn control chemical reactions: all of these data are used to explain the light emission.
-For extreme enough conditions in the bubble molecular dissociation, atomic/excited-state transitions, and electron-ion/electron-neutral atom bremsstrahlung all occur \cite{}.
-These are the processes which emit light \cite{}.
-
-In very viscous fluids, such as 85\% aqueous H$_2$SO$_4$, the pressure and temperature are lower and most light emission occurs from molecular dissociation and ...
-\cite{}.
-In water, the pressure in the bubble is considerably higher and most light emission is due the electron-ion and electron-neutral atom bremsstrahlung \cite{}.
-Since our goal in this paper is to explain \emph{shrimpoluminescence}, which occurs in water, we will mainly focus on the later emission processes.
-
-Firstly, we will look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
-Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
-After that, we will combine these two and discuss solutions of the full problem.
-The ultimate result of this will turn out to be an estimate of the \emph{temperature} and \emph{pressure} in the bubble.
-We will then look at theories presenting the origin of light from the very hot contents of the bubble.
-Throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
-Connections to more advanced treatments, and thier implications, are provided for completeness.
-\section{Bubble Dyanmics}
+\subsection{The Bubble Wall}
 As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
-This isn't particularly surprising since the light emission and bubble collapse occur at different timescales (the duration of maximum compressions is $\sim 0.003\%$ of the bubble's period) \cite{suslick2008inside,prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
+This isn't particularly surprising since the light emission and bubble collapse occur at different timescales (the duration of maximum compression's is $\sim 0.003\%$ of the bubble's period) \cite{suslick2008inside,prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
 The starting point for any theory of the bubble dynamics is Rayleigh's original work \cite{rayleigh1917pressure}, now expounded upon by many others.
@@
 However, for brevity's sake, we will only derive the RP equation here and merely quote the others as they are very similar in form to the RP equation.
-The field of \emph{bubble dynamics} is quite mature and attempting to review it here would be out of scope of this paper; instead, the reader is referred to numerous books and reviews \cite{}.
+The field of \emph{bubble dynamics} is quite mature and attempting to review it here would be out of scope of this paper; instead, the reader is referred to numerous books and reviews \cite{prosperetti1999old,brenner2002single,yasui2018acoustic,prosperetti1986bubble}.
 
@@
 gravity, electrostatic forces) which are usually from external sources and will be assumed negligible from here on.
-The subscrits on $\partial$ denote what deriviative is being taken and $\nabla$ and $\nabla^2$ have thier usual meanings.
+The subscripts on $\partial$ denote what derivative is being taken and $\nabla$ and $\nabla^2$ have their usual meanings.
 The first equation is \emph{conservation of momentum} for the fluid, i.e.
-Newton's $2^{nd}$ law; the second equation is \emph{mass convervation}.
+Newton's $2^{nd}$ law; the second equation is \emph{mass conservation}.
 The total-derivative of a moving fluid $\frac{d}{dt}\equiv D_t$ is sometimes called the \emph{material derivative}.
@@
 To make progress, we will make some drastic approximations that are more-or-less valid for our problem.
-Firslty, we will assume irrotational flow (i.e.
+Firstly, we will assume irrational flow (i.e.
 $\nabla\times\bm{u}\equiv 0$) so that there is only radial motion in the liquid (i.e.
@@
 
-Nextly, we assume that the viscous terms are negligible in the bulk dyanimcs of the liquid.
+Nextly, we assume that the viscous terms are negligible in the bulk dynamics of the liquid.
 For (relatively) low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation and is widely used \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
-We will acount, to some extent, for viscosity later when looking at the bubble-liquid interface.
+We will account, to some extent, for viscosity later when looking at the bubble-liquid interface.
 Moreover, we are assuming that the flow is \emph{isentropic}, i.e.
@@
 This is relevant to us since we are claiming the bubble wall moves so quickly that it is compressed to its maximum pressure before any heat can flow out.
-On the otherhand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
+On the other hand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
 This is, e.g.
-the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its groundstate at all times.})
+the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its ground state at all times.})
 
@@
 \end{equation}
-Eq. \ref{eq:press_vel} gives the pressure at cordinate $(r,t)$ in terms of the velocity of the fluid, the density, and the appied pressure.
+Eq. \ref{eq:press_vel} gives the pressure at coordinate $(r,t)$ in terms of the velocity of the fluid, the density, and the applied pressure.
 
@@
 Eliminating $h$ between the two and dropping small terms $\sim \frac{u}{c}$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.} with $c$ the speed of sound in the fluid.
-Again, we use incompressibilty by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$ \footnote{For our problem, it is true that $\left( \frac{1}{c^2} \partial_t \phi \right) / \nabla^2 \phi \sim \left( \frac{R}{\lambda} \right)^2 \ll 1$, i.e.
+Again, we use incompressibility by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$ \footnote{For our problem, it is true that $\left( \frac{1}{c^2} \partial_t \phi \right) / \nabla^2 \phi \sim \left( \frac{R}{\lambda} \right)^2 \ll 1$, i.e.
 the long wave-length approximation is accurate \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}.} .
@@
 The right hand side is the change in enthalpy (density) due to the fluctuating bubble, i.e.
-the dyanmics of the bubble given on the left hand side are determined by the enthalpy in the fluid due to the bubble's motion.
-In deriving this equation, we have assumed that the speed of propogation in the fluid is infinite.
+the dynamics of the bubble given on the left hand side are determined by the enthalpy in the fluid due to the bubble's motion.
+In deriving this equation, we have assumed that the speed of propagation in the fluid is infinite.
 In reality, this is not true but this approximation will be very good close to the bubble (in the \emph{near-field}) where retardation effects are negligible anyway.
@@
 Since the motion in the fluid is purely radial, we expect the only relevant stresses to be normal to the bubble wall.
-Let's denote the radial stress in the liquid at some point $r$ by $s_r(r)=-p(r,t)+2\mu \partial_r u$.
-Recall, $\mu$ is the fluid viscosity from earlier.
+Let's denote the radial stress in the liquid at some point $r$ by $s_r(r)=-p(r,t)+2\eta \partial_r u$.
+Recall, $\eta$ is the fluid viscosity from earlier.
 Using conservation of mass and momentum, and our expression above for the velocity, we can equate the normal stresses in the fluid to those in the bubble:
@@
 \begin{split}
-    s_r(R)=-p_B-4\frac{\mu \dot{R}}{R}=-p_g+2\frac{\sigma}{R}
+    s_r(R)=-p_B-4\frac{\eta \dot{R}}{R}=-p_g+2\frac{\sigma}{R}
 \end{split}
@@
 \begin{equation}
-    p_B(t)=p_g(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)
+    p_B(t)=p_g(t)-\frac{1}{R}\left( 2\sigma+4\eta \dot{R} \right)
     \label{eq:p_B}.
@@
 \begin{equation}
-    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) \right]
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\eta \dot{R} \right)-P(t) \right]
     \label{eq:RP_2}
@@
 \ref{eq:RP_2} is the equation of motion for a driven spherical bubble in an infinite, incompressible fluid.
-It is worth breifly summarizing this all again.
+It is worth briefly summarizing this all again.
 On the left hand side, $R$ is the radius of the bubble, $\dot{R}$ is the velocity of the bubble's radius i.e.
@@
 $p_g(t)$ is evaluated at the bubble wall.
-In the second term, the $\sim \sigma$ part gives the restoring force from surface tension stresses while the $\sim \mu$ term is due to the fluid viscos stresses: $\sigma$ and $\mu$ are the surface tension and viscosity coefficients respectively.
+In the second term, the $\sim \sigma$ part gives the restoring force from surface tension stresses while the $\sim \eta$ term is due to the fluid viscous stresses: $\sigma$ and $\eta$ are the surface tension and viscosity coefficients respectively.
 The last term on the right, $P(t)$, is the \emph{external pressure}.
@@
 
-Let us breifly digress to discuss retardation effects.
+Let us briefly digress to discuss retardation effects.
 If we had not assumed incompressibility, our solution would include a term arising from the sound radiation from the bubble itself.
@@
 For now, let's see what we else we can understand from this analytically by making some more approximations.
-If we take $p_B-(p_\infty+P) \equiv p_0$ with $p_0$ constant, we recover Rayleigh's equation discussed above \cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
-This is valid during the part of the bubble's cycle called \emph{The Rayleigh collapse}: when the bubble's wall is very fast, the $\dot{R}^2$ term will be much larger than the right hand side of eq.
+If we look at the interval where the bubble wall's velocity is very large, called \emph{The Rayleigh Collapse}, the $\dot{R}^2$ term will be much larger than the right hand side of eq.
 \ref{eq:RP_2} and we may neglect it.
+\cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
 Rayleigh looked at the same problem, i.e.
@@
 \end{equation}
-which has a solution that diverges as $R(t)\sim\left[(t_*-t)/t_* \right]^{2/5}$ \cite{brenner2002single}.
+and found the solution to be $R(t)\sim\left(t_*-t\right)^{2/5}$ \cite{brenner2002single}, with $t_*$ the time of collapse.
+The \emph{velocity}, $\dot{R}\sim\left(t_*-t \right)^{-3/5}$ diverges as the bubble collapses.
 This observation is what lead to Rayleigh's argument for the origin of cavitation damage to ship's propellers.
-Yasui gave a rather elegant heurestic explanation for the origin of \emph{caviation} in a fluid that is worth stating here \cite{yasui2018acoustic}.
-Condider the mass flowing through contentric shells around a bubble, $4\pi R_i^2 v_i \rho$.
+Yasui gave a rather elegant heuristic explanation for the origin of \emph{cavitation} in a fluid that is worth stating here \cite{yasui2018acoustic}.
+Consider the mass flowing through concentric shells around a bubble, $4\pi R_i^2 v_i \rho$.
 For a source free field, we require that neighboring shells satisfy conservation of mass: $4\pi R_1^2 v_1 = 4\pi R_2^2 v_2$.
@@
 In the full RP equation (eq.
-\ref{eq:RP_1}) the only term capable of compensating the bubble wall's motion is the divering pressure inside the bubble itself \cite{brenner2002single}.
+\ref{eq:RP_1}) the only term capable of compensating the bubble wall's motion is the diverging pressure inside the bubble itself \cite{brenner2002single}.
 Applying the same analysis to the family of solutions derived when including the bubble's sound radiation, it turns out that the most important term for slowing the bubble wall's diverging velocity is $\propto\dot{p}_g$.
@@
 \begin{equation}
-    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\mu \dot{R} \right)-P(t) +\frac{R}{c} \dot{p}_g \right].
+    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\eta \dot{R} \right)-P(t) +\frac{R}{c} \dot{p}_g \right].
     \label{eq:RP_3}
@@
 This equation is only slightly more complicated than the RP equation (eq.
-\ref{eq:RP_2}), the only chage being the added term $\propto \dot{p}_g$ which is determined from solving the decoupled problem of the gas inside the bubble.
+\ref{eq:RP_2}), the only change being the added term $\propto \dot{p}_g$ which is determined from solving the decoupled problem of the gas inside the bubble.
 Eq.
 \ref{eq:RP_3} is usually solved numerically; we will look at results of numerical solutions later when talking about experiments.
-For now, we note that the observed error between this equation and experiment is only significant in the interval during bubble collapse and even still, results of solving this equation are quantitavely in good agreement with measurements of the bubble's radius \cite{brenner2002single}.
+For now, we note that the observed error between this equation and experiment is only significant in the interval during bubble collapse and even still, results of solving this equation are quantitatively in good agreement with measurements of the bubble's radius \cite{brenner2002single}.
 
-\section{The Bubble's Interior}
-
-The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (from which we may calculate the temperature) are used to try to reproduce the easily measuremable dynamics of the bubble.
+\subsection{The Bubble's Interior}
+The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (from which we may calculate the temperature) are used to try to reproduce the easily measurable dynamics of the bubble.
 (ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{}.
@@
 $10^8$ K \cite{}.
-It was later discovered that SBSL is volume emission, indicating that black-body radiation is not valid \cite{}.
+It was later discovered that SBSL is volume emission, indicating that black-body radiation is not valid \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
 Little progress was made on this front for quite some time \cite{brenner2002single}.
@@
 
-SBSL was first measured in partially degassed water \cite{gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single} \footnote{Degassing lowers the ambient pressure, making it easier for bubbles to form.} Early attempts to understand the physics required exploring the phase-space over which SBSL occurs.
-Some attempts were made using non-aqueous host liquids e.g.
-alcohols, silicone oils \cite{weninger1995sonoluminescence,barber1997defining}, but experiments with air bubbles were not very sucessful \cite{barber1997defining}.
-Instead, different gases were tried in water.
+Some attempts were made to measure SBSL using non-aqueous host liquids e.g.
+alcohols, silicone oils \cite{weninger1995sonoluminescence,barber1997defining}, but experiments with air bubbles were not very successful \cite{barber1997defining}.
+Eventually, different gases were tried in water.
 Recall, air is $\sim 80\%$ N$_2$, $\sim 20\%$ O$_2$, and $\sim 1\% $ Ar.
@@
 These species then react with the water vapor radicals to form species that are soluble in the host fluid.
-As the pressure becomes very large during the compression stage of SBSL, the soluble materials leave the bubble and do not re-enter since thier solubility in water is enormous compared to the Ar content of the bubble \cite{lohse1997sonoluminescing}.
-Over many cycles, the contents of the bubble become nearly pure argon.
+As the pressure becomes very large during the compression stage of SBSL, the soluble materials leave the bubble and do not re-enter since their solubility in water is enormous compared to the Ar content of the bubble \cite{lohse1997sonoluminescing}.
+Over many cycles, the contents of an air bubble in water become nearly pure argon.
 Moreover, it was hypothesized that the SL would much more intense when the contents of the bubble are a pure inert gas: if the contents were e.g.
@@
 
-Most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurments of the light spectum \cite{}.
-More importantly, new measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressures than water.
+Still, most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurements of the light spectrum \cite{}.
+More importantly, new measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressure than water.
 Careful experiments revealed that the spectrum depended critically on the noble gas content and driving pressure.
 Suslick et.
-al measured the different emission spectra from Ar, Xe, and Kr bubbles in aqueous H$_2$SO$_4$ \cite{flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
+al measured the different emission spectra from Ar, Xe, and Kr bubbles in aqueous H$_2$SO$_4$, driven to have similar conditions to those in an air bubble in water \cite{flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
 Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
 Perhaps equally as crucially, they found the \emph{absence} of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas, experimentally confirming the argon rectification hypothesis.
-Moreover, the spectra at different pressures were in very good agreement with detailed calcutions of emission from noble gas plasmas \cite{an2009diagnosing,an2008spectral}.
+Moreover, the spectra at different pressures were in very good agreement with detailed calculations of emission from noble gas plasmas \cite{an2009diagnosing,an2008spectral}.
 Increased pressure results in increased collision rate between particles, broadening the excited state transition spectral lines \cite{an2008spectral,suslick2008inside,flannigan2005plasma,flannigan2006measurement}.
@@
 With this in mind, we choose only to focus on results known to be relevant to the dynamics of a bubble undergoing SBSL.
-Excellent books and modern reviews on modelling the gas dynamics exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.
+Excellent books and modern reviews on modeling the gas dynamics exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.
 We will hold off on discussing the dynamics in the plasma until a later section.
@@
 \subsection{Gas Dynamics}
-The most straighforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
+The most straightforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
 In fact, serious quantitative predictions of the conditions of the bubble's interior take this path \cite{an2006mechanism,an2008spectral,an2009diagnosing,flannigan2005plasma,flannigan2006measurement}.
 With simplifying assumptions, equations of motion with varying degrees of sophistication can be derived for the gas and the system of gas dynamical equations and the RP equations are solved numerically.
-The discussion proceeds in much the same way as deriving the RP equations above but is considerally more complicated \cite{brenner2002single,yasui2018acoustic}, so we will omit it here.
+The discussion proceeds in much the same way as deriving the RP equations above but is considerably more complicated \cite{brenner2002single,yasui2018acoustic}, so we will omit it here.
 An important aspect of these studies is the understanding of why relatively simple models work quite well.
@@
 Various other models fully including dissipation and heat transport into/out of the bubble showed that the bubble's compression is, to a good approximation, adiabatic and quasistatic \cite{}.
-Before we turn to discuss a model for the gas dynamics, let us pause to mention that another theoretically appealing method for modelling the gas dynamics is \emph{molecular dynamics} \cite{}.
+Before we turn to discuss a model for the gas dynamics, let us pause to mention that another theoretically appealing method for modeling the gas dynamics is \emph{molecular dynamics} \cite{}.
 While this method is attractive for a number of reasons, practical calculations containing realistic numbers of particles aren't possible, limiting the applicability \cite{brenner2002single}.
 
-A simple but reasonable model for the gas dyanmics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.
+A simple but reasonable model for the gas dynamics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.
 The modifications to the result for an ideal gas are simple \cite{sivasubramanian2002temperature}.}.
@@
     C_P &= C_V+N k_B \\
-    \frac{C_P}{C_V} &= 1+\frac{f}{2} \equiv \Gamma 
+    \frac{C_P}{C_V} &= 1+\frac{2}{f} \equiv \gamma 
 \end{split}
 \end{equation}
-with $C_V$ and $C_P$ the heat capcities at constant volume and pressure respectively.
-For a monatomic ideal gas, $f=3$ and $\Gamma=5/2$.
+with $C_V$ and $C_P$ the heat capacities at constant volume and pressure respectively.
+For a monatomic ideal gas, $f=3$ and $\gamma=5/3$.
 Now we can use the second law of thermodynamics to write $C_V dT = -P dV$.
@@
 \begin{equation}
-    \frac{dP}{P}=-\left(\frac{C_P}{C_V}\right) \frac{dV}{V} = -\Gamma \frac{dV}{V}.
+    \frac{dP}{P}=-\left(\frac{C_P}{C_V}\right) \frac{dV}{V} = -\gamma \frac{dV}{V}.
     \label{eq:dP}
@@
 \begin{equation}
-    P_2 = P_1 \left(\frac{V_1}{V_2}\right)^\Gamma
+    P_2 = P_1 \left(\frac{V_1}{V_2}\right)^\gamma
 \label{eq:P}
@@
 \begin{equation}
-    p_g(t) = p_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ \Gamma
+    p_g(t) = \left( p_\infty+2\frac{\sigma}{R_0} \right) \left[ \frac{R_0^3-h^3}{R^3(t)-h^3} \right] ^ \gamma
     \label{eq:p_g}
@@
 Detailed calculations starting from the full NS equations further tell us that the gas dynamics in the bubble can be regarded as \emph{quasistatic} \cite{}.
-During the expansion and afterbounce stages of the bubbles cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal}.
+During the expansion and after-bounce stages of the bubbles cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal}.
 This is true because, when the bubble's velocity is on the order of the \emph{heat diffusion} timescale, the temperature throughout the bubble is nearly equal to that in the liquid \cite{prosperetti1999old,brenner2002single,yasui2018acoustic}.
 The relevant modifications to eq.
-\ref{eq:p_g} turns out to be replacing $\Gamma\rightarrow 1$.
-A neat way of including both isothermal and adiabatic gas dynamics is continously varying the exponent, $\Gamma$ in eq.
+\ref{eq:p_g} turns out to be replacing $\gamma\rightarrow 1$.
+A neat way of including both isothermal and adiabatic gas dynamics is continuously varying the exponent, $\gamma$ in eq.
 \ref{eq:p_g}, between its adiabatic and isothermal values, $C_P/C_V$ and $1$, respectively.
@@
 \cite{brenner2002single} and \cite{prosperetti1999old}.
-It turns out that this process doesn't result in better quantitative predictions since, during the assumed isothermal part of the process, dissipation becomes imporant \cite{brenner2002single}.
-A simple alternative is choosing a constant exponent somewhere between the adiabatic $\Gamma=5/2$ and isothermal $\Gamma=1$ values \cite{hilgenfeldt1999simple}.
-It is also possible to include heat and mass diffusion in simple model forms and these do give slighlty improved quantitative agreement with experiment; however we choose to discuss solutions of the RP equations using the simpler form given in eq.
+It turns out that this process doesn't result in better quantitative predictions since, during the assumed isothermal part of the process, dissipation becomes important \cite{brenner2002single}.
+A simple alternative is choosing a constant exponent somewhere between the adiabatic $\gamma=5/3$ and isothermal $\gamma=1$ values \cite{hilgenfeldt1999simple}.
+It is also possible to include heat and mass diffusion in simple model forms and these do give slightly improved quantitative agreement with experiment; however we choose to discuss solutions of the RP equations using the simpler form given in eq.
 \ref{eq:p_g} as this will enable to understand the important features of the bubble's cycle.
@@
 
-\subsection{Experimental Results}
+Now recall that one of main goals of modeling the bubble dynamics in SBSL was to calculate the temperature inside the bubble in order to model the light emission.
+Following similar steps to deriving $p_g(t)$, we can write down an equation for the temperature $T$ \cite{barber1997defining,brenner2002single}:
+\begin{equation}
+    T(t) = T_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ {\gamma-1}.
+    \label{eq:T(t)}
+\end{equation}
+Earlier, we assumed that during the bubble's after-bounces and expansion, the gas dynamics were isothermal.
+Then we take $T_0$ to be the temperature of the liquid.
+The RP equation, eq.
+\ref{eq:RP_3}, combined with the gas dynamical equation, eq.
+\ref{eq:p_g}, and the known liquid and gas properties $\eta$, $\sigma$, and $h$, can be used to calculate the dynamics of an acoustically driven bubble, determining the time-dependence of the pressure and temperature in the bubble along the way.
 
+\subsection{Bubbles in the Lab}
 \begin{figure}
@@
 
-Creating stable, single bubbles in a laboratory experiment is not particulary challenging and can be done with standard and low-cost materials \footnote{Actually, the phase-space for SBSL is ``small".
+Creating stable, single bubbles in a laboratory experiment is not particularly challenging and can be done with standard and low-cost materials \footnote{Actually, the phase-space for SBSL is ``small".
 The coordinates being driving amplitude, driving frequency, fluid composition, temperature, ambient pressure, gas content and composition, etc...
@@
 
-Some authors have looked at corrections to the bubble's dynamics in the presence of bouyancy, which causes a net translation other than the Bjerknes force.
+Some authors have looked at corrections to the bubble's dynamics in the presence of buoyancy, which cause translational motion even for vanishing Bjerknes force.
 The result is the bubble oscillates around an equilibrium position near but away from the center of the flask \cite{matula2000single,matula1997bjerknes,matula1999inertial}.
-The translational motion of the bubble leads to shape distortion which is beleived to supress SL; this is confirmed by SBSL experiments in microgravity resulting in larger intensity light production \cite{matula2000single}.
+The translational motion of the bubble leads to shape distortion which is believed to suppress SL; this is confirmed by SBSL experiments in microgravity resulting in larger intensity light production \cite{matula2000single}.
 
@@
 
-Mie scattering measurements scatter a laser off a bubble assuming the bubble is a homogenous dielectric sphere \cite{}.
-The angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measurethe bubble's radius as a function of time \cite{}.
+Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
+Assuming the bubble is a homogeneous dielectric sphere \cite{}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{}.
 A typical Mie scattering experiment is shown in fig.
@@
 Two important sources of experimental error are changes to the dielectric properties of the bubble during its cycle \cite{} and the fact that cavitating bubbles can emit an outward going shock \cite{}: some light scatters off the outward-going spherical wave too \cite{}.
-Both effects are largest during the SL period of the bubble's cycle, leading to an important source of experimental error \cite{}.
+Both effects are largest during the SL period of the bubble's cycle, leading to an important source of error when fitting the calculated $R(t)$ curve \cite{}.
+Parameters used to calculate the gas dynamics in the bubble can then be fit to the measured radius to build a full model of the dynamics.
+In the case of eq.
+\ref{eq:p_g}, the quantity to fit is $R_0$.
+The results of a calculation using eq.
+\ref{eq:p_g} combined with the RP variant eq.
+\ref{eq:RP_3} and fit to experiment are shown in fig.
+\ref{fig:bubble_radius_1} (a).
+The ambient ambient radius is $R_0=1.45$ $\mu$m for driving pressure $P_0=1.45$ atm \cite{barber1997defining,barber1992light}.
+Note the large change in the calculated dynamics for relatively small changes in $R_0$ and $P_0$ in Fig.
+\ref{fig:bubble_radius_1} (b) and (c).
 
-Parameters used to calculate the gas dynamics in the bubble can then be fit to the radius to build a full model of the dyanmics.
-In the case of eq.
-\ref{eq:p_g}, the quantities are $P_0$ and $R_0$.
-The results of a calculation using eq.
-\ref{eq:p_g} combined with the eq.
-\ref{eq:RP_3} RP variant fit to experiment are shown in fig.
-\ref{fig:bubble_radius_1} (a).
-The ambient ambient radius and pressure were $R_0=1.45$ $\mu$m and $P_0=1.45$ atm \cite{barber1997defining,barber1992light}.
+More accurate treatments of the gas dynamics from numerically solving the gas's NS equations are qualitatively in good agreement with these simpler model calculations \cite{brenner2002single,yasui2018acoustic}.
+They are, however, in better quantitative agreement when their results are used to calculate the light spectrum \cite{an2009diagnosing,an2008spectral,an2006mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside}.
+Let us turn to discuss the light's spectrum now.
 
 \section{Let there be light!}
-The radiative processes occuring during \cite{hilgenfeldt1999simple}
+An important requirement for models of the light emission is \emph{local thermodynamic equilibrium} (LTE).
+LTE is required for many reasons, foremost being the definition of temperature.
+LTE is expected to be true in a bubble undergoing SBSL since the particle density, $\sim 10^{28}/m^3$, and temerature, $\sim 10^4$ K, guarantee collisions between particles occur so frequently that equilibrium is ensured \cite{brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism}.
+
+\emph{Any} ensemble of atoms at a well defined temperature will emit radiation according to Planck's law.
+In the case of SBSL, we will look for corrections due to interactions of the emitted photons with atoms on the way out of the bubble \cite{hilgenfeldt1999sonoluminescence}.
+Since we are assuming LTE, the intensity of emitted radiation depends instantaneoisly on the temperature-time curve, $T(t)$, that we calculated before.
+Planck's law is \cite{schroeder1999introduction}
+\begin{equation}
+    I_{B}(\lambda;T(t))=\frac{2 h c^2}{\lambda^5}\frac{1}{\exp(hc/\lambda k_B T(t))-1}
+\end{equation}
+The \emph{black-body model} assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out.
+Then the only emission that is measured comes from a thin layer on the bubbles surface.
+Calculations of SBSL assuming this over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}.
+Instead, we need to consider finite opacity of the bubble, i.e.
+some of the emitted radiation is absorbed on its way out.
+
+Let's call the absorption coefficient $\kappa(\lambda,T(t))$\footnote{Really, this is an \emph{effective} absortion coeffecient $\kappa = \kappa'(1-\exp\left[ hc/k_B T \lambda \right])$ with $\kappa'$ the actual absoption coefficient of the material.
+The effective coefficient $\kappa$ includes induced emission of photons as light passes through the medium which 'decreases' absorption.
+Physically, this can be viewed as radiation that is absorbed being immediately re-emitted or, equivalently, vice-versa \cite{zel2002physics,taylor1969experimental}.} where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$.
+Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics}
+\begin{equation}
+    I(\lambda;T(t))=I_{B}(\lambda;T(t))  \left( 1-\exp \left[ -\kappa(\lambda;T(t)) s \right] \right)
+    \label{eq:intensity}
+\end{equation}
+where have continued with the approximation that the temperature in the bubble is uniform \cite{hilgenfeldt1999sonoluminescence,hilgenfeldt1999simple}.
+In the limit that $\kappa\rightarrow\infty$, we recover the result for a black-body emitter.
+On the otherhand, for a strong absorber, $1-\exp(\kappa s)\approx\kappa s$ and $I\approx I_B \kappa s$, i.e.
+the number of photons emitted is greatly reduced, a result that we require.
+For more complicated $\kappa(\lambda;T(t))$ that may even depend on \emph{position} through $T(\bm{x},t)$ in a more accurate gas dynamical model, eq.
+\ref{eq:intensity} is integrated numerically.
+Moreover, integrating over the bubbles volume and surface, and over time, allow calculating the radiated power spectrum which may be directly compared to experiment.
+A great deal of work has been devoted to identifying the relevant absorption mechanisms, with most progress being made only after SBSL was discovered in aqueous H$_2$SO$_4$ \cite{hilgenfeldt1999simple,brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside,an2009diagnosing,an2008spectral,an2006mechanism}.
+
+What was discovered is that the radiative processes occurring during SBSL depend very sensitively on the fluid properties and driving pressure \cite{flannigan2006measurement,flannigan2005plasma}.
+Due to its different mechanical properties, H$_2$SO$_4$ can support SBSL over a larger range of driving pressures than water \cite{an2009diagnosing}.
+For low driving pressures, the temperature and pressure in the bubble are much lower than those encountered in bubbles in water and \emph{discrete} emission lines for molecular dissociation and electronic transitions are directly observable \cite{flannigan2005plasma,flannigan2006measurement}.
+At larger driving pressures leading to bubble conditions comparable to those in water, however, what was found is that the \emph{continous} processes electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination dominate \cite{hilgenfeldt1999simple,yasui2018acoustic,suslick2008inside,zel2002physics}.
+Moreover, it is now known that radiative (de)attachement of electrons and water-vapor molecules for SBSL in water is an important source of absorption, though this effect is limited by the small amount of these species in the bubble (see the argon rectification hypothesis above) \cite{an2006mechanism}.
+
+Let us discuss SBSL of a \emph{pure} Ar bubble in water.
+At the very high temperatures that occur in the bubble, a large fraction of (outer-shell) electrons are in excited states or are ionized from thier atoms completely \cite{zel2002physics}.
+Moreover, at the very high temperature and pressures in the bubble ($\sim 10^4$ K, $\sim$GPa), collisions between the particles are frequent.
+When an ionized electron passes near an atom or (not so near) an ion, it is deflected by the atom's potential.
+The deflection accelerates the electron, changing its kinetic energy.
+Since mass of the ion or atom is much greater than the mass of the electron, we can take the atom or ion to be fixed.
+Since the electron is charged (and energy is conserved), it emits radiation through \emph{bremsstrahlung} as its kinetic energy changes \cite{jackson1999classical,zel2002physics}.
+It is also possible for the electron to become \emph{bound} to the ion, emitting light to lower its kinetic energy.
+
+The \emph{Saha equation} of astrophysics allows us to estimate the fraction of particles that are ionized at a given temperature \cite{hilgenfeldt1999simple,an2006mechanism,zel2002physics}
+\begin{equation}
+    \alpha(T(t)) \approx \left( \frac{2 \pi m_e k_B T}{h^2} \right)^{3/4} \exp \left(-\frac{E_{ion}}{2 k_B T} \right)
+\end{equation}
+It turns out that at typical SBSL temperatures, only $\sim 1\%$ of the atoms are ionized and thus, since collisions are very frequent, electron-neutral atom bremsstrahlung is perhaps as important as electron-ion processes \cite{hilgenfeldt1999simple,an2006mechanism,flannigan2005plasma}.
+For Ar$^+$ and $e$, formulas for recombination and electron-ion bremsstrahlung can be found in ref.
+\cite{hilgenfeldt1999simple}.
+For electron-atom processes, most modern work follows that of Geltman \cite{geltman1973free,an2006mechanism,an2008spectral}.
+
+For electron-ion collision, the potential leading to bremsstrahlung is the long range Coulumb interaction and analytical calculations are available.
+For electron-atom collision, the potential is short range and the emission is very sensitive to the potential.
+The best route is numerical calculations with pseudopotentials \cite{geltman1973free}.
+
+Book \cite{zel2002physics} gives a very good heurestic explanation of the the absorptions processes in III.2.
+Also interesting: "Let us imagine an infinite medium in thermodynamic equilibrium at a constant temperature T.
+Under steady state conditions the radiation field will also be in equilibrium.
+Thermal radiative equilibrium is characterized by the fact that the number of photons or amount of radiant energy emitted by the medium per unit time per unit volume in a given frequency interval $d\nu$ and in a given differential solid angle $d\Omega$ is exactly equal to the number of absorbed photons or to the radiant energy absorbed by the medium in the same intervals $d\nu$ and $d\Omega$." Also "Kirchhoff's law expresses the general principle of detailed balancing applied to the emission and absorption of light.
+It allows us to calculate the emission coefficient of a substance if the absorption coefficient is known (and vice versa)."
+
+Gives the three important processes: (i) free-free transitions such as electron-ion and electron-atom \emph{bremsstrahlung}.
+(ii) bound-free transitions such as detachment/recombination of electrons and ions, and (iii) bound-bound transitions such as transitions between bound electronic states in an atom.
+Both (i) and (ii) produce continous spectra, while (iii) leads to \emph{lines} \cite{zel2002physics}.
+
+Early work ignored bound-bound transitions since lines weren't observed in spectrum.
+Argued that they would be very broad, but estimated the broadening was too small to lead to what was seen in experiment, concluding instead that these process don't occur and that the relevant species aren't present in the bubble \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
+Later work used more accurate gas dyanmical models to show which species are important \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of the spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
 
@@
 
-The \emph{total} change (in time) of mass in an arbitrary element of a fluid (whose coordinates and boundary may depend on time) is given by  $\frac{d}{dt} \int_{\Omega_t} \rho dV$ where $\Omega_t$ is the domain of the fluid element under consdiration and $\rho(\bm{x},t)$ is the local, time-dependent density of the fluid.
+The \emph{total} change (in time) of mass in an arbitrary element of a fluid (whose coordinates and boundary may depend on time) is given by  $\frac{d}{dt} \int_{\Omega_t} \rho dV$ where $\Omega_t$ is the domain of the fluid element under consideration and $\rho(\bm{x},t)$ is the local, time-dependent density of the fluid.
 Using the transport theorem of fluid dynamics \cite{mcdonough2009lectures}, Gauss's theorem, and (with only some loss of generality) taking $\Omega_t$ to be a \emph{fluid parcel} (so that the boundary of the fluid parcel, $\partial \Omega_t$, moves with the fluid's velocity, $\bm{u}(\bm{x},t)$), this becomes
@@
 We write the momentum-density as $\rho \bm{u}$ with momentum given by $\int_{\Omega_t} \rho \bm{u} dV$.
-To simiplify what follows, we focus on only the $i^{th}$ component of velocity at a time.
+To simplify what follows, we focus on only the $i^{th}$ component of velocity at a time.
 Then Newton's second law can be written as
@@
 We now turn to the forces.
-Let us seperate the forces into \emph{body-} and \emph{surface-forces}: $\bm{F}=\int_{\Omega_t} \bm{f}_B dV+ \int_{\partial \Omega_t} \bm{f}_S dA $.
+Let us separate the forces into \emph{body-} and \emph{surface-forces}: $\bm{F}=\int_{\Omega_t} \bm{f}_B dV+ \int_{\partial \Omega_t} \bm{f}_S dA $.
 The body-force density, $\bm{f}_B$, is due to forces acting in the bulk of the liquid e.g.
@@
 We won't be concerned with these forces here.
-Instead, let us concentrate on the suface part.
+Instead, let us concentrate on the surface part.
 If we only consider an ideal fluid, the surface part can be written as $\int_{\partial \Omega_t} \bm{f}_S dA =-\int_{\partial \Omega_t} p \bm{n} dA = -\int_{\Omega_t}\nabla p dV$ with $p$ the pressure  and the minus sign being due to the sign convention: \emph{out} of the surface $\partial \Omega_t$ is positive.
@@
 In general, $\hat{\bm{\mu}}$ is a symmetric rank-4 tensor.
-However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in SL experiments.}, we find that the the viscous-stress tensor can be seperated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
+However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in SL experiments.}, we find that the the viscous-stress tensor can be separated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
 \begin{equation}
@@
 Now we would rather have this explicitly in terms of the velocity $\bm{u}$.
-To that end, we may subsitute eq.
+To that end, we may substitute eq.
 \ref{eq:viscous_stress_tensor} into the momentum-continuity equation.

### 2022-01-14 sonoluminescence/paper/bak/v4/paper.tex
sdfadsfa

--- 2022-01-13 sonoluminescence/paper/paper.tex
@@
 Since we are assuming LTE, the intensity of emitted radiation depends instantaneoisly on the temperature-time curve, $T(t)$, that we calculated before.
-Planck's law is \cite{schroeder1999introduction}
+Planck's law gives the energy radiated per unit-time and unit-volume, $I_B$, \cite{schroeder1999introduction}
 \begin{equation}
-    I_{B}(\lambda;T(t))=\frac{2 h c^2}{\lambda^5}\frac{1}{\exp(hc/\lambda k_B T(t))-1}
+    I_{B}(\lambda;T(t))=\frac{2 h c^2}{\lambda^5}\frac{1}{\exp(hc/\lambda k_B T(t))-1}.
 \end{equation}
-The \emph{black-body model} assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out.
-Then the only emission that is measured comes from a thin layer on the bubbles surface.
-Calculations of SBSL assuming this over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}.
-Instead, we need to consider finite opacity of the bubble, i.e.
-some of the emitted radiation is absorbed on its way out.
+The subscript $B$ is to remind us this is for a \emph{black-body} which assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out.
+Then the only emission that is measured comes from a thin layer on the bubble's surface.
+Calculations of SBSL assuming black-body radiation over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}.
+Instead, we need to consider finite opacity of the bubble.
 
@@
 Physically, this can be viewed as radiation that is absorbed being immediately re-emitted or, equivalently, vice-versa \cite{zel2002physics,taylor1969experimental}.} where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$.
-Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics}
+Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics,hilgenfeldt1999simple}
 \begin{equation}
@@
 the number of photons emitted is greatly reduced, a result that we require.
-For more complicated $\kappa(\lambda;T(t))$ that may even depend on \emph{position} through $T(\bm{x},t)$ in a more accurate gas dynamical model, eq.
+For more complicated $\kappa(\lambda;T(t))$ (that may even depend on \emph{position} through $T(\bm{x},t)$ in a more accurate gas dynamical model) eq.
 \ref{eq:intensity} is integrated numerically.
-Moreover, integrating over the bubbles volume and surface, and over time, allow calculating the radiated power spectrum which may be directly compared to experiment.
-A great deal of work has been devoted to identifying the relevant absorption mechanisms, with most progress being made only after SBSL was discovered in aqueous H$_2$SO$_4$ \cite{hilgenfeldt1999simple,brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside,an2009diagnosing,an2008spectral,an2006mechanism}.
+Moreover, integrating $I(\lambda,T(t))$ over the bubbles and over time allows calculating the radiated power spectrum which may be directly compared to experiment.
+A great deal of work has been devoted to identifying the relevant absorption mechanisms leading to $\kappa$, with most progress being made only after SBSL was discovered in aqueous H$_2$SO$_4$ \cite{hilgenfeldt1999simple,brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside,an2009diagnosing,an2008spectral,an2006mechanism}.
 
-What was discovered is that the radiative processes occurring during SBSL depend very sensitively on the fluid properties and driving pressure \cite{flannigan2006measurement,flannigan2005plasma}.
+What is now known is that the radiative processes occurring during SBSL depend very sensitively on the fluid properties and driving pressure \cite{flannigan2006measurement,flannigan2005plasma}.
 Due to its different mechanical properties, H$_2$SO$_4$ can support SBSL over a larger range of driving pressures than water \cite{an2009diagnosing}.
@@
 At larger driving pressures leading to bubble conditions comparable to those in water, however, what was found is that the \emph{continous} processes electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination dominate \cite{hilgenfeldt1999simple,yasui2018acoustic,suslick2008inside,zel2002physics}.
-Moreover, it is now known that radiative (de)attachement of electrons and water-vapor molecules for SBSL in water is an important source of absorption, though this effect is limited by the small amount of these species in the bubble (see the argon rectification hypothesis above) \cite{an2006mechanism}.
+Moreover, it is now known that recombination of water-vapor molecules for SBSL in water is an important source of absorption, though this effect is limited by the small amount of these species in the bubble (see the argon rectification hypothesis above) \cite{an2006mechanism}.
 
@@
 \begin{equation}
-    \alpha(T(t)) \approx \left( \frac{2 \pi m_e k_B T}{h^2} \right)^{3/4} \exp \left(-\frac{E_{ion}}{2 k_B T} \right)
+    \frac{N_+ N_e}{N} \approx \left( \frac{2 \pi m_e k_B T}{h^2} \right)^{3/2} \exp \left(-\frac{E_{ion}}{ k_B T} \right)
+    \label{eq:saha}
 \end{equation}
@@
 
-For electron-ion collision, the potential leading to bremsstrahlung is the long range Coulumb interaction and analytical calculations are available.
-For electron-atom collision, the potential is short range and the emission is very sensitive to the potential.
-The best route is numerical calculations with pseudopotentials \cite{geltman1973free}.
+For electron-ion collision, the potential leading to bremsstrahlung is the long range Coulumb interaction.
+Below, we will derive a classical formula which is the most commonly used expressions in the context of SBSL.
+For electron-atom collisions, the potential is short range and the emission is very sensitive to the potential.
+The best route is numerical calculations with pseudopotentials \cite{geltman1973free}, though classical approximations similar to the one we will derive below have been used too \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,zel2002physics}.
+Full quantum calculations of electron-ion bremsstrahlung absorption have also been done (see ref.
+for a review \cite{eckert2015aether}).
 
-Book \cite{zel2002physics} gives a very good heurestic explanation of the the absorptions processes in III.2.
-Also interesting: "Let us imagine an infinite medium in thermodynamic equilibrium at a constant temperature T.
-Under steady state conditions the radiation field will also be in equilibrium.
-Thermal radiative equilibrium is characterized by the fact that the number of photons or amount of radiant energy emitted by the medium per unit time per unit volume in a given frequency interval $d\nu$ and in a given differential solid angle $d\Omega$ is exactly equal to the number of absorbed photons or to the radiant energy absorbed by the medium in the same intervals $d\nu$ and $d\Omega$." Also "Kirchhoff's law expresses the general principle of detailed balancing applied to the emission and absorption of light.
-It allows us to calculate the emission coefficient of a substance if the absorption coefficient is known (and vice versa)."
+The instantaneous radiation energy emitted by an electron is\footnote{This is evaluated on the surface of a sphere at a very large distance from the electron; the time, $t$, should be the retarded time.
+But since we are going to look at spectral properties, we won't worry about this detail.} \cite{zel2002physics,griffiths2005introduction}
+\begin{equation}
+    S(t)=\frac{2e^2}{3c^2}\bm{a}^2(t)
+\end{equation}
+where $\bm{a}$ is the acceleration and $e$ the electron charge (we are using CGS units here).
+This can be deduced by calculating the Poynting vector and finding the terms that don't vanish at infinite distance.
+If we Fourier transform $\bm{a}(t) \rightarrow \bm{a}(\nu)$, this becomes
+\begin{equation}
+    S(\nu)=\frac{16 \pi^2 e^2}{3c^2}\bm{a}^2(\nu)
+\end{equation}
+with $\nu=c/\lambda$ the ordinary frequency.
 
-Gives the three important processes: (i) free-free transitions such as electron-ion and electron-atom \emph{bremsstrahlung}.
-(ii) bound-free transitions such as detachment/recombination of electrons and ions, and (iii) bound-bound transitions such as transitions between bound electronic states in an atom.
-Both (i) and (ii) produce continous spectra, while (iii) leads to \emph{lines} \cite{zel2002physics}.
+Now consider a parallel beam of electrons with initial velocity $v$, and impact parameter $b$, incident from infinity with constant number density $N_e$.
+Integrating over the azimuthal angle, there are $v N_e \cdot  2\pi b db$ electrons passing per unit time through a ring with area $2\pi b db$ centered on the ion.
+Integrating over all possible orbits, the energy radiated \emph{per-ion} per frequency interval $d\nu$, for unit electron flux, $v N_e=1$, is the so-called ``effective radiation" $d\epsilon_\nu$ \cite{zel2002physics}
+\begin{equation}
+    d\epsilon_\nu = d\nu \int_0^\infty S(\nu) 2 \pi b db.
+    \label{eq:effective_radiation}
+\end{equation}
+This can be integrated for the motion of electron in the field of an ion with charge $Ze$ to give \cite{zel2002physics,landau2013classical}
+\begin{equation}
+    d\epsilon_\nu = \frac{32 \pi^2 }{3 \sqrt{3}} \frac{Z^2e^6}{m^2 c^3 v^2} d\nu
+    \label{eq:effective_radiation_2}
+\end{equation}
 
-Early work ignored bound-bound transitions since lines weren't observed in spectrum.
-Argued that they would be very broad, but estimated the broadening was too small to lead to what was seen in experiment, concluding instead that these process don't occur and that the relevant species aren't present in the bubble \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
-Later work used more accurate gas dyanmical models to show which species are important \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of the spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
+Now suppose that, in a unit volume of gas, there are $N_+$ ions with charge $Ze$ and $N_e$ electrons with velocities distributed according to $f(v)dv=4\pi(m/2\pi k_B T)^{3 / 2} \exp(-mv^2 / 2 k_B T) v^2 dv$, i.e.
+the \emph{Maxwell velocity distribution}, passing through the volume.
+The energy emitted per unit-volume and unit-time in the frequency interval $\nu+d\nu$ by the electrons moving throught the field of the ions is
+\begin{equation}
+    N_+ N_e f(v') dv' v' d\epsilon_\nu(v').
+    \label{eq:emission}
+\end{equation}
+Note, we are assuming the velocity of the ions is neglibigle and that there is no recoil.
+The energy emitted by all of the electrons is obtained by integrating from $v_{min}$ to $v=\infty$, with $v_{min}$ satisfying $1/2 m_e v^2_{min}=h\nu$.
+$v_{min}$ is the minimum velocity of an electron that is able to emit a photon.
+Using eq.
+\ref{eq:effective_radiation_2} and integrating, we find the emission coefficient, $j_\nu$, due to electron-ion bremsstrahlung:
+\begin{equation}
+    j_\nu d \nu=\frac{32\pi}{3}\left( \frac{2\pi}{3k_B T m_e} \right)^{1/2} \frac{Z^2 e^6}{m_e c^3} N_+ N_e \exp\left( -\frac{h\nu}{k_B T} \right)d\nu.
+\end{equation}
+
+This is a nice result, but we want the \emph{absorption coefficient}. Following nearly identical arguments, the energy in the interval $\nu+d\nu$ \emph{absorbed} per unit-volume and unit-time is
+\begin{equation}
+    N_+ N_e \cdot I_B(\nu;T) d\nu \cdot f(v) dv \cdot a_\nu \left[ 1-\exp\left( -\frac{h\nu}{k_B T} \right)\right].
+    \label{eq:absorption}
+\end{equation}
+Here, $a_\nu$ is the absorption per-ion per-electron with velocity $v$ and the factor $1-\exp(-h\nu/k_B T)$ accounts for effective decrease in absorption due to stimulated emission \cite{zel2002physics}.
+In equilibrium, which we assume in SBSL, emission is exactly cancled by absorption.
+Then we set eq.
+\ref{eq:emission} equal to eq.
+\ref{eq:absorption} to solve for $a_\nu$.
+Integrating eq.
+\ref{eq:absorption} over $v$, we finally arrive at the spectral absorption coeffecient for a gas of electrons and ions with number densisites $N_e$ and $N_+$ respectively
+\begin{equation}
+    \kappa_\nu = \frac{4}{3} \left( \frac{2\pi}{3 m_e k_B T(t)} \right)^{1/2} \frac{Z^2 e^6}{h c m_e \nu^3} N_+ N_e .
+\end{equation}
+In terms of the wavelength, $\lambda$, and in SI units, we arrive at the equation used to calculate the SBSL spectrum in nearly all modern work \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing}
+\begin{equation}
+    \kappa_\lambda = \frac{4}{3} \left( \frac{2\pi}{3 m_e k_B T(t)} \right)^{1/2} \frac{Z^2 e^6 \lambda^3}{ (4 \pi \epsilon_0)^3 h c^4 m_e } N_+ N_e .
+    \label{eq:absorption_coeff}
+\end{equation}
+
+We are finally have a complete model for SBSL!
+An RP equation, e.g.
+eq.
+\ref{eq:RP_3}, combined with a gas-dynamical equation, e.g.
+eq.
+\ref{eq:p_g}, can be used to solve for the dynamics of a bubble undergoing SBSL.
+This allow's us to calculate the temperature as a function of time, $T(t)$, according to e.g.
+eq.
+\ref{eq:T(t)}.
+Then eq.
+\ref{eq:saha} can be inserted into \ref{eq:absorption_coeff} and, with expressions for electron-atom bremsstrahlung \cite{geltman1973free} and electron-ion recombination \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing,zel2002physics}, eq.
+\ref{eq:intensity} can be integrated to calculate the power spectrum as a function of $T(t)$ and the power spectrum can be directly compared to experiment.
+
+Early work on SBSL ignored bound-bound transitions since lines weren't observed in the SBSL spectrum in water.
+It was that they would be very broad due to the extremely high pressure in the bubble, but estimated broadening was too small to lead to what was seen in experiment and it was concluded instead that these process don't occur \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
+Later work using more accurate gas dyanmical models have shown that bound-bound transitions are important under the right conditions \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of SBSL spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
+SBSL is a solved problem!
 

### 2022-01-14 sonoluminescence/paper/bak/v5/paper.tex
sdfadsfa

--- 2022-01-14 sonoluminescence/paper/bak/v4/paper.tex
@@
 
-\documentclass[rmp,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
+\documentclass[prb,aps,11pt,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
 \usepackage{graphicx}
@@
 \usepackage{bm}
-\bibliographystyle{apsrev4-2}
-\setcitestyle{numbers,square}
+%\bibliographystyle{apsrev4-2}
+%\setcitestyle{numbers,square}
 
@@
 
-\title{On the Origin of Shrimpoluminescence}
+\title{The Mechanism of Shrimpoluminescence}
 
@@
 %\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}
-
 \date{\today}
@@
 \begin{abstract}
-shrimpity shrimp shrimp shrimp.
-abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract
+Snapping shrimp can produce bubbles that emit light when they collapse.
+The light emission is known as \emph{shrimpoluminescence} in the case of shrimp, but more generally when an acoustically driven bubble emits light, it is called \emph{sonoluminescence}.
+We would like to understand how the light is emitted.
+In this paper, we will specialize to the case of a single, stable bubble driven to produce light over many cycles.
+We will derive the bubble's equation of motion and examine the conditions in its interior.
+It turns out that, when the bubble collapses in a process known as \emph{cavitation}, the gas trapped in the bubble becomes hot enough to ionize.
+Light is emitted as black-body radiation with electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination limiting the amount of light that escapes the bubble.
 \end{abstract}
@@
 
-%\listoffigures
-%\listoftables
-%\tableofcontents
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\emph{Frequently Used Abbreviations}
-%\begin{itemize}
-%    \item SL: sonoluminescence
-%    \item SBSL: single-bubble sonoluminesence
-%    \item MBSL: multi-bubble sonoluminescence
-%\end{itemize}
-
 \section{Introduction}
 Snapping shrimp, like our cute friend in Fig.
-\ref{fig:shrimp} (left), have long been known to produce cavitating bubbles by \emph{snapping} their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
+\ref{fig:shrimp} (left), can produce cavitating bubbles by \emph{snapping} their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
 They have a strong appendage (the \emph{dactyl}) in their claw [Fig.
-\ref{fig:shrimp} (center)] that is used to create a very high-velocity jet of water.
+\ref{fig:shrimp} (center)] that is used to create a high-velocity jet of water.
 The low pressure region in the jet's wake forms a bubble [Fig.
-\ref{fig:shrimp} (right)] that, among other things, produces an exceptionally loud noise when it collapses.
+\ref{fig:shrimp} (right)] that produces a loud noise when it collapses.
 The sound (i.e.
@@
 Navy used them as ``sonar-camouflage" in the Pacific ocean during World War II \cite{versluis2000snapping}.
+The shrimp were not patriots helping the war-effort, however; they snapped for food.
+The shock-wave produced by the cavitating bubble is used to stun or kill prey \cite{versluis2000snapping}.
+If the shrimp's prey had very\footnote{Shrimpoluminescence is too weak to be seen by the naked-eye.} sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.
+
+Sonoluminescence (SL) is the process by which a ``driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission" \cite{brenner2002single}.
+SL comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonoluminescence.
+Multi-bubble sonoluminescence (MBSL) consists of  ``the simultaneous creation and destruction of many separate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
+The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
+In SBSL, light emission from a single bubble is not complicated from scattering off other bubbles and the dynamics is not perturbed by interaction with other bubbles or, due to the bubble's tiny size, by interaction with the container walls.
+The theory is greatly simplified compared to MBSL.
+It is not an exaggeration to say that nearly all theoretical and experimental progress on SL has been made using SBSL, with some authors even calling it ``the hydrogen atom of SL" \cite{lohse2018bubble,crum1994sonoluminescence}.
 
@@
 
-The shrimp were not patriots helping the war-effort, however; they snapped for food.
-The shock-wave produced by the cavitating bubble is used to stun and even kill prey \cite{versluis2000snapping}.
-If the shrimp's prey had very\footnote{We stress \emph{very}: shrimpoluminescence is too weak to be seen by the naked-eye.} sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.
-
-Sonoluminescence (SL) is more precisely defined as the process by which a ``driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission" \cite{brenner2002single}.
-SL comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonoluminescence.
-The distinction is self-explanatory: multi-bubble sonoluminescence (MBSL) consists of  ``the simultaneous creation and destruction of many separate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
-The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
-In SBSL, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volumetric-light emission from the bubble cloud, etc.
-Similarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls.
-As a result, the theory is greatly simplified compared to MBSL.
-It is not an exaggeration to say that nearly all theoretical and experimental progress to explain SL has been made using SBSL, with some authors even calling it ``the hydrogen atom of SL" \cite{lohse2018bubble,crum1994sonoluminescence}.
-
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{figs/pulses}
-    \caption{\cite{brenner2002single,crum1994sonoluminescence}}
-\label{fig:pulses}
-\end{figure}
-
-Studying SBSL, while offering advantages, also comes with complications.
+Studying SBSL also comes with complications.
 Due to the tiny volume of the bubble, directly measuring the temperature or pressure \emph{inside} the bubble is practically impossible \cite{suslick2008inside}.
@@
 We will not pursue this; our goal in this paper is to explain \emph{shrimpoluminescence} so we will mainly focus on SBSL as it occurs in \emph{water} containing \emph{air bubbles}, though occasionally we will comment on SBSL in other fluids.
+
+\begin{figure}
+\includegraphics[width=0.7\linewidth]{figs/pulses}
+    \caption{Driving pressure, bubble radius, and SL intensity as a function of time for an air bubble in water.
+Note that the width of the light pulse at bubble collapse is \emph{very} short.
+From ref.
+\cite{crum1994sonoluminescence}}
+\label{fig:pulses}
+\end{figure}
 
@@
 Once it becomes positive, the bubble violently collapses to its minimum radius: during compression, the negative pressure in the bubble is very large and works with the driving pressure to rapidly accelerate the bubble's wall.
-Since the collapse is so quick, the bubble's contents are adiabatically compressed, resulting in very large temperature and pressure in the bubble.
-If everything is just right, the temperature and pressure are large enough for a burst of light to be produced as shown by the bottom curve in Fig.
+Since the collapse is so fast, the bubble's contents are adiabatically compressed, resulting in very large temperature and pressure in the bubble.
+If everything is just right, the temperature becomes so large that a burst of light is produced as shown by the bottom curve in Fig.
 \ref{fig:pulses}.
@@
 The result of this will turn out to be an estimate of the \emph{temperature} and \emph{pressure} in the bubble.
-We will then look at theories presenting the origin of light from the very hot contents of the bubble.
-For relatively mild conditions inside the bubble, most light emission occurs during molecular dissociation processes \cite{}.
-These are the conditions in e.g.
-weakly driven 85\% H$_2$SO$_4$.
-In water, the pressure in the bubble is considerably higher and most light emission is due the electron-ion and electron-neutral atom bremsstrahlung which occur at higher temperatures \cite{flannigan2005plasma,an2009diagnosing,suslick2008inside}.
-Recalling that we want to understand SBSL in water, we will mainly focus on these latter processes.
-Moreover, throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
+We will then look at the theory that describes the origin of the light.
+Throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
 Connections to more advanced treatments, and their implications, are provided for completeness.
@@
 \subsection{Historical Overview}
-
-MBSL was discovered by accident in 1933 by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz} \footnote{Neither the paper due to Marinesco et al.
-nor the one due to Frenzel et al.
-can be found by the author; in any case he can't read French or German so having the papers wouldn't be much use.
-As such, the story of how the effect was discovered is taken from more recent sources whose authors hopefully could read French and German \cite{brenner2002single,gaitan1990experimental,crum1994sonoluminescence}} Marinesco et al.
-were trying to accelerate photo development by \emph{insonating} developing fluid.
-They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy," which they attributed to exposure to light.
-Shortly after, Frenzel et al.
-repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
 
@@
 
+MBSL was discovered by accident in 1933 by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz}.
+Marinesco et al.
+were trying to accelerate photo development by \emph{insonating} developing fluid.
+They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy," which they attributed to exposure to light.
+Shortly after, Frenzel et al.
+repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
 This result was not particularly surprising to the community since it had been known for a while that cavitating bubbles, e.g.
-those generated by the sound field, could do tremendous damage to ship's propellers \footnote{A more modern example could be my van's oil pump, which is in need of replacing.}.
+those generated by the sound field, could do tremendous damage to ship's propellers.
 The discovery's impact is concisely summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
-In fact, cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
+
+Cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
 The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures could become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence}.
 Cavitation was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
-Shortly after, Lord Rayleigh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigorous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
+Shortly after, Lord Rayleigh wrote down and solved an equation of motion for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigorous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
 Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
@@
 
-SBSL, unlike previous studies on MBSL, allowed very precise control and measurement of the SL process.
-With unprecedented experimental control, many discrepancies in previous assumptions about SL were discovered.
 Experiments measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
@@
 New models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubbles $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
-At the same time, experimentalists fit the bubble's light emission spectra as a \emph{black body emitter} and concluded that the temperature in the bubble was at least 25,000 K \cite{hiller1992spectrum}.
-
-These very high estimates for the temperature in the bubbles led to a spark of interest more broadly: if the temperature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
+The high estimates for the temperature in the bubbles led to a spark of interest more broadly: if the temperature in the bubble were really so large, then it should be possible for nuclear \emph{fusion} to occur \cite{brenner2002single}.
 An action movie featuring Keanu Reeves was even made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
@@
 Unfortunately these results turned out to be fraudulent \cite{}.
-Shortly after this fiasco, evidenced emerged that there are shape instabilities in the bubbles dynamics indicating that the converging shock-wave hypothesis is probably not right.
-This severely reduced the estimates for the temperature to $\sim 10,000$ K, ruling out the possibility of fusion occurring at all \cite{}.
-Moreover, it was shown that the light is most likely emitted from the \emph{bulk} of the bubble as opposed to its surface, making fits to a black body spectrum (i.e.
-surface emission) incorrect \cite{}.
+Eventually evidenced emerged that shape instabilities in the bubbles dynamics prohibited converging, severely reducing the estimates for the temperature to $\sim 10,000$ K and ruling out the possibility of fusion occurring at all \cite{}.
 Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble at $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
@@
 
-\begin{figure}
-\includegraphics[width=0.5\linewidth]{figs/bubble_radius_2.pdf}
-    \caption{\cite{brenner2002single}}
-\label{fig:bubble_radius_2}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=0.5\linewidth]{figs/bubble_radius_2.pdf}
+%    \caption{\cite{brenner2002single}}
+%\label{fig:bubble_radius_2}
+%\end{figure}
 
@@
 Careful measurements of the emission from SBSL in water resulted in an almost featureless spectrum \cite{}.
-Early attempts were made to fit this as black-body radiation resulting in \emph{huge} temperature estimates, e.g.
-$10^8$ K \cite{}.
+Early attempts to fit this as black-body radiation were not very sucessful \cite{hiller1992spectrum}.
 It was later discovered that SBSL is volume emission, indicating that black-body radiation is not valid \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
@@
 \subsection{Bubbles in the Lab}
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{figs/bubble_radius_1.pdf}
-    \caption{\cite{barber1997defining,barber1992light}}
-\label{fig:bubble_radius_1}
-\end{figure}
+
+%\begin{figure}
+%\includegraphics[width=0.9\linewidth]{figs/bubble_radius_1.pdf}
+%    \caption{\cite{barber1997defining,barber1992light}}
+%\label{fig:bubble_radius_1}
+%\end{figure}
 
@@
 Early work on SBSL ignored bound-bound transitions since lines weren't observed in the SBSL spectrum in water.
-It was that they would be very broad due to the extremely high pressure in the bubble, but estimated broadening was too small to lead to what was seen in experiment and it was concluded instead that these process don't occur \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
+It was known that they would be very broad due to the extremely high pressure in the bubble, but estimated broadening was too small to lead to what was seen in experiment and it was concluded instead that these process don't occur \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
 Later work using more accurate gas dyanmical models have shown that bound-bound transitions are important under the right conditions \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of SBSL spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
@@
 \appendix
-
-\section{The Navier-Stokes Equation}
-The Rayleigh-Plesset (RP) equation can be deduced from the Navier-Stokes equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
-To that end, we remind ourselves of the Navier-Stokes equations here.
-
-The \emph{total} change (in time) of mass in an arbitrary element of a fluid (whose coordinates and boundary may depend on time) is given by  $\frac{d}{dt} \int_{\Omega_t} \rho dV$ where $\Omega_t$ is the domain of the fluid element under consideration and $\rho(\bm{x},t)$ is the local, time-dependent density of the fluid.
-Using the transport theorem of fluid dynamics \cite{mcdonough2009lectures}, Gauss's theorem, and (with only some loss of generality) taking $\Omega_t$ to be a \emph{fluid parcel} (so that the boundary of the fluid parcel, $\partial \Omega_t$, moves with the fluid's velocity, $\bm{u}(\bm{x},t)$), this becomes
-\begin{equation}
-    \frac{d}{dt} \int_{\Omega_t} \rho dV = \int_{\Omega_t} \partial_t \rho dV+\int_{\partial \Omega_t} \rho \bm{u} \cdot \bm{n} dA = \int_{\Omega_t} \partial_t \rho+\nabla \cdot(\rho \bm{u})dV.
-    \label{eq:mass_continuity}
-\end{equation}
-Now to have \emph{mass-continuity}, which we will require on physical grounds, we set eq.
-\ref{eq:mass_continuity} to 0.
-Basically, we are saying no additional material is being added or removed from the fluid.
-Since the fluid parcel's boundary $\Omega_t$ is arbitrary, we find that $\partial_t \rho+\nabla\cdot(\rho \bm{u})\equiv 0$ which is true for every point in the fluid volume.
-
-Newton's second law for fluids (\emph{momentum-continuity}) is defined similarly \cite{schoeffel2014lecture,mcdonough2009lectures}.
-We write the momentum-density as $\rho \bm{u}$ with momentum given by $\int_{\Omega_t} \rho \bm{u} dV$.
-To simplify what follows, we focus on only the $i^{th}$ component of velocity at a time.
-Then Newton's second law can be written as
-\begin{equation}
-    \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \partial_t (\rho u_i) dV+\int_{\partial \Omega_t} \rho u_i ( \bm{u} \cdot \bm{n}) dA = F_i.
-\end{equation}
-Now using Gauss's theorem, the product rule, and mass-continuity we find $ \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \rho \frac{d u_i}{dt} dV = F_i $. In vector notation this reads
-\begin{equation}
-    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt} = \bm{F}
-    \label{eq:mom_continuity}
-\end{equation}
-
-We now turn to the forces.
-Let us separate the forces into \emph{body-} and \emph{surface-forces}: $\bm{F}=\int_{\Omega_t} \bm{f}_B dV+ \int_{\partial \Omega_t} \bm{f}_S dA $.
-The body-force density, $\bm{f}_B$, is due to forces acting in the bulk of the liquid e.g.
-gravity, electrostatic forces, etc.
-and is typically from sources external to the liquid.
-We won't be concerned with these forces here.
-Instead, let us concentrate on the surface part.
-If we only consider an ideal fluid, the surface part can be written as $\int_{\partial \Omega_t} \bm{f}_S dA =-\int_{\partial \Omega_t} p \bm{n} dA = -\int_{\Omega_t}\nabla p dV$ with $p$ the pressure  and the minus sign being due to the sign convention: \emph{out} of the surface $\partial \Omega_t$ is positive.
-
-If instead the fluid is real, we can't neglect viscous forces.
-We suppose that the viscous surface term can be represented by a tensor, $\hat{\bm{\tau}}$ such that $\bm{F}_v=\int_{\partial \Omega_t} \hat{\bm{\tau}}\cdot \bm{n} dA = \int_{\Omega_t} \nabla \cdot \hat{\bm{\tau}} dV$.
-(Note, here the dot product is to be understood as matrix multiplication).
-To leading order in gradients of the velocity, (and under the physical constraints that the tensor be symmetric), the viscous-stress tensor is $\hat{\bm{\tau}}=\hat{\bm{\mu}}\cdot \hat{\bm{\epsilon}}$ where $\hat{\bm{\mu}}$ is the viscosity-tensor and $\hat{\bm{\epsilon}}$ is the strain-rate tensor: $\epsilon_{ij}=\frac{1}{2}\left[ \partial_i u_j+ \partial_j u_i \right]$.
-
-In general, $\hat{\bm{\mu}}$ is a symmetric rank-4 tensor.
-However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in SL experiments.}, we find that the the viscous-stress tensor can be separated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
-\begin{equation}
-\begin{split}
-    \hat{\bm{\tau}} & = \xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)} \\
-    \hat{\bm{\epsilon}}^{(v)} & = \delta_{ij}\epsilon_{kk} \\
-    \hat{\bm{\epsilon}}^{(s)} & = \hat{\bm{\epsilon}}-\frac{1}{3}\delta_{ij}\epsilon_{kk}.
-    \label{eq:viscous_stress_tensor}
-\end{split}
-\end{equation}
-Thus, we see that the viscosity-tensor only has two free components: $\xi$, the normal viscosity and $\mu$ the shear viscosity.
-It is usual to combine the pressure (i.e.
-elastic-stresses) and the viscous-stresses into the stress-tensor $\hat{\bm{\sigma}}=-p\hat{\bm{I}}+\xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)}$.
-Combined with this, Newton's second law becomes
-\begin{equation}
-    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt}-\bm{F} = \int_{\Omega_t} \rho \frac{d \bm{u}}{dt} - \nabla \cdot \hat{\bm{\sigma}}-\bm{f}_B dV = 0
-\end{equation}
-Again, since we allow the volume of integration to be arbitrary, the integrand must vanish everywhere: $\rho \frac{d \bm{u}}{dt} - \nabla \cdot \hat{\bm{\sigma}}-\bm{f}_B=0$.
-This is the momentum-continuity equation for a fluid.
-Now we would rather have this explicitly in terms of the velocity $\bm{u}$.
-To that end, we may substitute eq.
-\ref{eq:viscous_stress_tensor} into the momentum-continuity equation.
-For e.g.
-the $i^{th}$ component we may expand out the divergence part as
-\begin{equation}
-\begin{split}
-    \left(\nabla \cdot \hat{\bm{\sigma}}\right)_i = -\partial_i p + \left[\xi +\frac{2}{3}\mu \right] \partial_i(\nabla \cdot \bm{u})+\mu \nabla^2 u_i \equiv -\partial_i p + \eta \partial_i(\nabla \cdot \bm{u})+\mu \nabla^2 u_i
-\end{split}
-\end{equation}
-where we have introduced the \emph{bulk-viscosity} $\eta=\xi+\frac{2}{3}\mu$.
-Collecting all of the components into vector form and, combined with the mass-continuity equation \ref{eq:mass_continuity}, we arrive at the \emph{Navier-Stokes} (NS) equations for a compressible fluid:
-\begin{equation}
-\begin{split}
-    \rho \frac{d \bm{u}}{dt} = \rho \left[ \partial_t \bm{u}+\left(\bm{u}\cdot \nabla \right)\bm{u} \right] & =-\nabla p + \eta \nabla \left(\nabla \cdot \bm{u} \right)+\mu \nabla^2 \bm{u} + \bm{f}_B \\
-     \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
-     \label{eq:ns_equations_2}
-\end{split}
-\end{equation}
-
-Some comments are in order.
-The total-derivative of a moving fluid $\frac{d}{dt}\equiv D_t$ is sometimes called the \emph{material derivative}.
-In the case of velocity $D_t \bm{u}$ is the \emph{acceleration} of the fluid.
-It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies \emph{in space} as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
-The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
-On the other hand, the terms $\sim \bm{u}$ are viscous terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
-The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
-dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
-The body term $\bm{f}_B$ is from external sources.
-It is worth noting that solving eqs.
-\ref{eq:ns_equations_2} is a formidable task \footnote{In fact, it's not even clear that the NS equations \emph{can} be solved even for incompressible fluids.
-In 2000, it was announced that proving smooth, sensible solutions exist earns \$1,000,000 \cite{fefferman2006existence}.
-Better hurry though...
-inflation was 6.8\% in Nov.
-2021.} and is usually done using computers.
 

### 2022-01-14 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 
-\documentclass[rmp,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
+\documentclass[prb,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
 \usepackage{graphicx}
@@
 \usepackage{bm}
-\bibliographystyle{apsrev4-2}
-\setcitestyle{numbers,square}
 
@@
 
-%\listoffigures
-%\listoftables
-%\tableofcontents
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\emph{Frequently Used Abbreviations}
-%\begin{itemize}
-%    \item SL: sonoluminescence
-%    \item SBSL: single-bubble sonoluminesence
-%    \item MBSL: multi-bubble sonoluminescence
-%\end{itemize}
-
 \section{Introduction}
-Snapping shrimp, like our cute friend in Fig.
-\ref{fig:shrimp} (left), have long been known to produce cavitating bubbles by \emph{snapping} their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
-They have a strong appendage (the \emph{dactyl}) in their claw [Fig.
-\ref{fig:shrimp} (center)] that is used to create a very high-velocity jet of water.
-The low pressure region in the jet's wake forms a bubble [Fig.
-\ref{fig:shrimp} (right)] that, among other things, produces an exceptionally loud noise when it collapses.
-The sound (i.e.
-the pressure wave) produced by even a \emph{single} shrimp's snap is detectable over a mile away \cite{everest1948acoustical}.
+Snapping shrimp, like our cute friend in fig.
+\ref{fig:shrimp}, are known to produce cavitating bubbles by snapping their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
+They have a strong appendage (the \emph{dactyl}) in thier claw (fig.
+\ref{fig:shrimp} [center]) that is used to create a high-velocity jet of water.
+The low pressure region in the jet's wake forms a bubble (fig.
+\ref{fig:shrimp} [right]) that produces an exceptionally loud noise when it collapses.
+The sound produced by even one shrimp's snap is detectable over a mile away \cite{everest1948acoustical}.
 The noise produced by groups of shrimp is so intense that the U.S.
 Navy used them as ``sonar-camouflage" in the Pacific ocean during World War II \cite{versluis2000snapping}.
+The shrimp were not patriots helping the war-effort, however; they snapped for food.
+The sound wave produced by the cavitating bubble is used to stun or kill prey \cite{versluis2000snapping}.
+If the shrimp's prey had very\footnote{Shrimpoluminescence is too weak to be seen by the naked-eye.} sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.
 
@@
 
-The shrimp were not patriots helping the war-effort, however; they snapped for food.
-The shock-wave produced by the cavitating bubble is used to stun and even kill prey \cite{versluis2000snapping}.
-If the shrimp's prey had very\footnote{We stress \emph{very}: shrimpoluminescence is too weak to be seen by the naked-eye.} sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the pistol shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.
-
-Sonoluminescence (SL) is more precisely defined as the process by which a ``driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission" \cite{brenner2002single}.
+Sonoluminescence (SL) is defined as the process by which a ``driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission" \cite{brenner2002single}.
 SL comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonoluminescence.
-The distinction is self-explanatory: multi-bubble sonoluminescence (MBSL) consists of  ``the simultaneous creation and destruction of many separate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
-The discovery of MBSL predates SBSL by $\sim$60 years but due to the more-or-less random and fleeting nature of the bubbles, SL in general wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
-In SBSL, emission from a single bubble is not complicated from multiple scattering (off other bubbles), volumetric-light emission from the bubble cloud, etc.
-Similarly, the bubble is not perturbed by interaction with other bubbles and, due to its tiny size, by interaction with the container walls.
-As a result, the theory is greatly simplified compared to MBSL.
-It is not an exaggeration to say that nearly all theoretical and experimental progress to explain SL has been made using SBSL, with some authors even calling it ``the hydrogen atom of SL" \cite{lohse2018bubble,crum1994sonoluminescence}.
+Multi-bubble sonoluminescence (MBSL) consists of  ``the simultaneous creation and destruction of many separate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
+The discovery of MBSL predates SBSL by $\sim$60 years but due to the random and fleeting nature of the bubbles, SL wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
+In SBSL, emission from a single bubble is not complicated from scattering off other bubbles.
+Similarly, the bubble is not perturbed by interaction with other bubbles or, due to its tiny extent, by interaction with the container walls.
+The theory is greatly simplified compared to MBSL.
+It is not an exaggeration to say that nearly all progress to explain SL has been made using SBSL, with some authors even calling it ``the hydrogen atom of SL" \cite{lohse2018bubble,crum1994sonoluminescence}.
+The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are too many to try list and we won't be interested in most of them.
+References can be found in Gaitan \cite{gaitan1990experimental} and Brenner et al.
+\cite{brenner2002single}.} to more exotic explanations based of quantum field theory \cite{schwinger1992casimir1,schwinger1992casimir2,schwinger1993casimir1,schwinger1993casimir2,schwinger1993casimir3,schwinger1993casimir4,schwinger1994casimir,eberlein1996sonoluminescence,liberati2000sonoluminescence}.
 
 \begin{figure}
-\includegraphics[width=0.6\linewidth]{figs/pulses}
-    \caption{\cite{brenner2002single,crum1994sonoluminescence}}
+\includegraphics[width=0.7\linewidth]{figs/pulses}
+    \caption{Driving pressure, bubble radius, and SL intensity as a function of time for an air bubble in water.
+Note that the width of the light pulse at bubble collapse is \emph{very} short.
+From ref.
+\cite{crum1994sonoluminescence}}
 \label{fig:pulses}
@@
 
-Studying SBSL, while offering advantages, also comes with complications.
-Due to the tiny volume of the bubble, directly measuring the temperature or pressure \emph{inside} the bubble is practically impossible \cite{suslick2008inside}.
-Still, the discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are far too many to try list here and we will not be interested in the vast majority of early models which are now irrelevant.
-References to these can be found in Gaitan \cite{gaitan1990experimental} and Brenner et al.
-\cite{brenner2002single}.} to more exotic explanations based of quantum field theory \cite{schwinger1993casimir,eberlein1996sonoluminescence,liberati2000sonoluminescence}.
-Over time, the \emph{phase-space} of SBSL has been refined: SBSL only occurs for a narrow range of driving pressures and for a particular combination of host-liquid and gas content in the bubble.
-The mapping of the phase-space of SBSL is a very rich subject and merits a review in itself (several exist \cite{brenner2002single,yasui2018acoustic}).
-We will not pursue this; our goal in this paper is to explain \emph{shrimpoluminescence} so we will mainly focus on SBSL as it occurs in \emph{water} containing \emph{air bubbles}, though occasionally we will comment on SBSL in other fluids.
-
-Let us now sketch the SBSL process a little more precisely.
+Let us now sketch SBSL more precisely.
 An air bubble is trapped at the center of an acoustically driven flask containing a fluid.
-The pressure wave in the fluid, shown as a function of time in Fig.
-\ref{fig:pulses}, drives the expansion and contraction of the bubble and, in turn, the dynamics of the gas in the bubble responds to the bubble's oscillations.
-Within the phase-space relevant to SBSL, we get the very complicated non-linear response of the bubble's radius as shown in Fig.
+The pressure wave in the fluid, shown as a function of time as the top curve in fig.
+\ref{fig:pulses}, drives the expansion and contraction of the bubble and, in turn, the gas in the bubble responds to the bubble's oscillations.
+Under the right circumstances, we get the very complicated non-linear response of the bubble's radius shown as the middle curve in fig.
 \ref{fig:pulses}.
-Let's see how these dynamics lead to SBSL.
+Now let's see how these dynamics lead to SBSL.
 
@@
 Once it becomes positive, the bubble violently collapses to its minimum radius: during compression, the negative pressure in the bubble is very large and works with the driving pressure to rapidly accelerate the bubble's wall.
-Since the collapse is so quick, the bubble's contents are adiabatically compressed, resulting in very large temperature and pressure in the bubble.
-If everything is just right, the temperature and pressure are large enough for a burst of light to be produced as shown by the bottom curve in Fig.
+The collapse is so fast that the bubble's contents are adiabatically compressed, resulting in a very large temperature in the bubble.
+If everything is just right, the temperature is large enough that a pulse of light is produced as shown by the bottom curve in fig.
 \ref{fig:pulses}.
@@
 It turns out that the light pulse is very short (only $\sim0.003\%$ of the bubble's cycle \cite{suslick2008inside}) and its spectrum depends sensitively on both the mechanical and chemical properties of the host-liquid and gas in the bubble, as well as the driving pressure and frequency \cite{brenner2002single,suslick2008inside}.
-The mechanical properties determine the pressure and temperature inside the bubble which in turn control chemical reactions: all of these data are used to explain the light emission.
+The mechanical properties determine the pressure and temperature in the bubble which in turn control chemical reactions: all of these data are used to explain the light emission.
 For extreme enough conditions in the bubble, which occur only briefly, molecular dissociation and recombination, atomic/excited-state transitions, and electron-ion/electron-neutral atom bremsstrahlung are possible \cite{an2009diagnosing,an2008spectral,an2006mechanism,flannigan2005plasma,suslick2008inside,flannigan2006measurement}.
@@
 
-Our goal in this paper is to present the simplest theory needed to understand \emph{shrimpoluminescence}, though we will actually concentrate on the related physics of a bubble undergoing SBSL in the lab as it is apparently simpler to create and characterize SBSL without involving shrimp \footnote{ For example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}.
+Our goal in this paper is to present the simplest theory needed to understand shrimpoluminescence, though we will specialize to the related physics of a bubble undergoing SBSL in the lab as it is apparently simpler to create and characterize SBSL without involving shrimp\footnote{For example, convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}.
 Besides notable exceptions \cite{tang2019bioinspired}, practically all work to study SL has not involved shrimp.}.
-Firstly, we will look at the problem of the fluid containing the bubble; we will derive the equations of motion for the bubble's wall.
-Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble.
-The result of this will turn out to be an estimate of the \emph{temperature} and \emph{pressure} in the bubble.
-We will then look at theories presenting the origin of light from the very hot contents of the bubble.
-For relatively mild conditions inside the bubble, most light emission occurs during molecular dissociation processes \cite{}.
-These are the conditions in e.g.
-weakly driven 85\% H$_2$SO$_4$.
-In water, the pressure in the bubble is considerably higher and most light emission is due the electron-ion and electron-neutral atom bremsstrahlung which occur at higher temperatures \cite{flannigan2005plasma,an2009diagnosing,suslick2008inside}.
-Recalling that we want to understand SBSL in water, we will mainly focus on these latter processes.
-Moreover, throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
+Firstly, we will look at the problem of the fluid containing the bubble and derive the equations of motion for the bubble's wall.
+Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble, estimate the \emph{temperature} and \emph{pressure} in the gas.
+We will then discuss theories presenting the mechanisms of light emission/absorption in the bubble which, combined with the temperature, allow use to calculate the bubbles light spectrum.
+Throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
 Connections to more advanced treatments, and their implications, are provided for completeness.
@@
 \includegraphics[width=0.95\linewidth]{figs/flask.pdf}
-    \caption{(a) Schematic of a spherical acoustic levitation cell.
+    \caption{(a) Schematic of a spherical ``acoustic levitation cell".
 The piezoelectric transducers that drive the bubble trapped in the flask are labeled in the diagram.
@@
 (c) Schematic of a Mie scattering experimental setup.
-A similar acoustic levitation cell as (a) is shown with an added laser source and photomultiplier tubes (PMT) to analyze the scattered and emitted light.
+A similar levitation cell as (a) is shown with an added laser source and photomultiplier tubes (PMT) to analyze the scattered and emitted light.
 (a), (b), and (c) are from refs.
@@
 New models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubbles $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
-At the same time, experimentalists fit the bubble's light emission spectra as a \emph{black body emitter} and concluded that the temperature in the bubble was at least 25,000 K \cite{hiller1992spectrum}.
-
 These very high estimates for the temperature in the bubbles led to a spark of interest more broadly: if the temperature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
-An action movie featuring Keanu Reeves was even made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
-In the early 2000's, several papers were even published claiming that stable nuclear fusion was possible in the lab using a method based on SBSL \cite{}.
+An action movie featuring Keanu Reeves was made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
+and in the early 2000's, several papers were published claiming that stable nuclear fusion was possible in the lab using a method based on SBSL \cite{}.
 Unfortunately these results turned out to be fraudulent \cite{}.
-Shortly after this fiasco, evidenced emerged that there are shape instabilities in the bubbles dynamics indicating that the converging shock-wave hypothesis is probably not right.
-This severely reduced the estimates for the temperature to $\sim 10,000$ K, ruling out the possibility of fusion occurring at all \cite{}.
-Moreover, it was shown that the light is most likely emitted from the \emph{bulk} of the bubble as opposed to its surface, making fits to a black body spectrum (i.e.
-surface emission) incorrect \cite{}.
-Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble at $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
+Eventually evidenced emerged that there are shape instabilities in the bubbles dynamics prohibiting converging shock-waves and severely reducing estimates for the temperature to $\sim 10,000$ K, ruling out the possibility of fusion occurring \cite{brenner2002single}.
+Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble between $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
 
@@
 Firstly, we will derive the equations of motion for the bubble's wall.
-Nextly, we will review some important experimental developments that we will then use to argue for a simple model for the dynamics of the gas in the bubble.
+Next, we will review some important experimental developments that we will then use to argue for a simple model for the dynamics of the gas in the bubble.
 Lastly, we will compare results of a simulated bubble to experimental data.
@@
 \subsection{The Bubble Wall}
-As it turns out, the dynamics of the bubble in SBSL (regardless of the light emission) are quite well described both qualitatively and \emph{quantitatively} by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
-This isn't particularly surprising since the light emission and bubble collapse occur at different timescales (the duration of maximum compression's is $\sim 0.003\%$ of the bubble's period) \cite{suslick2008inside,prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble}.
-The starting point for any theory of the bubble dynamics is Rayleigh's original work \cite{rayleigh1917pressure}, now expounded upon by many others.
+It turns out the dynamics in SBSL are quite well described both qualitatively and quantitatively by \emph{the classical theory of bubble dynamics} \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
+The starting point for the theory is Rayleigh's original work \cite{rayleigh1917pressure}, now expounded upon by many others.
 The most widely known modern work is that of Plesset \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}, resulting in the so called ``the Rayleigh-Plesset" (RP) equation.
 The derivation assumes the host liquid is incompressible, ultimately resulting in us neglecting the effects of sound radiated by the bubble.
-Corrections exist that improve upon the RP equation in various ways and we will discuss these below \cite{}.
-However, for brevity's sake, we will only derive the RP equation here and merely quote the others as they are very similar in form to the RP equation.
-The field of \emph{bubble dynamics} is quite mature and attempting to review it here would be out of scope of this paper; instead, the reader is referred to numerous books and reviews \cite{prosperetti1999old,brenner2002single,yasui2018acoustic,prosperetti1986bubble}.
+Corrections exist that improve upon the RP equation in various ways and we will discuss this below.
+However, we will only derive the RP equation here and merely quote the others as they are similar in form.
+The field of bubble dynamics is quite mature and attempting to review it here would be out of scope of this paper; instead, the reader is referred elsewhere \cite{prosperetti1999old,brenner2002single,yasui2018acoustic,prosperetti1986bubble}.
 
-The RP equation can be deduced from the \emph{Navier-Stokes} (NS) equations \footnote{The NS equations are derived in the appendix} \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
+The RP equation is deduced from the \emph{Navier-Stokes} (NS) equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
 The NS equations for a compressible fluid are
@@
 
-Here, $\bm{u}(\bm{x},t)$ is the velocity of a fluid parcel at position $\bm{x}$ at time $t$, $\rho (\bm{x},t)$ is the fluid's density, $p$ is the pressure, $\eta \equiv \left(\xi + \frac{2}{3}\mu \right)$ is the \emph{bulk} viscosity with $\mu$ the \emph{shear} viscosity, and $\bm{f}_B$ are \emph{body} forces (e.g.
+Here, $\bm{u}(\bm{x},t)$ is the velocity of a fluid parcel at position $\bm{x}$ at time $t$, $\rho (\bm{x},t)$ is the fluid's density, $p$ is the pressure, $\eta \equiv \left(\xi + 2\mu/3 \right)$ is the \emph{bulk} viscosity with $\mu$ the \emph{shear} viscosity, and $\bm{f}_B$ are body forces (e.g.
 gravity, electrostatic forces) which are usually from external sources and will be assumed negligible from here on.
 The subscripts on $\partial$ denote what derivative is being taken and $\nabla$ and $\nabla^2$ have their usual meanings.
-The first equation is \emph{conservation of momentum} for the fluid, i.e.
-Newton's $2^{nd}$ law; the second equation is \emph{mass conservation}.
-The total-derivative of a moving fluid $\frac{d}{dt}\equiv D_t$ is sometimes called the \emph{material derivative}.
-In the case of velocity $D_t \bm{u}$ is the \emph{acceleration} of the fluid.
-It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies \emph{in space} as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
-The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
+The first equation is conservation of momentum for the fluid, i.e.
+Newton's $2^{nd}$ law; the second equation is mass conservation.
+The total-derivative of a moving fluid $d/dt\equiv D_t$ is sometimes called the \emph{material derivative}.
+In the case of velocity $D_t \bm{u}$ is the acceleration of the fluid.
+It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies in space as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
+The pressure term on the right hand side is to be understood as forces from the elastic energy.
 On the other hand, the terms $\sim \bm{u}$ are viscous (damping) terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
@@
 Solving eqs.
-\ref{eq:NS_equations} is an intractable problem.
-To make progress, we will make some drastic approximations that are more-or-less valid for our problem.
+\ref{eq:NS_equations} is intractable.
+To make progress, we make some drastic approximations that are more-or-less valid for our problem.
 Firstly, we will assume irrational flow (i.e.
@@
 $\bm{u}=u\bm{r}$).
-Note that, since the flow is assumed irrotational, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
-This amounts to assuming that the bubble is always spherical which seems like a rather drastic approximation but has been validated experimentally in many cases \cite{}.
-The fact that the bubble tends to remain spherical can be understood by accounting for surface-tension at the liquid-bubble interface \cite{prosperetti1999old}.
+Since the flow is assumed irrotational, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
+This amounts to assuming that the bubble is always spherical which seems like a rather drastic approximation but is validated experimentally.
+The fact that the bubble tends to remain spherical is understood by accounting for surface-tension at the liquid-bubble interface \cite{prosperetti1999old}.
 
-Nextly, we assume that the viscous terms are negligible in the bulk dynamics of the liquid.
-For (relatively) low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation and is widely used \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
-We will account, to some extent, for viscosity later when looking at the bubble-liquid interface.
-Moreover, we are assuming that the flow is \emph{isentropic}, i.e.
-that it is reversible (no damping) and that no heat is exchanged between fluid parcels.
-We will further assume that the liquid is isothermal, i.e.
-its temperature is constant (in space).
-Then pressure $p$ is determined from an instantaneous equation-of-state $p=p(\rho,T)$.
-In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and, as we will later see, heat-transport across the liquid-bubble interfacing is usually neglected anyway (i.e.
-the bubble is compressed adiabatically in the thermodynamic sense \footnote{According to Wikipedia, \emph{adiabatic} means \emph{fast} in thermodynamics lingo.
-This is relevant to us since we are claiming the bubble wall moves so quickly that it is compressed to its maximum pressure before any heat can flow out.
-On the other hand, the mechanics lingo implies \emph{adiabatic} to mean \emph{slow}.
-This is, e.g.
-the adiabatic theorem in quantum mechanics: a perturbation acts so slowly that the system is in its ground state at all times.})
+Next, we assume that the viscous terms are negligible in the bulk dynamics of the liquid.
+For low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
+We will account for viscosity later when looking at the bubble-liquid interface.
+We will also assume \emph{isentropic} and \emph{isothermal} flow, i.e.
+no damping and no heat flow between fluid parcels.
+The pressure $p$ is then determined from an instantaneous equation-of-state $p=p(\rho,T)$.
+In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and any heat-transport across the liquid-bubble interfacing is negligible \cite{prosperetti1999old}.
 
 Lastly, related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
-Of course the bubble-liquid interface enters both systems as a boundary condition \cite{prosperetti1986bubble}.
+Of course the bubble-liquid interface enters both systems as a boundary condition.
 
 With these simplifying assumptions (and a vector identity \footnote{$\left( \nabla \cdot \bm{u} \right) \bm{u} = \frac{1}{2} \nabla (u^2) - \bm{u}\times \left( \nabla \times \bm{u} \right)$}) we rewrite eqs.
-\ref{eq:NS_equations} as a system of $1+1$ dimensional partial differential equations:
+\ref{eq:NS_equations} as
 \begin{equation}
@@
 \end{equation}
-We can integrate the first equation above using the fact that the compressibility is negligible to move the density through the integration \cite{leighton2007derivation}
+We can integrate the first equation using the fact that the compressibility is negligible \cite{leighton2007derivation}.
 \begin{equation}
-    \int_{r_\infty}^{r} \frac{\partial}{\partial r'} \left[ \frac{\partial \phi}{\partial t} +\frac{1}{2} \left( \frac{\partial \phi}{\partial r'} \right)^2 \right] dr' = - \frac{1}{\rho} \int_{r_\infty}^{r} \frac{\partial p}{\partial r'} dr'.
+    \int_{r_\infty}^{r} \frac{\partial}{\partial r'} \left[ \frac{\partial \phi}{\partial t} +\frac{1}{2} \left( \frac{\partial \phi}{\partial r'} \right)^2 \right] dr' = - \frac{1}{\rho} \int_{r_\infty}^{r} \frac{\partial p}{\partial r'} dr'
 \end{equation}
 On the left side, we take the velocity to vanish at infinity.
-On the right side, we assume that only the static and unperturbed pressures, $p_\infty$ and $P(t)$ respectively, are relevant so that \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}
+On the right side, we assume that only the static and driving pressures, $p_\infty$ and $P(t)$ respectively, are relevant so that \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}
 \begin{equation}
@@
 
-Now, let us assume that the velocity potential satisfies a wave equation $\nabla^2 \phi - \frac{1}{c^2}\partial_t^2 \phi =0 $ \footnote{This doesn't come from thin air.
+Now, let us assume that the velocity potential satisfies a wave equation $\nabla^2 \phi - (1/c^2)\partial_t^2 \phi =0 $ \footnote{This doesn't come from thin air.
 We could write the r.h.s.
 of eq.
-\ref{eq:press_vel} as the enthalpy, $h\equiv \int \frac{dp}{\rho}$, and similarly introduce $dh=\frac{dp}{\rho}$ in the first line of eq.
+\ref{eq:press_vel} as the enthalpy, $h\equiv \int dp/\rho$, and similarly introduce $dh=dp/\rho$ in the first line of eq.
 \ref{eq:NS_1D}.
-Eliminating $h$ between the two and dropping small terms $\sim \frac{u}{c}$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.} with $c$ the speed of sound in the fluid.
-Again, we use incompressibility by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$ \footnote{For our problem, it is true that $\left( \frac{1}{c^2} \partial_t \phi \right) / \nabla^2 \phi \sim \left( \frac{R}{\lambda} \right)^2 \ll 1$, i.e.
-the long wave-length approximation is accurate \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}.} .
-In this case, we can ignore retardation effects and, remembering that we have spherical symmetry, write the solution of $\nabla^2 \phi$ as
+Eliminating $h$ between the two and dropping small terms $\sim u/c$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.} with $c$ the speed of sound in the fluid.
+Again, we use incompressibility by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$ \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}.
+Then we can ignore retardation effects and write the solution of $\nabla^2 \phi$ as
 \begin{equation}
@@
 \end{equation}
-with $\psi$ the time-dependent coefficient. We now use the boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, to determine $\psi$:
+with $\psi$ the time-dependent coefficient. Now use the boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, to determine $\psi$:
 \begin{equation}
@@
 \end{equation}  
-Thus, we find $\phi(r,t) = -\frac{R^2 \dot{R}}{r}$. Plugging this into the left hand side of eq. \ref{eq:press_vel} and evaluating at the bubble wall, $r\equiv R$
+We find $\phi(r,t) = -R^2 \dot{R}/r$. Plugging this into the left hand side of eq. \ref{eq:press_vel} and evaluating at the bubble wall, $r\equiv R$
 \begin{equation}
@@
 \end{equation}
-
 Finally, sticking all this together, we arrive at the Rayleigh-Plesset equation \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,plesset1949dynamics,plesset1977bubble}
@@
 with $p_B \equiv p(r=R,t)$ the pressure in the fluid at the bubble wall.
-Evaluated beyond the bubble wall, $r > R$, eq.
-\ref{eq:RP_1} gives the pressure radiated by the fluctuating bubble.
-Some comments are in order.
-The combination $p_\infty+P(t)$ is the \emph{ambient} pressure \cite{prosperetti1999old,prosperetti1986bubble} which gives the pressure in the fluid in the absence of the bubble.
+The combination $p_\infty+P(t)$ is the ambient pressure \cite{prosperetti1999old,prosperetti1986bubble} which gives the pressure in the fluid in the absence of the bubble.
 The left hand side of eq.
@@
 This approximation will be accurate near the bubble as well since the kinetic energy in this region will dominate \cite{prosperetti1999old}.
+
+If we had not assumed incompressibility, eq.
+\ref{eq:RP_1} would include a term arising from the sound radiation from the bubble itself.
+There are practical issues with this solution however; it is third order in time and has unphysical, exponentially diverging solutions \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,lezzi1987bubble}.
+Both issues are solved by imposing boundary conditions on the $\ddot{R}$ term.
+The procedure is difficult and apparently no unique way to do it exists \cite{prosperetti1986bubble,prosperetti1988nonlinear,keller1956damping}.
+Instead, there is a whole family of solutions that can be derived; different variations of the RP equation, which are collectively referred to as ``Rayleigh-Plesset equations" are derived by neglecting different terms in the full compressible solution.
+We will discuss this again below.
 
@@
 This was valid since only combinations of viscosity and compressibility $\nabla \cdot \bm{u}$ appeared, both of which are assumed small.
-Here, however, we opt to keep the viscosity terms coupling the bubble wall to the fluid.
-With this, we can rewrite the RP equation as
+Here, however, we keep the viscosity terms coupling the bubble wall to the fluid.
+The RP equation becomes
 \begin{equation}
@@
 a driving stress.
-This will be from the transducers driving the acoustic levitation cell.
-
-Eq.
-\ref{eq:RP_2} is the equation of motion for a driven spherical bubble in an infinite, incompressible fluid.
-It is worth briefly summarizing this all again.
-On the left hand side, $R$ is the radius of the bubble, $\dot{R}$ is the velocity of the bubble's radius i.e.
-the velocity of the \emph{bubble wall}, and $\ddot{R}$ is the bubble wall's acceleration.
-On the right hand side, $\rho$ is the (constant) density of the fluid.
-$\Delta p(t) = p_g(t)-p_\infty$ is the deviation of the gas pressure in the bubble, $p_g(t)$, from the static fluid pressure at infinite distance, $p_\infty$.
-$p_g(t)$ is evaluated at the bubble wall.
-In the second term, the $\sim \sigma$ part gives the restoring force from surface tension stresses while the $\sim \eta$ term is due to the fluid viscous stresses: $\sigma$ and $\eta$ are the surface tension and viscosity coefficients respectively.
-The last term on the right, $P(t)$, is the \emph{external pressure}.
-For problems relevant to SBSL, the driving pressure, $P(t)$ in the RP equations, is an acoustic plane wave.
-Keller and coworkers showed that the modification required to the RP equations result in the replacement $P(t) \rightarrow P_0 \sin(\omega t)$ \cite{keller1980bubble}.
+In SBSL, this is from the transducers driving the flask and $P(t)$ is an acoustic plane wave.
+Keller and coworkers showed that the modification required to the RP equation is the replacement $P(t) \rightarrow P_0 \sin(\omega t)$ \cite{keller1980bubble}.
 If we linearize in the bubble's radius $R(t)\approx R_0 \left( 1+x(t) \right)$ with $x(t) \ll 1$, we can derive a forced harmonic oscillator equation for $x(t)$ \cite{brennen2014cavitation,yasui2018acoustic}.
@@
 
-\begin{figure}
-\includegraphics[width=0.5\linewidth]{figs/bubble_radius_2.pdf}
-    \caption{\cite{brenner2002single}}
-\label{fig:bubble_radius_2}
-\end{figure}
-
-Let us briefly digress to discuss retardation effects.
-If we had not assumed incompressibility, our solution would include a term arising from the sound radiation from the bubble itself.
-There are practical issues with this solution however; it is \emph{third} order in time and has unphysical, exponentially diverging solutions \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,lezzi1987bubble}.
-Both issues are solved by imposing boundary conditions on the $\ddot{R}$ term.
-The method for doing so is also challenging and apparently no unique way to do it exists \cite{prosperetti1986bubble,prosperetti1988nonlinear,keller1956damping}.
-Instead, there is a whole family of solutions that can be derived; different variations of the improved RP equation, which are collectively referred to as ``Rayleigh-Plesset equations" are derived by neglecting different terms in the full compressible solution.
-We will discuss this again below.
+%\begin{figure}
+%\includegraphics[width=0.5\linewidth]{figs/bubble_radius_2.pdf}
+%    \caption{Solutions to eq.
+\ref{eq:RP_2} for different driving pressures.
+For driving at 1 atm, the dynamics are a simple pulsating bubble.
+For larger driving pressure, e.g.
+1.2 atm, we observe strongly non-linear behavior exhibiting cavitation and afterbounces.
+For even larger driving pressure, we encounter the regime where SBSL begins to occur.
+From ref.
+\cite{brenner2002single}}
+%\label{fig:bubble_radius_2}
+%\end{figure}
 
 The RP equation (eq.
-\ref{eq:RP_2}) is pretty but its analytical solution intractable.
+\ref{eq:RP_2}) is pretty but its analytical solution is intractable.
 Direct solutions of the RP equation or its variants are done numerically using e.g.
-the Euler method \cite{yasui2018acoustic,yasui2015dynamics}.
+Euler or Runge-Kutta methods \cite{yasui2018acoustic,yasui2015dynamics}.
 It is no great challenge computationally.
 For now, let's see what we else we can understand from this analytically by making some more approximations.
-If we look at the interval where the bubble wall's velocity is very large, called \emph{The Rayleigh Collapse}, the $\dot{R}^2$ term will be much larger than the right hand side of eq.
-\ref{eq:RP_2} and we may neglect it.
-\cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
-Rayleigh looked at the same problem, i.e.
-a cavitating \emph{void} in water, by replacing $p_0\rightarrow 0$ \cite{brenner2002single,rayleigh1917pressure}:
+If we look at the interval where the bubble wall's velocity is very large, called \emph{The Rayleigh Collapse}, the right hand side of eq.
+\ref{eq:RP_2} will be much smaller than the $\dot{R}^2$ term and we may neglect it \cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
+Rayleigh looked at the same problem, i.e. a cavitating void in water:
 \begin{equation}
@@
 and found the solution to be $R(t)\sim\left(t_*-t\right)^{2/5}$ \cite{brenner2002single}, with $t_*$ the time of collapse.
-The \emph{velocity}, $\dot{R}\sim\left(t_*-t \right)^{-3/5}$ diverges as the bubble collapses.
+The velocity, $\dot{R}\sim\left(t_*-t \right)^{-3/5}$, diverges as the bubble collapses.
 This observation is what lead to Rayleigh's argument for the origin of cavitation damage to ship's propellers.
-Yasui gave a rather elegant heuristic explanation for the origin of \emph{cavitation} in a fluid that is worth stating here \cite{yasui2018acoustic}.
+%Yasui gave a rather elegant heuristic explanation for the origin of \emph{cavitation} in a fluid that is worth stating here \cite{yasui2018acoustic}.
 Consider the mass flowing through concentric shells around a bubble, $4\pi R_i^2 v_i \rho$.
@@
 \ref{eq:RP_2}), the only change being the added term $\propto \dot{p}_g$ which is determined from solving the decoupled problem of the gas inside the bubble.
-Eq.
-\ref{eq:RP_3} is usually solved numerically; we will look at results of numerical solutions later when talking about experiments.
-For now, we note that the observed error between this equation and experiment is only significant in the interval during bubble collapse and even still, results of solving this equation are quantitatively in good agreement with measurements of the bubble's radius \cite{brenner2002single}.
+The observed error between this equation and experiment is only significant in the interval during bubble collapse and even still, results of solving this equation are quantitatively in good agreement with measurements of the bubble's radius \cite{brenner2002single}.
 
@@
 The theoretical progress on the bubble's interior since the discovery of SBSL in the late 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (from which we may calculate the temperature) are used to try to reproduce the easily measurable dynamics of the bubble.
-(ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{}.
+(ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{hiller1992spectrum,brenner2002single}.
 Obviously these are related topics, but the distinction in the context of SBSL is important as we will now see.
@@
 Most early progress on the bubble's interior depended on the former method: if we predict the right dynamics, then we might know the correct pressure and temperature in the bubble.
-These data are then used to try to describe the light emission \cite{}.
-This was preferred to studying the emitted spectra directly as careful measurements resulted in an almost featureless spectrum that could not be explained \cite{}.
+These data are then used to try to describe the light emission.
+This was preferred over studying the emitted spectra directly as careful measurements resulted in an almost featureless spectrum that could not be explained \cite{hiller1992spectrum}.
 Calculating the dynamics of the bubble's wall from the RP equations (eq.
@@
 For the first decade or so, SBSL experiments were done using partially degassed\footnote{Degassing lowers the ambient pressure, making it easier for bubbles to form.} water as the liquid medium \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}.
-Careful measurements of the emission from SBSL in water resulted in an almost featureless spectrum \cite{}.
-Early attempts were made to fit this as black-body radiation resulting in \emph{huge} temperature estimates, e.g.
-$10^8$ K \cite{}.
-It was later discovered that SBSL is volume emission, indicating that black-body radiation is not valid \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
+Careful measurements of the emission from SBSL in water resulted in an almost featureless spectrum and attempts to fit it as black-body radiation were not sucessful \cite{hiller1992spectrum}.
+It was later discovered that SBSL is volume emission, indicating that a black-body is not valid anyway \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
 Little progress was made on this front for quite some time \cite{brenner2002single}.
@@
 Considering the relatively large content of O$_2$ and N$_2$, degassed water \emph{regassed} with N$_2$ or O$_2$ or a mixture were checked first and found not to produce SBSL \cite{hiller1994effect}.
-What was discovered was that a small amount of \emph{noble gas} was required for SBSL to occur \cite{barber1997defining,hiller1994effect,brenner2002single}.
+What was discovered was that a small amount of noble gas was required for SBSL to occur \cite{barber1997defining,hiller1994effect,brenner2002single}.
 A result of this work was the \emph{argon rectification hypothesis} \cite{lohse1997sonoluminescing,brenner2002single,yasui2018acoustic,suslick2008inside}.
@@
 Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
-Perhaps equally as crucially, they found the \emph{absence} of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas, experimentally confirming the argon rectification hypothesis.
+Perhaps equally as crucially, they found the absence of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas, experimentally confirming the argon rectification hypothesis.
 Moreover, the spectra at different pressures were in very good agreement with detailed calculations of emission from noble gas plasmas \cite{an2009diagnosing,an2008spectral}.
@@
 Suppose that we have a volume, $V$, of an ideal gas containing $N$ particles at a temperature $T$.
-The ideal gas law tells us the equation of state is $PV=Nk_BT$ with $k_B$ \emph{Boltzmann's constant}.
+The ideal gas law tells us the equation of state is $PV=Nk_BT$ with $k_B$ Boltzmann's constant.
 Since we are assuming adiabatic compression, the second law of thermodynamics tells us that $dU=-PdV$ with $U$ the internal energy in the gas.
-Moreover, we may use the equipartition theorem to write $dU=N f k_B dT/2$ with $f$ the number of \emph{degrees of freedom} per particle \cite{}.
-Recall that for an ideal gas \cite{}:
-\begin{equation}
-\begin{split}
-    C_V &= \frac{N f k_B}{2} \\
-    C_P &= C_V+N k_B \\
-    \frac{C_P}{C_V} &= 1+\frac{2}{f} \equiv \gamma 
-\end{split}
-\end{equation}
-with $C_V$ and $C_P$ the heat capacities at constant volume and pressure respectively.
+Moreover, we may use the equipartition theorem to write $dU=N f k_B dT/2$ with $f$ the number of degrees-of-freedom per particle \cite{schroeder1999introduction}.
+Recall that for an ideal gas $C_V = N f k_B/2$, $C_P = C_V+N k_B$, and $C_P/C_V = 1+2/f \equiv \gamma$ with $C_V$ and $C_P$ the heat capacities at constant volume and pressure respectively.
 For a monatomic ideal gas, $f=3$ and $\gamma=5/3$.
@@
 Let us now connect this to the problem of SBSL.
-In our case, we assume that our bubble has an \emph{ambient} radius $R_0$ where the pressure is at its equilibrium value, $p_0$.
+In our case, we assume that our bubble has an ambient radius $R_0$ where the pressure is at its equilibrium value, $p_0$.
 The volume occupied by the gas is $V_0 =  4/3 \pi R_0^3$.
@@
 During the expansion and after-bounce stages of the bubbles cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal}.
-This is true because, when the bubble's velocity is on the order of the \emph{heat diffusion} timescale, the temperature throughout the bubble is nearly equal to that in the liquid \cite{prosperetti1999old,brenner2002single,yasui2018acoustic}.
+This is true because, when the bubble's velocity is on the order of the heat diffusion timescale, the temperature throughout the bubble is nearly equal to that in the liquid \cite{prosperetti1999old,brenner2002single,yasui2018acoustic}.
 The relevant modifications to eq.
@@
 Since we are assuming LTE, the intensity of emitted radiation depends instantaneoisly on the temperature-time curve, $T(t)$, that we calculated before.
-Planck's law is \cite{schroeder1999introduction}
+Planck's law gives the energy radiated per unit-time and unit-volume, $I_B$, \cite{schroeder1999introduction}
 \begin{equation}
-    I_{B}(\lambda;T(t))=\frac{2 h c^2}{\lambda^5}\frac{1}{\exp(hc/\lambda k_B T(t))-1}
+    I_{B}(\lambda;T(t))=\frac{2 h c^2}{\lambda^5}\frac{1}{\exp(hc/\lambda k_B T(t))-1}.
 \end{equation}
-The \emph{black-body model} assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out.
-Then the only emission that is measured comes from a thin layer on the bubbles surface.
-Calculations of SBSL assuming this over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}.
-Instead, we need to consider finite opacity of the bubble, i.e.
-some of the emitted radiation is absorbed on its way out.
+The subscript $B$ is to remind us this is for a black-body which assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out.
+Then the only emission that is measured comes from a thin layer on the bubble's surface.
+Calculations of SBSL assuming black-body radiation over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}.
+Instead, we need to consider finite opacity of the bubble.
 
@@
 \appendix
-
-\section{The Navier-Stokes Equation}
-The Rayleigh-Plesset (RP) equation can be deduced from the Navier-Stokes equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
-To that end, we remind ourselves of the Navier-Stokes equations here.
-
-The \emph{total} change (in time) of mass in an arbitrary element of a fluid (whose coordinates and boundary may depend on time) is given by  $\frac{d}{dt} \int_{\Omega_t} \rho dV$ where $\Omega_t$ is the domain of the fluid element under consideration and $\rho(\bm{x},t)$ is the local, time-dependent density of the fluid.
-Using the transport theorem of fluid dynamics \cite{mcdonough2009lectures}, Gauss's theorem, and (with only some loss of generality) taking $\Omega_t$ to be a \emph{fluid parcel} (so that the boundary of the fluid parcel, $\partial \Omega_t$, moves with the fluid's velocity, $\bm{u}(\bm{x},t)$), this becomes
-\begin{equation}
-    \frac{d}{dt} \int_{\Omega_t} \rho dV = \int_{\Omega_t} \partial_t \rho dV+\int_{\partial \Omega_t} \rho \bm{u} \cdot \bm{n} dA = \int_{\Omega_t} \partial_t \rho+\nabla \cdot(\rho \bm{u})dV.
-    \label{eq:mass_continuity}
-\end{equation}
-Now to have \emph{mass-continuity}, which we will require on physical grounds, we set eq.
-\ref{eq:mass_continuity} to 0.
-Basically, we are saying no additional material is being added or removed from the fluid.
-Since the fluid parcel's boundary $\Omega_t$ is arbitrary, we find that $\partial_t \rho+\nabla\cdot(\rho \bm{u})\equiv 0$ which is true for every point in the fluid volume.
-
-Newton's second law for fluids (\emph{momentum-continuity}) is defined similarly \cite{schoeffel2014lecture,mcdonough2009lectures}.
-We write the momentum-density as $\rho \bm{u}$ with momentum given by $\int_{\Omega_t} \rho \bm{u} dV$.
-To simplify what follows, we focus on only the $i^{th}$ component of velocity at a time.
-Then Newton's second law can be written as
-\begin{equation}
-    \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \partial_t (\rho u_i) dV+\int_{\partial \Omega_t} \rho u_i ( \bm{u} \cdot \bm{n}) dA = F_i.
-\end{equation}
-Now using Gauss's theorem, the product rule, and mass-continuity we find $ \frac{d}{dt} \int_{\Omega_t} \rho u_i dV =  \int_{\Omega_t} \rho \frac{d u_i}{dt} dV = F_i $. In vector notation this reads
-\begin{equation}
-    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt} = \bm{F}
-    \label{eq:mom_continuity}
-\end{equation}
-
-We now turn to the forces.
-Let us separate the forces into \emph{body-} and \emph{surface-forces}: $\bm{F}=\int_{\Omega_t} \bm{f}_B dV+ \int_{\partial \Omega_t} \bm{f}_S dA $.
-The body-force density, $\bm{f}_B$, is due to forces acting in the bulk of the liquid e.g.
-gravity, electrostatic forces, etc.
-and is typically from sources external to the liquid.
-We won't be concerned with these forces here.
-Instead, let us concentrate on the surface part.
-If we only consider an ideal fluid, the surface part can be written as $\int_{\partial \Omega_t} \bm{f}_S dA =-\int_{\partial \Omega_t} p \bm{n} dA = -\int_{\Omega_t}\nabla p dV$ with $p$ the pressure  and the minus sign being due to the sign convention: \emph{out} of the surface $\partial \Omega_t$ is positive.
-
-If instead the fluid is real, we can't neglect viscous forces.
-We suppose that the viscous surface term can be represented by a tensor, $\hat{\bm{\tau}}$ such that $\bm{F}_v=\int_{\partial \Omega_t} \hat{\bm{\tau}}\cdot \bm{n} dA = \int_{\Omega_t} \nabla \cdot \hat{\bm{\tau}} dV$.
-(Note, here the dot product is to be understood as matrix multiplication).
-To leading order in gradients of the velocity, (and under the physical constraints that the tensor be symmetric), the viscous-stress tensor is $\hat{\bm{\tau}}=\hat{\bm{\mu}}\cdot \hat{\bm{\epsilon}}$ where $\hat{\bm{\mu}}$ is the viscosity-tensor and $\hat{\bm{\epsilon}}$ is the strain-rate tensor: $\epsilon_{ij}=\frac{1}{2}\left[ \partial_i u_j+ \partial_j u_i \right]$.
-
-In general, $\hat{\bm{\mu}}$ is a symmetric rank-4 tensor.
-However, if we assume an isotropic fluid \footnote{Definitely true for water which is has been frequently used in SL experiments.}, we find that the the viscous-stress tensor can be separated into two irreducible parts: a scalar, $\hat{\bm{\epsilon}}^{(v)}$, and traceless symmetric part, $\hat{\bm{\epsilon}}^{(s)}$ \cite{zee2016group,landau1987fluid}:
-\begin{equation}
-\begin{split}
-    \hat{\bm{\tau}} & = \xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)} \\
-    \hat{\bm{\epsilon}}^{(v)} & = \delta_{ij}\epsilon_{kk} \\
-    \hat{\bm{\epsilon}}^{(s)} & = \hat{\bm{\epsilon}}-\frac{1}{3}\delta_{ij}\epsilon_{kk}.
-    \label{eq:viscous_stress_tensor}
-\end{split}
-\end{equation}
-Thus, we see that the viscosity-tensor only has two free components: $\xi$, the normal viscosity and $\mu$ the shear viscosity.
-It is usual to combine the pressure (i.e.
-elastic-stresses) and the viscous-stresses into the stress-tensor $\hat{\bm{\sigma}}=-p\hat{\bm{I}}+\xi \hat{\bm{\epsilon}}^{(v)}+\mu \hat{\bm{\epsilon}}^{(s)}$.
-Combined with this, Newton's second law becomes
-\begin{equation}
-    \int_{\Omega_t} \rho \frac{d \bm{u}}{dt}-\bm{F} = \int_{\Omega_t} \rho \frac{d \bm{u}}{dt} - \nabla \cdot \hat{\bm{\sigma}}-\bm{f}_B dV = 0
-\end{equation}
-Again, since we allow the volume of integration to be arbitrary, the integrand must vanish everywhere: $\rho \frac{d \bm{u}}{dt} - \nabla \cdot \hat{\bm{\sigma}}-\bm{f}_B=0$.
-This is the momentum-continuity equation for a fluid.
-Now we would rather have this explicitly in terms of the velocity $\bm{u}$.
-To that end, we may substitute eq.
-\ref{eq:viscous_stress_tensor} into the momentum-continuity equation.
-For e.g.
-the $i^{th}$ component we may expand out the divergence part as
-\begin{equation}
-\begin{split}
-    \left(\nabla \cdot \hat{\bm{\sigma}}\right)_i = -\partial_i p + \left[\xi +\frac{2}{3}\mu \right] \partial_i(\nabla \cdot \bm{u})+\mu \nabla^2 u_i \equiv -\partial_i p + \eta \partial_i(\nabla \cdot \bm{u})+\mu \nabla^2 u_i
-\end{split}
-\end{equation}
-where we have introduced the \emph{bulk-viscosity} $\eta=\xi+\frac{2}{3}\mu$.
-Collecting all of the components into vector form and, combined with the mass-continuity equation \ref{eq:mass_continuity}, we arrive at the \emph{Navier-Stokes} (NS) equations for a compressible fluid:
-\begin{equation}
-\begin{split}
-    \rho \frac{d \bm{u}}{dt} = \rho \left[ \partial_t \bm{u}+\left(\bm{u}\cdot \nabla \right)\bm{u} \right] & =-\nabla p + \eta \nabla \left(\nabla \cdot \bm{u} \right)+\mu \nabla^2 \bm{u} + \bm{f}_B \\
-     \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
-     \label{eq:ns_equations_2}
-\end{split}
-\end{equation}
-
-Some comments are in order.
-The total-derivative of a moving fluid $\frac{d}{dt}\equiv D_t$ is sometimes called the \emph{material derivative}.
-In the case of velocity $D_t \bm{u}$ is the \emph{acceleration} of the fluid.
-It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies \emph{in space} as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
-The pressure term on the right hand side is to be understood as forces from the \emph{elastic} energy.
-On the other hand, the terms $\sim \bm{u}$ are viscous terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
-The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
-dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
-The body term $\bm{f}_B$ is from external sources.
-It is worth noting that solving eqs.
-\ref{eq:ns_equations_2} is a formidable task \footnote{In fact, it's not even clear that the NS equations \emph{can} be solved even for incompressible fluids.
-In 2000, it was announced that proving smooth, sensible solutions exist earns \$1,000,000 \cite{fefferman2006existence}.
-Better hurry though...
-inflation was 6.8\% in Nov.
-2021.} and is usually done using computers.
-
 \section{Electrical Arguments}
-
 \section{The Casimir Argument}

### 2022-01-19 sonoluminescence/paper/bak/v6/paper.tex
sdfadsfa

--- 2022-01-14 sonoluminescence/paper/paper.tex
@@
 Keller and coworkers showed that the modification required to the RP equation is the replacement $P(t) \rightarrow P_0 \sin(\omega t)$ \cite{keller1980bubble}.
-If we linearize in the bubble's radius $R(t)\approx R_0 \left( 1+x(t) \right)$ with $x(t) \ll 1$, we can derive a forced harmonic oscillator equation for $x(t)$ \cite{brennen2014cavitation,yasui2018acoustic}.
-However, the low driving amplitude regime where this method is valid is irrelevant to SBSL so we won't pursue it here.
 
@@
 It is no great challenge computationally.
-For now, let's see what we else we can understand from this analytically by making some more approximations.
-If we look at the interval where the bubble wall's velocity is very large, called \emph{The Rayleigh Collapse}, the right hand side of eq.
+For now, let's see what we can understand analytically by making some more approximations.
+If we linearize in the bubble's radius $R(t)\approx R_0 \left( 1+x(t) \right)$ with $x(t) \ll 1$, we can derive a forced harmonic oscillator equation for $x(t)$ \cite{brennen2014cavitation,yasui2018acoustic}.
+However, the low driving amplitude regime where this method is valid is irrelevant to SBSL so we won't pursue it here.
+Instead, we can look at the interval where the bubble wall's velocity is very large, called \emph{The Rayleigh Collapse}; the right hand side of eq.
 \ref{eq:RP_2} will be much smaller than the $\dot{R}^2$ term and we may neglect it \cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
@@
 The observed error between this equation and experiment is only significant in the interval during bubble collapse and even still, results of solving this equation are quantitatively in good agreement with measurements of the bubble's radius \cite{brenner2002single}.
+
+Above, we assumed that the bubble always remains spherical.
+Of course this isn't always true and \emph{shape-instabilities} occur for the right combinations of mechanical properties and driving force/frequency.
+Usually shape-instabilities destroy the bubble and kill SBSL so we won't focus on these isues here.
+We will always assume we are in a parameter-regime where SBSL occurs.
+The reader is referred to one of several reviews on the \emph{phase-space} of SBSL for more info \cite{yasui2018acoustic,brenner2002single,hilgenfeldt1999sonoluminescence,an2009diagnosing}.
 
@@
 Considering the relatively large content of O$_2$ and N$_2$, degassed water \emph{regassed} with N$_2$ or O$_2$ or a mixture were checked first and found not to produce SBSL \cite{hiller1994effect}.
-What was discovered was that a small amount of noble gas was required for SBSL to occur \cite{barber1997defining,hiller1994effect,brenner2002single}.
+What was discovered was that a small amount of noble-gas was required for SBSL to occur \cite{barber1997defining,hiller1994effect,brenner2002single}.
 A result of this work was the \emph{argon rectification hypothesis} \cite{lohse1997sonoluminescing,brenner2002single,yasui2018acoustic,suslick2008inside}.
@@
 Over many cycles, the contents of an air bubble in water become nearly pure argon.
-Moreover, it was hypothesized that the SL would much more intense when the contents of the bubble are a pure inert gas: if the contents were e.g.
+Moreover, it was hypothesized that the SL would be much more intense when the contents of the bubble are a pure inert gas: if the contents were e.g.
 molecules, bond breaking/formation would alter the pressure of the gas, ultimately reducing the temperature and reducing the amount of light production \cite{brenner2002single,yasui2018acoustic,lohse2018bubble,suslick2008inside}.
@@
 the low temperature, in MBSL: the bubbles are transient and cannot remove a significant amount of O$_2$ or N$_2$ over a single cycle.
-The current belief is that the contents of a bubble undergoing SBSL in water are pure argon after $\sim 10^3$ cycles \cite{brenner2002single,yasui2018acoustic}.
+The current belief is that the contents of a bubble undergoing SBSL in water are (nearly) pure argon after $\sim 10^3$ cycles \cite{brenner2002single,yasui2018acoustic}.
 
-Still, most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurements of the light spectrum \cite{}.
+Still, most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurements of the light spectrum \cite{flannigan2005plasma,flannigan2006measurement}.
+The mechanical properties in aqueous H$_2$SO$_4$ allow stable SBSL with a larger bubble radii than in water, increasing the emitting volume.
 More importantly, new measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of pressure than water.
-Careful experiments revealed that the spectrum depended critically on the noble gas content and driving pressure.
+SBSL in aqueous H$_2$SO$_4$ could be driven at very low pressure, leading to much lower temperature bubbles, while driving at large pressure led to conditions similar to SBSL in water.
+
+Careful experiments revealed that the spectrum depended critically on the noble-gas content and driving pressure.
 Suslick et.
@@
 Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
-Perhaps equally as crucially, they found the absence of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas, experimentally confirming the argon rectification hypothesis.
-Moreover, the spectra at different pressures were in very good agreement with detailed calculations of emission from noble gas plasmas \cite{an2009diagnosing,an2008spectral}.
-Increased pressure results in increased collision rate between particles, broadening the excited state transition spectral lines \cite{an2008spectral,suslick2008inside,flannigan2005plasma,flannigan2006measurement}.
-Further calculations showed that the greatly increased viscosity and density of aqueous H$_2$SO$_4$, $\sim 10 \times$ and $\sim 1.3 \times$ larger than that that of water, lead to reduced (by orders of magnitude) pressure in the bubble at the same driving pressures \cite{an2009diagnosing}.
-The larger pressure in the bubbles in water was found to be consistent with the featureless emission spectrum \cite{suslick2008inside,yasui2018acoustic}.
+Perhaps equally as crucially, they found the absence of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble-gas, experimentally confirming the argon rectification hypothesis.
+The spectra at different pressures were in very good agreement with detailed calculations of emission from noble-gas plasmas \cite{an2009diagnosing,an2008spectral}.
+This work also explained why the spectrum in water is featureless (i.e.
+has no emission lines).
+The larger pressure in SBSL in water results in an increased collision rate between particles, broadening the spectral lines \cite{an2009diagnosing,suslick2008inside,flannigan2005plasma,flannigan2006measurement}.
 
-This argon rectification hypothesis greatly helped to refine models for the gas dynamics in the bubble and clarified several unexplained experimental observations \cite{suslick2008inside,brenner2002single}.
+This argon rectification hypothesis greatly helped to refine models for the gas dynamics in the bubbles and confirmed why simple models usually work well \cite{suslick2008inside,brenner2002single}.
 With this in mind, we choose only to focus on results known to be relevant to the dynamics of a bubble undergoing SBSL.
 Excellent books and modern reviews on modeling the gas dynamics exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.
-We will hold off on discussing the dynamics in the plasma until a later section.
+We will hold off on discussing the plasma until a later section.
 
 \subsection{Gas Dynamics}
-The most straightforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
-In fact, serious quantitative predictions of the conditions of the bubble's interior take this path \cite{an2006mechanism,an2008spectral,an2009diagnosing,flannigan2005plasma,flannigan2006measurement}.
+Of course the most straightforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
+In fact, serious quantitative predictions of the conditions of the bubble's interior still take this path \cite{an2006mechanism,an2008spectral,an2009diagnosing,flannigan2005plasma,flannigan2006measurement}.
 With simplifying assumptions, equations of motion with varying degrees of sophistication can be derived for the gas and the system of gas dynamical equations and the RP equations are solved numerically.
-The discussion proceeds in much the same way as deriving the RP equations above but is considerably more complicated \cite{brenner2002single,yasui2018acoustic}, so we will omit it here.
+The discussion proceeds in much the same way as deriving the RP equations above but is slighlty more complicated \cite{brenner2002single,yasui2018acoustic}, so we will omit it.
 An important aspect of these studies is the understanding of why relatively simple models work quite well.
+
 We will highlight the salient features before introducing a simple model for the gas that will enable us to gain some physical understanding analytically.
-An especially important result of direct numerical solution of the gas dynamics was (theoretical) confirmation of the argon rectification hypothesis \cite{}, reducing the possible gas dynamical models to those applicable to a monatomic gas.
-Various other models fully including dissipation and heat transport into/out of the bubble showed that the bubble's compression is, to a good approximation, adiabatic and quasistatic \cite{}.
-Before we turn to discuss a model for the gas dynamics, let us pause to mention that another theoretically appealing method for modeling the gas dynamics is \emph{molecular dynamics} \cite{}.
+An especially important result of direct numerical solution of the gas dynamics was (theoretical) confirmation of the argon rectification hypothesis, reducing the possible gas dynamical models to those applicable to a monatomic gas.
+Various other models fully including dissipation and heat transport into/out of the bubble showed that the bubble's compression is, to a good approximation, adiabatic and quasistatic.
+Before we turn to discuss a model for the gas dynamics, let us pause to mention that another theoretically appealing method for modeling the gas dynamics is \emph{molecular dynamics}.
 While this method is attractive for a number of reasons, practical calculations containing realistic numbers of particles aren't possible, limiting the applicability \cite{brenner2002single}.
 
-A simple but reasonable model for the gas dynamics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.
-The modifications to the result for an ideal gas are simple \cite{sivasubramanian2002temperature}.}.
-Let us remind ourselves what the pressure is in this case.
+Recalling that the bubble is (almost) purely noble-gas, a simple but reasonable model for the gas dynamics is that of an adiabatically, quasistatically compressed ideal\footnote{Actually a \emph{Van der Walls} gas model is better, but we will sneak this aspect in later.
+The modifications to the result for an ideal gas are simple \cite{sivasubramanian2002temperature}.} gas \cite{brenner2002single}.
+Let us remind ourselves what the pressure is.
 Suppose that we have a volume, $V$, of an ideal gas containing $N$ particles at a temperature $T$.
 The ideal gas law tells us the equation of state is $PV=Nk_BT$ with $k_B$ Boltzmann's constant.
-Since we are assuming adiabatic compression, the second law of thermodynamics tells us that $dU=-PdV$ with $U$ the internal energy in the gas.
-Moreover, we may use the equipartition theorem to write $dU=N f k_B dT/2$ with $f$ the number of degrees-of-freedom per particle \cite{schroeder1999introduction}.
-Recall that for an ideal gas $C_V = N f k_B/2$, $C_P = C_V+N k_B$, and $C_P/C_V = 1+2/f \equiv \gamma$ with $C_V$ and $C_P$ the heat capacities at constant volume and pressure respectively.
-For a monatomic ideal gas, $f=3$ and $\gamma=5/3$.
+Since we are assuming adiabatic compression, the second law of thermodynamics is $dU=-PdV$ with $U$ the internal energy in the gas.
+We may use the equipartition theorem to write $dU=N f k_B dT/2$ with $f$ the number of degrees-of-freedom per particle \cite{schroeder1999introduction}.
+Recall that for an ideal gas the specific heat at constant volume is $C_V = N f k_B/2$ and at constant pressure is $C_P = C_V+N k_B$, with $C_P/C_V = 1+2/f \equiv \gamma$.
+For a noble-gas, $f=3$ and $\gamma=5/3$.
 Now we can use the second law of thermodynamics to write $C_V dT = -P dV$.
@@
 \end{equation}
-Eliminating $dT$ and simplifying, 
+Inserting $dT$ into the second law and simplifying, 
 \begin{equation}
@@
 \ref{eq:p_g} is sensible during the bubble's collapse: the bubble wall's velocity is fast and very little heat can flow out during compression.
-Detailed calculations starting from the full NS equations further tell us that the gas dynamics in the bubble can be regarded as \emph{quasistatic} \cite{}.
+Detailed calculations starting from the full NS equations further tell us that the gas dynamics in the bubble can be regarded as \emph{quasistatic} \cite{brenner2002single}.
 During the expansion and after-bounce stages of the bubbles cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal}.
@@
 A simple alternative is choosing a constant exponent somewhere between the adiabatic $\gamma=5/3$ and isothermal $\gamma=1$ values \cite{hilgenfeldt1999simple}.
-It is also possible to include heat and mass diffusion in simple model forms and these do give slightly improved quantitative agreement with experiment; however we choose to discuss solutions of the RP equations using the simpler form given in eq.
-\ref{eq:p_g} as this will enable to understand the important features of the bubble's cycle.
+It is also possible to include heat and mass diffusion in simple model forms and these do give slightly improved quantitative agreement with experiment; however the simpler form given in eq.
+\ref{eq:p_g} gives close to the correct dynamics (see fig.
+\ref{fig:bubble_radius_1}).
 More advanced solutions using the full gas dynamical equations are qualitatively similar \cite{brenner2002single,yasui2018acoustic}.
@@
 \end{equation}
-Earlier, we assumed that during the bubble's after-bounces and expansion, the gas dynamics were isothermal.
-Then we take $T_0$ to be the temperature of the liquid.
+Earlier, we said that during the bubble's after-bounces and expansion, the gas dynamics is isothermal; then we take $T_0$ to be the temperature of the liquid.
 The RP equation, eq.
 \ref{eq:RP_3}, combined with the gas dynamical equation, eq.
-\ref{eq:p_g}, and the known liquid and gas properties $\eta$, $\sigma$, and $h$, can be used to calculate the dynamics of an acoustically driven bubble, determining the time-dependence of the pressure and temperature in the bubble along the way.
+\ref{eq:p_g}, and the known liquid and gas properties $\eta$, $\sigma$, and $h$, can be used to calculate the dynamics of an acoustically driven bubble, giving us the time-dependence of the pressure and temperature in the bubble.
 
@@
 \includegraphics[width=0.9\linewidth]{figs/bubble_radius_1.pdf}
-    \caption{\cite{barber1997defining,barber1992light}}
+    \caption{Numerical solution of the RP equation variant eq.
+\ref{eq:RP_3} with the gas dynamics given by eq.
+\ref{eq:p_g} compared to Mie scattering results \cite{barber1997defining,barber1992light}.}
 \label{fig:bubble_radius_1}
@@
 What we are claiming is simple is the experimental setup once the correct parameters are identified.}.
-A typical experimental setup is shown in Fig.
+A typical experimental setup is shown in fig.
 \ref{fig:flask}.
 Since we are mainly concerned with the explaining the light emission process itself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}.
-A sample of degassed liquid is placed into a flask \footnote{Actually, the degassing part is quite important.
-We focus on it elsewhere.}.
+A suitable sample of liquid is placed into a flask.
 Coupled to the outside of the flask are piezoelectric transducers of some sort.
@@
 
-Let us take a brief aside: in deriving the RP equations, we assumed the center of the bubble was stationary.
+%Let us take a brief aside: in deriving the RP equations, we assumed the center of the bubble was stationary.
 Actually, this is what we want in an SBSL experiment and it is worth discussing under what conditions this is true.
 In an ideal, incompressible fluid the net force acting on a bubble may be written as
-\begin{equation}
-    \bm{f}=-\int_{\partial \Omega} p \bm{n} dA = -\int_\Omega \nabla p dV
-\end{equation}
-Here, $\bm{f}$ is the net force, $p$ is the pressure at the bubble's surface, $\bm{n}$ is the unit normal vector point out of the bubbles surface, and $\partial \Omega$ is the bubble's boundary with $\Omega$ the bubble's domain.
+%\begin{equation}
+%    \bm{f}=-\int_{\partial \Omega} p \bm{n} dA = -\int_\Omega \nabla p dV
+%\end{equation}
+%Here, $\bm{f}$ is the net force, $p$ is the pressure at the bubble's surface, $\bm{n}$ is the unit normal vector point out of the bubbles surface, and $\partial \Omega$ is the bubble's boundary with $\Omega$ the bubble's domain.
 The last term is due to an application of Gauss's theorem.
 Bjerknes \cite{bjerknes1909kraftfelder} took $\nabla p \equiv \nabla p(r=0,t)$ to be constant and averaged over one period of oscillation to find
-\begin{equation}
-    \bm{f} = -\frac{4}{3}\pi \langle R^3 \nabla p \rangle.
-    \label{eq:Bj_force}
-\end{equation}
-This is the so called \emph{Bjerknes force}.
+%\begin{equation}
+%    \bm{f} = -\frac{4}{3}\pi \langle R^3 \nabla p \rangle.
+%    \label{eq:Bj_force}
+%\end{equation}
+%This is the so called \emph{Bjerknes force}.
 The point is that if the bubble is located at a pressure extrema, e.g.
@@
 
-Some authors have looked at corrections to the bubble's dynamics in the presence of buoyancy, which cause translational motion even for vanishing Bjerknes force.
+%Some authors have looked at corrections to the bubble's dynamics in the presence of buoyancy, which cause translational motion even for vanishing Bjerknes force.
 The result is the bubble oscillates around an equilibrium position near but away from the center of the flask \cite{matula2000single,matula1997bjerknes,matula1999inertial}.
@@
 
-Back to the experimental setup.
-The typical driving pressure relevant to SBSL ($\sim 1.5$ atm) is too small to cause bubbles to form spontaneously in the flask \cite{}.
+The typical driving pressure relevant to SBSL ($\sim 1.5$ atm) is too small to cause bubbles to form spontaneously in the flask \cite{brenner2002single}.
 Instead, the bubble is usually seeded somehow.
 In the original work of Gaitan, an air bubble was injected using a needle and syringe through a small hole in the flask's seal \cite{gaitan1992sonoluminescence}.
-More recently, seeding methods involve shooting a small jet of water into the flask (like our shrimp friends!) or blasting the water with a laser to boil a small volume \cite{}.
+More recently, seeding methods involve shooting a small jet of water into the flask (like our shrimp friends!) or blasting the water with a laser to boil a small volume \cite{suslick2008inside,yasui2018acoustic}.
 The later method is preferred as it enables more precise control over the initial bubble's radius.
-The small bubbles formed this way coalesce at the center of the flask into a single, stable bubble \cite{}.
 
-Now that we know about how single, stable bubbles are formed in the lab, we can discuss two experimentally accessible quantities that have turned out the be the most important: (i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations \cite{}.
+Now that we know about how single, stable bubbles are formed in the lab, we can discuss two experimentally accessible quantities that have turned out the be the most important: (i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations \cite{barber1992light,barber1997defining,brenner2002single}.
 This can tell us about the temperature and pressure in the bubble.
-(ii) The other quantity of interest is the light spectrum which contains information on the bubble's contents: in principle, we can learn about temperature and pressure inside the bubble, with spectral lines telling us about the matter content \cite{}.
-As we will see, it is not simple to understand the light spectrum \cite{} and we will devote a section of this paper to this topic.
+(ii) The other quantity of interest is the light spectrum which contains information on the bubble's contents: in principle, we can learn about temperature and pressure inside the bubble, with spectral lines telling us about the matter content \cite{hiller1992spectrum,hilgenfeldt1999simple,flannigan2005plasma,flannigan2006measurement}.
+As we will see, it is not simple to understand the light spectrum and we will devote a section of this paper to this topic.
 For now, let us see what measurements of the bubble's dynamics have told us.
 
-Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
-Assuming the bubble is a homogeneous dielectric sphere \cite{}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{}.
+%Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
+Assuming the bubble is a homogeneous dielectric sphere \cite{jackson1999classical}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{brenner2002single}.
 A typical Mie scattering experiment is shown in fig.
@@
 Both effects are largest during the SL period of the bubble's cycle, leading to an important source of error when fitting the calculated $R(t)$ curve \cite{}.
-Parameters used to calculate the gas dynamics in the bubble can then be fit to the measured radius to build a full model of the dynamics.
+Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
+Assuming the bubble is a homogeneous dielectric sphere \cite{jackson1999classical}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{brenner2002single}.
+A typical Mie scattering experiment is shown in fig.
+\ref{fig:flask} (c).
+Parameters used to calculate the gas dynamics in the bubble are fit to the measured radius to build a full model of the dynamics.
 In the case of eq.
@@
 In the case of SBSL, we will look for corrections due to interactions of the emitted photons with atoms on the way out of the bubble \cite{hilgenfeldt1999sonoluminescence}.
-Since we are assuming LTE, the intensity of emitted radiation depends instantaneoisly on the temperature-time curve, $T(t)$, that we calculated before.
+Since we are assuming LTE, the intensity of emitted radiation depends instantaneoisly on the temperature-time curve, $T(t)$, that we just learned how to calculate.
 Planck's law gives the energy radiated per unit-time and unit-volume, $I_B$, \cite{schroeder1999introduction}
@@
 \end{equation}
-The subscript $B$ is to remind us this is for a black-body which assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out.
-Then the only emission that is measured comes from a thin layer on the bubble's surface.
+The subscript $B$ is to remind us this is for a black-body; this is an \emph{optically thick} thick bubble that absorbs all radiation.
+Then there is perfect emission from the bubble's surface.
 Calculations of SBSL assuming black-body radiation over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}.
@@
 
-Let's call the absorption coefficient $\kappa(\lambda,T(t))$\footnote{Really, this is an \emph{effective} absortion coeffecient $\kappa = \kappa'(1-\exp\left[ hc/k_B T \lambda \right])$ with $\kappa'$ the actual absoption coefficient of the material.
-The effective coefficient $\kappa$ includes induced emission of photons as light passes through the medium which 'decreases' absorption.
-Physically, this can be viewed as radiation that is absorbed being immediately re-emitted or, equivalently, vice-versa \cite{zel2002physics,taylor1969experimental}.} where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$.
-Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics,hilgenfeldt1999simple}
+Let's call the absorption coefficient $\kappa(\lambda,T(t))$ where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$.
+Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics,hilgenfeldt1999simple,taylor1969experimental}
 \begin{equation}
@@
 In the limit that $\kappa\rightarrow\infty$, we recover the result for a black-body emitter.
-On the otherhand, for a strong absorber, $1-\exp(\kappa s)\approx\kappa s$ and $I\approx I_B \kappa s$, i.e.
+On the otherhand, for a nearly tranparent bubble, $1-\exp(\kappa s)\approx\kappa s$ and $I\approx I_B \kappa s$, i.e.
 the number of photons emitted is greatly reduced, a result that we require.
@@
 \ref{eq:intensity} is integrated numerically.
-Moreover, integrating $I(\lambda,T(t))$ over the bubbles and over time allows calculating the radiated power spectrum which may be directly compared to experiment.
+Integrating over the bubbles volume and over time allows calculating the radiated power spectrum which may be directly compared to experiment.
 A great deal of work has been devoted to identifying the relevant absorption mechanisms leading to $\kappa$, with most progress being made only after SBSL was discovered in aqueous H$_2$SO$_4$ \cite{hilgenfeldt1999simple,brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside,an2009diagnosing,an2008spectral,an2006mechanism}.
@@
 Due to its different mechanical properties, H$_2$SO$_4$ can support SBSL over a larger range of driving pressures than water \cite{an2009diagnosing}.
-For low driving pressures, the temperature and pressure in the bubble are much lower than those encountered in bubbles in water and \emph{discrete} emission lines for molecular dissociation and electronic transitions are directly observable \cite{flannigan2005plasma,flannigan2006measurement}.
+For low driving pressures, the temperature and pressure are much lower than those encountered in bubbles in water and \emph{discrete} emission lines for molecular dissociation and electronic transitions are directly observable \cite{flannigan2005plasma,flannigan2006measurement}.
 At larger driving pressures leading to bubble conditions comparable to those in water, however, what was found is that the \emph{continous} processes electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination dominate \cite{hilgenfeldt1999simple,yasui2018acoustic,suslick2008inside,zel2002physics}.
@@
 The deflection accelerates the electron, changing its kinetic energy.
-Since mass of the ion or atom is much greater than the mass of the electron, we can take the atom or ion to be fixed.
+Since the mass of the ion or atom is much greater than the mass of the electron, we can take the atom or ion to be fixed.
 Since the electron is charged (and energy is conserved), it emits radiation through \emph{bremsstrahlung} as its kinetic energy changes \cite{jackson1999classical,zel2002physics}.
@@
 
-The \emph{Saha equation} of astrophysics allows us to estimate the fraction of particles that are ionized at a given temperature \cite{hilgenfeldt1999simple,an2006mechanism,zel2002physics}
+The Saha equation of astrophysics allows us to estimate the fraction of particles that are ionized at a given temperature \cite{hilgenfeldt1999simple,an2006mechanism,zel2002physics}
 \begin{equation}
@@
 \end{equation}
-It turns out that at typical SBSL temperatures, only $\sim 1\%$ of the atoms are ionized and thus, since collisions are very frequent, electron-neutral atom bremsstrahlung is perhaps as important as electron-ion processes \cite{hilgenfeldt1999simple,an2006mechanism,flannigan2005plasma}.
-For Ar$^+$ and $e$, formulas for recombination and electron-ion bremsstrahlung can be found in ref.
-\cite{hilgenfeldt1999simple}.
-For electron-atom processes, most modern work follows that of Geltman \cite{geltman1973free,an2006mechanism,an2008spectral}.
+It turns out that at typical SBSL temperatures, only $\sim 1\%$ of the atoms are ionized and thus, since collisions are very frequent, electron-atom bremsstrahlung is as important as electron-ion processes \cite{hilgenfeldt1999simple,an2006mechanism,flannigan2005plasma}.
 
@@
 \end{equation}
-where $\bm{a}$ is the acceleration and $e$ the electron charge (we are using CGS units here).
+where $\bm{a}$ is the acceleration, $e$ the electron charge, and $c$ the speed of light (we are using CGS units here).
 This can be deduced by calculating the Poynting vector and finding the terms that don't vanish at infinite distance.
@@
 Integrating over the azimuthal angle, there are $v N_e \cdot  2\pi b db$ electrons passing per unit time through a ring with area $2\pi b db$ centered on the ion.
-Integrating over all possible orbits, the energy radiated \emph{per-ion} per frequency interval $d\nu$, for unit electron flux, $v N_e=1$, is the so-called ``effective radiation" $d\epsilon_\nu$ \cite{zel2002physics}
+Integrating over all possible orbits, the energy radiated \emph{per-ion} per frequency interval $\nu+d\nu$, for unit electron flux, $v N_e=1$, is the so-called ``effective radiation" $d\epsilon_\nu$ \cite{zel2002physics}
 \begin{equation}
@@
 
-Now suppose that, in a unit volume of gas, there are $N_+$ ions with charge $Ze$ and $N_e$ electrons with velocities distributed according to $f(v)dv=4\pi(m/2\pi k_B T)^{3 / 2} \exp(-mv^2 / 2 k_B T) v^2 dv$, i.e.
-the \emph{Maxwell velocity distribution}, passing through the volume.
+Now suppose that, in a unit-volume of gas, there are $N_+$ ions with charge $Ze$ and $N_e$ electrons with velocities distributed according to $f(v)dv=4\pi(m/2\pi k_B T)^{3 / 2} \exp(-mv^2 / 2 k_B T) v^2 dv$, i.e.
+the \emph{Maxwell velocity distribution}.
 The energy emitted per unit-volume and unit-time in the frequency interval $\nu+d\nu$ by the electrons moving throught the field of the ions is
 \begin{equation}
-    N_+ N_e f(v') dv' v' d\epsilon_\nu(v').
+    N_+ N_e v f(v) dv d\epsilon_\nu.
     \label{eq:emission}
@@
 The energy emitted by all of the electrons is obtained by integrating from $v_{min}$ to $v=\infty$, with $v_{min}$ satisfying $1/2 m_e v^2_{min}=h\nu$.
-$v_{min}$ is the minimum velocity of an electron that is able to emit a photon.
+$v_{min}$ is the minimum velocity of an electron that won't become bound.
 Using eq.
@@
 \end{equation}
-Here, $a_\nu$ is the absorption per-ion per-electron with velocity $v$ and the factor $1-\exp(-h\nu/k_B T)$ accounts for effective decrease in absorption due to stimulated emission \cite{zel2002physics}.
-In equilibrium, which we assume in SBSL, emission is exactly cancled by absorption.
+Here, $a_\nu$ is the absorption per-ion for electonrs velocity $v$ and the factor $(1-\exp[-h\nu/k_B T])$ accounts for effective decrease in absorption due to stimulated emission \cite{zel2002physics}.
+In equilibrium, which we assume in SBSL, total emission is exactly cancled by absorption.
 Then we set eq.
@@
 
-We are finally have a complete model for SBSL!
-An RP equation, e.g.
+With this (and expressions for other absorption mechanisms \cite{an2009diagnosing,hilgenfeldt1999simple,zel2002physics}) we finally have a complete model for SBSL in water!
+
+\section{Conclusion}
+We could now perform a full calculation of the light spectrum in SBSL; let's sketch it here: an RP equation (e.g.
 eq.
-\ref{eq:RP_3}, combined with a gas-dynamical equation, e.g.
+\ref{eq:RP_3}) is combined with a gas-dynamical equation (e.g.
 eq.
-\ref{eq:p_g}, can be used to solve for the dynamics of a bubble undergoing SBSL.
-This allow's us to calculate the temperature as a function of time, $T(t)$, according to e.g.
+\ref{eq:p_g}) and the radius-time curve $R(t)$ of bubble undergoing SBSL is calculated numerically, also giving us the temperature as a function of time, $T(t)$ (e.g.
 eq.
-\ref{eq:T(t)}.
-Then eq.
-\ref{eq:saha} can be inserted into \ref{eq:absorption_coeff} and, with expressions for electron-atom bremsstrahlung \cite{geltman1973free} and electron-ion recombination \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing,zel2002physics}, eq.
-\ref{eq:intensity} can be integrated to calculate the power spectrum as a function of $T(t)$ and the power spectrum can be directly compared to experiment.
+\ref{eq:T(t)}).
+We use eq.
+\ref{eq:saha} to determing the fraction of ions and electrons present and eq.
+\ref{eq:absorption_coeff} gives us the amount of light absorbed by electron-ion bremsstrahlung, with similar expressions for electron-atom bremsstrahlung \cite{geltman1973free} and electron-ion recombination \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing,zel2002physics}.
+These equations and eq.
+\ref{eq:intensity} for the energy radiated in the bubble give us the power spectrum as a function of $T(t)$ and the power spectrum can be directly compared to experiment.
+Calculations from almost exactly these lines are how SBSL was explained in water \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,brenner2002single,yasui2018acoustic}.
 
+Assuming different absorption mechanisms for lower temperatures and directly solving the full gas dynamical NS equations for SBSL in 85\% aqueous H$_2$SO$_4$, the SBSL has also been calculated nearly perfectly \cite{an2006mechanism,an2008spectral,an2009diagnosing}.
 Early work on SBSL ignored bound-bound transitions since lines weren't observed in the SBSL spectrum in water.
 It was that they would be very broad due to the extremely high pressure in the bubble, but estimated broadening was too small to lead to what was seen in experiment and it was concluded instead that these process don't occur \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
-Later work using more accurate gas dyanmical models have shown that bound-bound transitions are important under the right conditions \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of SBSL spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
+Later work using more accurate gas dyanmical models have shown that bound-bound transitions are important too, with emission lines visible at the lower pressures in SBSL in aqueous H$_2$SO$_4$ \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of SBSL spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
 SBSL is a solved problem!
-
-\section{Summary}
 

### 2022-01-19 sonoluminescence/paper/paper.tex
sdfadsfa

@@
 
-\documentclass[prb,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
+\documentclass[11pt,prb,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2} 
 \usepackage{graphicx}
@@
 \begin{abstract}
-Snapping shrimp can produce bubbles that emit light when they collapse.
+%Snapping shrimp can produce bubbles that emit light when they collapse.
 The light emission is known as \emph{shrimpoluminescence} in the case of shrimp, but more generally when an acoustically driven bubble emits light, it is called \emph{sonoluminescence}.
@@
 Snapping shrimp, like our cute friend in fig.
-\ref{fig:shrimp}, are known to produce cavitating bubbles by snapping their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
-They have a strong appendage (the \emph{dactyl}) in thier claw (fig.
+\ref{fig:shrimp}, produce cavitating bubbles by snapping their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}.
+They have a strong appendage called the \emph{dactyl} (fig.
 \ref{fig:shrimp} [center]) that is used to create a high-velocity jet of water.
 The low pressure region in the jet's wake forms a bubble (fig.
-\ref{fig:shrimp} [right]) that produces an exceptionally loud noise when it collapses.
-The sound produced by even one shrimp's snap is detectable over a mile away \cite{everest1948acoustical}.
+\ref{fig:shrimp} [right]) that, when it collapses, produces a noise loud enough to be detected over a mile away \cite{everest1948acoustical}.
 The noise produced by groups of shrimp is so intense that the U.S.
@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{figs/shrimp.pdf}
+\includegraphics[width=0.9\linewidth]{figs/shrimp.pdf}
     \caption{(left) Snapping shrimp (\emph{Alpheus heterochaelis}).
@@
 Multi-bubble sonoluminescence (MBSL) consists of  ``the simultaneous creation and destruction of many separate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}, whereas in single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}.
-The discovery of MBSL predates SBSL by $\sim$60 years but due to the random and fleeting nature of the bubbles, SL wasn't well studied until the early 1990's when it was discovered that single bubbles could be created and periodically driven to produce light with very high precision \cite{crum1994sonoluminescence,gaitan1990experimental,gaitan1992sonoluminescence,brenner2002single}.
-In SBSL, emission from a single bubble is not complicated from scattering off other bubbles.
+
+MBSL was discovered by accident in 1933 when engineers who were trying to accelerate photo-development noticed that a photo-sensitive plate immersed in an ``insonated" fluid became foggy from exposure to light \cite{marinesco1933actions,frenzel1934luminescenz}.
+This result was not particularly important at the time since there was no practical use for the light emission and it was already known that cavitating bubbles, e.g.
+those generated by the sound field, could do trememdous damage to ship's propellers.
+The discovery's impact is summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
+The theory of caviation was already quite mature.
+Euler (as early as 1754) hypothesized that if the velocity in a fluid was large enough, negative pressures could become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence}.
+In the early 1900's, Lord Rayleigh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigorous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
+He found that the bubble wall diverges during collapse, giving rise to ``cavitation".
+Since then, the theory of \emph{bubble dyanmics} has been refined considerably and we will devote a large portion of the paper to it \cite{prosperetti1999old,plesset1977bubble,plesset1977bubble,brenner2002single,lofstedt1995sonoluminescing,barber1992resolving}.
+
+Still, very little information was accessible about the light emission until the 1990's when it was discovered that stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing cavitation.
+A stable bubble persists through multiple cycles of cavitation, oscillating non-linearly around an equilibrium size.
+Transient bubbles appear and then collapse to disappear}, single bubbles could be created and driven to emit light in SBSL \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
+SBSL, unlike previous studies on MBSL, allowed very precise control and measurement of the SL process.
+Practically all progress on understanding SL in general has been based on SBSL, with some authors even calling it ``the hydrogen atom of sonoluminescence" \cite{lohse2018bubble,crum1994sonoluminescence}.
+Light emission from an isolated bubble is not complicated from scattering off other bubbles.
 Similarly, the bubble is not perturbed by interaction with other bubbles or, due to its tiny extent, by interaction with the container walls.
-The theory is greatly simplified compared to MBSL.
-It is not an exaggeration to say that nearly all progress to explain SL has been made using SBSL, with some authors even calling it ``the hydrogen atom of SL" \cite{lohse2018bubble,crum1994sonoluminescence}.
-The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are too many to try list and we won't be interested in most of them.
+Both theory and experiments are greatly simplified compared to MBSL.
+Measurments of the SBSL spectrum resulted in a nearly featureless continuum, leading to speculation that the light emission was very high-temperature thermal radiation \cite{hiller1992spectrum,hiller1994effect}.
+Around the same time, it was found that the light pulse was orders or magnitude shorter than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
+This discovery implied that SL was nearly decoupled from the bubble's dynamics and models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubble at $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
+These very high estimates for the temperature in the bubble finally led to a spark of interest more broadly: if the temperature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur.
+An action movie featuring Keanu Reeves was made with this concept at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
+and in the early 2000's, several papers were published claiming that stable nuclear fusion was possible in the lab using a method based on SBSL \cite{}.
+Unfortunately these studies turned out to be fraud and evidence eventually emerged that shape instabilities prevent converging shock-waves, severely reducing estimates for the temperature to $\sim 10,000$ K \cite{brenner2002single}.
+Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble between $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui20185acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
+
+%The discovery of SBSL created a rush of effort to explain the phenomena with arguments ranging from the simplest models \footnote{There are too many to try list and we won't be interested in most of them.
 References can be found in Gaitan \cite{gaitan1990experimental} and Brenner et al.
@@
 An air bubble is trapped at the center of an acoustically driven flask containing a fluid.
-The pressure wave in the fluid, shown as a function of time as the top curve in fig.
+The pressure wave in the fluid, shown as a function of time in the top curve in fig.
 \ref{fig:pulses}, drives the expansion and contraction of the bubble and, in turn, the gas in the bubble responds to the bubble's oscillations.
-Under the right circumstances, we get the very complicated non-linear response of the bubble's radius shown as the middle curve in fig.
+Under the right circumstances, we get the very complicated non-linear response of the bubble's radius shown in the middle curve in fig.
 \ref{fig:pulses}.
 Now let's see how these dynamics lead to SBSL.
-
 First, as the pressure decreases and becomes negative, the bubble expands gradually: surface tension and the negative pressure in the bubble tend to slow this down.
@@
 Firstly, we will look at the problem of the fluid containing the bubble and derive the equations of motion for the bubble's wall.
-Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble, estimate the \emph{temperature} and \emph{pressure} in the gas.
-We will then discuss theories presenting the mechanisms of light emission/absorption in the bubble which, combined with the temperature, allow use to calculate the bubbles light spectrum.
+Nextly, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble, estimating the \emph{temperature} and \emph{pressure} along they way.
+Finally, we will discuss theories presenting the mechanisms of light emission/absorption in the bubble which, combined with the temperature, allow use to calculate the bubble's light spectrum.
 Throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need.
@@
 
-\subsection{Historical Overview}
-
-MBSL was discovered by accident in 1933 by Marinesco and Trillat \cite{marinesco1933actions} and was subsequently characterized by Frenzel and Schultes \cite{frenzel1934luminescenz} \footnote{Neither the paper due to Marinesco et al.
-nor the one due to Frenzel et al.
-can be found by the author; in any case he can't read French or German so having the papers wouldn't be much use.
-As such, the story of how the effect was discovered is taken from more recent sources whose authors hopefully could read French and German \cite{brenner2002single,gaitan1990experimental,crum1994sonoluminescence}} Marinesco et al.
-were trying to accelerate photo development by \emph{insonating} developing fluid.
-They discovered that a photosensitive plate immersed in the insonated fluid became ``foggy," which they attributed to exposure to light.
-Shortly after, Frenzel et al.
-repeated the experiment and confirmed that the insonated fluid emits light in the form of a faintly glowing cloud of bubbles.
-
-\begin{figure}
-\includegraphics[width=0.95\linewidth]{figs/flask.pdf}
-    \caption{(a) Schematic of a spherical ``acoustic levitation cell".
-The piezoelectric transducers that drive the bubble trapped in the flask are labeled in the diagram.
-(b) Photograph of the same.
-Here, the fluid is aqueous H$_2$SO$_4$ instead of water.
-A sonoluminescing bubble is clearly visible in the center of the flask.
-The photo was taken in a fully lit room with exposure time 2 s.
-(c) Schematic of a Mie scattering experimental setup.
-A similar levitation cell as (a) is shown with an added laser source and photomultiplier tubes (PMT) to analyze the scattered and emitted light.
-(a), (b), and (c) are from refs.
-\cite{brenner2002single}, \cite{suslick2008inside}, and \cite{gompf2000mie} respectively.}
-\label{fig:flask}
-\end{figure}
-
-This result was not particularly surprising to the community since it had been known for a while that cavitating bubbles, e.g.
-those generated by the sound field, could do tremendous damage to ship's propellers \footnote{A more modern example could be my van's oil pump, which is in need of replacing.}.
-The discovery's impact is concisely summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.
-In fact, cavitating bubbles had been of interest to engineers working on fluid mechanics for a while.
-The discovery of cavitation is credited to Euler (as early as 1754) who hypothesized that if the velocity in a fluid was large enough, negative pressures could become so large as to ``break the fluid" \cite{li2015introduction,gaitan1992sonoluminescence}.
-Cavitation was confirmed to exist (and named ``cavitation") in 1895 by engineers studying the failure of a British Navy ship's propeller \cite{li2015introduction}.
-Shortly after, Lord Rayleigh wrote down and solved the differential equation for a vapor filled cavity collapsing in water (the so-called Rayleigh equation), giving the first rigorous theoretical treatment of cavitation \cite{rayleigh1917pressure,plesset1977bubble}.
-Rayleigh found that, for a bubble at a lower pressure than the surrounding fluid (and both pressures held constant), the bubble wall diverges during collapse.
-The theory has been refined considerably since then \cite{prosperetti1999old,plesset1977bubble,plesset1977bubble,brenner2002single,lofstedt1995sonoluminescing,barber1992resolving} and we will devote a large portion of this paper to this subject.
-Still, very little information was accessible about the light emission until the 1990's when stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing cavitation.
-A stable bubble persists through multiple cycles of cavitation, oscillating non-linearly around an equilibrium size.
-Transient bubbles appear and then collapse to disappear}, single bubbles could be created and driven to emit light \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}.
-With this discovery, serious interest took root.
-
-SBSL, unlike previous studies on MBSL, allowed very precise control and measurement of the SL process.
-With unprecedented experimental control, many discrepancies in previous assumptions about SL were discovered.
-Experiments measuring the duration of the light pulse found that it was orders or magnitude smaller than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}.
-This discovery implied that SL was nearly decoupled from the bubble's dynamics.
-New models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at the center of bubbles $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}.
-These very high estimates for the temperature in the bubbles led to a spark of interest more broadly: if the temperature in the bubble were really that large, then it should be possible for nuclear \emph{fusion} to occur \cite{}.
-An action movie featuring Keanu Reeves was made with this at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.}.
-and in the early 2000's, several papers were published claiming that stable nuclear fusion was possible in the lab using a method based on SBSL \cite{}.
-Unfortunately these results turned out to be fraudulent \cite{}.
-Eventually evidenced emerged that there are shape instabilities in the bubbles dynamics prohibiting converging shock-waves and severely reducing estimates for the temperature to $\sim 10,000$ K, ruling out the possibility of fusion occurring \cite{brenner2002single}.
-Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble between $10,000-30,000$ K \cite{flannigan2005plasma,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.
-
 \section{Bubble Dynamics}
-In this section, we will present the theory needed to model and understand the dynamics of a bubble undergoing SBSL.
+In this section, we present the theory needed to model the dynamics of a bubble undergoing SBSL.
 We will split this into parts.
-Firstly, we will derive the equations of motion for the bubble's wall.
+First, we will derive the equations of motion for the bubble's wall.
 Next, we will review some important experimental developments that we will then use to argue for a simple model for the dynamics of the gas in the bubble.
@@
 \subsection{The Bubble Wall}
-It turns out the dynamics in SBSL are quite well described both qualitatively and quantitatively by \emph{the classical theory of bubble dynamics} \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
+It turns out the dynamics in SBSL are quite well described both qualitatively and quantitatively by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}.
 The starting point for the theory is Rayleigh's original work \cite{rayleigh1917pressure}, now expounded upon by many others.
 The most widely known modern work is that of Plesset \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}, resulting in the so called ``the Rayleigh-Plesset" (RP) equation.
-The derivation assumes the host liquid is incompressible, ultimately resulting in us neglecting the effects of sound radiated by the bubble.
-Corrections exist that improve upon the RP equation in various ways and we will discuss this below.
-However, we will only derive the RP equation here and merely quote the others as they are similar in form.
 The field of bubble dynamics is quite mature and attempting to review it here would be out of scope of this paper; instead, the reader is referred elsewhere \cite{prosperetti1999old,brenner2002single,yasui2018acoustic,prosperetti1986bubble}.
 
-The RP equation is deduced from the \emph{Navier-Stokes} (NS) equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
-The NS equations for a compressible fluid are
-\begin{equation}
-\begin{split}
-    \rho \frac{d \bm{u}}{dt} = \rho \left[ \partial_t \bm{u}+\left(\bm{u}\cdot \nabla \right)\bm{u} \right] & =-\nabla p + \eta \nabla \left(\nabla \cdot \bm{u} \right)+\mu \nabla^2 \bm{u} + \bm{f}_B \\
-     \partial_t \rho+\nabla\cdot(\rho \bm{u}) & = 0.
-     \label{eq:NS_equations}
-\end{split}
-\end{equation}
-
-Here, $\bm{u}(\bm{x},t)$ is the velocity of a fluid parcel at position $\bm{x}$ at time $t$, $\rho (\bm{x},t)$ is the fluid's density, $p$ is the pressure, $\eta \equiv \left(\xi + 2\mu/3 \right)$ is the \emph{bulk} viscosity with $\mu$ the \emph{shear} viscosity, and $\bm{f}_B$ are body forces (e.g.
-gravity, electrostatic forces) which are usually from external sources and will be assumed negligible from here on.
-The subscripts on $\partial$ denote what derivative is being taken and $\nabla$ and $\nabla^2$ have their usual meanings.
-The first equation is conservation of momentum for the fluid, i.e.
-Newton's $2^{nd}$ law; the second equation is mass conservation.
-The total-derivative of a moving fluid $d/dt\equiv D_t$ is sometimes called the \emph{material derivative}.
-In the case of velocity $D_t \bm{u}$ is the acceleration of the fluid.
-It records not only explicit time dependence of the fluid's velocity $\partial_t \bm{u}$ but also how the fluid's velocity varies in space as it moves past us: $\left( \bm{u}\cdot\nabla \right) \bm{u}$.
-The pressure term on the right hand side is to be understood as forces from the elastic energy.
-On the other hand, the terms $\sim \bm{u}$ are viscous (damping) terms that tend to make the velocity field spatially uniform: they vanish when the (spatial) derivatives of the velocity vanish.
-The bulk-viscosity term $\sim \eta$ damps radial changes in the fluids velocity (e.g.
-dilation/contraction) while the shear-viscosity term $\sim\mu$ resists shearing.
-
-Solving eqs.
-\ref{eq:NS_equations} is intractable.
-To make progress, we make some drastic approximations that are more-or-less valid for our problem.
-Firstly, we will assume irrational flow (i.e.
-$\nabla\times\bm{u}\equiv 0$) so that there is only radial motion in the liquid (i.e.
-$\bm{u}=u\bm{r}$).
-Since the flow is assumed irrotational, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=\partial_r \phi \bm{r}$ \cite{jackson1999classical}.
+With suitable approximations, the RP equation is deduced from the \emph{Navier-Stokes} equations \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic}.
+We assume the host liquid is incompressible, ultimately resulting in neglecting effects of sound radiated by the bubble.
+We also assume irrotational flow so that there is only radial motion in the liquid, i.e.
+$\bm{u}=u\bm{r}$ is the fluid's velocity with $u$ the speed and $\bm{r}$ the radial unit-vector.
+For irrotational flow, we can represent the velocity as the gradient of a scalar function: $u\bm{r}=(\partial_r \phi) \bm{r}$ \cite{jackson1999classical}.
 This amounts to assuming that the bubble is always spherical which seems like a rather drastic approximation but is validated experimentally.
@@
 
-Next, we assume that the viscous terms are negligible in the bulk dynamics of the liquid.
-For low viscosity fluids such as water which is also nearly incompressible, this is an accurate approximation \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
+Next, we assume that \emph{viscosity} is negligible in the bulk dynamics of the liquid which, in low viscosity fluids such as water which, is an accurate approximation \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}.
 We will account for viscosity later when looking at the bubble-liquid interface.
-We will also assume \emph{isentropic} and \emph{isothermal} flow, i.e.
+We will also assume reversible and isothermal flow, i.e.
 no damping and no heat flow between fluid parcels.
-The pressure $p$ is then determined from an instantaneous equation-of-state $p=p(\rho,T)$.
+The pressure $p$ is determined from an instantaneous equation-of-state, $p=p(\rho,T)$, with $\rho$ the density and $T$ the temperature.
 In the case of SBSL, this is valid since the bubble makes up a \emph{tiny} fraction of the total volume and any heat-transport across the liquid-bubble interfacing is negligible \cite{prosperetti1999old}.
-
-Lastly, related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
+Related to the fact that the bubble is tiny, we assume that the extent of the liquid is so large compared to the bubble that we may consider the dynamics of the liquid as if there were no bubble present; similarly, we consider the dynamics of a bubble in an infinite, isotropic medium.
 Of course the bubble-liquid interface enters both systems as a boundary condition.
 
-With these simplifying assumptions (and a vector identity \footnote{$\left( \nabla \cdot \bm{u} \right) \bm{u} = \frac{1}{2} \nabla (u^2) - \bm{u}\times \left( \nabla \times \bm{u} \right)$}) we rewrite eqs.
-\ref{eq:NS_equations} as
+With these simplifying assumptions the Navier-Stokes equations for the fluid are \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,brennen2014cavitation}
 \begin{equation}
@@
 \end{equation}
+The first equation is conservation of momentum; the second is mass-continuity.
 We can integrate the first equation using the fact that the compressibility is negligible \cite{leighton2007derivation}.
-\begin{equation}
-    \int_{r_\infty}^{r} \frac{\partial}{\partial r'} \left[ \frac{\partial \phi}{\partial t} +\frac{1}{2} \left( \frac{\partial \phi}{\partial r'} \right)^2 \right] dr' = - \frac{1}{\rho} \int_{r_\infty}^{r} \frac{\partial p}{\partial r'} dr'
-\end{equation}
 On the left side, we take the velocity to vanish at infinity.
@@
 \end{equation}
-Eq. \ref{eq:press_vel} gives the pressure at coordinate $(r,t)$ in terms of the velocity of the fluid, the density, and the applied pressure.
+%Eq. \ref{eq:press_vel} gives the pressure at coordinate $(r,t)$ in terms of the velocity of the fluid, the density, and the applied pressure.
 
-Now, let us assume that the velocity potential satisfies a wave equation $\nabla^2 \phi - (1/c^2)\partial_t^2 \phi =0 $ \footnote{This doesn't come from thin air.
+Now let us assume that the velocity potential satisfies a wave equation $\nabla^2 \phi - (1/c^2)\partial_t^2 \phi =0 $ with $c$ the speed of sound in the fluid \footnote{This doesn't come from thin air.
 We could write the r.h.s.
 of eq.
-\ref{eq:press_vel} as the enthalpy, $h\equiv \int dp/\rho$, and similarly introduce $dh=dp/\rho$ in the first line of eq.
+\ref{eq:press_vel} as the enthalpy, $h\equiv \int dp/\rho$, and similarly introduce $dh=dp/\rho$ in the second line of eq.
 \ref{eq:NS_1D}.
-Eliminating $h$ between the two and dropping small terms $\sim u/c$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.} with $c$ the speed of sound in the fluid.
+Eliminating $h$ between the two and dropping small terms $\sim u/c$, we arrive at the homogeneous wave equation \cite{leighton2007derivation,brenner2002single,prosperetti1999old}.}.
 Again, we use incompressibility by recalling that for an incompressible fluid $c\rightarrow \infty$ which implies $\nabla^2 \phi =0$ \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}.
-Then we can ignore retardation effects and write the solution of $\nabla^2 \phi$ as
-\begin{equation}
-    \phi = \frac{\psi(t)}{r}
-\end{equation}
-with $\psi$ the time-dependent coefficient. Now use the boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, to determine $\psi$:
-\begin{equation}
-    \dot{R} = -\frac{\psi}{R^2} \Rightarrow \psi = -R^2 \dot{R}.
-\end{equation}  
-We find $\phi(r,t) = -R^2 \dot{R}/r$. Plugging this into the left hand side of eq. \ref{eq:press_vel} and evaluating at the bubble wall, $r\equiv R$
-\begin{equation}
-\begin{split}
-    \partial_t \phi \vert_{r=R} & = -2\dot{R}^2-R\ddot{R} \\
-    \partial_r \phi \vert_{r=R} & = \dot{R} \equiv u(R)
-\end{split}
-\end{equation}
-Finally, sticking all this together, we arrive at the Rayleigh-Plesset equation \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,plesset1949dynamics,plesset1977bubble}
+Then we can ignore retardation effects and write the solution of $\nabla^2 \phi$ as $\phi = \psi(t)/r$ with $\psi$ a time-dependent coefficient.
+The boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, gives $\psi= -R^2 \dot{R}$ and $\phi = -R^2 \dot{R}/r$.
+Plugging this into the left hand side of eq.
+\ref{eq:press_vel} and evaluating at the bubble wall, we arrive at the Rayleigh-Plesset equation \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,plesset1949dynamics,plesset1977bubble}
 \begin{equation}
@@
 
-If we had not assumed incompressibility, eq.
-\ref{eq:RP_1} would include a term arising from the sound radiation from the bubble itself.
-There are practical issues with this solution however; it is third order in time and has unphysical, exponentially diverging solutions \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,lezzi1987bubble}.
-Both issues are solved by imposing boundary conditions on the $\ddot{R}$ term.
-The procedure is difficult and apparently no unique way to do it exists \cite{prosperetti1986bubble,prosperetti1988nonlinear,keller1956damping}.
-Instead, there is a whole family of solutions that can be derived; different variations of the RP equation, which are collectively referred to as ``Rayleigh-Plesset equations" are derived by neglecting different terms in the full compressible solution.
-We will discuss this again below.
-
-We can refine eq.
-\ref{eq:RP_1} a little further.
+Eq.
+\ref{eq:RP_1} can be refined a little further.
 Since the motion in the fluid is purely radial, we expect the only relevant stresses to be normal to the bubble wall.
-Let's denote the radial stress in the liquid at some point $r$ by $s_r(r)=-p(r,t)+2\eta \partial_r u$.
-Recall, $\eta$ is the fluid viscosity from earlier.
-Using conservation of mass and momentum, and our expression above for the velocity, we can equate the normal stresses in the fluid to those in the bubble:
-\begin{equation}
-\begin{split}
-    s_r(R)=-p_B-4\frac{\eta \dot{R}}{R}=-p_g+2\frac{\sigma}{R}
-\end{split}
-\end{equation}
-where $p_g$ is the pressure in the gas and $\sigma$ is the surface tension. Solving for $p_B(t)$, we have the relation \cite{brenner2002single,prosperetti1999old,prosperetti1986bubble}
+The normal stress combined with the pressure in the gas, $p_g(t)$, gives us \cite{brenner2002single,leighton2007derivation,prosperetti1999old,prosperetti1986bubble}
 \begin{equation}
@@
 \end{equation}
-Note that in deriving the equations \ref{eq:NS_1D} we neglected the viscosity terms.
-This was valid since only combinations of viscosity and compressibility $\nabla \cdot \bm{u}$ appeared, both of which are assumed small.
-Here, however, we keep the viscosity terms coupling the bubble wall to the fluid.
-The RP equation becomes
+The term $\sim\sigma$ is called the \emph{Laplace pressure} and is from surface tension. The $\sim\eta$ term is due to the \emph{bulk viscosity} in the fluid. The RP equation becomes
 \begin{equation}
@@
 It is no great challenge computationally.
-For now, let's see what we else we can understand from this analytically by making some more approximations.
-If we look at the interval where the bubble wall's velocity is very large, called \emph{The Rayleigh Collapse}, the right hand side of eq.
+For now, let's see what we can understand analytically by making some more approximations.
+If we linearize in the bubble's radius $R(t)\approx R_0 \left( 1+x(t) \right)$ with $x(t) \ll 1$, we can derive a forced harmonic oscillator equation for $x(t)$ \cite{brennen2014cavitation,yasui2018acoustic}.
+However the low driving amplitude regime where this method is valid is irrelevant to SBSL so we won't pursue it here.
+Instead, we can look at the interval where the bubble wall's velocity is very large, called \emph{The Rayleigh Collapse}; the right hand side of eq.
 \ref{eq:RP_2} will be much smaller than the $\dot{R}^2$ term and we may neglect it \cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}.
-Rayleigh looked at the same problem, i.e. a cavitating void in water:
+Rayleigh looked at the same problem, i.e.
+a cavitating void in water:
 \begin{equation}
@@
 \ref{eq:RP_1}) the only term capable of compensating the bubble wall's motion is the diverging pressure inside the bubble itself \cite{brenner2002single}.
-Applying the same analysis to the family of solutions derived when including the bubble's sound radiation, it turns out that the most important term for slowing the bubble wall's diverging velocity is $\propto\dot{p}_g$.
-Keeping only this term, we arrive at the most popular variant of the RP equation \cite{lofstedt1995sonoluminescing,barber1997defining}:
-\begin{equation}
-    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ \Delta p(t)-\frac{1}{R}\left( 2\sigma+4\eta \dot{R} \right)-P(t) +\frac{R}{c} \dot{p}_g \right].
-    \label{eq:RP_3}
-\end{equation}
+If we had not assumed incompressibility, eq.
+\ref{eq:RP_1} would include a term arising from the sound radiation from the bubble itself.
+A family of solutions, collectively called ``Rayleigh-Plesset equations," are derived in this way \cite{prosperetti1986bubble,prosperetti1988nonlinear,keller1956damping,lezzi1987bubble}.
+It turns out that the most important term for slowing the bubble wall's diverging velocity is $\propto\dot{p}_g$.
+Keeping only this term provides the most popular variant of the RP equation \cite{lofstedt1995sonoluminescing,barber1997defining}.
 This equation is only slightly more complicated than the RP equation (eq.
@@
 
+Above, we assumed that the bubble always remains spherical.
+Of course this isn't always true and \emph{shape-instabilities} occur for the right combinations of mechanical properties and driving force/frequency.
+Usually shape-instabilities destroy the bubble and kill SBSL so we won't focus on these isues here.
+We will always assume we are in a parameter-regime where SBSL occurs.
+The reader is referred to one of several reviews on the \emph{phase-space} of SBSL for more info \cite{yasui2018acoustic,brenner2002single,hilgenfeldt1999sonoluminescence,an2009diagnosing}.
+
 \subsection{The Bubble's Interior}
@@
 (ii) We can model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{hiller1992spectrum,brenner2002single}.
-Obviously these are related topics, but the distinction in the context of SBSL is important as we will now see.
+%Obviously these are related topics, but the distinction in the context of SBSL is important as we will now see.
 
@@
 These data are then used to try to describe the light emission.
-This was preferred over studying the emitted spectra directly as careful measurements resulted in an almost featureless spectrum that could not be explained \cite{hiller1992spectrum}.
-Calculating the dynamics of the bubble's wall from the RP equations (eq.
-\ref{eq:RP_2} or eq.
-\ref{eq:RP_3}) requires the pressure inside the bubble as input.
-We would like to find a suitable form for $\Delta p(t) \sim p_g(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating bubble.
+Calculating the dynamics of the bubble's wall from one of the RP equations (e.g.
+eq.
+\ref{eq:RP_2}) requires the pressure inside the bubble as input.
+We would like to find a suitable form for $p_g(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating bubble.
 It turns out that this is a very complicated problem: gas diffusion and rectification (one-directional diffusion) between the bubble and liquid vary the number of particles present and, to make things worse, the conditions inside the bubble facilitate chemical reactions between the air and water-vapor changing the properties of gas dynamically \cite{brenner2002single}.
@@
 The other method to understand the bubble's interior is its light spectrum.
-For the first decade or so, SBSL experiments were done using partially degassed\footnote{Degassing lowers the ambient pressure, making it easier for bubbles to form.} water as the liquid medium \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}.
+For the first decade or so, SBSL experiments were done using water as the liquid medium \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}.
 Careful measurements of the emission from SBSL in water resulted in an almost featureless spectrum and attempts to fit it as black-body radiation were not sucessful \cite{hiller1992spectrum}.
-It was later discovered that SBSL is volume emission, indicating that a black-body is not valid anyway \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
 Little progress was made on this front for quite some time \cite{brenner2002single}.
@@
 Importantly, they were able to identify spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma.
-Perhaps equally as crucially, they found the absence of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble gas, experimentally confirming the argon rectification hypothesis.
-Moreover, the spectra at different pressures were in very good agreement with detailed calculations of emission from noble gas plasmas \cite{an2009diagnosing,an2008spectral}.
-Increased pressure results in increased collision rate between particles, broadening the excited state transition spectral lines \cite{an2008spectral,suslick2008inside,flannigan2005plasma,flannigan2006measurement}.
-Further calculations showed that the greatly increased viscosity and density of aqueous H$_2$SO$_4$, $\sim 10 \times$ and $\sim 1.3 \times$ larger than that that of water, lead to reduced (by orders of magnitude) pressure in the bubble at the same driving pressures \cite{an2009diagnosing}.
-The larger pressure in the bubbles in water was found to be consistent with the featureless emission spectrum \cite{suslick2008inside,yasui2018acoustic}.
+Perhaps equally as crucially, they found the absence of emission lines from components of aqueous H$_2$SO$_4$ vapor \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble-gas, experimentally confirming the argon rectification hypothesis.
+The spectra at different pressures were in very good agreement with detailed calculations of emission from noble-gas plasmas \cite{an2009diagnosing,an2008spectral}.
+This work also explained why the spectrum in water is featureless (i.e.
+has no emission lines).
+The larger pressure in SBSL in water results in an increased collision rate between particles, broadening the spectral lines \cite{an2009diagnosing,suslick2008inside,flannigan2005plasma,flannigan2006measurement}.
 
-This argon rectification hypothesis greatly helped to refine models for the gas dynamics in the bubble and clarified several unexplained experimental observations \cite{suslick2008inside,brenner2002single}.
+This argon rectification hypothesis greatly helped to refine models for the gas dynamics in the bubbles and confirmed why simple models usually work well \cite{suslick2008inside,brenner2002single}.
 With this in mind, we choose only to focus on results known to be relevant to the dynamics of a bubble undergoing SBSL.
 Excellent books and modern reviews on modeling the gas dynamics exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.
-We will hold off on discussing the dynamics in the plasma until a later section.
+We will hold off on discussing the plasma until a later section.
 
 \subsection{Gas Dynamics}
-The most straightforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
-In fact, serious quantitative predictions of the conditions of the bubble's interior take this path \cite{an2006mechanism,an2008spectral,an2009diagnosing,flannigan2005plasma,flannigan2006measurement}.
-With simplifying assumptions, equations of motion with varying degrees of sophistication can be derived for the gas and the system of gas dynamical equations and the RP equations are solved numerically.
-The discussion proceeds in much the same way as deriving the RP equations above but is considerably more complicated \cite{brenner2002single,yasui2018acoustic}, so we will omit it here.
-An important aspect of these studies is the understanding of why relatively simple models work quite well.
-We will highlight the salient features before introducing a simple model for the gas that will enable us to gain some physical understanding analytically.
-An especially important result of direct numerical solution of the gas dynamics was (theoretical) confirmation of the argon rectification hypothesis \cite{}, reducing the possible gas dynamical models to those applicable to a monatomic gas.
-Various other models fully including dissipation and heat transport into/out of the bubble showed that the bubble's compression is, to a good approximation, adiabatic and quasistatic \cite{}.
-Before we turn to discuss a model for the gas dynamics, let us pause to mention that another theoretically appealing method for modeling the gas dynamics is \emph{molecular dynamics} \cite{}.
+Of course the most straightforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}.
+In fact, serious quantitative predictions of the conditions of the bubble's interior still take this path \cite{an2006mechanism,an2008spectral,an2009diagnosing,flannigan2005plasma,flannigan2006measurement}.
+Equations of motion with varying degrees of sophistication are derived for the gas and the system of gas dynamical equations and the RP equations are solved numerically.
+The discussion proceeds in much the same way as deriving the RP equations above but is slighlty more complicated \cite{brenner2002single,yasui2018acoustic}, so we will omit it.
+Instead, let us look at a simple model and discuss how it can be extended to more acurately represent the dynamics.
+
+%Salient features before introducing a simple model for the gas that will enable us to gain some physical understanding analytically.
+An especially important result of direct numerical solution of the gas dynamics was (theoretical) confirmation of the argon rectification hypothesis, reducing the possible gas dynamical models to those applicable to a monatomic gas.
+Various other models fully including dissipation and heat transport into/out of the bubble showed that the bubble's compression is, to a good approximation, adiabatic and quasistatic.
+Before we turn to discuss a model for the gas dynamics, let us pause to mention that another theoretically appealing method for modeling the gas dynamics is \emph{molecular dynamics}.
 While this method is attractive for a number of reasons, practical calculations containing realistic numbers of particles aren't possible, limiting the applicability \cite{brenner2002single}.
 
-A simple but reasonable model for the gas dynamics in the bubble is that of an adiabatically, quasistatically compressed ideal gas \cite{brenner2002single} \footnote{Actually we want to study a \emph{Van der Walls} gas, but we will sneak this aspect in later.
-The modifications to the result for an ideal gas are simple \cite{sivasubramanian2002temperature}.}.
-Let us remind ourselves what the pressure is in this case.
-Suppose that we have a volume, $V$, of an ideal gas containing $N$ particles at a temperature $T$.
-The ideal gas law tells us the equation of state is $PV=Nk_BT$ with $k_B$ Boltzmann's constant.
-Since we are assuming adiabatic compression, the second law of thermodynamics tells us that $dU=-PdV$ with $U$ the internal energy in the gas.
-Moreover, we may use the equipartition theorem to write $dU=N f k_B dT/2$ with $f$ the number of degrees-of-freedom per particle \cite{schroeder1999introduction}.
-Recall that for an ideal gas $C_V = N f k_B/2$, $C_P = C_V+N k_B$, and $C_P/C_V = 1+2/f \equiv \gamma$ with $C_V$ and $C_P$ the heat capacities at constant volume and pressure respectively.
-For a monatomic ideal gas, $f=3$ and $\gamma=5/3$.
-Now we can use the second law of thermodynamics to write $C_V dT = -P dV$.
-From the equation of state, we have
-\begin{equation}
-    dT = \frac{PdV+VdP}{Nk_B}=\frac{PdV+VdP}{C_P-C_V}.
-\end{equation}
-Eliminating $dT$ and simplifying, 
-\begin{equation}
-    \frac{dP}{P}=-\left(\frac{C_P}{C_V}\right) \frac{dV}{V} = -\gamma \frac{dV}{V}.
-    \label{eq:dP}
-\end{equation}
-Integrating eq. \ref{eq:dP}, we find 
+Recalling that the bubble is (almost) purely noble-gas, a simple but reasonable model for the gas dynamics, at least during the collapse, is that of an adiabatically, quasistatically compressed Van der Walls gas \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward,hilgenfeldt1999simple}.
+In the context of SBSL, the interaction term in the Van der Walls (VdW) equation of state is usually neglected and the only modification to the ideal gas result is the excluded volume of the real gas molecules, $V_h = 4 \pi h^3/3$, with $h$ the VdW \emph{hard-core radius}.
+The pressure in an adiabatically compressed \emph{ideal} gas at volume $V_2$ is \cite{schroeder1999introduction}
 \begin{equation}
@@
 \end{equation}
-where $P_2$ is the pressure in a gas at volume $V_2$ that was initially contained in a volume $V_1$ at pressure $P_1$. Importantly we have not made any assumptions about the shape of the container.
-
-Let us now connect this to the problem of SBSL.
+where $P_1$ is the initial pressure at volume $V_1$ and $\gamma=C_P/C_V$ is the ratio of the constant-pressure and constant-volume specific heats respectively.
 In our case, we assume that our bubble has an ambient radius $R_0$ where the pressure is at its equilibrium value, $p_0$.
-The volume occupied by the gas is $V_0 =  4/3 \pi R_0^3$.
+The ambient volume of the bubble is $V_0 =  4 \pi R_0^3/3$.
 We know that $p_0=p_\infty+2\sigma /R$ from earlier.
-What we want to calculate is the pressure in the gas as a function of the bubble's instantaneous radius $R(t)$ determined from the RP equations (eq.
-\ref{eq:RP_2} or eq.
-\ref{eq:RP_3}).
-Earlier we called the pressure in the gas $p_g(t)$ and we will stick with this here.
+We want to calculate the pressure in the gas, $p_g(t)$, as a function of the bubble's instantaneous radius $R(t)$ determined from an RP equation (e.g.
+eq.
+\ref{eq:RP_2}).
 Plugging these into eq.
-\ref{eq:P}, (and sneaking in a modification) we arrive at a very commonly used equation in the context of SBSL \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward,hilgenfeldt1999simple}:
+\ref{eq:P}, and subtracting the excluded volume, we arrive at a very commonly used equation in the context of SBSL \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward,hilgenfeldt1999simple,sivasubramanian2002temperature}:
 \begin{equation}
@@
 \end{equation}
-The new variable $h$ that was added willy-nilly is called the \emph{Van der Walls hard-core radius} and its purpose is to make sure that, if the bubble collapses so strongly that its contents become incompressible (i.e.
+It's clear that the purpose of the excluded volume is to make sure that if the bubble collapses so strongly that its contents become incompressible (i.e.
 $R(t)\rightarrow h$), the pressure diverges \cite{lofstedt1993toward,brenner2002single}.
 Eq.
-\ref{eq:p_g} can be inserted into an RP equation which can then be integrated numerically to determine the bubble's dynamics.
+\ref{eq:p_g} can be inserted into an RP equation which can then be integrated numerically to determine the bubble's complete dynamics for all time.
 
@@
 \ref{eq:p_g} is sensible during the bubble's collapse: the bubble wall's velocity is fast and very little heat can flow out during compression.
-Detailed calculations starting from the full NS equations further tell us that the gas dynamics in the bubble can be regarded as \emph{quasistatic} \cite{}.
-During the expansion and after-bounce stages of the bubbles cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal}.
+During the expansion and after-bounce stages of the bubbles cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal} \cite{brenner2002single}.
 This is true because, when the bubble's velocity is on the order of the heat diffusion timescale, the temperature throughout the bubble is nearly equal to that in the liquid \cite{prosperetti1999old,brenner2002single,yasui2018acoustic}.
@@
 \ref{eq:p_g}, between its adiabatic and isothermal values, $C_P/C_V$ and $1$, respectively.
-A summary of this work is given in refs.
-\cite{brenner2002single} and \cite{prosperetti1999old}.
-It turns out that this process doesn't result in better quantitative predictions since, during the assumed isothermal part of the process, dissipation becomes important \cite{brenner2002single}.
-A simple alternative is choosing a constant exponent somewhere between the adiabatic $\gamma=5/3$ and isothermal $\gamma=1$ values \cite{hilgenfeldt1999simple}.
-It is also possible to include heat and mass diffusion in simple model forms and these do give slightly improved quantitative agreement with experiment; however we choose to discuss solutions of the RP equations using the simpler form given in eq.
-\ref{eq:p_g} as this will enable to understand the important features of the bubble's cycle.
-More advanced solutions using the full gas dynamical equations are qualitatively similar \cite{brenner2002single,yasui2018acoustic}.
+This technique has been commonly used in calculations of SBSL in water \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,prosperetti1986bubble,brenner2002single}.
+These simple forms, even using the constant adiabatic exponent $\gamma=5/3$, give close to the correct bubble dynamics (see fig.
+\ref{fig:bubble_radius_1}) with more advanced solutions using the full gas dynamical equations improving only on estimates for the conditions inside the bubble \cite{brenner2002single,yasui2018acoustic}.
 
-Now recall that one of main goals of modeling the bubble dynamics in SBSL was to calculate the temperature inside the bubble in order to model the light emission.
-Following similar steps to deriving $p_g(t)$, we can write down an equation for the temperature $T$ \cite{barber1997defining,brenner2002single}:
+Recalling that one of main goals of modeling the bubble dynamics in SBSL was to calculate the temperature inside the bubble, we can also derive an an equation for $T(t)$ \cite{barber1997defining,brenner2002single,sivasubramanian2002temperature}:
 \begin{equation}
@@
 \end{equation}
-Earlier, we assumed that during the bubble's after-bounces and expansion, the gas dynamics were isothermal.
-Then we take $T_0$ to be the temperature of the liquid.
-The RP equation, eq.
-\ref{eq:RP_3}, combined with the gas dynamical equation, eq.
-\ref{eq:p_g}, and the known liquid and gas properties $\eta$, $\sigma$, and $h$, can be used to calculate the dynamics of an acoustically driven bubble, determining the time-dependence of the pressure and temperature in the bubble along the way.
+Earlier, we said that during the bubble's after-bounces and expansion, the gas dynamics is isothermal; then we take $T_0$ to be the temperature of the liquid.
+An RP equation, e.g.
+eq.
+\ref{eq:RP_2}, combined with the gas dynamical equation, eq.
+\ref{eq:p_g}, and the known liquid and gas properties $\eta$, $\sigma$, and $h$, can be used to calculate the dynamics of an acoustically driven bubble, giving us the time-dependence of the pressure and temperature in the bubble.
 
 \subsection{Bubbles in the Lab}
+
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{figs/flask.pdf}
+    \caption{(a) Schematic of a spherical ``acoustic levitation cell".
+The piezoelectric transducers that drive the bubble trapped in the flask are labeled in the diagram.
+(b) Photograph of the same.
+Here, the fluid is aqueous H$_2$SO$_4$ instead of water.
+A sonoluminescing bubble is clearly visible in the center of the flask.
+The photo was taken in a fully lit room with exposure time 2 s.
+(c) Schematic of a Mie scattering experimental setup.
+A similar levitation cell as (a) is shown with an added laser source and photomultiplier tubes (PMT) to analyze the scattered and emitted light.
+(a), (b), and (c) are from refs.
+\cite{brenner2002single}, \cite{suslick2008inside}, and \cite{gompf2000mie} respectively.}
+\label{fig:flask}
+\end{figure}
+
+Creating stable, single bubbles in a laboratory experiment is not particularly challenging and can be done with standard and low-cost materials.
+A typical experimental setup is shown in fig.
+\ref{fig:flask}.
+Since we are concerned with explaining the light emission process itself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}.
+A suitable sample of liquid is placed into a flask.
+Stuck to the outside are piezoelectric transducers driven sinusoidally to excite an acoustic resonance of the flask: a typical flask is a few cm across with a resonance at $\approx 20$ kHz \cite{brenner2002single}.
+The frequency is chosen to form a pressure \emph{antinode} at the center of the flask, trapping a bubble there.
+
+%Let us take a brief aside: in deriving the RP equations, we assumed the center of the bubble was stationary.
+Actually, this is what we want in an SBSL experiment and it is worth discussing under what conditions this is true.
+In an ideal, incompressible fluid the net force acting on a bubble may be written as
+%\begin{equation}
+%    \bm{f}=-\int_{\partial \Omega} p \bm{n} dA = -\int_\Omega \nabla p dV
+%\end{equation}
+%Here, $\bm{f}$ is the net force, $p$ is the pressure at the bubble's surface, $\bm{n}$ is the unit normal vector point out of the bubbles surface, and $\partial \Omega$ is the bubble's boundary with $\Omega$ the bubble's domain.
+The last term is due to an application of Gauss's theorem.
+Bjerknes \cite{bjerknes1909kraftfelder} took $\nabla p \equiv \nabla p(r=0,t)$ to be constant and averaged over one period of oscillation to find
+%\begin{equation}
+%    \bm{f} = -\frac{4}{3}\pi \langle R^3 \nabla p \rangle.
+%    \label{eq:Bj_force}
+%\end{equation}
+%This is the so called \emph{Bjerknes force}.
+The point is that if the bubble is located at a pressure extrema, e.g.
+the antinode at the center of the flask, then this force should vanish and the bubble will remain stationary, an essential result for SBSL experiments.
+
+%Some authors have looked at corrections to the bubble's dynamics in the presence of buoyancy, which cause translational motion even for vanishing Bjerknes force.
+The result is the bubble oscillates around an equilibrium position near but away from the center of the flask \cite{matula2000single,matula1997bjerknes,matula1999inertial}.
+The translational motion of the bubble leads to shape distortion which is believed to suppress SL; this is confirmed by SBSL experiments in microgravity resulting in larger intensity light production \cite{matula2000single}.
+
+The driving pressures relevant to SBSL ($\sim 1.5$ atm) are too small to cause bubbles to form spontaneously \cite{brenner2002single}.
+Instead, a bubble is usually \emph{seeded} somehow.
+In the original work of Gaitan, an air bubble was injected using a syringe \cite{gaitan1992sonoluminescence}.
+More recently, seeding methods involve shooting a small jet of water into the flask (like our shrimp friends!) or blasting the water with a laser to boil a small volume \cite{suslick2008inside,yasui2018acoustic}.
+The later method is preferred as it enables more precise control over the initial bubble's radius.
+
+Now let's discuss two experimentally accessible quantities that have turned out the be important.
+(i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations, allowing us to infer $T(t)$ by fitting to experiment \cite{barber1992light,barber1997defining,brenner2002single}.
+(ii) The the light can be analyzed with a spectrometer, giving us information on the bubble's contents: in principle, we can learn about temperature and pressure inside the bubble, with spectral lines telling us about matter \cite{hiller1992spectrum,hilgenfeldt1999simple,flannigan2005plasma,flannigan2006measurement}.
+We will look at this in the next section.
+
 \begin{figure}
 \includegraphics[width=0.9\linewidth]{figs/bubble_radius_1.pdf}
-    \caption{\cite{barber1997defining,barber1992light}}
+    \caption{Numerical solution of the $\propto \dot{p}_g$ RP equation variant discussed earlier with the gas dynamics given by the simple form in eq.
+\ref{eq:p_g}.
+The dots are experimental results from Mie scattering \cite{barber1997defining,barber1992light}.}
 \label{fig:bubble_radius_1}
@@
 
-Creating stable, single bubbles in a laboratory experiment is not particularly challenging and can be done with standard and low-cost materials \footnote{Actually, the phase-space for SBSL is ``small".
-The coordinates being driving amplitude, driving frequency, fluid composition, temperature, ambient pressure, gas content and composition, etc...
-What we are claiming is simple is the experimental setup once the correct parameters are identified.}.
-A typical experimental setup is shown in Fig.
-\ref{fig:flask}.
-Since we are mainly concerned with the explaining the light emission process itself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}.
-A sample of degassed liquid is placed into a flask \footnote{Actually, the degassing part is quite important.
-We focus on it elsewhere.}.
-Coupled to the outside of the flask are piezoelectric transducers of some sort.
-The piezos are driven by a sinusoidal function generator to excite an acoustic resonance of the flask: a typical flask is a few cm across with a resonance at $\approx 20$ kHz \cite{brenner2002single}.
-The frequency is chosen to form a pressure \emph{antinode} at the center of the flask.
-
-Let us take a brief aside: in deriving the RP equations, we assumed the center of the bubble was stationary.
-Actually, this is what we want in an SBSL experiment and it is worth discussing under what conditions this is true.
-In an ideal, incompressible fluid the net force acting on a bubble may be written as
-\begin{equation}
-    \bm{f}=-\int_{\partial \Omega} p \bm{n} dA = -\int_\Omega \nabla p dV
-\end{equation}
-Here, $\bm{f}$ is the net force, $p$ is the pressure at the bubble's surface, $\bm{n}$ is the unit normal vector point out of the bubbles surface, and $\partial \Omega$ is the bubble's boundary with $\Omega$ the bubble's domain.
-The last term is due to an application of Gauss's theorem.
-Bjerknes \cite{bjerknes1909kraftfelder} took $\nabla p \equiv \nabla p(r=0,t)$ to be constant and averaged over one period of oscillation to find
-\begin{equation}
-    \bm{f} = -\frac{4}{3}\pi \langle R^3 \nabla p \rangle.
-    \label{eq:Bj_force}
-\end{equation}
-This is the so called \emph{Bjerknes force}.
-The point is that if the bubble is located at a pressure extrema, e.g.
-the antinode at the center of the flask, then this force should vanish and the bubble will remain stationary, an essential result for SBSL experiments.
-
-Some authors have looked at corrections to the bubble's dynamics in the presence of buoyancy, which cause translational motion even for vanishing Bjerknes force.
-The result is the bubble oscillates around an equilibrium position near but away from the center of the flask \cite{matula2000single,matula1997bjerknes,matula1999inertial}.
-The translational motion of the bubble leads to shape distortion which is believed to suppress SL; this is confirmed by SBSL experiments in microgravity resulting in larger intensity light production \cite{matula2000single}.
-
-Back to the experimental setup.
-The typical driving pressure relevant to SBSL ($\sim 1.5$ atm) is too small to cause bubbles to form spontaneously in the flask \cite{}.
-Instead, the bubble is usually seeded somehow.
-In the original work of Gaitan, an air bubble was injected using a needle and syringe through a small hole in the flask's seal \cite{gaitan1992sonoluminescence}.
-More recently, seeding methods involve shooting a small jet of water into the flask (like our shrimp friends!) or blasting the water with a laser to boil a small volume \cite{}.
-The later method is preferred as it enables more precise control over the initial bubble's radius.
-The small bubbles formed this way coalesce at the center of the flask into a single, stable bubble \cite{}.
-
-Now that we know about how single, stable bubbles are formed in the lab, we can discuss two experimentally accessible quantities that have turned out the be the most important: (i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations \cite{}.
-This can tell us about the temperature and pressure in the bubble.
-(ii) The other quantity of interest is the light spectrum which contains information on the bubble's contents: in principle, we can learn about temperature and pressure inside the bubble, with spectral lines telling us about the matter content \cite{}.
-As we will see, it is not simple to understand the light spectrum \cite{} and we will devote a section of this paper to this topic.
-For now, let us see what measurements of the bubble's dynamics have told us.
-
-Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
-Assuming the bubble is a homogeneous dielectric sphere \cite{}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{}.
+%Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
+Assuming the bubble is a homogeneous dielectric sphere \cite{jackson1999classical}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{brenner2002single}.
 A typical Mie scattering experiment is shown in fig.
@@
 Both effects are largest during the SL period of the bubble's cycle, leading to an important source of error when fitting the calculated $R(t)$ curve \cite{}.
-Parameters used to calculate the gas dynamics in the bubble can then be fit to the measured radius to build a full model of the dynamics.
-In the case of eq.
-\ref{eq:p_g}, the quantity to fit is $R_0$.
-The results of a calculation using eq.
-\ref{eq:p_g} combined with the RP variant eq.
-\ref{eq:RP_3} and fit to experiment are shown in fig.
-\ref{fig:bubble_radius_1} (a).
-The ambient ambient radius is $R_0=1.45$ $\mu$m for driving pressure $P_0=1.45$ atm \cite{barber1997defining,barber1992light}.
-Note the large change in the calculated dynamics for relatively small changes in $R_0$ and $P_0$ in Fig.
-\ref{fig:bubble_radius_1} (b) and (c).
-
+Mie scattering measurements scatter a laser off a bubble and analyze the scattered intensity.
+Assuming the bubble is a homogeneous dielectric sphere \cite{jackson1999classical}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{brenner2002single}.
+A typical Mie scattering experiment is shown in fig.
+\ref{fig:flask} (c).
+Parameters used to calculate the gas dynamics in the bubble are fit to the measured radius to build a full model of the dynamics.
+This procedure allows us to ``measure" the temperature of a bubble undergoing SBSL.
 More accurate treatments of the gas dynamics from numerically solving the gas's NS equations are qualitatively in good agreement with these simpler model calculations \cite{brenner2002single,yasui2018acoustic}.
@@
 \end{equation}
-The subscript $B$ is to remind us this is for a black-body which assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out.
-Then the only emission that is measured comes from a thin layer on the bubble's surface.
+The subscript $B$ is to remind us this is for a black-body; this is an \emph{optically thick} thick bubble with ``perfect" emission from the bubble's surface.
 Calculations of SBSL assuming black-body radiation over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}.
-Instead, we need to consider finite opacity of the bubble.
+Instead, we need to consider that the bubble has finite opacity.
 
-Let's call the absorption coefficient $\kappa(\lambda,T(t))$\footnote{Really, this is an \emph{effective} absortion coeffecient $\kappa = \kappa'(1-\exp\left[ hc/k_B T \lambda \right])$ with $\kappa'$ the actual absoption coefficient of the material.
-The effective coefficient $\kappa$ includes induced emission of photons as light passes through the medium which 'decreases' absorption.
-Physically, this can be viewed as radiation that is absorbed being immediately re-emitted or, equivalently, vice-versa \cite{zel2002physics,taylor1969experimental}.} where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$.
-Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics,hilgenfeldt1999simple}
+Let's call the absorption coefficient $\kappa(\lambda,T(t))$ where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$.
+Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics,hilgenfeldt1999simple,taylor1969experimental}
 \begin{equation}
@@
 
-We are finally have a complete model for SBSL!
-An RP equation, e.g.
+With this (and expressions for other absorption mechanisms \cite{an2009diagnosing,hilgenfeldt1999simple,zel2002physics}) we finally have a complete model for SBSL in water!
+
+\section{Summary and Outlook}
+We could now perform a full calculation of the light spectrum in SBSL; let's sketch it here: an RP equation (e.g.
 eq.
-\ref{eq:RP_3}, combined with a gas-dynamical equation, e.g.
+\ref{eq:RP_3}) is combined with a gas-dynamical equation (e.g.
 eq.
-\ref{eq:p_g}, can be used to solve for the dynamics of a bubble undergoing SBSL.
-This allow's us to calculate the temperature as a function of time, $T(t)$, according to e.g.
+\ref{eq:p_g}) and the radius-time curve $R(t)$ of bubble undergoing SBSL is calculated numerically, also giving us the temperature as a function of time, $T(t)$ (e.g.
 eq.
-\ref{eq:T(t)}.
-Then eq.
-\ref{eq:saha} can be inserted into \ref{eq:absorption_coeff} and, with expressions for electron-atom bremsstrahlung \cite{geltman1973free} and electron-ion recombination \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing,zel2002physics}, eq.
-\ref{eq:intensity} can be integrated to calculate the power spectrum as a function of $T(t)$ and the power spectrum can be directly compared to experiment.
+\ref{eq:T(t)}).
+We use eq.
+\ref{eq:saha} to determing the fraction of ions and electrons present and eq.
+\ref{eq:absorption_coeff} gives us the amount of light absorbed by electron-ion bremsstrahlung, with similar expressions for electron-atom bremsstrahlung \cite{geltman1973free} and electron-ion recombination \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing,zel2002physics}.
+These equations and eq.
+\ref{eq:intensity} for the energy radiated in the bubble give us the power spectrum as a function of $T(t)$ and the power spectrum can be directly compared to experiment.
+Calculations from almost exactly these lines are how SBSL was explained in water \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,brenner2002single,yasui2018acoustic}.
 
+Assuming different absorption mechanisms for lower temperatures and directly solving the full gas dynamical NS equations for SBSL in 85\% aqueous H$_2$SO$_4$, the SBSL has also been calculated nearly perfectly \cite{an2006mechanism,an2008spectral,an2009diagnosing}.
 Early work on SBSL ignored bound-bound transitions since lines weren't observed in the SBSL spectrum in water.
 It was that they would be very broad due to the extremely high pressure in the bubble, but estimated broadening was too small to lead to what was seen in experiment and it was concluded instead that these process don't occur \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.
-Later work using more accurate gas dyanmical models have shown that bound-bound transitions are important under the right conditions \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of SBSL spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
+Later work using more accurate gas dyanmical models have shown that bound-bound transitions are important too, with emission lines visible at the lower pressures in SBSL in aqueous H$_2$SO$_4$ \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of SBSL spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.
 SBSL is a solved problem!
-
-\section{Summary}
 

### 2022-01-31 sonoluminescence/paper/bak/v7/paper.tex
some bug fixes

\documentclass[11pt,prb,aps,nofootinbib,superscriptaddress,floatfix]{revtex4-2}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natmove}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{bm}

\begin{document}

\title{The Mechanism of Shrimpoluminescence}

\author{Tyler C. Sterling}
\email{ty.sterling@colorado.edu}
%\affiliation{Department of Physics, University of Colorado at Boulder, Boulder, Colorado 80309, USA}

\date{\today}

\begin{abstract}
Snapping shrimp produce bubbles that emit light when they collapse. When a bubble collapses so strongly that it emits light, the light emission is usually called sonoluminescence. In the case of the shrimp, the light emission is called \emph{shrimpoluminescence}. The bubble collapses so fast that no heat can escape and the gas trapped in the bubble becomes hot enough to ionize. Light is emitted through electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination. In this paper, we study the dynamics of a sonoluminescing bubble and learn how to calculate the spectrum of emitted light, allowing us to explain the physical mechanisms of shrimpoluminescence.
\end{abstract}

\maketitle

\section{Introduction}

Snapping shrimp, like our cute friend in Fig. \ref{fig:shrimp}, produce cavitating bubbles by snapping their claws \cite{versluis2000snapping,lohse2001snapping,tang2019bioinspired}. They have a strong appendage called the \emph{dactyl} [Fig. \ref{fig:shrimp} (center)] that is used to create a high-velocity jet of water. The low pressure region in the jet's wake forms a bubble [Fig. \ref{fig:shrimp} (right)] that, when it collapses, produces a noise loud enough to be detected over a mile away \cite{everest1948acoustical}. The sound wave produced by the collapsing bubble is used to stun or kill prey \cite{versluis2000snapping}. If the shrimp's prey had very sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect known as ``shrimpoluminescence" in the case of the shrimp \cite{lohse2001snapping}, but more generally called \emph{sonoluminescence}.
%The noise produced by groups of shrimp is so intense that the U.S. Navy used them as ``sonar-camouflage" in the Pacific ocean during World War II \cite{versluis2000snapping}. The shrimp were not patriots helping the war-effort however; they snapped for food. The sound wave produced by the cavitating bubble is used to stun or kill prey \cite{versluis2000snapping}. If the shrimp's prey had very sensitive eyes (and also were not dead) they might notice a flash of light is also produced through an effect referred to as ``shrimpoluminescence" in the case of the shrimp \cite{lohse2001snapping}, but more generally known as \emph{sonoluminescence}.

\begin{figure}
\includegraphics[width=0.85\linewidth]{figs/shrimp.pdf}
    \caption{(left) Snapping shrimp (\emph{Alpheus heterochaelis}). (center) Blown-up view of the shrimp's claw. The \emph{plunger} (pl) on the \emph{dactyl} (d) rapidly enters the \emph{socket} (s), ejecting a high-velocity jet of water. A time near the bubble collapse is shown on the right. The light emission is too dim to be seen by the naked eye. Adapted from refs. \cite{versluis2000snapping} and \cite{lohse2001snapping}.}
\label{fig:shrimp}
\end{figure}

Sonoluminescence (SL) is defined as the process by which a ``driven gas bubble collapses so strongly that the energy focusing at collapse leads to light emission" \cite{brenner2002single}. SL comes in two forms: (i) single-bubble sonoluminescence and (ii) multi-bubble sonoluminescence. Multi-bubble sonoluminescence (MBSL) consists of  ``the simultaneous creation and destruction of many separate, individual cavitation bubbles" \cite{crum1994sonoluminescence,brenner2002single}. In single-bubble sonoluminescence (SBSL), rather obviously, only a single bubble is present \cite{gaitan1992sonoluminescence}. In contrast to MBSL, there are no other bubbles present for the emitted light to scatter from. Similarly, the bubble does not interact with other bubbles and, due to its tiny extent, does not interact with the container walls. Both theory and experiments are greatly simplified compared to MBSL. Practically all progress on understanding SL has come from SBSL, with some authors even calling it ``the hydrogen atom of sonoluminescence" \cite{lohse2018bubble,crum1994sonoluminescence}.

With this in mind, let's use SBSL as an idealization of shrimpoluminescence. The discussion will be much simpler if we further specialize to \emph{stable} SBSL. In stable SBSL the bubble is driven to produce light but does not disappear after collapsing; rather, it is periodically driven and emits light every cycle. This is in contrast to shrimpoluminescence, where the bubble is transient; i.e. it disappears after emitting light. The significance of this contradiction is addressed later.

In a SBSL\footnote{From here on,``SBSL" means \emph{stable} SBSL unless explicitly specified.} experiment, a bubble filled with noble-gas is trapped at the center of a flask containing water [Fig. \ref{fig:summary_fig} (a)]. A transducer is connected to the flask and is tuned to excite one of its vibrational modes. The flask's oscillations induce oscillating pressure in the fluid, $P(t)$, shown as a function of time in Fig. \ref{fig:summary_fig} (b). The oscillating pressure drives the bubble's expansion/contraction. For just the right driving pressure, we get the complicated behavior of the bubble's radius, $R(t)$, shown in Fig. \ref{fig:summary_fig} (b). To understand how these dynamics lead to SBSL, let's look at the blown-up snapshots of the bubble at the times $t_1$, $t_2$ and $t_3$:
\begin{itemize}
    \item [$\bm{t}$=$\bm{t_1}$] The driving pressure in the fluid is negative and the bubble is expanding. Surface-tension opposes the driving pressure, slowing the bubble's expansion.
    \item [$\bm{t}$=$\bm{t_2}$] The driving pressure is now positive and the bubble's radius is very large. The gas is dilute and offers little resistance to compression. Surface-tension works with the positive driving pressure to shrink the bubble: the bubble violently collapses in a process called \emph{cavitation}. 
    \item [$\bm{t}$=$\bm{t_3}$] The bubble is at its minimum radius. During cavitation, the gas was compressed so quickly that heat could not flow out of the bubble. For a very short time ($ 0.003\% $ of the bubble's cycle \cite{suslick2008inside}), the temperature in the bubble can reach $10,000-30,000$ K \cite{an2006mechanism,hilgenfeldt1999simple,an2008spectral,an2009diagnosing}. This is when light is emitted from the bubble. After the light pulse, there is a period of ``after-bouncing" until the driving pressure becomes negative and the cycle begins again.
\end{itemize}

\begin{figure}
\includegraphics[width=0.9\linewidth]{figs/summary_fig}
    \caption{Schematic of the physics of SBSL. (a) Diagram of an SBSL experiment with a zoomed-in region around the bubble at different times shown to the right. A transducer, labeled as ``T", drives the flask and bubble. (b) Time-dependence of the driving pressure $P(t)$, bubble radius $R(t)$, and number of photons emitted $I(t)$, during SBSL. $P_0$ is the driving pressure amplitude, $R_0$ is the ``ambient" radius of the bubble, and $I_0$ is the peak number of photons emitted during the light pulse. (c) Inside the bubble during the light pulse at $t=t_3$. Electrons, ions, and atoms are labeled in the legend. The mechanisms of light emission are (i) electron-ion bremsstrahlung, (ii) electron-ion recombination, and (iii) electron-atom bremsstrahlung. }
\label{fig:summary_fig}
\end{figure}

The interior of the bubble is depicted at $t=t_3$ in Fig. \ref{fig:summary_fig} (c). When the temperature is $\sim 10,000$ K, about $1 \%$ of the atoms in the gas ionize \cite{hilgenfeldt1999simple,yasui1999mechanism}, creating pairs of free electrons and positively charged ions. When a free electron passes near an ion, it is accelerated by the Coulomb force. The electron's kinetic energy changes and the lost energy is converted into radiation \cite{zel2002physics,griffiths2005introduction,jackson1999classical}. This process is called \emph{electron-ion bremsstrahlung} (``bremsstrahlung" means ``braking-radiation") and is labeled as (i) in Fig. \ref{fig:summary_fig} (c). Similarly, if an electron and an ion combine to form a neutral atom, light is emitted as their energies change: \emph{electron-ion recombination} is labeled as (ii). 
The Coulomb force is long-ranged so even at low densities of electrons and ions, electron-ion bremsstrahlung is important \cite{zel2002physics}. Still, most of the particles are neutral atoms. Ordinarily, electron-atom scattering would be irrelevant since the electron-neutral particle interaction is very short-ranged. However, the pressure and density in the bubble at maximum compression are so large that electrons frequently collide with atoms, emitting radiation through \emph{electron-atom bremsstrahlung}. This mechanism is labeled as (iii) in Fig. \ref{fig:summary_fig} (c). Electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination are the most important mechanisms that produce light during SBSL \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,an2006mechanism,an2008spectral,an2009diagnosing,suslick2008inside}.

The goal of the rest of this paper is to present the theory needed to understand shrimpoluminescence in detail, though we will mainly focus on SBSL \footnote{It is apparently simpler to characterize SBSL without involving shrimp; convincing the shrimp to snap requires tickling them \cite{lohse2001snapping,versluis2000snapping,lohse2018bubble}. Besides notable exceptions \cite{tang2019bioinspired}, practically all work on SBSL has not involved shrimp.}. First, we will look at the problem of the fluid containing the bubble and derive the equations of motion for the bubble's wall. Next, we will use our results for the bubble's wall to understand the dynamics of the gas trapped in the bubble, calculating the temperature along the way. Finally, we will see how knowing the mechanisms of light emission/absorption in the bubble allow us to calculate the wave-length dependence of the bubble's light spectrum. Throughout the course of our journey, we will stick to the simplest results that still contain the essential physics we need. Connections to more advanced treatments, and their implications, are provided for completeness.

%MBSL was discovered by accident in 1933 when engineers trying to accelerate photo-development noticed that a photo-sensitive plate immersed in an ``insonated" fluid became foggy from exposure to light \cite{marinesco1933actions,frenzel1934luminescenz}. This wasn't ground-breaking since there was no practical use for the light emission and it was already known that cavitating bubbles, e.g. those generated by the sound field, could do tremendous damage to ship's propellers. The discovery's impact is summarized by Brenner: ``if the cloud [of cavitating bubbles] collapses violently enough to break molecular bonds in a solid, why should it \emph{not} emit photons" \cite{brenner2002single}.

%Still, little information was accessible about the light emission until the 1990's when it was discovered that stable \footnote{\emph{Stable} is compared with \emph{transient} when discussing cavitation. A stable bubble persists through multiple cycles of cavitation, oscillating non-linearly around an equilibrium size. Transient bubbles appear and then collapse to disappear} single bubbles could be created and driven to emit light in SBSL \cite{gaitan1990experimental,gaitan1992sonoluminescence,crum1994sonoluminescence}. SBSL, unlike previous studies on MBSL, allowed very precise control and measurement of the SL process. Practically all progress on understanding SL has come from SBSL, with some authors even calling it ``the hydrogen atom of sonoluminescence" \cite{lohse2018bubble,crum1994sonoluminescence}. In SBSL, light emission from an isolated bubble is not complicated by scattering off other bubbles. Similarly, the bubble is not perturbed by interaction with other bubbles or, due to its tiny extent, by interaction with the container walls. Both theory and experiments are greatly simplified compared to MBSL.

%SBSL still remained mysterious for some time, however. Measurements of the light spectrum from SBSL resulted in a nearly featureless continuum [Fig. \ref{fig:exp}, (bottom left)], leading to speculation that the light emission was very high-temperature thermal radiation \cite{hiller1992spectrum,hiller1994effect}. Around the same time, it was found that the light pulse was orders or magnitude shorter than the time in which the bubble was compressed to its smallest radius \cite{barber1992resolving,barber1991observation}. This discovery implied that SL was nearly decoupled from the bubble's dynamics and models were proposed based on \emph{converging shock waves} at the bubble's center with estimates of the temperature at $\sim10^8$ K \cite{wu1993shock,greenspan1993sonoluminescence}. These very high estimates for the temperature in the bubble finally led to a spark of interest more broadly: if the temperature were really that large, then it should be possible for nuclear \emph{fusion} to occur. An action movie featuring Keanu Reeves was made with this concept at the center of the plot \footnote{It is called \emph{Chain Reaction} and I am unwilling to watch it.} and in the early 2000's several papers were published claiming that nuclear fusion was possible in the lab using methods based on SBSL \cite{taleyarkhan2002evidence,taleyarkhan2004additional}. Unfortunately these studies turned out to be fraud \cite{reich2009bubble} and evidence eventually emerged that bubble distortions prevent converging shock-waves, severely reducing estimates for the temperature \cite{brenner2002single}. Current estimates for the temperature, which we will study in the rest of this paper, put the center of a sonoluminescing bubble between $10,000-30,000$ K \cite{flannigan2005plasma,flannigan2005plasma1,suslick2008inside,yasui2018acoustic,an2009diagnosing,an2008spectral,an2006mechanism}.

\section{Bubble Dynamics}

Lord Rayleigh solved the problem of a vapor filled cavity collapsing in water (the so-called Rayleigh collapse), giving the first rigorous theoretical treatment of cavitation in the early 1900's \cite{rayleigh1917pressure,plesset1977bubble}. He found that the bubble wall's velocity diverges during collapse, giving rise to ``cavitation." Since then, the theory of \emph{bubble dynamics} has been refined considerably and we will devote a large portion of the paper to it \cite{prosperetti1999old,plesset1977bubble,plesset1977bubble,brenner2002single,lofstedt1995sonoluminescing,barber1992resolving}. We split this is split into parts. First, we derive the equations of motion for the bubble's wall. Then we review some important experimental developments that allow us to use a simple model for the dynamics of the gas in the bubble. Finally, we compare results of simulated bubble dynamics to experimental data. The connection to experiment is what allows us to calculate the temperature in real bubbles.

\subsection{The Bubble Wall}
It turns out the dynamics in SBSL are quite well described both qualitatively and quantitatively by the classical theory of bubble dynamics \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,plesset1977bubble,suslick2008inside,yasui2018acoustic,brennen2014cavitation}. The most widely known modern work is that of Plesset \cite{plesset1977bubble,plesset1949dynamics,prosperetti1986bubble}, resulting in the so called ``Rayleigh-Plesset" (RP) equation which we derive in this section. The field of bubble dynamics is quite mature and is too large to review here; instead, the reader is referred elsewhere \cite{prosperetti1999old,brenner2002single,prosperetti1986bubble,brennen2014cavitation}.

We start with some assumptions. We assume the host liquid is incompressible, ultimately resulting in neglecting effects of sound radiated by the bubble. We also assume irrotational flow so that there is only radial motion in the liquid, i.e. $\bm{u}=u\bm{r}$ is the fluid's velocity with $u$ the speed and $\bm{r}$ the radial unit-vector. Then we can represent the velocity as the gradient of a scalar function: $u\bm{r}=(\partial_r \phi) \bm{r}$ \cite{jackson1999classical}. This amounts to assuming that the bubble is always spherical which seems like a rather drastic approximation but is validated experimentally \cite{prosperetti1999old,brenner2002single}.  %The fact that the bubble tends to remain spherical is understood by accounting for surface-tension at the liquid-bubble interface \cite{prosperetti1999old}.

Next, we assume that viscosity is negligible in the bulk dynamics of the liquid. In nearly incompressible, low-viscosity fluids such as water, this is a good approximation \cite{prosperetti1986bubble,prosperetti1999old,brenner2002single,plesset1977bubble}. We will account for viscosity later when looking at the bubble-liquid interface. Furthermore, since the bubble ($\sim\mu$m) is tiny compared to the flask ($\sim$ cm), we treat the dynamics of the liquid as if there is no bubble present. Similarly, we imagine the bubble is in an infinite, isotropic medium. The bubble-liquid interface enters both systems as a boundary condition later.

With these simplifying assumptions, the Navier-Stokes equation for momentum conservation in the fluid is \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,brennen2014cavitation}
\begin{equation}
\begin{split}
     \partial_r \left[ \partial_t \phi +\frac{1}{2} \left( \partial_r \phi \right)^2 \right] & = - \frac{1}{\rho} \partial_r p  \\ 
     %\partial_t \rho+ \partial_r \phi \partial_r \rho + \rho \partial^2_r \phi & = 0.
     \label{eq:NS_1D}
\end{split}
\end{equation}
We can integrate this using the fact that the compressibility is negligible \cite{leighton2007derivation}. On the left side, we take the velocity to vanish at infinity. On the right side, we assume that only the static and driving pressures, $p_\infty$ and $P(t)$ respectively, are relevant so that \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation} 
\begin{equation}
    \partial_t \phi + \frac{1}{2}\left( \partial_r \phi \right)^2 = \frac{(p_\infty+P(t))-p(t,r)}{\rho}.
    \label{eq:press_vel}
\end{equation}

The velocity potential, $\phi$, approximately satisfies the wave equation $\nabla^2 \phi - (1/c^2)\partial_t^2 \phi =0 $ with $c$ the speed of sound in the fluid (this is exact in the limit $u/c\rightarrow 0$) \cite{leighton2007derivation,brenner2002single,prosperetti1999old}. We use incompressibility again by recalling that for an incompressible fluid $c\rightarrow \infty$. Then $\nabla^2 \phi =0$ and we can ignore retardation effects, writing the solution of $\nabla^2 \phi$ as $\phi = \psi(t)/r$ with $\psi$ a time-dependent coefficient \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation}. The boundary condition for the velocity at the bubble wall, $u(R)\equiv \dot{R}$ with $R$ the bubble radius, gives $\psi= -R^2 \dot{R}$ and $\phi = -R^2 \dot{R}/r$. Plugging this into the left hand side of eq. \ref{eq:press_vel} and evaluating at the bubble wall, we arrive at the Rayleigh-Plesset equation \cite{prosperetti1999old,prosperetti1986bubble,leighton2007derivation,plesset1949dynamics,plesset1977bubble}
\begin{equation}
    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{p_B(t)-(p_\infty+P(t))}{\rho}
    \label{eq:RP_1}
\end{equation}
with $p_B \equiv p(r=R,t)$ the pressure in the fluid at the bubble wall. The combination $p_\infty+P(t)$ is the ambient pressure \cite{prosperetti1999old,prosperetti1986bubble} which gives the pressure in the absence of the bubble. The left hand side of eq. \ref{eq:RP_1} may be regarded as the kinetic energy (density). The right hand side is the change in enthalpy (density) due to the fluctuating bubble, i.e. the dynamics of the bubble given on the left hand side are determined by the enthalpy in the fluid due to the bubble's motion. In deriving this equation, we have assumed that the speed of propagation in the fluid is infinite. In reality, this is not true but this approximation will be very good close to the bubble (in the \emph{near-field}) where retardation effects are negligible anyway. In a similar way, we have assumed that the energy in the fluid due to distorting its volume is negligible. This approximation will be accurate near the bubble as well since the kinetic energy in this region will dominate \cite{prosperetti1999old}.

Eq. \ref{eq:RP_1} can be refined a little further. Since the motion in the fluid is purely radial, we expect the only relevant stresses to be normal to the bubble wall. Balancing these stresses at the bubble wall gives \cite{brenner2002single,leighton2007derivation,prosperetti1999old,prosperetti1986bubble}
\begin{equation}
    p_B(t)=p_g(t)-\frac{1}{R}\left( 2\sigma+4\eta \dot{R} \right)
    \label{eq:p_B}
\end{equation}
with $p_g(t)$ the pressure in the gas at the bubble wall. The term $\sim\sigma$ is called the \emph{Laplace pressure} and is from surface tension at the bubble-liquid interface. The $\sim\eta$ term is due to the \emph{bulk viscosity} in the fluid. The surface tension and viscosity coefficients, $\sigma$ and $\eta$ respectively, are assumed known. The RP equation becomes
\begin{equation}
    R\ddot{R}+\frac{3}{2}\dot{R}^2 = \frac{1}{\rho} \left[ p_g(t)-p_\infty-\frac{1}{R}\left( 2\sigma+4\eta \dot{R} \right)-P(t) \right]
    \label{eq:RP_2}
\end{equation}
With the above assumption that the compression is adiabatic, the conditions inside the bubble are decoupled from the liquid (besides the coupling through the RP equation) and we temporarily consider $p_g(t)$ to be a known quantity. $P(t)$, the time-dependent part of the pressure in the absence of the bubble, is from an external source. In SBSL, the sources are the transducers driving the flask and $P(t)$ is an acoustic plane wave. Keller and coworkers showed that the modification required to the RP equation is the replacement $P(t) \rightarrow P_0 \sin(\omega t)$ \cite{keller1980bubble}. $P_0$ is the driving pressure amplitude and $\omega$ is the driving frequency.

%\begin{figure}
%\includegraphics[width=0.5\linewidth]{figs/bubble_radius_2.pdf}
%    \caption{Solutions to eq. \ref{eq:RP_2} for different driving pressures. For driving at 1 atm, the dynamics are a simple pulsating bubble. For larger driving pressure, e.g. 1.2 atm, we observe strongly non-linear behavior exhibiting cavitation and afterbounces. For even larger driving pressure, we encounter the regime where SBSL begins to occur. From ref. \cite{brenner2002single}}
%\label{fig:bubble_radius_2}
%\end{figure}

The RP equation (eq. \ref{eq:RP_2}) is pretty but its analytical solution is intractable. Direct solutions of the RP equation or its variants are done numerically using Euler or Runge-Kutta methods \cite{yasui2018acoustic,yasui2015dynamics}. It is no great challenge computationally. For now, let's see what we can understand analytically by making some more approximations. We look at the interval where the bubble wall's velocity is very large, called \emph{the Rayleigh collapse}. Here, the right hand side of eq. \ref{eq:RP_2} will be much smaller than the $\dot{R}^2$ term and we may neglect it \cite{rayleigh1917pressure,plesset1949dynamics,prosperetti1999old,brenner2002single}. Rayleigh looked at the same problem:
\begin{equation}
    \ddot{R}=-\frac{3}{2} \frac{\dot{R}^2}{R}.
    \label{eq:Rayleigh}
\end{equation}
The right hand side of eq. \ref{eq:Rayleigh} is always positive, so the acceleration is always negative. The bubble wall's \emph{speed} gets larger, causing larger negative acceleration, which increases the speed... in this way, the bubble collapses. Eq. \ref{eq:Rayleigh} can be solved and the result is $R(t)\sim\left(1-t/\tau \right)^{2/5}$, with $\tau$ satisfying $R(\tau)=0$; the velocity, $\dot{R}\sim\left(1-t / \tau \right)^{-3/5}$, diverges. These observations are what lead to Rayleigh's argument for the origin of cavitation damage to ship's propellers. %Yasui gave a rather elegant heuristic explanation for the origin of \emph{cavitation} in a fluid that is worth stating here \cite{yasui2018acoustic}. Consider the mass flowing through concentric shells around a bubble, $4\pi R_i^2 v_i \rho$. For a source free field, we require that neighboring shells satisfy conservation of mass: $4\pi R_1^2 v_1 = 4\pi R_2^2 v_2$. For $R_2 > R_1$, we see that $v_1/v_2 = \left(R_2/R_1\right)^2$. The velocity for a smaller radius is always larger, diverging towards the origin.

In reality, the bubble's velocity is always finite and in the case of \emph{stable} SBSL, so is the radius. In the full RP equation (eq. \ref{eq:RP_1}) the only term capable of compensating the bubble wall's motion is the diverging pressure inside the bubble \cite{brenner2002single}. If we had not assumed incompressibility, eq. \ref{eq:RP_1} would also include terms from the sound radiated by the bubble. A family of solutions, collectively called ``Rayleigh-Plesset equations," are derived in this way \cite{prosperetti1986bubble,prosperetti1988nonlinear,keller1956damping,lezzi1987bubble}. It turns out that the most important term for slowing the bubble wall's diverging velocity is $\propto\dot{p}_g$. Keeping only this term results in another popular variant of the RP equation that is only slightly more complicated than eq. \ref{eq:RP_2} \cite{lofstedt1995sonoluminescing,barber1997defining}. The observed error between this equation and experiment is only significant shortly after the Rayleigh collapse and it is quantitatively in good agreement with measurements of the bubble's radius during the rest of the cycle \cite{brenner2002single}.

Above, we assumed that the bubble always remains spherical. Of course this isn't always true and \emph{shape-instabilities} occur for the right combinations of mechanical properties and driving force/frequency. Usually shape-instabilities destroy the bubble and kill SBSL so we won't focus on these issues here. We will always assume we are in a parameter-regime where SBSL occurs. The reader is referred to one of several reviews on the \emph{phase-space} of SBSL for details \cite{yasui2018acoustic,brenner2002single,hilgenfeldt1999sonoluminescence,an2009diagnosing}.

\subsection{The Bubble's Interior}
Progress on understanding the bubble's \emph{interior} since the discovery of SBSL in the 1990's \cite{gaitan1992sonoluminescence} follows two paths \cite{brenner2002single,suslick2008inside,yasui2018acoustic}: (i) model calculations based on the RP equations and an equation of state for the pressure (from which we may calculate the temperature) are used to try to reproduce the easily measurable dynamics of the bubble. (ii) We model the light emission itself and compare our calculation to spectroscopic measurements of SBSL \cite{hiller1992spectrum,brenner2002single}. 
%Obviously these are related topics, but the distinction in the context of SBSL is important as we will now see.

Most early progress on the bubble's interior depended on the former method: if we predict the right dynamics, then we might know the correct pressure and temperature in the bubble. These data are then used to try to describe the light emission. Calculating the dynamics of the bubble's wall from one of the RP equations (e.g. eq. \ref{eq:RP_2}) requires the pressure inside the bubble as input. We need a suitable form for $p_g(t)$ in the RP equations that reproduces the measured radius-time curve $R(t)$ for a stable cavitating bubble. It turns out that this is a very complicated problem: gas diffusion between the bubble and liquid varies the number of particles present and, to make things worse, the conditions inside the bubble facilitate chemical reactions between the air and water-vapor, changing the properties of gas dynamically \cite{brenner2002single}. Brenner et. al elegantly summarized the significance of this problem \cite{brenner2002single}: ``one of the exciting features of modern research on SBSL is that it is a testing ground for how well mathematical models can deal with such a complicated situation."

\begin{figure}
\includegraphics[width=0.9\linewidth]{figs/exp.pdf}
    \caption{(left) Emission spectra as a function of wave-length, $\lambda$, for Ar and Xe bubbles in water and 85\% aqueous H$_2$SO$_4$. The pressure was optimized to give maximum intensity of emitted light. (right) Emission spectra as a function of wave-length, $\lambda$, for an Ar bubble in 85\% aqueous H$_2$SO$_4$ at different driving pressures, $P_0$. The spectral lines around $\sim 800$ nm are from Ar $4s\leftrightarrow4p$ transitions and disappear with increasing driving pressure. From ref. \cite{suslick2008inside}}
\label{fig:exp}
\end{figure}

The other way to understand the bubble's interior is its light spectrum. For the first decade or so, SBSL experiments were mainly done with \emph{air} bubbles in water \cite{suslick2008inside,brenner2002single,gaitan1992sonoluminescence}. Careful measurements of light emission from SBSL in water resulted in an almost featureless spectrum (see Fig. \ref{fig:exp}) and attempts to fit it as black-body radiation were not successful \cite{hiller1992spectrum}. Little progress was made on this front for quite some time \cite{brenner2002single}. The utility of analyzing the light spectrum of SBSL in water is succinctly described by Suslick \cite{suslick2008inside}: ``because of the inherent ambiguity associated with the analysis of featureless spectra of unknown origin, a more rigorous explanation [of SBSL] is unlikely to be generated".

Some attempts were made to measure SBSL using non-aqueous liquids (e.g. alcohol, silicone oil) but experiments were not very successful \cite{weninger1995sonoluminescence,barber1997defining,barber1997defining}. Eventually, different gases were tried in water. Air is $\sim 80\%$ N$_2$, $\sim 20\%$ O$_2$, and $\sim 1\% $ Ar so, naturally, pure O$_2$ or N$_2$ were tried first. Degassed water \emph{regassed} with N$_2$, O$_2$, or even a mixture of O$_2$ and N$_2$ did not to produce SBSL \cite{hiller1994effect}. It was discovered that a noble-gas was required for SBSL to occur \cite{barber1997defining,hiller1994effect,brenner2002single}. This resulted in the \emph{argon rectification hypothesis}, which claims that all species in the air inside the bubble besides Ar are gradually ejected until all that remains is pure Ar \cite{lohse1997sonoluminescing,brenner2002single,yasui2018acoustic,suslick2008inside}. The hypothesis is based on the fact that, at the high temperature inside the bubble, dissociation of O$_2$ and N$_2$ is possible. These species react with dissociated water vapor radicals to form new species that are soluble in water. As the pressure becomes very large during the compression stage of SBSL, the soluble materials leave the bubble and do not re-enter since their solubility in water is enormous compared to the Ar content of the bubble \cite{lohse1997sonoluminescing}. Over many cycles, the contents of an air bubble in water become nearly pure Ar. It was also realized that SL would be much more intense if the contents of the bubble are a pure inert gas: if the contents are e.g. molecules, the specific heat of the gas is larger, resulting in lower temperature and decreased light production from the bubble \cite{brenner2002single,yasui2018acoustic,lohse2018bubble,suslick2008inside}. This mechanism is used to explain the relatively weak light observed in MBSL: the bubbles are transient and cannot eject a significant amount of O$_2$ or N$_2$ over a single cycle. The current belief is that the contents of a bubble undergoing SBSL in water are (nearly) pure Ar after $\sim 10^3$ cycles \cite{brenner2002single,yasui2018acoustic}.

Most studies continued to use water as the host liquid until it was realized that SBSL in aqueous H$_2$SO$_4$ produces light $10^3$ times brighter than in water, allowing more precise measurements of the light spectrum \cite{flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement}. The mechanical properties in aqueous H$_2$SO$_4$ allow stable SBSL with larger bubble radii than in water, increasing the emitting volume. More importantly, new measurements revealed the phase space of SBSL in aqueous H$_2$SO$_4$ included a much larger range of driving pressure than water [Fig. \ref{fig:exp} (right)]. SBSL in aqueous H$_2$SO$_4$ could be driven at very low pressure, leading to much lower temperature bubbles, while driving at large pressure led to conditions similar to SBSL in water.

Careful experiments revealed that the spectrum depended critically on the noble-gas content and driving pressure. Suslick et. al measured the emission spectra from Ar, Xe, and Kr bubbles in aqueous H$_2$SO$_4$ \cite{flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement,suslick2008inside}. At low driving pressure, they identified spectral lines from Ar$^+$, Xe$^+$, and Kr$^+$ excited state transitions, proving that the core of the bubble is plasma. Perhaps equally as crucially, they found the absence of emission lines from components of aqueous H$_2$SO$_4$ vapor at large driving pressure \cite{suslick2008inside,flannigan2006measurement,flannigan2006measurement}: the plasma contained only the noble-gas, experimentally confirming the argon rectification hypothesis. This work also explained why the spectrum in water is featureless; the large pressure in SBSL in water results in an increased collision rate between particles, broadening the spectral lines until they are no longer visible  \cite{an2009diagnosing,suslick2008inside,flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement}.

The argon rectification hypothesis greatly helped to refine models for the gas dynamics in the bubbles and confirmed why simple models usually work well \cite{suslick2008inside,brenner2002single}. With this in mind, we will only focus on a simple model for the gas dynamics in this paper. Modern reviews on modelling the bubble's interior in SBSL exist elsewhere \cite{brenner2002single,yasui2018acoustic,brennen2014cavitation}.

\subsection{Gas Dynamics}
The most straightforward way to model the gas dynamics in the bubble is through direct solution of the Navier-Stokes equations for the gas \cite{brenner2002single}. In fact, the most accurate quantitative predictions of the conditions of the bubble's interior take this path \cite{an2006mechanism,an2008spectral,an2009diagnosing,flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement}. Equations of motion with varying degrees of sophistication are derived and the coupled system of gas dynamical and RP equations are solved numerically. The discussion proceeds in much the same way as deriving the RP equations above but is considerably more complicated since we have to account for mass and heat transfer across the bubble wall \cite{brenner2002single,yasui2018acoustic}. Instead, we will look at a simple model and discuss how it can be extended to more accurately represent the dynamics.

Recalling that the bubble is (almost) purely noble-gas, a simple but reasonable model for the gas dynamics, at least during the collapse, is that of an adiabatically, quasistatically compressed Van der Waals gas \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward,hilgenfeldt1999simple}. In the context of SBSL, the interaction term in the Van der Waals (VdW) equation of state is usually neglected and the only modification to the ideal gas result is the excluded volume of the real gas molecules, $V_h = 4 \pi h^3/3$, with $h$ the VdW \emph{hard-core radius}. The pressure in an adiabatically compressed \emph{ideal} gas at volume $V_2$ is \cite{schroeder1999introduction}
\begin{equation}
    P_2 = P_1 \left(\frac{V_1}{V_2}\right)^\gamma
\label{eq:P}
\end{equation}
where $P_1$ is the initial pressure at volume $V_1$ and $\gamma=C_P/C_V$ is the ratio of the constant-pressure and constant-volume specific heats respectively. In our case, we assume that our bubble has an ambient radius, $R_0$, where the pressure is at its equilibrium value, $p_0$. The ambient volume of the bubble is $V_0 =  4 \pi R_0^3/3$. It follows from our discussion of the RP equation that $p_0=p_\infty+2\sigma /R_0$. We want to calculate the pressure in the gas, $p_g(t)$, as a function of the bubble's radius, $R(t)$, determined from the RP equation. Plugging these data into eq. \ref{eq:P}, and subtracting the excluded volume, we arrive at a very commonly used equation in the context of SBSL \cite{brenner2002single,lofstedt1995sonoluminescing,barber1997defining,lofstedt1993toward,hilgenfeldt1999simple,sivasubramanian2002temperature}:
\begin{equation}
    p_g(t) = \left( p_\infty+2\frac{\sigma}{R_0} \right) \left[ \frac{R_0^3-h^3}{R^3(t)-h^3} \right] ^ \gamma
    \label{eq:p_g}
\end{equation}
It's clear that the purpose of the excluded volume is to make sure that if the bubble collapses so strongly that its contents become incompressible (i.e. $R(t)\rightarrow h$), the pressure diverges \cite{lofstedt1993toward,brenner2002single}.

Physical intuition tells us why eq. \ref{eq:p_g} is sensible during the bubble's collapse: the bubble wall's velocity is fast and very little heat can flow out during compression. During the expansion and after-bounce stages of the bubble's cycle, which comprise its majority, the gas dynamics should instead be regarded as \emph{isothermal} \cite{brenner2002single}. This is true because, when the bubble's velocity is comparable to the heat diffusion timescale, the temperature throughout the bubble is nearly equal to that in the liquid \cite{prosperetti1999old,brenner2002single,yasui2018acoustic}. The relevant modification to eq. \ref{eq:p_g} is replacing $\gamma\rightarrow 1$. A neat way of including both isothermal and adiabatic gas dynamics is continuously varying the exponent, $\gamma$ in eq. \ref{eq:p_g}, between its adiabatic and isothermal values, $C_P/C_V$ and $1$, respectively. This technique has been commonly used in calculations of SBSL in water \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,prosperetti1986bubble,brenner2002single}. Simple forms, e.g. eq. \ref{eq:p_g} with the constant adiabatic exponent $\gamma=5/3$, give close to the correct bubble dynamics with the most obvious error occurring during the isothermal after-bounces (see Fig. \ref{fig:mie}).

%more advanced solutions of the gas Navier-Stokes equations improving only on estimates for the conditions inside the bubble \cite{brenner2002single,yasui2018acoustic}.

Recalling that one of main goals of modeling the bubble dynamics in SBSL was to calculate the temperature inside the bubble, we can also derive an equation for $T(t)$ \cite{barber1997defining,brenner2002single,sivasubramanian2002temperature}:
\begin{equation}
    T(t) = T_0 \left( \frac{R_0^3-h^3}{R^3(t)-h^3} \right)^ {\gamma-1}
    \label{eq:T(t)}
\end{equation}
We said that for most of the bubble's cycle, the gas dynamics is isothermal: $T_0$ is the temperature of the liquid. The only ``unknown" quantity still preventing us from simulating the bubble's dynamics and calculating the temperature is $R_0$ in eqs. \ref{eq:p_g} and \ref{eq:T(t)}. We learn how to determine $R_0$ in the next section.

\subsection{Bubbles in the Lab}

%\begin{figure}
%\includegraphics[width=0.9\linewidth]{figs/flask.pdf}
%    \caption{(a) Schematic of a spherical ``acoustic levitation cell". The piezoelectric transducers that drive the bubble trapped in the flask are labeled in the diagram. (b) Photograph of a sonoluminescing bubble in 85\% aqueous H$_2$SO$_4$. (c) Schematic of a Mie scattering experiment. Panels (a), (b), and (c) are from refs. \cite{brenner2002single}, \cite{suslick2008inside}, and \cite{gompf2000mie} respectively.}
%\label{fig:flask}
%\end{figure}

Creating stable, single bubbles in an experiment is not particularly challenging and can be done with standard and low-cost materials. A typical experimental setup is shown in Fig. \ref{fig:summary_fig} (a). Since we are concerned with explaining the light emission process itself, we will only sketch the experimental setup; it can be summarized as follows \cite{lentz1995mie,gaitan1990experimental,gaitan1992sonoluminescence,gompf2000mie,brenner2002single,yasui2018acoustic,brennen2014cavitation,suslick2008inside}. A suitable sample of liquid is placed into a flask. Stuck to the outside are piezoelectric transducers driven sinusoidally to excite a resonance of the flask: a typical flask is a few cm across with a resonance at $\approx 20$ kHz \cite{brenner2002single}. The frequency is chosen to form a pressure \emph{antinode} at the center of the flask, trapping a bubble there.

%Let us take a brief aside: in deriving the RP equations, we assumed the center of the bubble was stationary. Actually, this is what we want in an SBSL experiment and it is worth discussing under what conditions this is true. In an ideal, incompressible fluid the net force acting on a bubble may be written as 
%\begin{equation}
%    \bm{f}=-\int_{\partial \Omega} p \bm{n} dA = -\int_\Omega \nabla p dV
%\end{equation}
%Here, $\bm{f}$ is the net force, $p$ is the pressure at the bubble's surface, $\bm{n}$ is the unit normal vector point out of the bubbles surface, and $\partial \Omega$ is the bubble's boundary with $\Omega$ the bubble's domain. The last term is due to an application of Gauss's theorem. Bjerknes \cite{bjerknes1909kraftfelder} took $\nabla p \equiv \nabla p(r=0,t)$ to be constant and averaged over one period of oscillation to find 
%\begin{equation}
%    \bm{f} = -\frac{4}{3}\pi \langle R^3 \nabla p \rangle.
%    \label{eq:Bj_force}
%\end{equation}
%This is the so called \emph{Bjerknes force}. The point is that if the bubble is located at a pressure extrema, e.g. the antinode at the center of the flask, then this force should vanish and the bubble will remain stationary, an essential result for SBSL experiments.

%Some authors have looked at corrections to the bubble's dynamics in the presence of buoyancy, which cause translational motion even for vanishing Bjerknes force. The result is the bubble oscillates around an equilibrium position near but away from the center of the flask \cite{matula2000single,matula1997bjerknes,matula1999inertial}. The translational motion of the bubble leads to shape distortion which is believed to suppress SL; this is confirmed by SBSL experiments in microgravity resulting in larger intensity light production \cite{matula2000single}.

The driving pressures relevant to SBSL ($\sim 1.1-1.5$ atm) are too small to cause bubbles to form spontaneously \cite{brenner2002single}. Instead, a bubble is usually \emph{seeded} somehow. In the original work of Gaitan, an air bubble was injected using a syringe \cite{gaitan1992sonoluminescence}. More recently, seeding methods involve shooting a small jet of water into the flask (like our shrimp friends!) or blasting the water with a laser to boil a small volume \cite{suslick2008inside,yasui2018acoustic}. The later method is preferred as it enables more precise experimental control over the bubble's ambient radius, $R_0$.

Now let's discuss two experimentally accessible quantities that have turned out the be important. (i) The bubble's radius may be measured by \emph{Mie scattering} and compared directly to solutions of the RP equations, allowing us to ``measure" $T(t)$. (ii) The light can be analyzed with a spectrometer, giving us information on the bubble's contents: in principle, we can learn about temperature and pressure inside the bubble \cite{hiller1992spectrum,hilgenfeldt1999simple,flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement}. We look at this in the next section.

\begin{figure}
\includegraphics[width=0.6\linewidth]{figs/mie.pdf}
    \caption{Numerical solution of the $\propto \dot{p}_g$ RP equation variant discussed earlier with the gas dynamics given by the simple form in eq. \ref{eq:p_g}. The dots are experimental results from Mie scattering. The fit resulted in $R_0=4.5$ $\mu m$ and max temperature $8,500$ K. From ref. \cite{lofstedt1993toward}.}
\label{fig:mie}
\end{figure}

In Mie scattering, a laser is shot at a bubble and the scattered intensity is analyzed to determine $R(t)$ \cite{matula1999inertial,lentz1995mie}. Modeling the bubble as a homogeneous dielectric sphere \cite{jackson1999classical}, the angular distribution of light scattered from the bubble at \emph{fixed radius} and the time-dependent scattering off the bubble at \emph{fixed angle} allow use to accurately measure the bubble's radius as a function of time \cite{brenner2002single}. Results from a Mie scattering experiment are shown as dots in Fig. \ref{fig:mie}. Parameters used to calculate the gas dynamics in the bubble are fit to the measured radius to build a full model of the bubble dynamics. For the gas model in the last section, eq. \ref{eq:p_g}, the parameter to fit is $R_0$. Once we fit it, we know the temperature in the bubble from eq. \ref{eq:T(t)}. For the calculation in Fig. \ref{fig:mie}, the dynamics fit the experiment well, giving $R_0=4.5~ \mu$m and putting the ``measured" temperature at $8,500$ K.

Before discussing the light spectrum, we note that it is possible to calculate the temperature from ``first-principles" without fitting to experiment. This can done by direct numerical solution of the gas's Navier-Stokes equations. The results are qualitatively similar to simple model calculations, but are in slightly better quantitative agreement when used to calculate the light spectrum \cite{an2009diagnosing,an2008spectral,an2006mechanism}.

\section{Let there be light!}

Now that we know the temperature in the bubble, $T(t)$, we are finally ready to discuss the light emission. A requirement for models of light emission in SBSL is \emph{local thermodynamic equilibrium} (LTE). LTE is expected to be true in a bubble undergoing SBSL since the particle density ($\sim 10^{28}/m^3$) and temperature ($\sim 10^4$ K) are very large, causing collisions between particles to occur so frequently that equilibrium is guaranteed \cite{brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism}.

Any ordinary matter at finite temperature will continuously and spontaneously emit radiation. In LTE, the rate energy is emitted equals the rate it is absorbed \cite{zel2002physics}. This means that we only need to know \emph{either} the emission coefficient \emph{or} the absorption coefficient of a material to know the intensity that it emits radiation. The intensity, $I(\lambda,T)$, is the amount of energy passing per unit-time and unit-solid angle through a unit-area oriented perpendicular to the light's propagation direction. An ideal emitter/absorber is called a ``black-body."  For a black-body, the intensity is given by Planck's law: \cite{schroeder1999introduction}
\begin{equation}
    I_{B}(\lambda,T(t))=\frac{2 h c^2}{\lambda^5}\left[\exp\left(\frac{hc}{\lambda k_B T(t)}\right)-1\right]^{-1}
\end{equation}
The subscript $B$ is to remind us this is for a black-body. $\lambda$ is the wave-length of the light, $c$ is the speed of light in vacuum (we assume the index of refraction is $n\approx 1$ \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,zel2002physics}), $h$ is Planck's constant, and $k_B$ is Boltzmann's constant. We can use the temperature, $T(t)$, from eq. \ref{eq:T(t)}, to calculate the instantaneous intensity radiated in the bubble. Integrating over the bubble's volume and over time, we can calculate the number of photons produced during SBSL; a quantity that can be directly compared to experiment \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,an2006mechanism,an2008spectral,an2009diagnosing}. Calculations that assume the bubble is a black-body over-estimate the number of photons by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}. We need to use a model that includes finite opacity.

Let's call the wave-length and temperature dependent absorption coefficient $\kappa(\lambda,T)$. For a ray of light emitted somewhere in the bubble, the intensity after travelling a distance $s$ is \cite{zel2002physics,hilgenfeldt1999simple,taylor1969experimental}
\begin{equation}
    I(\lambda,T(t))=I_{B}(\lambda,T(t))  \left[ 1-\exp \left( -\kappa[\lambda,T(t)] s \right) \right].
    \label{eq:intensity}
\end{equation}
We are assuming that that the temperature in the bubble is spatially uniform so that $\kappa$ is too \cite{hilgenfeldt1999sonoluminescence,hilgenfeldt1999simple}. In the limit that $\kappa \rightarrow\infty$, i.e. an opaque bubble, we recover the result for a black-body emitter. On the other hand, for a nearly transparent bubble, $\kappa \ll 1$ and $I\approx I_B \kappa s$. The number of photons emitted is greatly reduced; a result that what we require. A great deal of work has been devoted to identifying the relevant absorption mechanisms leading to $\kappa(\lambda,T)$ \cite{hilgenfeldt1999simple,brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism,flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement,suslick2008inside,an2009diagnosing,an2008spectral,an2006mechanism}.

\begin{figure}
\includegraphics[width=0.9\linewidth]{figs/light.pdf}
    \caption{Simulations of an Ar bubble in 85\% aqueous H$_2$SO$_4$ (top-row) and in water (bottom-row). The gas Navier-Stokes equations were solved directly. The $1^{st}$ column shows the emitted spectrum as a function of wave-length, $\lambda$. The $2^{nd}$ and $3^{rd}$ columns shows the temperature and pressure as a function of position inside the bubble at the time the bubble is compressed to its minimum size. In the top row, the temperature was 20 $^o$C and the bubble was driven at $P_0$=1.5, 1.7, 2.0, and 4.0 atm for curves A, B, C, and D respectively. In the bottom row, A and B are driven at $P_0=$1.22 and 1.32 atm, respectively, at 20 $^o$C. Curve C is at $P_0$=1.32 atm and 0 $^0$C. All data are from ref. \cite{an2009diagnosing}.}
\label{fig:light}
\end{figure}

Let us first discuss SBSL of a \emph{pure} Ar bubble in water. Due to the absence of lines in SBSL spectra, early models only considered \emph{continuous} emission processes \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}. The derivations of the formulae for absorption coefficients are too long to fit in this paper, so we will simply highlight what is important; we include one of them as an example. At the very high temperatures that occur in the bubble ($\sim10,000-30,000$ K, see Fig. \ref{fig:light}), it is possible for electrons to become ionized from their atoms \cite{zel2002physics}. When a free electron passes near an ion, it is deflected by the Coulomb force. The electron is accelerated, changing its kinetic energy. Since the electron is charged, it emits radiation through bremsstrahlung \cite{jackson1999classical,griffiths2005introduction,zel2002physics}. A \emph{classical} expression for absorption due to electron-ion bremsstrahlung is derived in ref. \cite{zel2002physics} and is frequently used in the context of SBSL \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,an2006mechanism,an2008spectral,an2009diagnosing}:
\begin{equation}
    \kappa(\lambda,T(t)) = \frac{4}{3} \left( \frac{2\pi}{3 m_e k_B T(t)} \right)^{1/2} \frac{Z^2 e^6 \lambda^3}{ (4 \pi \epsilon_0)^3 h c^4 m_e } N_+ N_e .
    \label{eq:absorption_coeff}
\end{equation}
In eq. \ref{eq:absorption_coeff}, $m_e$ is the electron rest mass, $e$ is the elementary charge, $Z$ is the ``effective" charge of the ion (usually taken to be $Z=1$), $\epsilon_0$ is the vacuum permittivity, and $N_+$ and $N_e$ are, respectively, the number density of ions and electrons. Usually $N_+=N_e$ and $N_+N_e=N_e^2$. The fact that $\kappa \propto N_e N_+$ is intuitive: the amount of energy emitted/absorbed by electrons scattering off ions ought to be proportional to the number of electron and ions present. $N_e$ is easily calculated from the ``Saha equation" of astrophysics \cite{hilgenfeldt1999simple,an2006mechanism,zel2002physics}. Quantum mechanical calculations of $\kappa$ for electron-ion bremsstrahlung have also been done (see ref. for a review \cite{eckert2015aether}), but the formulae are considerably more complicated and are seldom used in the context of SBSL. If the electron becomes bound instead of being deflected, a different expression for the absorption has to be used \cite{zel2002physics}. The coefficient depends on which Ar atomic level the electron is in, but the free electron energy is continuous and electron-ion \emph{recombination} radiation is continuous, consistent with the continuous spectrum measured in water \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}.

We're not done. It turns out that at the temperatures relevant to SBSL ($\sim 10^4$ K), the fraction of atoms that are ionized is only about $1\%$. The pressures that occur in SBSL ($\sim10^{12}$, see Fig. \ref{fig:light}) are so large that collisions between electrons and neutral atoms (which are otherwise rare) are frequent enough to be as important as electron-ion collisions \cite{zel2002physics,hilgenfeldt1999simple,frommhold1998electron}. A classical formula for the absorption coefficient for electron-atom bremsstrahlung is also available in ref. \cite{zel2002physics}, though more recent works \cite{an2006mechanism,an2008spectral,an2009diagnosing} use numerical results from quantum mechanical pseudopotential calculations \cite{geltman1973free}.

Eq. \ref{eq:absorption_coeff}, combined with analogous formulae for electron-atom bremsstrahlung \cite{geltman1973free} and electron-ion recombination \cite{zel2002physics} were used to accurately calculate the spectrum of SBSL from Ar in water \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,hammer2002light,yasui2001effect,yasui1999mechanism}. The only remaining error was small disagreement in the number of photons emitted \cite{an2006mechanism}, which was later accounted for after SBSL was discovered in aqueous H$_2$SO$_4$ \cite{flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement,suslick2008inside}. This later work discovered that emission \emph{lines} for Ar and Ar$^+$ excited state transitions [$\sim 400$ nm and $\sim 800$ nm, respectively, in Fig. \ref{fig:light} (left)] and continuous emission from electron-OH$^{+}$ recombination are present. However, at the conditions encountered in SBSL in water [Fig. \ref{fig:exp} (right)], the emission lines are concealed by the dominant continuum emission processes at high temperature while the electron-OH$^{+}$ recombination contribution diminishes as the remaining vapor content is driven out of the bubble at high pressure \cite{suslick2008inside,flannigan2005plasma,flannigan2005plasma1,flannigan2006measurement}. More recent calculations of SBSL in both water and aqueous H$_2$SO$_4$ that include all of these mechanisms are in excellent agreement with experiment, as can be seen by comparing Fig. \ref{fig:exp} and Fig. \ref{fig:light}.

\section{Summary and Outlook}
We now have everything we need to describe SBSL will-nilly! We can use the RP equation, eq. \ref{eq:RP_2}, or one of it's variants to calculate the radius time curve, $R(t)$, of a bubble in SBSL. With a suitable model for the gas dynamics in the bubble (e.g. eq. \ref{eq:p_g}) the coupled RP and gas equations are solved numerically and fit to measured $R(t)$. This procedure allows us to ``measure" the temperature in the bubble. With $T(t)$, and knowing the correct mechanisms for emission/absorption in the bubble (which are electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination), we can calculate the radiated intensity as a function of time. The results of this procedure are the wave-length dependence of the emitted radiation, the pulse width (by solving as a function of time), and the number of photons (by integrating over the bubble and over time). The fact that all of these quantities are in excellent agreement with experiments has led to the popular opinion that SBSL is a ``solved" problem \cite{yasui2018acoustic,lohse2018bubble}.

It is interesting to note other distinct models for SBSL. A class of ``fractoluminescence" models have persisted in spite of the success of the work presented in this paper. These are based on the hypothesis that shape instabilities cause the bubble to violently disintegrate, resulting in velocities faster than the fluid can flow \cite{prosperetti1997new}. The fluid ``fractures" and plasma forms during dielectric breakdown across the fracture, emitting light \cite{borisenok2020mechanisms,borissenok2008sonoluminescence}. These models haven't gained much traction, however, since the requirement that the bubble disintegrates contradicts the fact that they don't...

An even more exotic theory supposes SBSL is from ``quantum vacuum radiation" \cite{schwinger1992casimir1,schwinger1992casimir2,schwinger1993casimir1,schwinger1993casimir2,schwinger1993casimir3,schwinger1993casimir4,schwinger1994casimir}. Schwinger proposed that shrinking a cavity would change the number of electromagnetic modes confined in it, changing the energy expectation value; the lost energy would be carried away by photons. Direct calculation of the number of photons was carried out by others \cite{eberlein1996sonoluminescence}. It was found that the number of emitted photons would be $10^{-10}$ times too small \cite{unnikrishnan1996comment,lambrecht1997comment,garcia1997comment,milton1996comment,liberati2000sonoluminescence}. The quantum vacuum radiation theory has been abandoned.

The theory discussed in this paper most likely explains \emph{shrimpoluminescence}, though some discrepancies should be addressed. The foremost being that shrimpoluminescence is too dim to be seen by the naked eye \cite{lohse2001snapping},  while SBSL in the lab is not. Recall that shrimpoluminescence is transient, while the theory developed in this paper is for stable bubbles. For a transient bubble, it is not possible for air molecules or water vapor to be ejected from the bubble, a process which occurs over $\sim10^3$ cycles. These extra degrees of freedom prevent the temperature from becoming as large as in \emph{stable} SBSL and the amount of emitted light is suppressed.
Another issue is that the bubble produced by a snapping shrimp is not stationary \cite{lohse1997sonoluminescing,versluis2000snapping}. Translating bubbles are not quite spherical \cite{brenner2002single}, and small shape distortions could decrease the peak pressure and temperature in the bubble, suppressing the amount of light produced \cite{matula2000single}. Still, the essential physics we have learned from SBSL in this paper at least gives us the tools we need to \emph{shine light} on shrimpoluminescence.

%\raggedbottom

\bibliography{ref}

\end{document}

### 2022-10-04 bak/dft/notes/notes.tex
asdasd

--- 2021-08-07 dft/notes/notes.tex
@@
 
+There are apparently many different numerical integration methods for solving the radial Schr\"odinger's equation(let's call it the SE from now).
+Generally, these methods assume that space can be discretized into a finite grid and, assuming initial conditions at a few points, the solution can be solved by integrating across the domain.
+The method I keep coming across is \emph{Numerov's method}, which is allegedly pretty accurate while being very simple.
+Another one I have seen is the \emph{Runge-Kutta method} (well actually it's a class of methods) that is attractive because the discretized grid can be adjusted willy-nilly.
+This is an example of an 'adaptive grid' that can be adjusted on-the-fly to be fine in regions where the solutions vary rapidly, but coarse in regions where the solutions are smooth.
+
+Numerov's method is applicate to (among others) differential equations of the form 
+\begin{equation}
+    \frac{d^2y}{dx^2} = -g(x)y(x)+s(x)
+\end{equation}
+The SE is an example of this with $s(x)=0$ and $g(x)=2m/\hbar^2 \left[ E-V(x) \right]$.
+For the radial equation in 3D, $V(x)=V^l_{eff}(x)$, but this is valid for any 1D SE with well behaved $V$.
+Poisson's equation from electricity and magnetism is another example: $d^2\phi(x)/dx^2=-4\pi\rho(x)$ with $\phi$ the electrostatic potential and $\rho$ the charge density.
+In this case, $s(x)=-4\pi\rho(x)$ and $g(x)=0$.
+
+Suppose our differential equation is defined over some finite regions $x_{min}\leq x \leq x_{max}$.
+We further assume that the solution decays to negligibly small at $x_{max}$.
+We then divide the interval into $N$ points of equal spacing that are $\Delta x$ wide (it is possible and even useful to use \emph{unevenly} spaced grids, but we will get to that later).
+We label each point $x_i$ and the corresponding solution etc.
+by $y_i=y(x_i)$, $g_i=g(x_i)$, and $s_i=s(x_i)$.
+Then to discretize the differential equation, we Taylor expand the solution around a particular point $x_n$:
+\begin{equation}
+    y_{n-1}=y_n-y^{(1)}_n\Delta x + \frac{1}{2} y^{(2)}_n(\Delta x)^2 - \frac{1}{6} y^{(3)}_n(\Delta x)^3 + \frac{1}{24} y^{(4)}_n(\Delta x)^4 - \frac{1}{120} y^{(5)}_n(\Delta x)^5+\mathcal{O}\left[ (\Delta x)^6 \right]
+\end{equation}
+\begin{equation}
+    y_{n+1}=y_n+y^{(1)}_n\Delta x - \frac{1}{2} y^{(2)}_n(\Delta x)^2 + \frac{1}{6} y^{(3)}_n(\Delta x)^3 - \frac{1}{24} y^{(4)}_n(\Delta x)^4 + \frac{1}{120} y^{(5)}_n(\Delta x)^5+\mathcal{O}\left[ (\Delta x)^6 \right]
+\end{equation}
+Adding $y_{n+1}+y_{n-1}$, we get 
+\begin{equation}
+    y_{n+1}+y_{n-1}=2y_n+y^{(2)}(\Delta x)^2 + \frac{1}{12} y^{(4)}_n(\Delta x)^4 + \mathcal{O}\left[ (\Delta x)^6 \right]
+    \label{eq:numerov_1}
+\end{equation}
+But we know that $y^{(2)}_n=-g_ny_n+s_n\equiv z_n$. Note that eq. \ref{eq:numerov_1} is applicatble to $z_n$ too if replace $y$ by $z$. If we plug $z_n$ into \ref{eq:numerov_1}, we get:
+\begin{equation}
+    y^{(4)}_n=\equiv z^{(2)}_n = \frac{z_{n+1}+z_{n-1}-2z_n}{(\Delta x)^2}+\mathcal{O}\left[ (\Delta x)^2 \right]
+\end{equation}
+
+Plugging this back into \ref{eq:numerov_1}, we get 
+\begin{multline}
+    y_{n+1}=2y_n - y_{n-1} + (-g_n y_n+s_n)(\Delta x)^2 + \\ \frac{1}{12} (-g_{n+1}y_{n+1} + s_{n+1} - g_{n-1}y_{n-1}+s_{n-1}+2g_n y_n -2s_n)(\Delta x)^2 + \mathcal{O}[(\Delta x)^6]
+\end{multline}
+and after some manipulation we arrive at \emph{Numerov's formula}:
+\begin{multline}
+    y_{n+1}\left[1+g_{n+1}\frac{(\Delta x)^2}{12} \right]= 2y_n \left[ 1 - 5 g_n \frac{(\Delta x)^2}{12} \right] - y_{n-1} \left[1+g_{n-1}\frac{(\Delta x)^2}{12} \right] + \\ (s_{n+1}+10 s_n +s_{n-1})\frac{(\Delta x)^2}{12}+\mathcal{O}[(\Delta x)^6]
+    \label{eq:numerov}
+\end{multline}
+
+Apparently the error in our integration is order $\mathcal{O}[(\Delta x)^6]$ in the step size.
+Staring at this equation for little while, we see that if we know $y_n$ and $y_n+1$, we can iteratively determine the solution at other $y_i$.
+Importantly, we can integrate in both the positive and negative $x$ directions.
+Recall that in the SE, the $s_n$ terms are missing and Numerov's formula is somewhat simpler.
+
 \subsection{Further reading}
@@
 \begin{itemize}
+  \item \emph{Numerical Methods in Quantum Mechanics (2019/2020)} by Paolo Gianozzi
   \item \emph{Introduction to Quantum Mechanics} by David Griffiths

### 2022-10-04 tinydft/conftest.py

@@ -1,38 +1,0 @@
-#!/usr/bin/env python3
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Configuration of pytest."""
-
-
-import pytest
-
-from tinygrid import setup_grid
-from tinybasis import Basis
-
-
-@pytest.fixture(scope="session")
-def grid_basis():
-    """Set up default integration grid and basis set.
-
-    This fixture makes it possible to initiliase and precompute these objects
-    only once for all tests.
-    """
-    grid = setup_grid()
-    basis = Basis(grid)
-    return grid, basis

### 2022-10-04 tinydft/program_evn.py

@@ -1,57 +1,0 @@
-#!/usr/bin/env python3
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Plot the energy as function of number of electrons for Carbon: N = 0 to 6."""
-
-import matplotlib.pyplot as plt
-
-from tinybasis import Basis
-from tinydft import scf_atom
-from tinygrid import setup_grid
-from program_mendelejev import klechkowski, interpret_econf
-
-
-def main():
-    """Make a plot of the energy of Carbon as function of the number of electrons."""
-    # Compute the energies.
-    atnum = 6
-    grid = setup_grid()
-    basis = Basis(grid)
-    energies_scan = [0.0]
-    for nelec in range(1, 7):
-        econf = klechkowski(nelec)
-        occups = interpret_econf(econf)
-        energies, _rhos = scf_atom(atnum, occups, grid, basis)
-        energies_scan.append(energies[0])
-
-    # Make a nice plot.
-    plt.rcParams['font.sans-serif'] = ["Fira Sans"]
-    fig, ax = plt.subplots(figsize=(4, 2.25))
-    ax.plot(energies_scan, "o")
-    ax.set_xlabel("Number of electrons")
-    ax.set_ylabel(r"Electronic energy [$\mathrm{E_h}$]")
-    ax.set_title("Carbon (HFS functional, TinyDFT)")
-    ax.spines['right'].set_visible(False)
-    ax.spines['top'].set_visible(False)
-    fig.tight_layout(pad=0.3)
-    fig.savefig("energy_verus_nelec_carbon.png", dpi=300)
-
-
-if __name__ == '__main__':
-    main()

### 2022-10-04 tinydft/test_tinybasis.py

@@ -1,71 +1,0 @@
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Unit tests for Tiny DFT."""
-
-import numpy as np
-from numpy.testing import assert_allclose
-from scipy.linalg import eigh
-import pytest
-
-from tinybasis import Basis
-from tinygrid import setup_grid
-from test_tinygrid import get_hydrogenic_solutions
-
-
-@pytest.mark.parametrize("atnum", [1, 11, 21, 31, 41, 51, 61, 71, 81, 91, 101, 111])
-@pytest.mark.parametrize("angqn", [0, 1, 2, 3, 4, 5, 6])
-def test_hydrogenic_op(atnum, angqn, grid_basis):
-    grid, basis = grid_basis
-    # Create modified copies of the operators. Do not modify in place to
-    # avoid side-effects.
-    kin = basis.kin_rad.copy()
-    if angqn > 0:
-        angmom_factor = (angqn * (angqn + 1)) / 2
-        kin += basis.kin_ang * angmom_factor
-    ext = basis.ext * atnum
-    evals, evecs = eigh(kin + ext, basis.olp)
-    psis = get_hydrogenic_solutions(grid, atnum, angqn)
-    for i, (priqn, factor, psi) in enumerate(psis):
-        case = "i={} priqn={}".format(i, priqn)
-        dot = abs(grid.integrate(psi, np.dot(evecs[:, i], basis.fnvals)))
-        norm = np.einsum('i,ij,j', evecs[:, i], basis.olp, evecs[:, i])
[…]

### 2022-10-04 tinydft/test_tinydft.py

@@ -1,105 +1,0 @@
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Unit tests for Tiny DFT."""
-
-import numpy as np
-from numpy.testing import assert_allclose
-import pytest
-
-from tinydft import scf_atom, build_rho, solve_poisson
-from tinygrid import setup_grid
-from program_mendelejev import char2angqn, klechkowski, interpret_econf
-
-
-def test_char2angqn():
-    assert char2angqn('s') == 0
-    assert char2angqn('S') == 0
-    assert char2angqn('p') == 1
-    assert char2angqn('P') == 1
-    assert char2angqn('d') == 2
-    assert char2angqn('D') == 2
-
-
-@pytest.mark.parametrize("atnum", [1, 11, 21, 31, 41, 51, 61, 71, 81, 91, 101, 111])
-def test_poisson(atnum):
-    # Numerically solve the electrostatic potential of an S-type Slater Density
-    grid = setup_grid()
-    alpha = 2 * atnum
-    aux = np.exp(-alpha * grid.points)
-    rho = aux * alpha**3 / (8 * np.pi)
-    assert_allclose(grid.integrate(4 * np.pi * grid.points**2 * rho), 1.0, atol=1e-11, rtol=0)
-    vnum = solve_poisson(grid, rho)
-    vann = (1 - aux) / grid.points - alpha * aux / 2
-    assert_allclose(vnum[-1], 1.0 / grid.points[-1], atol=1e-11, rtol=0)
[…]

### 2022-10-04 tinydft/test_tinygrid.py

@@ -1,242 +1,0 @@
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Unit tests for Tiny Grid."""
-
-import numpy as np
-from numpy.testing import assert_allclose
-from numpy.polynomial.legendre import legval
-import pytest
-from scipy.special import eval_genlaguerre
-
-from tinygrid import LegendreGrid, TransformedGrid
-
-
-@pytest.mark.parametrize("npoint", range(10, 20))
-def test_low_grid_basics(npoint):
-    grid = LegendreGrid(npoint)
-    assert grid.basis.shape == (npoint, npoint)
-    assert grid.basis_inv.shape == (npoint, npoint)
-    fn1 = np.random.uniform(0, 1, npoint)
-    coeffs = grid.tocoeffs(fn1)
-    fn2 = grid.tofnvals(coeffs)
-    assert_allclose(fn1, fn2, atol=1e-14)
-    fn3 = legval(grid.points, coeffs)
-    assert_allclose(fn1, fn3, atol=1e-14)
-
-
-def test_low_grid_basics_vectorized():
-    shape = (5, 4, 8)
-    npoint = 7
-    extshape = shape + (npoint, )
-    grid = LegendreGrid(npoint)
-    fn1 = np.random.uniform(0, 1, extshape)
-    coeffs = grid.tocoeffs(fn1)
-    assert coeffs.shape == extshape
-    fn2 = grid.tofnvals(coeffs)
-    assert fn2.shape == extshape
[…]

### 2022-10-04 tinydft/tinybasis.py

@@ -1,137 +1,0 @@
-#!/usr/bin/env python3
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Radial orbital basis of Gaussian primitives and their integrals for TinyDFT."""
-
-
-from functools import wraps
-
-import numpy as np
-from numpy.testing import assert_allclose
-
-from tinygrid import TransformedGrid
-
-
-def memoize(method):
-    """Wrap a function such that its result is computed only once.
-
-    See https://en.wikipedia.org/wiki/Memoization
-    """
-    @wraps(method)
-    def wrapper(obj) -> np.ndarray:
-        attrname = '_' + method.__name__
-        result = getattr(obj, attrname, None)
-        if result is None:
-            result = method(obj)
-        setattr(obj, attrname, result)
-        return result
-    return wrapper
-
-
-class Basis:
-    """Gaussian radial orbital basis set, precomputed integrals and integral evaluation.
-
-    Results of integrals that take no arguments (overlap, kin_rad, kin_ang and
-    ext) are memoized. After construction, it is assumed that the basis
-    functions are not modified.
-
-    Attributes
-    ----------
-    grid
-        The radial numerical integration grid.
-    alphas
-        Gaussian exponents.
-    fnvals
-        Normalized basis functions evaluated on the grid.
-    normalizations
-        Vector with normalization constants for the Gaussians.
-    olp
-        Overlap matrix
-    kin_rad
-        Radial kinetic energy operator.
-    kin_ang
-        Angular kinetic energy operator for angqn=1.
-    ext
[…]

### 2022-10-04 tinydft/tinydft.py

@@ -1,220 +1,0 @@
-#!/usr/bin/env python3
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Tiny DFT is a minimalistic atomic DFT implementation.
-
-It only supports closed-shell calculations with pure local (not even semi-local)
-functionals. One has to fix the occupations numbers of the atomic orbitals a
-priori.
-
-Atomic units are used throughout.
-"""
-
-from typing import List, Tuple
-
-import numpy as np
-from numpy.testing import assert_allclose
-from scipy.linalg import eigh
-
-from autograd import elementwise_grad
-import autograd.numpy as agnp
-
-from tinybasis import Basis
-from tinygrid import TransformedGrid
-
-
-__all__ = ['scf_atom', 'solve_poisson', 'xcfunctional']
-
-
-# pylint: disable=too-many-statements
-def scf_atom(atnum: float, occups: List[np.ndarray], grid: TransformedGrid,
-             basis: Basis, nscf: int = 25, mixing: float = 0.5) \
-        -> Tuple[np.ndarray, List[Tuple[np.ndarray, np.ndarray]]]:
-    """Perform a self-consistent field atomic calculation.
-
-    Parameters
-    ----------
-    atnum
-        The nuclear charge.
-    occups
-        Occupation numbers, see interpret_econf.
-    grid
-        A radial grid.
-    basis
-        The radial orbital basis.
-    nscf
-        The number of SCF cycles.
-    mixing
-        The SCF mixing parameter. 1 means the old Fock operator is not mixed in.
-
-    Returns
-    -------
-    energies
-        The atomic energy and its contributions.
-    eps_orbs_u
[…]

### 2022-10-04 tinydft/tinygrid.py

@@ -1,162 +1,0 @@
-# Tiny DFT is a minimalistic atomic DFT implementation.
-# Copyright (C) 2019 The Tiny DFT Development Team
-#
-# This file is part of Tiny DFT.
-#
-# Tiny DFT is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 3
-# of the License, or (at your option) any later version.
-#
-# Tiny DFT is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, see <http://www.gnu.org/licenses/>
-# --
-"""Tiny numerical integration library for Tiny DFT."""
-
-from typing import List, Callable
-
-import numpy as np
-from numpy.polynomial.legendre import legvander, legder, legint, leggauss
-
-from autograd import elementwise_grad
-import autograd.numpy as agnp
-
-
-__all__ = ['LegendreGrid', 'TransformedGrid', 'setup_grid']
-
-
-class BaseGrid:
-    """Basic grid API, only for (vectorized) definite integrals."""
-
-    def __init__(self, points: np.ndarray, weights: np.ndarray):
-        self.points = points
-        self.weights = weights
-
-    def integrate(self, *multi_fnvals: np.ndarray) -> np.ndarray:
-        """Compute the definite integrals (of products) of functions.
-
-        Parameters
-        ----------
-        fnvals1, fnvals2, ...
-            Each argument is an N-dimensional array of function values, where
-            the last index corresponds to the grid points. The arrays are
-            contracted over their last index and the resulting functions are
-            all integrated. This allows many products of functions to be
-            integrated in one shot.
-
-        Returns
-        -------
-        integrals
-            All integrals of products of functions. The shape is
-            ``sum(fnvals.shape[:-1] for fnvals in multi_fnvals)``.
-
-        """
-        args = [self.weights, [0]]
-        sublistout: List[int] = []
-        counter = 1
-        for fnvals in multi_fnvals:
-            args.append(fnvals)
-            nkeep = fnvals.ndim - 1
-            sublist = list(range(counter, counter + nkeep)) + [0]
[…]
